### 2010-09-13 paper/main.tex
Initial commit of paper files

\documentclass[english,12pt,a4paper]{report}
%[twoside] ?
\usepackage[T1]{fontenc} % In case we want special characters
\usepackage[utf8]{inputenc} % We are all writing in UTF-8

\usepackage{graphicx} % We _may_ use images
% Lager automagisk interne lenker i dokumentet.
\usepackage{hyperref} % Internal links is nice
\hypersetup{pdfborder=0 0 0} % ..especially without red borders

\author{Eirik Haver and Pål Ruud}
\title{Project assignment - Tahoe-LAFS with SHA-3 candidates}
\date{\today}

\begin{document}

\include{title}

\section*{Summary}

\section*{Foreword}

\tableofcontents
\newpage

\listoftables
\listoffigures

% Acronyms?

\chapter{SHA-3 in Tahoe-LAFS}

\section{Introduction}

\section{The candiates}

\end{document}

### 2010-09-13 paper/title.tex
Initial commit of paper files

% Latex-versjon av ITEM rapportmal.
% Lagd av <lasse.karstensen@gmail.com>, desember 2009.
% Lisens: public domain. 
%
\begin{titlepage}
\begin{center}
\textsc{NORWEGIAN UNIVERSITY OF SCIENCE AND TECHNOLOGY\\
FACULTY OF  INFORMATION TECHNOLOGY, MATHEMATICS AND ELECTRICAL ENGINEERING} \\
\vspace{0.5cm} 
% crop-et fra http://www.ntnu.no/infoavdelingen/selvhjelp/logoer/ntnu/NTNU_engelsk_RGB.png
\includegraphics[scale=0.5]{NTNU-logo} \\

\vspace{1.0cm}
{\Huge{PROJECT ASSIGNMENT}}
\vspace{1.0cm}

\begin{tabular}{ p{4cm} p{11cm}}

Student's name:	& Eirik Haver and Pål Ruud \\
Course: & TTM13 \\
Title: & A somewhat long and describing title \\
%\vspace{1cm}
Description: & \\
\end{tabular}
{\small{\begin{tabular}{p{15cm}}
\vspace{0.2cm}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nunc nibh quam, posuere quis dignissim vel, pulvinar vel leo. Morbi pellentesque est a magna rutrum sollicitudin. Duis nec orci porttitor tellus viverra aliquet gravida nec tellus. Aenean aliquet nulla sit amet metus sodales mattis. In pulvinar euismod consectetur. Maecenas felis nunc, suscipit ut semper et, consequat et est. 
\\\\
Donec vitae ipsum tortor. Nulla at tincidunt erat. Nunc vel risus vel mi ornare molestie non at nunc. Phasellus posuere cursus odio, nec consectetur dolor tempus quis. Proin non erat erat. Aenean et leo eros, et congue sem. Nunc auctor porta magna, ac fermentum augue malesuada vel. Nam aliquet augue eu nibh facilisis ac placerat felis vehicula. Quisque faucibus placerat vulputate. Mauris nunc turpis, venenatis ac tincidunt condimentum, tincidunt id magna. Mauris aliquet convallis accumsan. 
\\\\
\end{tabular}  }}

\begin{tabular}{ p{4cm} p{11cm}}
Deadline: & 2010-12-xx \\
Submission date: & 2010-12-xx \\
Department: & Department of Telematics \\
Supervisor: & Danilo Gligoroski \\\\
\end{tabular}
\vspace{0.5cm}

Trondheim, \today

\vspace{0.4cm}
\line(1,0){150} \\
Danilo Gligoroski, NTNU/ITEM.

\end{center}
\end{titlepage}

### 2010-09-20 python-bindings/bmw.pyx
Added digest function

@@ -62,11 +62,18 @@ cdef class bmw256:
 
         return s
 
+    def digest(self):
+        if not self.finished:
+            self.final()
+
+        digest = [self.hashval[i] for i from 0 <= i < self.hashbitlen / 8]
+        return ''.join(map(chr, digest))
+
     def hexdigest(self):
         if not self.finished:
             self.final()
 
         ret = [self.hashval[i] for i from 0 <= i < self.hashbitlen / 8]
 
-        # Return a hex string
+        # Return a hex string using str format specification
         return ''.join(['%02x' % i for i in ret])

### 2010-09-20 paper/problem_description.tex
Updated problem description

\documentclass[english,a4paper,12pt,titlepage,twoside]{report}
\usepackage[T1]{fontenc} % In case we want special characters
\usepackage[utf8]{inputenc} % We are all writing in UTF-8
%\usepackage{babel,color,cite}
\usepackage{babel}
\topmargin=5.mm
\oddsidemargin=0.mm
\evensidemargin=0.mm
\textheight=220.mm
\textwidth=170.mm

\usepackage{graphicx}

\begin{document}
% Latex-versjon av ITEM rapportmal.
% Lagd av <lasse.karstensen@gmail.com>, desember 2009.
% Lisens: public domain.
%
\begin{titlepage}
\begin{center}
\textsc{NORWEGIAN UNIVERSITY OF SCIENCE AND TECHNOLOGY\\
FACULTY OF  INFORMATION TECHNOLOGY, MATHEMATICS AND ELECTRICAL ENGINEERING} \\
\vspace{0.5cm}
% crop-et fra http://www.ntnu.no/infoavdelingen/selvhjelp/logoer/ntnu/NTNU_engelsk_RGB.png
\includegraphics[scale=0.5]{NTNU-logo} \\

\vspace{1.0cm}
{\Huge{PROJECT ASSIGNMENT}}
\vspace{1.0cm}

\begin{tabular}{ p{4cm} p{11cm}}

Students' name:	& Eirik Haver and Pål Ruud \\
Course: & TTM13 \\
Title: & Experimenting with SHA-3 candidates in Tahoe-LAFS \\
%\vspace{1cm}
Description: & \\
\end{tabular}
{\small{\begin{tabular}{p{15cm}}
\vspace{0.2cm}
Tahoe-LAFS is a Free Software/Open Source decentralized data store. It
distributes your filesystem across multiple servers, and even if some of the
servers fail or are taken over by an attacker, the entire filesystem continues
to work correctly and to preserve your privacy and security.
\\\\
One of the basic security components used in Tahoe-LAFS is the cryptographic
hash function SHA-256.
\\\\
In the light of the worldwide SHA-3 hash competition, this task is about
making a reproducible, automated benchmark which shows how the performance of
Tahoe-LAFS is affected by the performance of the different SHA-3 candidate hash
functions. Before any testing can be done, Python bindings to the C
implementations of the SHA-3 candidates have to be made, since Tahoe-LAFS is
written in the Python programming language.
\\\\
\end{tabular}  }}

\begin{tabular}{ p{4cm} p{11cm}}
Deadline: & 2010-12-20 \\
Submission date: & 2010-12-xx \\
Department: & Department of Telematics \\
Supervisor: & Danilo Gligoroski \\\\
\end{tabular}
\vspace{0.5cm}

Trondheim, \today

\vspace{0.4cm}
\line(1,0){150} \\
Danilo Gligoroski, NTNU/ITEM.

\end{center}
\end{titlepage}
\end{document}

### 2010-09-20 paper/simple_problem_desc.tex
Updated problem description

--- 2010-09-20 paper/problem_description.tex
@@
-\documentclass[english,a4paper,12pt,titlepage,twoside]{report}
+\documentclass[english,12pt,a4paper]{report}
 \usepackage[T1]{fontenc} % In case we want special characters
 \usepackage[utf8]{inputenc} % We are all writing in UTF-8
-%\usepackage{babel,color,cite}
-\usepackage{babel}
-\topmargin=5.mm
-\oddsidemargin=0.mm
-\evensidemargin=0.mm
-\textheight=220.mm
-\textwidth=170.mm
 
-\usepackage{graphicx}
+\author{Eirik Haver \and Pål Ruud}
+\title{Project assignment - Tahoe-LAFS with SHA-3 candidates}
+\date{\today}
 
 \begin{document}
-% Latex-versjon av ITEM rapportmal.
-% Lagd av <lasse.karstensen@gmail.com>, desember 2009.
-% Lisens: public domain.
-%
-\begin{titlepage}
-\begin{center}
-\textsc{NORWEGIAN UNIVERSITY OF SCIENCE AND TECHNOLOGY\\
-FACULTY OF  INFORMATION TECHNOLOGY, MATHEMATICS AND ELECTRICAL ENGINEERING} \\
-\vspace{0.5cm}
-% crop-et fra http://www.ntnu.no/infoavdelingen/selvhjelp/logoer/ntnu/NTNU_engelsk_RGB.png
-\includegraphics[scale=0.5]{NTNU-logo} \\
 
-\vspace{1.0cm}
-{\Huge{PROJECT ASSIGNMENT}}
-\vspace{1.0cm}
+\chapter*{Problem Description}
 
-\begin{tabular}{ p{4cm} p{11cm}}
-
-Students' name:	& Eirik Haver and Pål Ruud \\
-Course: & TTM13 \\
-Title: & Experimenting with SHA-3 candidates in Tahoe-LAFS \\
-%\vspace{1cm}
-Description: & \\
-\end{tabular}
-{\small{\begin{tabular}{p{15cm}}
-\vspace{0.2cm}
 Tahoe-LAFS is a Free Software/Open Source decentralized data store. It
@@
 to work correctly and to preserve your privacy and security.
-\\\\
+
 One of the basic security components used in Tahoe-LAFS is the cryptographic
 hash function SHA-256.
-\\\\
+
 In the light of the worldwide SHA-3 hash competition, this task is about
@@
 written in the Python programming language.
-\\\\
-\end{tabular}  }}
 
-\begin{tabular}{ p{4cm} p{11cm}}
-Deadline: & 2010-12-20 \\
-Submission date: & 2010-12-xx \\
-Department: & Department of Telematics \\
-Supervisor: & Danilo Gligoroski \\\\
-\end{tabular}
-\vspace{0.5cm}
+\vspace{10mm}
+Responsible professor: Danilo Gligoroski danilog@item.ntnu.no
 
-Trondheim, \today
-
-\vspace{0.4cm}
-\line(1,0){150} \\
-Danilo Gligoroski, NTNU/ITEM.
-
-\end{center}
-\end{titlepage}
 \end{document}

### 2010-09-20 paper/title.tex
Added more layout, and a problem description
Updated problem description

@@
 
-Student's name:	& Eirik Haver and Pål Ruud \\
+Students' name:	& Eirik Haver and Pål Ruud \\
 Course: & TTM13 \\
-Title: & A somewhat long and describing title \\
+Title: & Experimenting with SHA-3 candidates in Tahoe-LAFS \\
 %\vspace{1cm}
@@
 \vspace{0.2cm}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nunc nibh quam, posuere quis dignissim vel, pulvinar vel leo.
-Morbi pellentesque est a magna rutrum sollicitudin.
-Duis nec orci porttitor tellus viverra aliquet gravida nec tellus.
-Aenean aliquet nulla sit amet metus sodales mattis.
-In pulvinar euismod consectetur.
-Maecenas felis nunc, suscipit ut semper et, consequat et est.
+
+Tahoe-LAFS is a Free Software/Open Source decentralized data store. It
+distributes your filesystem across multiple servers, and even if some of the
+servers fail or are taken over by an attacker, the entire filesystem continues
+to work correctly and to preserve your privacy and security.
 \\\\
-Donec vitae ipsum tortor.
-Nulla at tincidunt erat.
-Nunc vel risus vel mi ornare molestie non at nunc.
-Phasellus posuere cursus odio, nec consectetur dolor tempus quis.
-Proin non erat erat.
-Aenean et leo eros, et congue sem.
-Nunc auctor porta magna, ac fermentum augue malesuada vel.
-Nam aliquet augue eu nibh facilisis ac placerat felis vehicula.
-Quisque faucibus placerat vulputate.
-Mauris nunc turpis, venenatis ac tincidunt condimentum, tincidunt id magna.
-Mauris aliquet convallis accumsan.
+One of the basic security components used in Tahoe-LAFS is the cryptographic
+hash function SHA-256.
+\\\\
+In the light of the worldwide SHA-3 hash competition, this task is about
+making a reproducible, automated benchmark which shows how the performance of
+Tahoe-LAFS is affected by the performance of the different SHA-3 candidate hash
+functions. Before any testing can be done, Python bindings to the C
+implementations of the SHA-3 candidates have to be made, since Tahoe-LAFS is
+written in the Python programming language.
 \\\\

### 2010-09-22 paper/problem_description.tex
Updated BibTeX use and the Makefile

@@
 Deadline: & 2010-12-20 \\
-Submission date: & 2010-12-xx \\
 Department: & Department of Telematics \\

### 2010-10-06 python-bindings/setup.py
added 32-bit optimized blake-sha3. NIST round2 version

@@ -20,6 +20,8 @@ ext_modules = [
         [hashpath+"groestl/groestl.pyx", hashpath+'groestl/Groestl-opt.c'],),
     Extension("sha3lib.hash_functions.echo.echo",
         [hashpath+"echo/echo.pyx", hashpath+'echo/echo32.c'],),
+    Extension("sha3lib.hash_functions.blake.blake",
+        [hashpath+"blake/blake.pyx", hashpath+'blake/blake_opt32.c'],),
     ]
 
 class RunTests(Command):
@@ -99,6 +101,7 @@ setup(
     ext_modules = ext_modules,
     packages =
     ['sha3lib','sha3lib.hash_functions','sha3lib.hash_functions.bmw',
-        'sha3lib.hash_functions.echo', 'sha3lib.hash_functions.groestl'],
+        'sha3lib.hash_functions.echo', 'sha3lib.hash_functions.groestl',
+        'sha3lib.hash_functions.blake'],
     #py_modules = ['sha3lib'],    
 )

### 2010-10-06 python-bindings/sha3lib/__init__.py

@@ -1,1 +1,1 @@
-from _sha3lib import bmw256,groestl256,echo256
+from _sha3lib import bmw256,groestl256,echo256,blake256

### 2010-10-06 python-bindings/sha3lib/_sha3lib.py

@@ -1,5 +1,5 @@
 # coding: utf-8
-from hash_functions import bmw,groestl,echo
+from hash_functions import bmw,groestl,echo,blake
 
 class bmw256(bmw):
     def __init__(self, initial=None):
@@ -12,3 +12,7 @@ class groestl256(groestl):
 class echo256(echo):
     def __init__(self, initial=None):
         super(echo256, self).__init__(256, initial)
+
+class blake256(blake):
+    def __init__(self, initial=None):
+        super(blake256, self).__init__(256, initial)

### 2010-10-06 python-bindings/sha3lib/hash_functions/__init__.py

@@ -1,9 +1,11 @@
 from bmw.bmw import bmw
 from groestl.groestl import groestl
 from echo.echo import echo
+from blake.blake import blake
 
 from bmw.bmw import hash as bmw256hash
 from groestl.groestl import hash as groestl256hash
 from echo.echo import hash as echo256hash
+from blake.blake import hash as blake256hash
 
-HASHES = ('bmw', 'echo', 'groestl',)
+HASHES = ('bmw', 'echo', 'groestl', 'blake')

### 2010-10-06 python-bindings/sha3lib/hash_functions/blake/blake.c

@@ -1,0 +1,2702 @@
+/* Generated by Cython 0.13 on Wed Oct  6 12:16:19 2010 */
+
+#define PY_SSIZE_T_CLEAN
+#include "Python.h"
+#ifndef Py_PYTHON_H
+    #error Python headers needed to compile C extensions, please install development version of Python.
+#else
+
+#include <stddef.h> /* For offsetof */
+#ifndef offsetof
+#define offsetof(type, member) ( (size_t) & ((type*)0) -> member )
+#endif
+
+#if !defined(WIN32) && !defined(MS_WINDOWS)
+  #ifndef __stdcall
+    #define __stdcall
+  #endif
+  #ifndef __cdecl
+    #define __cdecl
+  #endif
+  #ifndef __fastcall
+    #define __fastcall
+  #endif
+#endif
+
+#ifndef DL_IMPORT
+  #define DL_IMPORT(t) t
+#endif
+#ifndef DL_EXPORT
+  #define DL_EXPORT(t) t
+#endif
+
+#ifndef PY_LONG_LONG
+  #define PY_LONG_LONG LONG_LONG
+#endif
+
+#if PY_VERSION_HEX < 0x02040000
+  #define METH_COEXIST 0
+  #define PyDict_CheckExact(op) (Py_TYPE(op) == &PyDict_Type)
+  #define PyDict_Contains(d,o)   PySequence_Contains(d,o)
+#endif
+
+#if PY_VERSION_HEX < 0x02050000
+  typedef int Py_ssize_t;
+  #define PY_SSIZE_T_MAX INT_MAX
+  #define PY_SSIZE_T_MIN INT_MIN
+  #define PY_FORMAT_SIZE_T ""
+  #define PyInt_FromSsize_t(z) PyInt_FromLong(z)
+  #define PyInt_AsSsize_t(o)   PyInt_AsLong(o)
+  #define PyNumber_Index(o)    PyNumber_Int(o)
+  #define PyIndex_Check(o)     PyNumber_Check(o)
+  #define PyErr_WarnEx(category, message, stacklevel) PyErr_Warn(category, message)
+#endif
+
+#if PY_VERSION_HEX < 0x02060000
+  #define Py_REFCNT(ob) (((PyObject*)(ob))->ob_refcnt)
+  #define Py_TYPE(ob)   (((PyObject*)(ob))->ob_type)
+  #define Py_SIZE(ob)   (((PyVarObject*)(ob))->ob_size)
+  #define PyVarObject_HEAD_INIT(type, size) \
[…]

### 2010-10-06 python-bindings/sha3lib/hash_functions/blake/blake.pyx

@@ -1,0 +1,87 @@
+# encoding: utf-8
+cimport blake_h
+cimport cython
+
+from libc.stdlib cimport *
+
+def hash(int hashbitlen, bytes data, int in_length):
+    '''
+    Used for testing purposes.
+    '''
+    cdef char* c_string = <char *> data
+    cdef blake_h.DataLength datalen = in_length
+    cdef blake_h.BitSequence *hashval = <blake_h.BitSequence *> malloc(hashbitlen*8)
+
+    blake_h.Hash(hashbitlen, <blake_h.BitSequence *> c_string, datalen, hashval)
+
+    try:
+        digest = [hashval[i] for i from 0 <= i < hashbitlen / 8]
+        digest = ''.join(['%02X' % i for i in digest])
+    finally:
+        free(hashval)
+
+    return digest
+
+cdef class blake:
+    '''
+    A class that tries to mimic the behaviour of hashlib, i.e. keeping state
+    so that one can update the hashing procedure instead of doing it from
+    scratch.
+    '''
+    cdef blake_h.BitSequence *hashval
+    cdef blake_h.hashState previous_state
+    cdef blake_h.hashState state
+    cdef int finished
+    cdef int hashbitlen
+
+    def __init__(self, int in_hashbitlen, bytes initial=None):
+        self.finished = 0
+        self.hashbitlen = in_hashbitlen
+        blake_h.Init(&self.state, self.hashbitlen)
+
+        if initial:
+            self.update(initial)
+
+    cpdef update(self, bytes in_data):
+        cdef char* data = <char *> in_data
+        cdef int data_len = len(in_data)*8
+
+        if self.finished:
+            self.finished = 0
+            self.state = self.previous_state
+
+        blake_h.Update(&self.state, <blake_h.BitSequence *> data, data_len)
+
+    cpdef final(self):
+        self.hashval = <blake_h.BitSequence *> malloc(self.hashbitlen*8)
+
+        # We copy the state so that we can continue to update.
+        # This equals hashlibs functionality, but not pycryptopp.
+        self.previous_state = self.state
+
+        blake_h.Final(&self.state, self.hashval)
+        self.finished = 1
+
+    cpdef copy(self):
+        s = blake(self.hashbitlen)
+        s.state = self.state
+        s.previous_state = self.previous_state
+        s.finished = self.finished
+
+        return s
+
+    def digest(self):
+        if not self.finished:
+            self.final()
+
[…]

### 2010-10-06 python-bindings/sha3lib/hash_functions/blake/blake_opt32.c

@@ -1,0 +1,770 @@
+#include <string.h>
+#include <stdio.h>
+#include "blake_opt32.h"
+
+
+static  HashReturn compress32( hashState * state, const BitSequence * datablock ) {
+
+#define ROT32(x,n) (((x)<<(32-n))|( (x)>>(n)))
+#define ADD32(x,y)   ((u32)((x) + (y)))
+#define XOR32(x,y)    ((u32)((x) ^ (y)))
+
+#define G32(a,b,c,d,i) \
+  do {\
+    v[a] = XOR32(m[sigma[round][i]], c32[sigma[round][i+1]])+ADD32(v[a],v[b]);\
+    v[d] = ROT32(XOR32(v[d],v[a]),16);\
+    v[c] = ADD32(v[c],v[d]);\
+    v[b] = ROT32(XOR32(v[b],v[c]),12);\
+    v[a] = XOR32(m[sigma[round][i+1]], c32[sigma[round][i]])+ADD32(v[a],v[b]); \
+    v[d] = ROT32(XOR32(v[d],v[a]), 8);\
+    v[c] = ADD32(v[c],v[d]);\
+    v[b] = ROT32(XOR32(v[b],v[c]), 7);\
+  } while (0)
+
+  u32 v[16];
+  u32 m[16];
+  int round;
+
+  /* get message */
+  m[ 0] = U8TO32_BE(datablock + 0);
+  m[ 1] = U8TO32_BE(datablock + 4);
+  m[ 2] = U8TO32_BE(datablock + 8);
+  m[ 3] = U8TO32_BE(datablock +12);
+  m[ 4] = U8TO32_BE(datablock +16);
+  m[ 5] = U8TO32_BE(datablock +20);
+  m[ 6] = U8TO32_BE(datablock +24);
+  m[ 7] = U8TO32_BE(datablock +28);
+  m[ 8] = U8TO32_BE(datablock +32);
+  m[ 9] = U8TO32_BE(datablock +36);
+  m[10] = U8TO32_BE(datablock +40);
+  m[11] = U8TO32_BE(datablock +44);
+  m[12] = U8TO32_BE(datablock +48);
+  m[13] = U8TO32_BE(datablock +52);
+  m[14] = U8TO32_BE(datablock +56);
+  m[15] = U8TO32_BE(datablock +60);
+
+
+  /* initialization */
+  v[ 0] = state->h32[0];
+  v[ 1] = state->h32[1];
[…]

### 2010-10-06 paper/main.tex
Added more layout, and a problem description
Updated BibTeX use and the Makefile

@@
-\documentclass[english,12pt,a4paper]{report}
-%[twoside] ?
+\documentclass[english,12pt,a4paper]{book}
 \usepackage[T1]{fontenc} % In case we want special characters
@@
 
-\usepackage{graphicx} % We _may_ use images
-% Lager automagisk interne lenker i dokumentet.
-\usepackage{hyperref} % Internal links is nice
+\usepackage{appendix} % Fixes formatting of appendicies
+\usepackage{graphicx} % We *may* use images
+\usepackage{hyperref} % Internal and external links is nice
 \hypersetup{pdfborder=0 0 0} % ..especially without red borders
 
-\author{Eirik Haver and Pål Ruud}
+% We need this to make it easier to ensure we start chapters on odd pages.
+\newcommand{\ensureodd}{
+    \newpage
+%    \thispagestyle{empty}
+    \mbox{}
+}
+
+\author{Eirik Haver \and Pål Ruud}
 \title{Project assignment - Tahoe-LAFS with SHA-3 candidates}
@@
 \include{title}
+\pagestyle{empty}
 
-\section*{Summary}
+\chapter*{Abstract}
+\addcontentsline{toc}{chapter}{Abstract}
+\pagestyle{plain}
+\pagenumbering{Roman}
+\setcounter{page}{1}
 
-\section*{Foreword}
+\chapter*{Preface}
+\addcontentsline{toc}{chapter}{Preface}
 
 \tableofcontents
-\newpage
 
-\listoftables
+\addcontentsline{toc}{chapter}{\listfigurename} % Manual, because report style
 \listoffigures
 
-% Acronyms?
+\addcontentsline{toc}{chapter}{\listtablename}
+\listoftables
 
-\chapter{SHA-3 in Tahoe-LAFS}
+\chapter*{Acronyms}
+\addcontentsline{toc}{chapter}{Acronyms}
 
-\section{Introduction}
+% Insert acronyms here
 
-\section{The candiates}
+\chapter{Introduction}
+\pagenumbering{arabic}
+\setcounter{page}{1}
+
+Thorough problem description
+
+\section{Method}
+
+Technical procedure, Cython, two persons, test environment, how we test,
+GitHub?
+
+\section{Scope and objectives}
+
+Should we include this?
+
+\section{Outline}
+
+What follows in this document, chapter by chapter
+
+\chapter{Background technologies}
+% Find new and better title
+
+\section{NIST SHA-3}
+
+Hash functions, NIST Comptetition, Candidates, SuperCop?
+
+\section{Tahoe-LAFS}
+
+What is it, how/where are hash functions used, Python, pycryptopp
+
+\section{Python bindings using Cython}
+
+What is it, how, why, alternatives?
+
+\chapter{Procedure}
+
+TODO: We have to figure out a new title
+Keywords: Python bindings in detail, unit tests, using it in Tahoe-LAFS
+
+A bit more technical description of how we tested the candidates in Tahoe-LAFS
+
+Limitations/assumptions
+
+Optimized candidates.
+
+\chapter{Results}
+
+What is tested and what is not tested?
+
+tables, graphs?
+
+\chapter{Discussion}
+
+What could have been done differently?
+
+Compare results with general (supercop) results.
+
+Optimized candidates.
+
+\cite{nistround2}
+
+\chapter{Conclusion}
+
+A somewhat short conclusion summing up results and discussion
+
+\section{Future work}
+
+New SHA-3 capability in Tahoe-LAFS (GSOC ref) to get our code to trunk
+
+% BibTeX bibliography lives in external file
+\bibliographystyle{plain}
+\bibliography{ref_sha3}
+
+\appendix
+\appendixpage
+\addappheadtotoc
+
+% Use ordinary \chapters from here on..
 

### 2010-10-08 python-bindings/setup.py
added timing of hash-functions in unittests and support for 'extemely long test vectors'

@@ -74,6 +74,7 @@ class RealClean(Command):
     user_options = []
 
     def initialize_options(self):
+        print ""
         self._dir = os.getcwd()
         self._clean_me = []
         for root, dirs, files in os.walk('.'):
@@ -101,11 +102,12 @@ class RealClean(Command):
                 #print clean_me
             except OSError:
                 try:
-                    os.rmdir(clean_me)
-                except:
+                    import shutil
+                    shutil.rmtree(clean_me)
+                except Exception,e:
                     pass
-            except:
-                pass
+            except Exception,e:
+                print e
 
 
 

### 2010-10-08 python-bindings/tests/test_extremelylongkat.py

@@ -1,0 +1,57 @@
+#coding: utf-8
+import unittest
+from sha3lib import hash_functions
+from sha3lib import *
+from os import path
+import time
+
+class TestExtremelyLongKat(unittest.TestCase):
+
+    def init_functions(self):
+        workload = {}
+        files = []
+        functions = []
+        hash_path = path.dirname(path.abspath(hash_functions.__file__))
+        for r in hash_functions.HASHES:
+            k = {}
+            k['function'] = eval(r+'256')
+            k['file'] = path.join(hash_path,r,'kat/ExtremelyLongMsgKAT_256.txt')
+            workload[r] = k
+
+        self.workload = workload
+
+    def parse_kat_file(self, filename):
+        k = open(filename,'r')
+        attribs = ('Repeat','Text','MD')
+        props = {}
+        for line in k:
+            for attrib in attribs:
+                if line.startswith(attrib):
+                    props[attrib] = line[line.find(' = ')+3:].replace(
+                            '\r',"").replace('\n',"")
+        return props 
+            
+
+    def test_extremely_long_kat(self):
+        self.init_functions()
+        k = self.workload
+        for function in self.workload:
+            t1 = time.time()
+            props = self.parse_kat_file(k[function]['file'])
+            s = k[function]['function']() 
+            for l in xrange(int(props['Repeat'])):
+                s.update(props['Text'])
+            
+            output = s.hexdigest()
+            t2 = time.time()
+            self.assertEquals(output, props['MD'].lower(),
+                    'Mismatch:\n%s\n%s\n%s ' %
+                    (function,output,props['MD'].lower()))
+
+            print '%s: Tested ExtremelyLong ok took %0.3f ms' % (function,
+                    (t2-t1)*1000.0)
+            
+
+
+if __name__ == '__main__':
+    unittest.main()

### 2010-10-08 python-bindings/tests/test_longkat.py

@@ -5,6 +5,7 @@ from os import path
 from sha3lib import hash_functions
 
 from katparser import KAT
+import time
 
 class TestLongKat(unittest.TestCase):
     def run_katparser(self, hash_function):
@@ -18,6 +19,7 @@ class TestLongKat(unittest.TestCase):
         known_answers = self.run_katparser(function_str)
         hash_function = '%s256hash' % function_str
         counter = 0
+        t1 = time.time()
         for kat in known_answers:
             counter += 1
             to_test = "hash_functions.%s(256, %r, %s)" % \
@@ -25,11 +27,12 @@ class TestLongKat(unittest.TestCase):
             output_digest = eval(to_test).upper()
             self.assertEquals(output_digest, kat[2],
                 'Mismatch:\n%s\n%s' % (to_test, kat[2]))
-
-        print '%s: Tested %i KATs' % (function_str, counter)
+        t2 = time.time()
+        print '%s: Tested %i KATs took %0.3f ms' % (function_str, 
+                counter,(t2-t1)*1000.0)
 
     def test_if_the_kats_match(self):
-        print "\n"
+        print ""
         for hash_function in hash_functions.HASHES:
             self.run_kat_test(hash_function)
 

### 2010-10-08 python-bindings/tests/test_shortkat.py

@@ -6,6 +6,8 @@ from sha3lib import hash_functions
 
 from katparser import KAT
 
+import time
+
 class TestShortKat(unittest.TestCase):
     def run_katparser(self, hash_function):
         hash_path = path.dirname(path.abspath(hash_functions.__file__))
@@ -18,6 +20,7 @@ class TestShortKat(unittest.TestCase):
         known_answers = self.run_katparser(function_str)
         hash_function = '%s256hash' % function_str
         counter = 0
+        t1 = time.time()
         for kat in known_answers:
             counter += 1
             to_test = "hash_functions.%s(256, %r, %s)" % \
@@ -25,11 +28,12 @@ class TestShortKat(unittest.TestCase):
             output_digest = eval(to_test).upper()
             self.assertEquals(output_digest, kat[2],
                 'Mismatch:\n%s\n%s' % (to_test, kat[2]))
-
-        print '%s: Tested %i KATs' % (function_str, counter)
+        t2 = time.time()
+        print '%s: Tested %i KATs took %0.3f ms' % (function_str, 
+                counter, (t2-t1)*1000.0)
 
     def test_if_the_kats_match(self):
-        print "\n"
+        print ""
         for hash_function in hash_functions.HASHES:
             self.run_kat_test(hash_function)
 

### 2010-10-08 paper/main.tex
Started on Tahoe in general and added some references.

@@
 
-\usepackage{appendix} % Fixes formatting of appendicies
+\usepackage{appendix} % Fixes formatting of appendices
 \usepackage{graphicx} % We *may* use images
@@
 \section{Tahoe-LAFS}
+%What is it, how/where are hash functions used, Python, pycryptopp
+%comparison with RAID-6
 
-What is it, how/where are hash functions used, Python, pycryptopp
+The Tahoe Least-Authority Filesystem is a system for secure,
+distributed data storage \cite{t_tahoe}. Tahoe was originally developed with
+funding from the former web backup service provider Allmydata, but is now a
+stand-alone Open Source project \cite{t_ars}.
+
+New files saved using Tahoe follow this pattern:
+\begin{enumerate}
+\item The file is encrypted using AES.
+\item The file is split up into \emph{N} parts using \emph{erasure coding}.
+\item The parts are distributed to \emph{N} storage servers.
+\end{enumerate}
+
+The use of erasure coding enables Tahoe to recover a file using only a
+predefined subset of the parts distributed to the storage servers.
+
+% TODO: comparison with RAID?
+
+\subsection{Advantages using a secure, distributed file system}
+% Use cases?
 
@@
 
-\cite{nistround2}
+\cite{s_nistround2}
 
@@
 \bibliographystyle{plain}
-\bibliography{ref_sha3}
+\bibliography{ref_sha3,ref_tahoe}
 

### 2010-10-12 paper/main.tex
Started cryptograhic hash functions and Cython-section

@@
 
-\section{NIST SHA-3}
+\section{Cryptographic Hash Functions}
+A Cryptographic Hash function is a deterministic mathematical procedure which
+takes an arbritary block of data and outputs a fixed-size bit string. The
+output is refered to as the hash value, message digest or simply digest.
+Another property of a cryptographic hash function is that a small change in the
+input data (just 1 bit) should completly change the output of the hash
+function. In other words it should be infeasible to find the reverse of a
+cryptographic hash function. It should also be infeasible to find two blocks of
+data which produce the same hash value (a collision).
 
-Hash functions, NIST Comptetition, Candidates, SuperCop?
+\subsection{NIST SHA-3 Competition}
+% Hash functions, NIST Comptetition, Candidates, SuperCop?
+The Secure Hash Function version 3 (SHA-3) is a comming standard cryptographic
+hash function set to replace the current standards SHA-1 and SHA-2.
+It will be selected from a competition held by the National
+Institute of Standards and Technology (NIST). The competition started in 2006
+with submission deadline for candidates at the end of 2008. 51 candidates made
+it into round 1, however just 14 passed to the second round. The reason for not
+passing round 1 are either design weaknesses, substantial cryptographic
+weaknesses or performance issues. The winner is sceduled to be selected by the 
+2nd quarter of 2012.
+
+The submissions for the various rounds of the SHA-3 competition include
+reference implementations, optimized implementations, test-vectors to verify
+implementations and documentations explaining the submission.
+
+\subsection{SuperCop}
+Supercop is a toolkit developed by VAMPIRE lab for measuring performance of
+cryptographic software, among them the SHA-3 candidates. While security is a
+primary goal of a hash function, performance can not be neglected if the hash
+function should be taken into use. The difference between the implementations
+in the official submissions and in supercop is first that the supercop
+implementation are continiously updated and that there are more specific
+hardwareimplementations available in supercop.
+
+Results by the supercop are published at (URL).
 
@@
 \section{Python bindings using Cython}
+%What is it, how, why, alternatives?
+% TODO: Hvorfor er nedenfornevnte i ASM eller C?
+TAHOE-LAFS is written in python, but the SHA-3 candidate implementations found
+in both the official NIST submissions and supercop are either in C or in
+assembly. Therefore we need a way to make the implementations available in
+Python. There a several ways to do this. Python itself includes a C-API,  
+other options include SWIG, Cython and possible others.
 
-What is it, how, why, alternatives?
+\subsection(Python C-API)
+The Python C-API is the official method of creating C-bindings for Python. The
+procedure to do this includes writing entrypoints or wrappers for your C
+functions in C. Then, through compiling you can acess your C-libraries and
+functions through Python. The drawback with this approach, when faced with an
+already existing C-library is that you have to actually write the
+wrapper-functions or modify the C-library.
+
+\subsection{Cython}
+Cython tries to solve the problem of binding C-libraries and functions by the
+use of Python instead of the use of C which the official API does. This is done
+by only defining the entrypoints to the required C-functions and datatypes in a
+Python/C/C++ hybrid syntax. Cython also allows creating new C/C++ functions,
+available or not available to Python, in the same syntax. Cython works by
+transforming the hybrid syntax into C/C++-code which utilises the official
+Python C-API and compiling that. An expected tradeoff by using Cython is a
+performance drop due to the extra layer of abstraction.
 

### 2010-10-12 python-bindings/setup.py
Added sim256 hash function

@@ -29,6 +29,9 @@ ext_modules = [
         [hashpath+"luffa/luffa.pyx", hashpath+'luffa/luffa_for_32.c'],),
     Extension("sha3lib.hash_functions.shavite3.shavite3",
         [hashpath+"shavite3/shavite3.pyx", hashpath+'shavite3/SHAvite3.c'],),
+    Extension("sha3lib.hash_functions.simd.simd",
+        [hashpath+"simd/simd.pyx",hashpath+"simd/optimized.c",
+        hashpath+"simd/nist.c"],),
     ]
 
 class RunTests(Command):
@@ -88,7 +91,10 @@ class RealClean(Command):
                 './sha3lib/hash_functions/groestl/groestl.c',
                 './sha3lib/hash_functions/skein/skein_hash.c',
                 './sha3lib/hash_functions/echo/echo.c',
-                './sha3lib/hash_functions/shavite3.c']
+                './sha3lib/hash_functions/shavite3/shavite3.c'
+                './sha3lib/hash_functions/blake/blake.c',
+                './sha3lib/hash_functions/simd/simd.c',
+                ]
         self._clean_me.extend(manual_files)
 
 
@@ -119,9 +125,15 @@ setup(
     cmdclass = {'build_ext': build_ext, 'test': RunTests, 'realclean':RealClean},
     ext_modules = ext_modules,
     packages =
-        ['sha3lib','sha3lib.hash_functions','sha3lib.hash_functions.bmw',
-        'sha3lib.hash_functions.echo', 'sha3lib.hash_functions.groestl',
-        'sha3lib.hash_functions.blake', 'sha3lib.hash_functions.skein', 
-        'sha3lib.hash_functions.luffa', 'sha3lib.hash_functions.shavite3'],
+        ['sha3lib','sha3lib.hash_functions',
+        'sha3lib.hash_functions.bmw',
+        'sha3lib.hash_functions.echo', 
+        'sha3lib.hash_functions.groestl',
+        'sha3lib.hash_functions.blake', 
+        'sha3lib.hash_functions.skein', 
+        'sha3lib.hash_functions.luffa', 
+        'sha3lib.hash_functions.shavite3',
+        'sha3lib.hash_functions.simd',
+        ],
     #py_modules = ['sha3lib'],    
 )

### 2010-10-12 python-bindings/sha3lib/__init__.py

@@ -1,2 +1,2 @@
 from _sha3lib import (bmw256,groestl256,echo256,blake256,skein256,
-                      luffa256,shavite3256)
+                      luffa256,shavite3256,simd256)

### 2010-10-12 python-bindings/sha3lib/_sha3lib.py

@@ -1,5 +1,5 @@
 # coding: utf-8
-from hash_functions import bmw,groestl,echo,blake,skein,luffa,shavite3
+from hash_functions import bmw,groestl,echo,blake,skein,luffa,shavite3,simd
 
 class bmw256(bmw):
     def __init__(self, initial=None):
@@ -28,3 +28,7 @@ class luffa256(luffa):
 class shavite3256(shavite3):
     def __init__(self, initial=None):
         super(shavite3256, self).__init__(256, initial)
+
+class simd256(simd):
+    def __init__(self, initial=None):
+        super(simd256, self).__init__(256, initial)

### 2010-10-12 python-bindings/sha3lib/hash_functions/__init__.py

@@ -5,6 +5,7 @@ from blake.blake import blake
 from skein.skein_hash import skein
 from luffa.luffa import luffa
 from shavite3.shavite3 import shavite3
+from simd.simd import simd
 
 from bmw.bmw import hash as bmw256hash
 from groestl.groestl import hash as groestl256hash
@@ -13,6 +14,7 @@ from blake.blake import hash as blake256hash
 from skein.skein_hash import hash as skein256hash
 from luffa.luffa import hash as luffa256hash
 from shavite3.shavite3 import hash as shavite3256hash
+from simd.simd import hash as simd256hash
 
 # skein ignored due to possible wrong kat-files or implementation
-HASHES = ('bmw', 'echo', 'groestl', 'blake', 'luffa', 'shavite3')
+HASHES = ('bmw', 'echo', 'groestl', 'blake', 'luffa', 'shavite3', 'simd')

### 2010-10-12 python-bindings/sha3lib/hash_functions/simd/simd.pyx

@@ -1,0 +1,87 @@
+# encoding: utf-8
+cimport simd_h
+cimport cython
+
+from libc.stdlib cimport *
+
+def hash(int hashbitlen, bytes data, int in_length):
+    '''
+    Used for testing purposes.
+    '''
+    cdef char* c_string = <char *> data
+    cdef simd_h.DataLength datalen = in_length
+    cdef simd_h.BitSequence *hashval = <simd_h.BitSequence *> malloc(hashbitlen*8)
+
+    simd_h.Hash(hashbitlen, <simd_h.BitSequence *> c_string, datalen, hashval)
+
+    try:
+        digest = [hashval[i] for i from 0 <= i < hashbitlen / 8]
+        digest = ''.join(['%02X' % i for i in digest])
+    finally:
+        free(hashval)
+
+    return digest
+
+cdef class simd:
+    '''
+    A class that tries to mimic the behaviour of hashlib, i.e. keeping state
+    so that one can update the hashing procedure instead of doing it from
+    scratch.
+    '''
+    cdef simd_h.BitSequence *hashval
+    cdef simd_h.hashState previous_state
+    cdef simd_h.hashState state
+    cdef int finished
+    cdef int hashbitlen
+
+    def __init__(self, int in_hashbitlen, bytes initial=None):
+        self.finished = 0
+        self.hashbitlen = in_hashbitlen
+        simd_h.Init(&self.state, self.hashbitlen)
+
+        if initial:
+            self.update(initial)
+
+    cpdef update(self, bytes in_data):
+        cdef char* data = <char *> in_data
+        cdef int data_len = len(in_data)*8
+
+        if self.finished:
+            self.finished = 0
+            self.state = self.previous_state
+
+        simd_h.Update(&self.state, <simd_h.BitSequence *> data, data_len)
+
+    cpdef final(self):
+        self.hashval = <simd_h.BitSequence *> malloc(self.hashbitlen*8)
+
+        # We copy the state so that we can continue to update.
+        # This equals hashlibs functionality, but not pycryptopp.
+        self.previous_state = self.state
+
+        simd_h.Final(&self.state, self.hashval)
+        self.finished = 1
+
+    cpdef copy(self):
+        s = simd(self.hashbitlen)
+        s.state = self.state
+        s.previous_state = self.previous_state
+        s.finished = self.finished
+
+        return s
+
+    def digest(self):
+        if not self.finished:
+            self.final()
+
[…]

### 2010-10-12 python-bindings/setup.py
added Keccak hash function

@@ -32,6 +32,9 @@ ext_modules = [
     Extension("sha3lib.hash_functions.simd.simd",
         [hashpath+"simd/simd.pyx",hashpath+"simd/optimized.c",
         hashpath+"simd/nist.c"],),
+    Extension("sha3lib.hash_functions.keccak.keccak",
+        [hashpath+"keccak/keccak.pyx",hashpath+"keccak/KeccakF-1600-opt32.c",
+        hashpath+"keccak/KeccakSponge.c"],),
     ]
 
 class RunTests(Command):
@@ -94,6 +97,7 @@ class RealClean(Command):
                 './sha3lib/hash_functions/shavite3/shavite3.c'
                 './sha3lib/hash_functions/blake/blake.c',
                 './sha3lib/hash_functions/simd/simd.c',
+                './sha3lib/hash_functions/keccak/keccak.c',
                 ]
         self._clean_me.extend(manual_files)
 
@@ -120,7 +124,7 @@ class RealClean(Command):
 setup(
     name='SHA3lib',
     version='0.1',
-    description='Python bindings for SHA3 using cython',
+    description='Python bindings for SHA3-256 using Cython',
     author='Eirik Haver and Pål Ruud',
     cmdclass = {'build_ext': build_ext, 'test': RunTests, 'realclean':RealClean},
     ext_modules = ext_modules,
@@ -134,6 +138,7 @@ setup(
         'sha3lib.hash_functions.luffa', 
         'sha3lib.hash_functions.shavite3',
         'sha3lib.hash_functions.simd',
+        'sha3lib.hash_functions.keccak',
         ],
     #py_modules = ['sha3lib'],    
 )

### 2010-10-12 python-bindings/sha3lib/__init__.py

@@ -1,2 +1,2 @@
 from _sha3lib import (bmw256,groestl256,echo256,blake256,skein256,
-                      luffa256,shavite3256,simd256)
+                      luffa256,shavite3256,simd256,keccak256)

### 2010-10-12 python-bindings/sha3lib/_sha3lib.py

@@ -1,5 +1,6 @@
 # coding: utf-8
-from hash_functions import bmw,groestl,echo,blake,skein,luffa,shavite3,simd
+from hash_functions import(bmw,groestl,echo,blake,skein, 
+        luffa,shavite3,simd,keccak)
 
 class bmw256(bmw):
     def __init__(self, initial=None):
@@ -32,3 +33,7 @@ class shavite3256(shavite3):
 class simd256(simd):
     def __init__(self, initial=None):
         super(simd256, self).__init__(256, initial)
+
+class keccak256(keccak):
+    def __init__(self, initial=None):
+        super(keccak256, self).__init__(256, initial)

### 2010-10-12 python-bindings/sha3lib/hash_functions/__init__.py

@@ -6,6 +6,7 @@ from skein.skein_hash import skein
 from luffa.luffa import luffa
 from shavite3.shavite3 import shavite3
 from simd.simd import simd
+from keccak.keccak import keccak
 
 from bmw.bmw import hash as bmw256hash
 from groestl.groestl import hash as groestl256hash
@@ -15,6 +16,8 @@ from skein.skein_hash import hash as skein256hash
 from luffa.luffa import hash as luffa256hash
 from shavite3.shavite3 import hash as shavite3256hash
 from simd.simd import hash as simd256hash
+from keccak.keccak import hash as keccak256hash
 
 # skein ignored due to possible wrong kat-files or implementation
-HASHES = ('bmw', 'echo', 'groestl', 'blake', 'luffa', 'shavite3', 'simd')
+HASHES = ('bmw', 'echo', 'groestl', 'blake', 'luffa', 'shavite3', 'simd',
+'keccak')

### 2010-10-12 python-bindings/sha3lib/hash_functions/keccak/api.h

@@ -1,0 +1,2 @@
+#define CRYPTO_BYTES 32
+#define CRYPTO_VERSION "2.2.0"

### 2010-10-12 python-bindings/sha3lib/hash_functions/keccak/keccak.pyx

@@ -1,0 +1,87 @@
+# encoding: utf-8
+cimport keccak_h
+cimport cython
+
+from libc.stdlib cimport *
+
+def hash(int hashbitlen, bytes data, int in_length):
+    '''
+    Used for testing purposes.
+    '''
+    cdef char* c_string = <char *> data
+    cdef keccak_h.DataLength datalen = in_length
+    cdef keccak_h.BitSequence *hashval = <keccak_h.BitSequence *> malloc(hashbitlen*8)
+
+    keccak_h.Hash(hashbitlen, <keccak_h.BitSequence *> c_string, datalen, hashval)
+
+    try:
+        digest = [hashval[i] for i from 0 <= i < hashbitlen / 8]
+        digest = ''.join(['%02X' % i for i in digest])
+    finally:
+        free(hashval)
+
+    return digest
+
+cdef class keccak:
+    '''
+    A class that tries to mimic the behaviour of hashlib, i.e. keeping state
+    so that one can update the hashing procedure instead of doing it from
+    scratch.
+    '''
+    cdef keccak_h.BitSequence *hashval
+    cdef keccak_h.hashState previous_state
+    cdef keccak_h.hashState state
+    cdef int finished
+    cdef int hashbitlen
+
+    def __init__(self, int in_hashbitlen, bytes initial=None):
+        self.finished = 0
+        self.hashbitlen = in_hashbitlen
+        keccak_h.Init(&self.state, self.hashbitlen)
+
+        if initial:
+            self.update(initial)
+
+    cpdef update(self, bytes in_data):
+        cdef char* data = <char *> in_data
+        cdef int data_len = len(in_data)*8
+
+        if self.finished:
+            self.finished = 0
+            self.state = self.previous_state
+
+        keccak_h.Update(&self.state, <keccak_h.BitSequence *> data, data_len)
+
+    cpdef final(self):
+        self.hashval = <keccak_h.BitSequence *> malloc(self.hashbitlen*8)
+
+        # We copy the state so that we can continue to update.
+        # This equals hashlibs functionality, but not pycryptopp.
+        self.previous_state = self.state
+
+        keccak_h.Final(&self.state, self.hashval)
+        self.finished = 1
+
+    cpdef copy(self):
+        s = keccak(self.hashbitlen)
+        s.state = self.state
+        s.previous_state = self.previous_state
+        s.finished = self.finished
+
+        return s
+
+    def digest(self):
+        if not self.finished:
+            self.final()
+
[…]

### 2010-10-12 paper/main.tex
Added acronyms-package and more text to the tahoe-chapter

@@
 \usepackage{appendix} % Fixes formatting of appendices
+\usepackage[printonlyused]{acronym} % Package to handle the acronym list
 \usepackage{graphicx} % We *may* use images
+\graphicspath{{images/}} % and it is clean to put them in a separate dir
 \usepackage{hyperref} % Internal and external links is nice
@@
 
-% Insert acronyms here
+\begin{acronym}
+\acro{AES}{Advanced Encryption Standard}
+\acro{LAFS}{Least-Authority Filesystem}
+\acro{FEC}{Forward Error Correction}
+\end{acronym}
 
@@
 
+Paralellizing
+
 \section{Tahoe-LAFS}
 %What is it, how/where are hash functions used, Python, pycryptopp
-%comparison with RAID-6
+%comparison with RAID-6, mutable/immutable?
 
-The Tahoe Least-Authority Filesystem is a system for secure,
-distributed data storage \cite{t_tahoe}. Tahoe was originally developed with
-funding from the former web backup service provider Allmydata, but is now a
-stand-alone Open Source project \cite{t_ars}.
+The Tahoe \ac{LAFS} is a system for secure,
+distributed data storage. Files are encrypted client side, then
+split up, before each part is sent to other nodes in the grid. The integrity and
+confidentiality of the files are guaranteed by the algorithms used on the
+client, and is independent of the storage servers, which may be operated by
+untrusted people. This is defined as \emph{provider-independent security}
+\cite{t_tahoe}.
 
-New files saved using Tahoe follow this pattern:
-\begin{enumerate}
-\item The file is encrypted using AES.
-\item The file is split up into \emph{N} parts using \emph{erasure coding}.
-\item The parts are distributed to \emph{N} storage servers.
-\end{enumerate}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{Tahoe-newfile.pdf}
+    \caption{Tahoe-LAFS: Insertion of new file}
+    \label{fig:tahoeinsertion}
+\end{figure}
 
-The use of erasure coding enables Tahoe to recover a file using only a
-predefined subset of the parts distributed to the storage servers.
+Tahoe was originally developed with funding from the former commercial web
+backup service provider Allmydata, but is now a stand-alone Open
+Source\footnote{GNU General Public License (GPL) version 2} project \cite{t_ars}.
+It is written in the Python programming language with the Twisted framework, and
+can run on Windows, Mac OSX, Linux, Solaris and more.
+
+\subsection{Erasure coding}
+
+The use of \emph{erasure coding} enables Tahoe to recover a file using only a
+predefined subset of the parts distributed to the storage servers, i.e. the
+other nodes in the grid. Erasure coding is a type of \ac{FEC} code, which
+extends a message with $C$ characters into a longer message with $N$ symbols.
+The original $C$ characters can then be recovered from a subset of the $N$
+symbols.
 
@@
 
-\subsection{Advantages using a secure, distributed file system}
-% Use cases?
+\subsection{Use of secure hashes in Tahoe-LAFS}
+
+Where?
 
@@
 
+Error sources.
+
 \chapter{Results}

### 2010-10-22 paper/main.tex
Proper margins
Added hashing in Tahoe figure and tiny bit of text

@@
 
-% We need this to make it easier to ensure we start chapters on odd pages.
-\newcommand{\ensureodd}{
-    \newpage
-%    \thispagestyle{empty}
-    \mbox{}
-}
+% Set equal margins on book style
+% \usepackage{layout} % Use \layout to print out the margins (debug)
+\usepackage{geometry}
+\geometry{bindingoffset=1cm}
 
@@
 distributed data storage. Files are encrypted client side, then
-split up, before each part is sent to other nodes in the grid. The integrity and
-confidentiality of the files are guaranteed by the algorithms used on the
-client, and is independent of the storage servers, which may be operated by
-untrusted people. This is defined as \emph{provider-independent security}
-\cite{t_tahoe}.
+split up, before each part is sent to other nodes in the grid, as depicted in
+figure \ref{fig:tahoeinsertion}. The integrity and confidentiality of the files
+are guaranteed by the algorithms used on the client, and is independent of the
+storage servers, which may be operated by untrusted people. This is defined as
+\emph{provider-independent security} \cite{t_tahoe}.
 
@@
 
+The properties of erasure coding can be seen up on as those of recovery in RAID
+systems.
+
 % TODO: comparison with RAID?
@@
 
-Where?
+% Describe hash trees in further detail?
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{Tahoe-hashing.pdf}
+    \caption{Tahoe-LAFS: Location of hashing}
+    \label{fig:tahoehashing}
+\end{figure}
+
+As seen in figure \ref{fig:tahoehashing}, the usage of secure hashing is
+extensive in Tahoe, and has to be considered as a key part of the functionality.
 
@@
 
-\subsection(Python C-API)
+\subsection{Python C-API}
 The Python C-API is the official method of creating C-bindings for Python. The

### 2010-10-25 paper/main.tex
More text on Tahoe architecture

@@
 backup service provider Allmydata, but is now a stand-alone Open
-Source\footnote{GNU General Public License (GPL) version 2} project \cite{t_ars}.
-It is written in the Python programming language with the Twisted framework, and
-can run on Windows, Mac OSX, Linux, Solaris and more.
+Source\footnote{GNU General Public License (GPL) version 2} project
+\cite{t_ars}.  It is written in the Python programming language with the Twisted
+framework, and can run on Windows, Mac OSX, Linux, Solaris and more.
 
-\subsection{Erasure coding}
+\subsection{Architecture}
 
+Tahoe has a three layer architecture: the key-value store, the filesystem, and
+the application.
+
+The \textbf{key-value store}, or the ``capability-data bytes'' store, is the
+lowest layer and is implemented by a grid of Tahoe-LAFS storage servers. Data is
+kept on the storage servers in the form of ``shares'', which are encrypted and
+encoded parts of files. Capabilities are short ASCII strings, containing
+information on where to \emph{find} a file, and how to \emph{verify} it.
+Nodes in the grid learn about each other through an ``introducer'', which
+roughly relates to a tracker in the BitTorrent\footnote{See
+\url{http://www.bittorrent.org/beps/bep\_0003.html}} protocol.
+
+The \textbf{filesystem} layer is responsible for mapping human-meaningful
+pathnames to pieces of data. Each directory contains a table of capabilities
+for its children, i.e. subdirectories or files. Two forms of capabilities is
+available for each file, read-only and read-write, and these can be shared to
+provide shared/published directory structures with friends.
+
+Since it is not practical for users to remember strings containing random
+characters, the \textbf{application} layer is used for providing a user-friendly
+interface to the directories and files.
+
+\subsection{Related technologies}
+
+Tahoe-\ac{LAFS} uses a number of techniques to realize the functionality
+demanded by a secure, distributed filesystem.
+
+\paragraph{Erasure coding.}
+
+When a client puts a file on the grid, it first encrypts the file, before
+breaking the file into small segments. The segments are then erasure coded.
 The use of \emph{erasure coding} enables Tahoe to recover a file using only a
@@
 other nodes in the grid. Erasure coding is a type of \ac{FEC} code, which
-extends a message with $C$ characters into a longer message with $N$ symbols.
+extends a message with $C$ characters into a longer message with $N$ symbols
+\cite{t_reed-solomon}.
 The original $C$ characters can then be recovered from a subset of the $N$
@@
 The properties of erasure coding can be seen up on as those of recovery in RAID
-systems.
+systems. \cite{t_erasure} % TODO: comparison with RAID
 
-% TODO: comparison with RAID?
+\paragraph{Merkle hash trees.}
+
+Describe hash trees in further detail?
 
@@
 
-% Describe hash trees in further detail?
-
-\begin{figure}[h!]
+\begin{figure}[h]
     \centering

### 2010-10-25 paper/main.tex
More text on fec vs raid

@@
 
+\usepackage[numbers]{natbib}
 \usepackage{appendix} % Fixes formatting of appendices
@@
 Tahoe-\ac{LAFS} uses a number of techniques to realize the functionality
-demanded by a secure, distributed filesystem.
+needed by a secure, distributed filesystem.
 
@@
 When a client puts a file on the grid, it first encrypts the file, before
-breaking the file into small segments. The segments are then erasure coded.
-The use of \emph{erasure coding} enables Tahoe to recover a file using only a
-predefined subset of the parts distributed to the storage servers, i.e. the
-other nodes in the grid. Erasure coding is a type of \ac{FEC} code, which
-extends a message with $C$ characters into a longer message with $N$ symbols
-\cite{t_reed-solomon}.
-The original $C$ characters can then be recovered from a subset of the $N$
-symbols.
+breaking the file into small segments. The segments are then \emph{erasure
+coded}.  The use of the Solomon-Reed erasure coding scheme, enables Tahoe to
+recover a file using only a predefined subset of the parts distributed to the
+storage servers, i.e. the other nodes in the grid. Erasure coding is a type of
+\ac{FEC} code, which extends a message with $C$ characters into a longer message
+with $N$ symbols \cite{t_reed-solomon}.  The original $C$ characters can then be
+recovered from a subset of the $N$ symbols.
 
-The properties of erasure coding can be seen up on as those of recovery in RAID
-systems. \cite{t_erasure} % TODO: comparison with RAID
+The properties of erasure coding can be thought of as those of parity bits in
+RAID\footnote{Redundant Array of Independent Disks} systems. \citet*{t_erasure}
+compare erasure coding and plain replication, and conclude that ``\emph{...
+erasure codes have mean time to failures many orders of magnitude higher than
+replicated systems with similar storage and bandwidth requirements.}''
 
@@
 
-Describe hash trees in further detail?
+%TODO: Describe hash trees in further detail?
 
@@
 % BibTeX bibliography lives in external file
-\bibliographystyle{plain}
+\bibliographystyle{plainnat}
 \bibliography{ref_sha3,ref_tahoe}

### 2010-10-26 paper/main.tex
Added text on Merkle Trees

@@
 
-\paragraph{Merkle hash trees.}
+\paragraph{Hash Trees.}
 
-%TODO: Describe hash trees in further detail?
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{Tahoe-MerkleTree.pdf}
+    \caption{Example of a Merkle tree.}
+    \label{fig:tahoemerkletree}
+\end{figure}
+
+A hash tree, also known as \emph{Merkle tree}, is a type of data structure
+which can be described as a tree with nodes that can verify all information
+below in the hierarchy, as depicted in figure \ref{fig:tahoemerkletree}. This
+enables Tahoe to verify small segments of a file at the time, and this can be
+used for instance to start playing a movie file while it is still being
+downloaded.
 

### 2010-10-27 paper/main.tex
Added NIST evaluation criterias

--- 2010-10-25 paper/main.tex
@@
 \acro{AES}{Advanced Encryption Standard}
+\acro{FEC}{Forward Error Correction}
 \acro{LAFS}{Least-Authority Filesystem}
-\acro{FEC}{Forward Error Correction}
+\acro{NIST}{National Institute of Standards and Technology}
+\acro{RAM}{Random Access Memory}
+\acro{SHA}{Secure Hash Algorithm}
 \end{acronym}
@@
 \section{Cryptographic Hash Functions}
+% Get a source for this %
 A Cryptographic Hash function is a deterministic mathematical procedure which
@@
 \subsection{NIST SHA-3 Competition}
-% Hash functions, NIST Comptetition, Candidates, SuperCop?
-The Secure Hash Function version 3 (SHA-3) is a comming standard cryptographic
-hash function set to replace the current standards SHA-1 and SHA-2.
-It will be selected from a competition held by the National
-Institute of Standards and Technology (NIST). The competition started in 2006
-with submission deadline for candidates at the end of 2008. 51 candidates made
-it into round 1, however just 14 passed to the second round. The reason for not
-passing round 1 are either design weaknesses, substantial cryptographic
-weaknesses or performance issues. The winner is sceduled to be selected by the 
-2nd quarter of 2012.
+The \ac{SHA} version 3 is a comming standard set to superceed the current
+standards \ac{SHA}-1 and \ac{SHA}-2 families. The hash function to become
+\ac{SHA}-3 will be decided by \ac{NIST} and choosen between the submitted
+contestants to the \ac{NIST} hash competition. The current status of the 
+competition is officially called Round 2, with 14 of 64 candidates having 
+survived round 1\cite{s_fedreg}.
 
-The submissions for the various rounds of the SHA-3 competition include
-reference implementations, optimized implementations, test-vectors to verify
-implementations and documentations explaining the submission.
+\subsubsection{\ac{NIST} evaluation criterias for \ac{SHA}-3}
+
+\paragraph{Security} is the most important criteria for the SHA-3
+candidates\cite{s_nistround2}. The full description of security criterias are
+mentioned in \cite{s_fedreg}, however the most noteworthy criteria in regards
+to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
+resistance against length-extension attacks. Length-extension attacks is
+something that both \ac{SHA}-1 and the \ac{SHA2}-family suffers
+from\cite{s_mdre}.
+
+\paragraph{Cost and Performance} are considered to be the 2nd most important
+criteria. The absolute minimum for performance is that the SHA-3 candidate
+should be faster than the functions in the \ac{SHA}-2 family. Cost is a measure
+of how much memory an implementation requires in software, both the 
+implementation itself and the use of \ac{RAM} during runtime. Another measure
+of cost is how many logic gates it takes to implement the function in hardware.
+
+\paragraph{Algorithm and implementation characteristics} is the third criteria.
+By this criteria \ac{NIST} emphasis that algorithms with greater flexbility
+will be given preference over other algorithms\cite{s_nistround2}. By flexible
+they imply the possiblity of the function to run efficiently on a varity of
+plattforms, to use parallelism or instruction set extensions. Another key point
+of a flexible hash function is that it should have a simple and elegant design
+to encourage understanding, analysis and design confidence.
 
@@
 backup service provider Allmydata, but is now a stand-alone Open
-Source\footnote{GNU General Public License (GPL) version 2} project
-\cite{t_ars}.  It is written in the Python programming language with the Twisted
-framework, and can run on Windows, Mac OSX, Linux, Solaris and more.
+Source\footnote{GNU General Public License (GPL) version 2} project \cite{t_ars}.
+It is written in the Python programming language with the Twisted framework, and
+can run on Windows, Mac OSX, Linux, Solaris and more.
 
-\subsection{Architecture}
+\subsection{Erasure coding}
 
-Tahoe has a three layer architecture: the key-value store, the filesystem, and
-the application.
-
-The \textbf{key-value store}, or the ``capability-data bytes'' store, is the
-lowest layer and is implemented by a grid of Tahoe-LAFS storage servers. Data is
-kept on the storage servers in the form of ``shares'', which are encrypted and
-encoded parts of files. Capabilities are short ASCII strings, containing
-information on where to \emph{find} a file, and how to \emph{verify} it.
-Nodes in the grid learn about each other through an ``introducer'', which
-roughly relates to a tracker in the BitTorrent\footnote{See
-\url{http://www.bittorrent.org/beps/bep\_0003.html}} protocol.
-
-The \textbf{filesystem} layer is responsible for mapping human-meaningful
-pathnames to pieces of data. Each directory contains a table of capabilities
-for its children, i.e. subdirectories or files. Two forms of capabilities is
-available for each file, read-only and read-write, and these can be shared to
-provide shared/published directory structures with friends.
-
-Since it is not practical for users to remember strings containing random
-characters, the \textbf{application} layer is used for providing a user-friendly
-interface to the directories and files.
-
-\subsection{Related technologies}
-
-Tahoe-\ac{LAFS} uses a number of techniques to realize the functionality
-demanded by a secure, distributed filesystem.
-
-\paragraph{Erasure coding.}
-
-When a client puts a file on the grid, it first encrypts the file, before
-breaking the file into small segments. The segments are then erasure coded.
 The use of \emph{erasure coding} enables Tahoe to recover a file using only a
@@
 other nodes in the grid. Erasure coding is a type of \ac{FEC} code, which
-extends a message with $C$ characters into a longer message with $N$ symbols
-\cite{t_reed-solomon}.
+extends a message with $C$ characters into a longer message with $N$ symbols.
 The original $C$ characters can then be recovered from a subset of the $N$
@@
 The properties of erasure coding can be seen up on as those of recovery in RAID
-systems. \cite{t_erasure} % TODO: comparison with RAID
+systems.
 
-\paragraph{Merkle hash trees.}
-
-Describe hash trees in further detail?
+% TODO: comparison with RAID?
 
@@
 
-\begin{figure}[h]
+% Describe hash trees in further detail?
+
+\begin{figure}[h!]
     \centering

### 2010-10-27 paper/main.tex
Fixed typos

@@
 The \ac{SHA} version 3 is a comming standard set to superceed the current
-standards \ac{SHA}-1 and \ac{SHA}-2 families. The hash function to become
+standards \ac{SHA}-1 and the \ac{SHA}-2 family. The hash function to become
 \ac{SHA}-3 will be decided by \ac{NIST} and choosen between the submitted
@@
 to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
-resistance against length-extension attacks. Length-extension attacks is
-something that both \ac{SHA}-1 and the \ac{SHA2}-family suffers
-from\cite{s_mdre}.
+resistance against length-extension attacks. Both \ac{SHA}-1 and the \ac{SHA}-2
+family are vulnerable to this attack\cite{s_mdre}.
 

### 2010-10-27 paper/main.tex
Added supercop description, with sources

@@
 \acro{SHA}{Secure Hash Algorithm}
+\acro{SSE2}{Streaming SIMD Extensions 2}
+\acro{SSSE3}{Supplemental Streaming SIMD Extensions 3}
+\acro{SUPERCOP}{System for Unified Performance Evaluation Related to
+Cryptographic Operations and Primitives}
 \end{acronym}
@@
 
-\subsection{SuperCop}
-Supercop is a toolkit developed by VAMPIRE lab for measuring performance of
-cryptographic software, among them the SHA-3 candidates. While security is a
-primary goal of a hash function, performance can not be neglected if the hash
-function should be taken into use. The difference between the implementations
-in the official submissions and in supercop is first that the supercop
-implementation are continiously updated and that there are more specific
-hardwareimplementations available in supercop.
+\subsection{\ac{SUPERCOP}}
+% TODO: Sjekk hvordan det er med sitering her
+\ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
+of cryptographic software\cite{s_supercop}. In relation to hash functions
+\ac{SUPERCOP} measures the following:
 
-Results by the supercop are published at (URL).
+\begin{itemize}
+    \item Time to hash a very short packet of data
+    \item Time to hash a typical-size Internet packet
+    \item Time to hash a long message
+    \item Length of the hash output
+\end{itemize}
 
-Paralellizing
+The round 2 \ac{SHA}-3 candidates are all included in the toolkit, with a
+number of different optimalizations for each candidate. Optimalizations range
+from 32-/64-bit optimalizations, the use of extended instruction sets such as
+\ac{SSE2} and \ac{SSSE3}, optimalizations for different number of cores and
+others. Also the toolkit will try different compiler-optimalizations to get the
+best results possible for each function.
+% Skal jeg sitere kildekode her?
+
+The benchmarking results for the \ac{SUPERCOP} toolkit of \ac{SHA}-3 candidates
+and \ac{SHA}-2 functions on a number of different plattforms and architectures 
+are available at
+\footnote{\url{http://bench.cr.yp.to/results-sha3.html}}
 

### 2010-10-27 paper/main.tex
Added possibly better explanation of python and cython

@@
 
-\section{Python bindings using Cython}
+\section{Python and Cython}
 %What is it, how, why, alternatives?
-% TODO: Hvorfor er nedenfornevnte i ASM eller C?
-TAHOE-LAFS is written in python, but the SHA-3 candidate implementations found
+% TODO: Finn en kilde på pythonting, om det trengs?
+Python\footnote{\url{http://www.python.org}} is a high-level general-purpose
+programming language. Python is also an interpreted language, which means that
+python programs are compiled at runtime.  There exists multiple implementations
+of Python, however the most common, CPython, is implemented in C. The fact that
+Python is a high-level language usually means that there is performance penalty
+in contrast with lower level languages, shuch as C. To remedy this problem it
+is possible to write extensions to the language by using the official C-API.
+This approach has been used by several cryptographic libraries for python such
+as the official hashlib\footnote{\url{http://code.krypto.org/python/hashlib/}}
+and pycryptopp\footnote{\url{http://tahoe-lafs.org/trac/pycryptopp}}.
+
+Tahoe-\ac{LAFS} is written in python, but the SHA-3 candidate implementations found
 in both the official NIST submissions and supercop are either in C or in
-assembly. Therefore we need a way to make the implementations available in
-Python. There a several ways to do this. Python itself includes a C-API,  
-other options include SWIG, Cython and possible others.
-
-\subsection{Python C-API}
-The Python C-API is the official method of creating C-bindings for Python. The
-procedure to do this includes writing entrypoints or wrappers for your C
-functions in C. Then, through compiling you can acess your C-libraries and
-functions through Python. The drawback with this approach, when faced with an
-already existing C-library is that you have to actually write the
-wrapper-functions or modify the C-library.
+assembly. Therefore it is necessary to either implement the functions in python
+or make extensions to python through the C-API. Where extensions should give
+the best performance. % TODO: KILDE?
 
 \subsection{Cython}
-Cython tries to solve the problem of binding C-libraries and functions by the
-use of Python instead of the use of C which the official API does. This is done
-by only defining the entrypoints to the required C-functions and datatypes in a
-Python/C/C++ hybrid syntax. Cython also allows creating new C/C++ functions,
-available or not available to Python, in the same syntax. Cython works by
-transforming the hybrid syntax into C/C++-code which utilises the official
-Python C-API and compiling that. An expected tradeoff by using Cython is a
-performance drop due to the extra layer of abstraction.
+Cython\footnote{\url{http://www.cython.org}} is a language for writing Python
+extensions in a language that closely resembles the Python language itself.
+Basically what Cython does is to parse cython files and generate C source files
+that utilises the Python C-API. Afterwards the generated source files can be
+compiled as if they were written directly in C. A small performance tradeoff is
+to be expected from this, due to the extra layer of abstraction.
 

### 2010-10-27 paper/main.tex
Moved something and wrote two more words.

@@
 Tahoe has a three layer architecture: the key-value store, the filesystem, and
-the application.
+the application \cite{t_tahoe}.
 
@@
 
-\subsection{Related technologies}
+\paragraph{File types.}
 
-Tahoe-\ac{LAFS} uses a number of techniques to realize the functionality
-needed by a secure, distributed filesystem.
+There are two kinds of files in the Tahoe-\ac{LAFS} -- \textbf{immutable} and
+\textbf{mutable} files.
 
@@
 
+\subsection{Use of secure hashes in Tahoe-LAFS}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{Tahoe-hashing.pdf}
+    \caption{Tahoe-LAFS: Location of hashing.}
+    \label{fig:tahoehashing}
+    \emph{Figure from
+     \href{http://tahoe-lafs.org/source/tahoe/trunk/docs/specifications/mut.svg}
+     {tahoe-lafs.org}}
+\end{figure}
+
+As seen in figure \ref{fig:tahoehashing}, the usage of secure hashing is
+extensive in Tahoe, and has to be considered as a key part of the functionality.
+
 \paragraph{Hash Trees.}
@@
 
-\subsection{Use of secure hashes in Tahoe-LAFS}
+\paragraph{Encryption key.}
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.9\columnwidth]{Tahoe-hashing.pdf}
-    \caption{Tahoe-LAFS: Location of hashing}
-    \label{fig:tahoehashing}
-\end{figure}
+\paragraph{Ciphertext and shares.}
 
-As seen in figure \ref{fig:tahoehashing}, the usage of secure hashing is
-extensive in Tahoe, and has to be considered as a key part of the functionality.
+Stored in Capability/URI Extension Block.
+URI Extension Block $=$ Root hash $+$ file size $+$ encoding parameters.
+URI / File read cap = Hash of UEB $+ K_c$
+
+\paragraph{Storage Index.}
+
+file-encoding1.svg
 

### 2010-10-29 python-bindings/tests/test_extremelylongkat.py
Added crude sha256 benchmark to extremelylongkat

@@ -4,6 +4,7 @@ from sha3lib import hash_functions
 from sha3lib import *
 from os import path
 import time
+from pycryptopp.hash.sha256 import SHA256 as sha256
 
 class TestExtremelyLongKat(unittest.TestCase):
 
@@ -54,8 +55,17 @@ class TestExtremelyLongKat(unittest.TestCase):
 
             print '\t%s: Tested ExtremelyLong ok took %0.3f ms' % (function,
                     (t2-t1)*1000.0)
-            
 
+        # Faker en sha256 for å sjekke output
+        s = sha256()
+        t1 = time.time()
+        for l in xrange(int(props['Repeat'])):
+            s.update(props['Text'])
+        output = s.hexdigest()
+        t2 = time.time()
+       
+        print '\t%s: Tested ExtremelyLong ok took %0.3f ms' % ('sha256',
+                    (t2-t1)*1000.0)
 
 if __name__ == '__main__':
     unittest.main()

### 2010-10-29 paper/main.tex
Added short text on mutable vs immutable. Needs more work.
Added figure of an UEB and some text on capabilities in relation to hashing.

@@
 
-\usepackage[numbers]{natbib}
+\usepackage[numbers]{natbib} % We need to tweak our referencing a bit.
 \usepackage{appendix} % Fixes formatting of appendices
@@
 \acro{AES}{Advanced Encryption Standard}
+\acro{CEB}{Capability Extension Block}
 \acro{FEC}{Forward Error Correction}
@@
 \acro{RAM}{Random Access Memory}
+\acro{RAID}{Redundant Array of Independent Disks}
 \acro{SHA}{Secure Hash Algorithm}
@@
 Cryptographic Operations and Primitives}
+\acro{UEB}{URI Extension Block}
 \end{acronym}
@@
 There are two kinds of files in the Tahoe-\ac{LAFS} -- \textbf{immutable} and
-\textbf{mutable} files.
+\textbf{mutable} files. An immutable file is created exactly once, i.e. it
+cannot be modified, and can be read repeatedly. Mutable files can be modified,
+and everyone who has access to the \emph{signing key} can make new versions of
+the mutable file.
+%More text here?
 
@@
 
-The properties of erasure coding can be thought of as those of parity bits in
-RAID\footnote{Redundant Array of Independent Disks} systems. \citet*{t_erasure}
-compare erasure coding and plain replication, and conclude that ``\emph{...
-erasure codes have mean time to failures many orders of magnitude higher than
-replicated systems with similar storage and bandwidth requirements.}''
+The properties of erasure coding can be thought of as those of replication in
+\acsu{RAID} systems. \citet*{t_erasure} compare erasure coding and plain
+replication, and conclude that ``\emph{...  erasure codes have mean time to
+failures many orders of magnitude higher than replicated systems with similar
+storage and bandwidth requirements.}''
 
@@
 
-\begin{figure}[h]
+\begin{figure}[!h]
     \centering
     \includegraphics[width=0.9\columnwidth]{Tahoe-hashing.pdf}
-    \caption{Tahoe-LAFS: Location of hashing.}
+    \caption{Tahoe-LAFS: Hashing operations in mutable files.}
     \label{fig:tahoehashing}
@@
 As seen in figure \ref{fig:tahoehashing}, the usage of secure hashing is
-extensive in Tahoe, and has to be considered as a key part of the functionality.
+extensive in Tahoe, and has to be considered as a key part of the functionality,
+and thus affecting the performance.
 
@@
 
-A hash tree, also known as \emph{Merkle tree}, is a type of data structure
+A hash tree, also known as a \emph{Merkle tree}, is a type of data structure
 which can be described as a tree with nodes that can verify all information
@@
 
-\paragraph{Encryption key.}
+These secure hashes of the shares of a file, are contained within a small
+datastructure named the \emph{\ac{CEB}}.
 
-\paragraph{Ciphertext and shares.}
+\paragraph{Capabilities.}
 
-Stored in Capability/URI Extension Block.
-URI Extension Block $=$ Root hash $+$ file size $+$ encoding parameters.
-URI / File read cap = Hash of UEB $+ K_c$
+A \textbf{capability} (or an URI) contains the encryption key, and a hash of the
+\emph{\ac{UEB}}. The \ac{UEB} for each file is a data structure containing the
+hash of the \ac{CEB}, the size of the file, any encoding parameters necessary to
+perform the erasure decoding, and a hash of the plaintext and the encrypted
+text. This is illustrated in figure \ref{fig:tahoeueb}.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{Tahoe-UEB.pdf}
+    \caption{The composition of an \ac{UEB}.}
+    \label{fig:tahoeueb}
+\end{figure}
 
@@
 
-file-encoding1.svg
+A hash of the encryption key is used to form the "storage index", which is
+used for both server selection and to index shares within each storage node.
+% More here?
 

### 2010-11-01 paper/main.tex
Updated tahoe hashing example image

@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{Tahoe-hashing.pdf}
-    \caption{Tahoe-LAFS: Hashing operations in mutable files.}
+    \includegraphics[width=0.9\columnwidth]{Tahoe-hashes.pdf}
+    \caption{Tahoe-LAFS: Example of hashing operations.}
     \label{fig:tahoehashing}
-    \emph{Figure from
-     \href{http://tahoe-lafs.org/source/tahoe/trunk/docs/specifications/mut.svg}
-     {tahoe-lafs.org}}
+    \emph{Figure based on
+     \href{http://tahoe-lafs.org/source/tahoe/trunk/docs/specifications/CHK-hashes.svg}
+     {CHK-hashes.svg from tahoe-lafs.org}}
+     %TODO: Footnote with full URL using footnotemark
 \end{figure}

### 2010-11-09 python-bindings/setup.py
Added compile flags to groestl

@@ -19,7 +19,8 @@ ext_modules = [
         extra_compile_args=['-funroll-loops','-m32','-march=pentium4','-O3','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.groestl.groestl_hash",
         [hashpath+"groestl/groestl_hash.pyx", hashpath+'groestl/sha3nist.c',
-            hashpath+"groestl/groestl.c"],),
+            hashpath+"groestl/groestl.c"],
+        extra_compile_args=['-m32','-march=pentium3','-O','-fomit-frame-pointer'],),
     Extension("sha3lib.hash_functions.echo.echo_hash",
         [hashpath+"echo/echo_hash.pyx", hashpath+'echo/sha3nist.c',
             hashpath+"echo/echo.c"],

### 2010-11-09 python-bindings/sha3lib/hash_functions/groestl/sph_types.h

@@ -1,4 +1,4 @@
-/* $Id: sph_types.h 230 2010-06-16 21:39:32Z tp $ */
+/* $Id: sph_types.h 250 2010-09-10 20:23:02Z tp $ */
 /**
  * Basic type definitions.
  *
@@ -985,6 +985,8 @@ typedef long long sph_s64;
  * SPH_I386_MSVC        x86-compatible (32-bit) with Microsoft Visual C
  * SPH_AMD64_GCC        x86-compatible (64-bit) with gcc
  * SPH_AMD64_MSVC       x86-compatible (64-bit) with Microsoft Visual C
+ * SPH_PPC32_GCC        PowerPC, 32-bit, with gcc
+ * SPH_PPC64_GCC        PowerPC, 64-bit, with gcc
  *
  * TODO: enhance automatic detection, for more architectures and compilers.
  * Endianness is the most important. SPH_UNALIGNED and SPH_UPTR help with
@@ -1075,6 +1077,21 @@ typedef long long sph_s64;
  */
 #elif defined __powerpc__ || defined __POWERPC__ || defined __ppc__ \
 	|| defined _ARCH_PPC
+
+/*
+ * Note: we do not declare cross-endian access to be "fast": even if
+ * using inline assembly, implementation should still assume that
+ * keeping the decoded word in a temporary is faster than decoding
+ * it again.
+ */
+#if defined __GNUC__
+#if SPH_64_TRUE
+#define SPH_DETECT_PPC64_GCC         1
+#else
+#define SPH_DETECT_PPC32_GCC         1
+#endif
+#endif
+
 #if defined __BIG_ENDIAN__ || defined _BIG_ENDIAN
 #define SPH_DETECT_BIG_ENDIAN        1
 #elif defined __LITTLE_ENDIAN__ || defined _LITTLE_ENDIAN
@@ -1142,6 +1159,12 @@ typedef long long sph_s64;
 #endif
 #if defined SPH_DETECT_AMD64_MSVC && !defined SPH_AMD64_MSVC
 #define SPH_AMD64_MSVC        SPH_DETECT_AMD64_MSVC
+#endif
+#if defined SPH_DETECT_PPC32_GCC && !defined SPH_PPC32_GCC
+#define SPH_PPC32_GCC         SPH_DETECT_PPC32_GCC
+#endif
[…]

### 2010-11-09 paper/main.tex

@@
 
+%**************************************%
 \chapter{Introduction}
+%**************************************%
 \pagenumbering{arabic}
@@
 
-Thorough problem description
+Thorough problem description. \\
+Stepwise, what will we actually do?
 
@@
 
+%**************************************%
 \chapter{Background technologies}
-% Find new and better title
+%**************************************%
+% FIXME: Find new and better title
 
@@
 \subsection{\ac{SUPERCOP}}
-% TODO: Sjekk hvordan det er med sitering her
+% TODO: Mention that some SUPERCOP candidates are missing UPDATE-functionality.
 \ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
@@
 % Skal jeg sitere kildekode her?
+% TODO: Insert image to visualize the forest of different implementations.
 
@@
 
+%**************************************%
 \chapter{Procedure}
+%**************************************%
 
-TODO: We have to figure out a new title
-Keywords: Python bindings in detail, unit tests, using it in Tahoe-LAFS
+% FIXME: A more descriptive/better chapter title
+%
+%A bit more technical description of how we tested the candidates in Tahoe-LAFS
+% - Limitations/assumptions
+% - Optimized candidates.
+% - Error sources.
+% The points above should be taken into consideration on each section in this
+% chapter.
 
-A bit more technical description of how we tested the candidates in Tahoe-LAFS
+In this chapter we will outline the procedure followed to complete our task.
 
-Limitations/assumptions
+\section{Choosing implementations of the SHA-3 Candidates}
 
-Optimized candidates.
+% sphlib, SUPERCOP, NIST
 
-Error sources.
+\section{Python bindings}
 
-\chapter{Results}
+Since Tahoe-\ac{LAFS} is written in Python, we needed some way of interacting
+with the C implementations of the NIST candidates.
 
+\subsection{Unit tests}
+% Include the use of Known Answer Tests
+
+\section{Modifications in the Tahoe-LAFS Code}
+
+% This should be a short text, but maybe drag in a discussion of how we SHOULD
+% have implemented it if we had the time, ref. capabilities.
+
+\section{Test vectors and Configuration}
+
+\subsection{Physical Setup}
+
+\section{Distribution of the Code}
+% Distutil, bash scripts
+
+%**************************************%
+\chapter{Measurements and Results}
+%**************************************%
+
+% Maybe cite
+% \url{http://tahoe-lafs.org/trac/tahoe-lafs/wiki/Performance?version=32}
 What is tested and what is not tested?
 
-tables, graphs?
+How much time is spent on hashing versus upload over network?
 
+How much gain can Tahoe-LAFS get by switching to SHA-3?
+
+Plenty of tables, graphs
+
+%**************************************%
 \chapter{Discussion}
+%**************************************%
 
@@
 
-\cite{s_nistround2}
+%**************************************%
+\chapter{Conclusion and Future Work}
+%**************************************%
 
-\chapter{Conclusion}
-
-A somewhat short conclusion summing up results and discussion
-
-\section{Future work}
+A somewhat short conclusion summing up what has been done, and results and
+discussion
 

### 2010-11-10 python-bindings/setup.py
Added hamsi simd-1 version

@@ -55,8 +55,8 @@ ext_modules = [
         [hashpath+"cubehash/cubehash_hash.pyx",hashpath+"cubehash/sha3nist.c",
             hashpath+"cubehash/cubehash.c"],),
     Extension("sha3lib.hash_functions.hamsi.hamsi_hash",
-        [hashpath+"hamsi/hamsi_hash.pyx",hashpath+"hamsi/sha3nist.c",
-            hashpath+"hamsi/hamsi.c",hashpath+"hamsi/hamsi_helper.c"],),
+        [hashpath+"hamsi/hamsi_hash.pyx",hashpath+"hamsi/hamsi.c"],
+        extra_compile_args=['-msse2','-funroll-loops','-m32','-march=pentium-m','-03','-fomit-frame-pointer']),
     ]
 
 class RunTests(Command):

### 2010-11-10 python-bindings/sha3lib/hash_functions/hamsi/hamsi.c

@@ -1,859 +1,387 @@
-/* $Id: hamsi.c 251 2010-10-19 14:31:51Z tp $ */
-/*
- * Hamsi implementation.
- *
- * ==========================(LICENSE BEGIN)============================
- *
- * Copyright (c) 2007-2010  Projet RNRT SAPHIR
- * 
- * Permission is hereby granted, free of charge, to any person obtaining
- * a copy of this software and associated documentation files (the
- * "Software"), to deal in the Software without restriction, including
- * without limitation the rights to use, copy, modify, merge, publish,
- * distribute, sublicense, and/or sell copies of the Software, and to
- * permit persons to whom the Software is furnished to do so, subject to
- * the following conditions:
- * 
- * The above copyright notice and this permission notice shall be
- * included in all copies or substantial portions of the Software.
- * 
- * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
- * EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
- * MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT.
- * IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY
- * CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT,
- * TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE
- * SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
- *
- * ===========================(LICENSE END)=============================
- *
- * @author   Thomas Pornin <thomas.pornin@cryptolog.com>
- */
+#include <emmintrin.h>
 
-#include <stddef.h>
-#include <string.h>
+#include "hamsi.h"
+#include "hamsi-tables.h"
 
[…]

### 2010-11-10 python-bindings/sha3lib/hash_functions/hamsi/hamsi_hash.pyx

@@ -11,6 +11,10 @@ def hash(int hashbitlen, bytes data, int in_length):
     cdef char* c_string = <char *> data
     cdef hamsi_hash_h.DataLength datalen = in_length
     cdef hamsi_hash_h.BitSequence *hashval = <hamsi_hash_h.BitSequence *> malloc(hashbitlen*8)
+    
+    #cdef hamsi_hash_h.hashState state
+    #hamsi_hash_h.Init(&state, 256)
+    #return 0
 
     hamsi_hash_h.Hash(hashbitlen, <hamsi_hash_h.BitSequence *> c_string, datalen, hashval)
 
@@ -38,8 +42,10 @@ cdef class hamsi:
     def __init__(self, int in_hashbitlen, bytes initial=None):
         self.finished = 0
         self.hashbitlen = in_hashbitlen
-        hamsi_hash_h.Init(&self.state, self.hashbitlen)
-
+        cdef hamsi_hash_h.hashState state
+        hamsi_hash_h.Init(&state, 256)
+        self.state = state
+        
         if initial:
             self.update(initial)
 
@@ -51,7 +57,12 @@ cdef class hamsi:
             self.finished = 0
             self.state = self.previous_state
 
-        hamsi_hash_h.Update(&self.state, <hamsi_hash_h.BitSequence *> data, data_len)
+        self.c_update(data, data_len)
+
+    cdef c_update(self, char* data, int data_len):
+        cdef hamsi_hash_h.hashState state = self.state
+        hamsi_hash_h.Update(&state, <hamsi_hash_h.BitSequence *> data, data_len)
+        self.state = state
 
     cpdef final(self):
         cdef hamsi_hash_h.BitSequence *hashval = <hamsi_hash_h.BitSequence *> malloc(self.hashbitlen*8)
@@ -59,11 +70,13 @@ cdef class hamsi:
         # We copy the state so that we can continue to update.
         # This equals hashlibs functionality, but not pycryptopp.
         self.previous_state = self.state
-
-        hamsi_hash_h.Final(&self.state, hashval)
+        
+        cdef hamsi_hash_h.hashState state = self.state
+        hamsi_hash_h.Final(&state, hashval)
         self.finished = 1
         
         self.hashval = [hashval[i] for i from 0 <= i < self.hashbitlen / 8]
+        self.state = state
         free(hashval)
 
     cpdef copy(self):

### 2010-11-10 python-bindings/sha3lib/hash_functions/hamsi/sha3nist.h

@@ -1,96 +1,0 @@
-/* $Id: sha3nist.h 170 2010-05-07 13:27:24Z tp $ */
-/**
- * NIST SHA-3 competition compatibility header.
- *
- * This file defines the API specified by NIST for the SHA-3 competition
- * candidates; it is applied over one of the compatible implementations
- * in this library. This header must first be modified to select a
- * function (inclusion of the relevant header files, and value of the
- * <code>SPH_NIST</code> macro).
- *
- * ==========================(LICENSE BEGIN)============================
- *
- * Copyright (c) 2007-2010  Projet RNRT SAPHIR
- * 
- * Permission is hereby granted, free of charge, to any person obtaining
- * a copy of this software and associated documentation files (the
- * "Software"), to deal in the Software without restriction, including
- * without limitation the rights to use, copy, modify, merge, publish,
- * distribute, sublicense, and/or sell copies of the Software, and to
- * permit persons to whom the Software is furnished to do so, subject to
- * the following conditions:
- * 
- * The above copyright notice and this permission notice shall be
- * included in all copies or substantial portions of the Software.
- * 
- * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
- * EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
- * MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT.
- * IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY
- * CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT,
- * TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE
- * SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
- *
- * ===========================(LICENSE END)=============================
- *
- * @file     sha3nist.h
- * @author   Thomas Pornin <thomas.pornin@cryptolog.com>
- */
-
-#ifndef SPH_SHA3NIST_H__
-#define SPH_SHA3NIST_H__
-
-#ifndef DOXYGEN_IGNORE
-
[…]

### 2010-11-10 python-bindings/sha3lib/hash_functions/hamsi/sph_sha3.h

@@ -1,60 +1,0 @@
-/* $Id: sph_sha3.h 208 2010-06-02 20:33:00Z tp $ */
-/**
- * SHA-3 is the colloquial codename for the future standard hash function
- * from NIST. SHA-3 will be "API compatible" (same range of output sizes)
- * with the SHA-2 family (SHA-224, SHA-256, SHA-384 and SHA-512). This
- * library implements a number of SHA-3 candidates (functions submitted
- * to the NIST competition for selection of SHA-3). This header file
- * includes the header files for SHA-2 and all the implemented SHA-3
- * candidates.
- *
- * ==========================(LICENSE BEGIN)============================
- *
- * Copyright (c) 2007-2010  Projet RNRT SAPHIR
- * 
- * Permission is hereby granted, free of charge, to any person obtaining
- * a copy of this software and associated documentation files (the
- * "Software"), to deal in the Software without restriction, including
- * without limitation the rights to use, copy, modify, merge, publish,
- * distribute, sublicense, and/or sell copies of the Software, and to
- * permit persons to whom the Software is furnished to do so, subject to
- * the following conditions:
- * 
- * The above copyright notice and this permission notice shall be
- * included in all copies or substantial portions of the Software.
- * 
- * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
- * EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
- * MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT.
- * IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY
- * CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT,
- * TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE
- * SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
- *
- * ===========================(LICENSE END)=============================
- *
- * @file     sph_sha3.h
- * @author   Thomas Pornin <thomas.pornin@cryptolog.com>
- */
-
-#ifndef SPH_SHA3_H__
-#define SPH_SHA3_H__
-
[…]

### 2010-11-10 paper/main.tex

@@
 \section{Cryptographic Hash Functions}
-% Get a source for this %
-A Cryptographic Hash function is a deterministic mathematical procedure which
-takes an arbritary block of data and outputs a fixed-size bit string. The
-output is refered to as the hash value, message digest or simply digest.
-Another property of a cryptographic hash function is that a small change in the
-input data (just 1 bit) should completly change the output of the hash
+
+A cryptographic hash function is a deterministic mathematical procedure which
+takes an arbitrary block of data and outputs a fixed-size bit string. The output
+is referred to as the hash value, message digest or simply digest.
+Another property of a cryptographic hash function is that the smallest change in
+the input data (e.g. one bit) should completely change the output of the hash
 function. In other words it should be infeasible to find the reverse of a
-cryptographic hash function. It should also be infeasible to find two blocks of
-data which produce the same hash value (a collision).
+cryptographic hash function \cite{stallings}. It should also be infeasible to
+find two blocks of data which produce the same hash value (a \emph{collision}).
 
 \subsection{NIST SHA-3 Competition}
-The \ac{SHA} version 3 is a comming standard set to superceed the current
-standards \ac{SHA}-1 and the \ac{SHA}-2 family. The hash function to become
-\ac{SHA}-3 will be decided by \ac{NIST} and choosen between the submitted
-contestants to the \ac{NIST} hash competition. The current status of the 
-competition is officially called Round 2, with 14 of 64 candidates having 
-survived round 1\cite{s_fedreg}.
+The \ac{SHA} version 3 is a coming standard set to supersede the current
+standards that the \ac{SHA}-1 and the \ac{SHA}-2 family has become. The hash
+function that will be known as \ac{SHA}-3 will be decided by the \ac{NIST} and
+chosen between the submitted contestants to the \ac{NIST} hash competition.
+At the time of writing, the current status of the competition is officially
+called Round 2, with 14 of 64 candidates having ``survived'' Round 1
+\cite{s_fedreg}.
 
-\subsubsection{\ac{NIST} evaluation criterias for \ac{SHA}-3}
+\subsubsection{\ac{NIST} evaluation criteria for \ac{SHA}-3}
 
-\paragraph{Security} is the most important criteria for the SHA-3
-candidates\cite{s_nistround2}. The full description of security criterias are
-mentioned in \cite{s_fedreg}, however the most noteworthy criteria in regards
-to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
+\paragraph{Security.} The most important criteria for the SHA-3
+candidates\cite{s_nistround2} is security. The full description of security
+criteria are mentioned in \cite{s_fedreg}, however the most noteworthy criterion
+in regards to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
 resistance against length-extension attacks. Both \ac{SHA}-1 and the \ac{SHA}-2
-family are vulnerable to this attack\cite{s_mdre}.
+family are vulnerable to this attack\cite{s_mdre}, and this requires Tahoe to
+run the algorithm twice, as described by .
+%TODO: cite Practical cryptography/Niels Ferguson, Bruce Schneier
 
-\paragraph{Cost and Performance} are considered to be the 2nd most important
-criteria. The absolute minimum for performance is that the SHA-3 candidate
-should be faster than the functions in the \ac{SHA}-2 family. Cost is a measure
-of how much memory an implementation requires in software, both the 
-implementation itself and the use of \ac{RAM} during runtime. Another measure
-of cost is how many logic gates it takes to implement the function in hardware.
+\paragraph{Cost and Performance.} Cost and Performance are considered to be the
+2nd most important criterion. The absolute minimum for performance is that the
+SHA-3 candidate should be faster than the functions in the \ac{SHA}-2 family.
+Cost is a measure of how much memory an implementation requires in software,
+both the implementation itself and the use of \ac{RAM} during runtime. Another
+measure of cost is how many logic gates it takes to implement the function in
+hardware.
 
-\paragraph{Algorithm and implementation characteristics} is the third criteria.
-By this criteria \ac{NIST} emphasis that algorithms with greater flexbility
-will be given preference over other algorithms\cite{s_nistround2}. By flexible
-they imply the possiblity of the function to run efficiently on a varity of
-plattforms, to use parallelism or instruction set extensions. Another key point
-of a flexible hash function is that it should have a simple and elegant design
-to encourage understanding, analysis and design confidence.
+\paragraph{Algorithm and implementation characteristics.} By this 3rd criteria,
+\ac{NIST} emphasis that algorithms with greater flexibility will be given
+preference over other algorithms \cite{s_nistround2}. By flexible they imply the
+possibility of the function to run efficiently on a variety of platforms, and to
+use parallelism or instruction set extensions. Another key point of a flexible
+hash function is that it should have a simple and elegant design to encourage
+understanding, analysis and design confidence.
 
@@
 \ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
-of cryptographic software\cite{s_supercop}. In relation to hash functions
+of cryptographic software \cite{s_supercop}. In relation to hash functions
 \ac{SUPERCOP} measures the following:
@@
 \begin{itemize}
-    \item Time to hash a very short packet of data
-    \item Time to hash a typical-size Internet packet
-    \item Time to hash a long message
-    \item Length of the hash output
+    \item Time to hash a very short packet of data.
+    \item Time to hash a typical-size Internet packet.
+    \item Time to hash a long message.
+    \item Length of the hash output.
 \end{itemize}
 
-The round 2 \ac{SHA}-3 candidates are all included in the toolkit, with a
-number of different optimalizations for each candidate. Optimalizations range
-from 32-/64-bit optimalizations, the use of extended instruction sets such as
-\ac{SSE2} and \ac{SSSE3}, optimalizations for different number of cores and
-others. Also the toolkit will try different compiler-optimalizations to get the
+The Round 2 \ac{SHA}-3 candidates are all included in the toolkit, with a number
+of different optimizations for each candidate. Optimizations range from
+32-/64-bit specific implementations, the use of extended instruction sets such
+as \ac{SSE2} and \ac{SSSE3}, optimizations for different number of cores and
+others. The toolkit will also try different compiler optimizations to get the
 best results possible for each function.
-% Skal jeg sitere kildekode her?
 % TODO: Insert image to visualize the forest of different implementations.
 
-The benchmarking results for the \ac{SUPERCOP} toolkit of \ac{SHA}-3 candidates
-and \ac{SHA}-2 functions on a number of different plattforms and architectures 
-are available at
+The benchmarking results for the \ac{SUPERCOP} toolkit of the \ac{SHA}-3
+candidates and \ac{SHA}-2 functions on a number of different plattforms and
+architectures are available at
 \footnote{\url{http://bench.cr.yp.to/results-sha3.html}}
+% TODO: Link to appendix with tahoe04 SUPERCOP results.
 

### 2010-11-10 python-bindings/setup.py
Added keccackc512 sseu2

@@ -40,8 +40,9 @@ ext_modules = [
         [hashpath+"simd/simd.pyx",hashpath+"simd/optimized.c",
             hashpath+"simd/nist.c"],),
     Extension("sha3lib.hash_functions.keccak.keccak",
-        [hashpath+"keccak/keccak.pyx",hashpath+"keccak/KeccakF-1600-opt32.c",
-            hashpath+"keccak/KeccakSponge.c"],),
+        [hashpath+"keccak/keccak.pyx",hashpath+"keccak/KeccakF-1600-opt64.c",
+            hashpath+"keccak/KeccakSponge.c"],
+        extra_compile_args=['-msse2','-march=barcelona','-O3','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.jh.jh_hash",
         [hashpath+"jh/jh_hash.pyx",hashpath+"jh/sha3nist.c",hashpath+"jh/jh.c"],),
     Extension("sha3lib.hash_functions.fugue.fugue_hash",

### 2010-11-10 python-bindings/sha3lib/hash_functions/keccak/KeccakF-1600-opt32-settings.h

@@ -1,2 +1,0 @@
-#define Unrolling 2
-#define UseSchedule 3

### 2010-11-10 python-bindings/sha3lib/hash_functions/keccak/keccak.pyx

@@ -38,8 +38,10 @@ cdef class keccak:
     def __init__(self, int in_hashbitlen, bytes initial=None):
         self.finished = 0
         self.hashbitlen = in_hashbitlen
-        keccak_h.Init(&self.state, self.hashbitlen)
-
+        cdef keccak_h.hashState state
+        keccak_h.Init(&state, self.hashbitlen)
+        self.state = state
+    
         if initial:
             self.update(initial)
 
@@ -51,7 +53,9 @@ cdef class keccak:
             self.finished = 0
             self.state = self.previous_state
 
-        keccak_h.Update(&self.state, <keccak_h.BitSequence *> data, data_len)
+        cdef keccak_h.hashState state = self.state
+        keccak_h.Update(&state, <keccak_h.BitSequence *> data, data_len)
+        self.state = state
 
     cpdef final(self):
         cdef keccak_h.BitSequence *hashval = <keccak_h.BitSequence *> malloc(self.hashbitlen*8)
@@ -60,7 +64,9 @@ cdef class keccak:
         # This equals hashlibs functionality, but not pycryptopp.
         self.previous_state = self.state
 
-        keccak_h.Final(&self.state, hashval)
+        cdef keccak_h.hashState state = self.state
+        keccak_h.Final(&state, hashval)
+        self.state = state
         self.finished = 1
         
         self.hashval = [hashval[i] for i from 0 <= i < self.hashbitlen / 8]

### 2010-11-10 python-bindings/setup.py
Added luffa256_sse2

@@ -33,7 +33,8 @@ ext_modules = [
         [hashpath+"skein/skein_hash.pyx", hashpath+'skein/SHA3api_ref.c',
             hashpath+"skein/skein_block.c",hashpath+"skein/skein.c"],),
     Extension("sha3lib.hash_functions.luffa.luffa",
-        [hashpath+"luffa/luffa.pyx", hashpath+'luffa/luffa_for_32.c'],),
+        [hashpath+"luffa/luffa.pyx", hashpath+'luffa/luffa_for_sse2.c'],
+        extra_compile_args=['-msse2','-funroll-loops','-march=k8','-O3','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.shavite3.shavite3",
         [hashpath+"shavite3/shavite3.pyx", hashpath+'shavite3/SHAvite3.c'],),
     Extension("sha3lib.hash_functions.simd.simd",

### 2010-11-10 python-bindings/sha3lib/hash_functions/luffa/hash.c

@@ -1,9 +1,0 @@
-#include "crypto_hash.h"
-#include "luffa_for_32.h"
-
-int crypto_hash(unsigned char *out, const unsigned char *in, unsigned long long inlen)
-{
-	if (Hash(crypto_hash_BYTES * 8,in,inlen * 8,out) == SUCCESS) return 0;
-	return -1;
-}
-

### 2010-11-10 python-bindings/sha3lib/hash_functions/luffa/luffa.pyx

@@ -38,7 +38,10 @@ cdef class luffa:
     def __init__(self, int in_hashbitlen, bytes initial=None):
         self.finished = 0
         self.hashbitlen = in_hashbitlen
-        luffa_h.Init(&self.state, self.hashbitlen)
+        
+        cdef luffa_h.hashState state
+        luffa_h.Init(&state, self.hashbitlen)
+        self.state = state
 
         if initial:
             self.update(initial)
@@ -50,8 +53,10 @@ cdef class luffa:
         if self.finished:
             self.finished = 0
             self.state = self.previous_state
-
-        luffa_h.Update(&self.state, <luffa_h.BitSequence *> data, data_len)
+        
+        cdef luffa_h.hashState state = self.state
+        luffa_h.Update(&state, <luffa_h.BitSequence *> data, data_len)
+        self.state = state
 
     cpdef final(self):
         cdef luffa_h.BitSequence *hashval = <luffa_h.BitSequence *> malloc(self.hashbitlen*8)
@@ -59,8 +64,10 @@ cdef class luffa:
         # We copy the state so that we can continue to update.
         # This equals hashlibs functionality, but not pycryptopp.
         self.previous_state = self.state
-
-        luffa_h.Final(&self.state, hashval)
+        
+        cdef luffa_h.hashState state = self.state
+        luffa_h.Final(&state, hashval)
+        self.state = state
         self.finished = 1
         
         self.hashval = [hashval[i] for i from 0 <= i < self.hashbitlen / 8]

### 2010-11-10 python-bindings/sha3lib/hash_functions/luffa/luffa_for_32.c

@@ -1,916 +1,0 @@
-/*
- * luffa_for_32.c
- * Version 2.0 (Sep 15th 2009)
- *
- * Copyright (C) 2008-2009 Hitachi, Ltd. All rights reserved.
- *
- * Hitachi, Ltd. is the owner of this software and hereby grant
- * the U.S. Government and any interested party the right to use
- * this software for the purposes of the SHA-3 evaluation process,
- * notwithstanding that this software is copyrighted.
- *
- * THE SOFTWARE IS PROVIDED "AS IS" AND THE AUTHOR DISCLAIMS ALL WARRANTIES
- * WITH REGARD TO THIS SOFTWARE INCLUDING ALL IMPLIED WARRANTIES OF
- * MERCHANTABILITY AND FITNESS. IN NO EVENT SHALL THE AUTHOR BE LIABLE FOR
- * ANY SPECIAL, DIRECT, INDIRECT, OR CONSEQUENTIAL DAMAGES OR ANY DAMAGES
- * WHATSOEVER RESULTING FROM LOSS OF USE, DATA OR PROFITS, WHETHER IN AN
- * ACTION OF CONTRACT, NEGLIGENCE OR OTHER TORTIOUS ACTION, ARISING OUT OF
- * OR IN CONNECTION WITH THE USE OR PERFORMANCE OF THIS SOFTWARE.
- */
-
-#include "luffa_for_32.h"
-
-
-#ifdef HASH_BIG_ENDIAN
-# define BYTES_SWAP32(x) x
-# define BYTES_SWAP64(x) x
-#else
-# define BYTES_SWAP32(x)                                                \
-    ((x << 24) | ((x & 0x0000ff00) << 8) | ((x & 0x00ff0000) >> 8) | (x >> 24))
-
-# define BYTES_SWAP64(x)                                                \
-    (((u64)(BYTES_SWAP32((u32)(x))) << 32) | (u64)(BYTES_SWAP32((u32)((x) >> 32))))
-#endif /* HASH_BIG_ENDIAN */
-
-/* BYTES_SWAP256(x) stores each 32-bit word of 256 bits data in little-endian convention */
-#define BYTES_SWAP256(x) {                                      \
-    int _i = 8; while(_i--){x[_i] = BYTES_SWAP32(x[_i]);}   \
-    }
-
-/* BYTES_SWAP512(x) stores each 64-bit word of 512 bits data in little-endian convention */
[…]

### 2010-11-10 python-bindings/sha3lib/hash_functions/luffa/luffa_for_32.h

@@ -1,66 +1,0 @@
-/*
- * luffa_for_32.h
- * Version 2.0 (Sep 15th 2009)
- *
- * Copyright (C) 2008-2009 Hitachi, Ltd. All rights reserved.
- *
- * Hitachi, Ltd. is the owner of this software and hereby grant
- * the U.S. Government and any interested party the right to use
- * this software for the purposes of the SHA-3 evaluation process,
- * notwithstanding that this software is copyrighted.
- *
- * THE SOFTWARE IS PROVIDED "AS IS" AND THE AUTHOR DISCLAIMS ALL WARRANTIES
- * WITH REGARD TO THIS SOFTWARE INCLUDING ALL IMPLIED WARRANTIES OF
- * MERCHANTABILITY AND FITNESS. IN NO EVENT SHALL THE AUTHOR BE LIABLE FOR
- * ANY SPECIAL, DIRECT, INDIRECT, OR CONSEQUENTIAL DAMAGES OR ANY DAMAGES
- * WHATSOEVER RESULTING FROM LOSS OF USE, DATA OR PROFITS, WHETHER IN AN
- * ACTION OF CONTRACT, NEGLIGENCE OR OTHER TORTIOUS ACTION, ARISING OUT OF
- * OR IN CONNECTION WITH THE USE OR PERFORMANCE OF THIS SOFTWARE.
- */
-
-/* The length of digests*/
-#define DIGEST_BIT_LEN_224 224
-#define DIGEST_BIT_LEN_256 256
-#define DIGEST_BIT_LEN_384 384
-#define DIGEST_BIT_LEN_512 512
-
-/*********************************/
-/* The parameters of Luffa       */
-#define MSG_BLOCK_BIT_LEN 256  /*The bit length of a message block*/
-#define MSG_BLOCK_BYTE_LEN (MSG_BLOCK_BIT_LEN >> 3) /* The byte length
-                                                     * of a message block*/
-
-/* The number of blocks in Luffa */
-#define WIDTH_224 3
-#define WIDTH_256 3
-#define WIDTH_384 4
-#define WIDTH_512 5
-
-/* The limit of the length of message */
-#define LIMIT_224 64
-#define LIMIT_256 64
-#define LIMIT_384 128
-#define LIMIT_512 128
-/*********************************/
-
-
-typedef unsigned char BitSequence;
-typedef unsigned long long DataLength;
[…]

### 2010-11-10 python-bindings/setup.py
Added new optimized shavite3-256, that does not conform to KAT

@@ -36,7 +36,8 @@ ext_modules = [
         [hashpath+"luffa/luffa.pyx", hashpath+'luffa/luffa_for_sse2.c'],
         extra_compile_args=['-msse2','-funroll-loops','-march=k8','-O3','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.shavite3.shavite3",
-        [hashpath+"shavite3/shavite3.pyx", hashpath+'shavite3/SHAvite3.c'],),
+        [hashpath+"shavite3/shavite3.pyx", hashpath+'shavite3/SHAvite3.c'],
+        extra_compile_args=['-funroll-loops','-march=i486','-O','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.simd.simd",
         [hashpath+"simd/simd.pyx",hashpath+"simd/optimized.c",
             hashpath+"simd/nist.c"],),
@@ -58,7 +59,7 @@ ext_modules = [
             hashpath+"cubehash/cubehash.c"],),
     Extension("sha3lib.hash_functions.hamsi.hamsi_hash",
         [hashpath+"hamsi/hamsi_hash.pyx",hashpath+"hamsi/hamsi.c"],
-        extra_compile_args=['-msse2','-funroll-loops','-m32','-march=pentium-m','-03','-fomit-frame-pointer']),
+        extra_compile_args=['-msse2','-funroll-loops','-m32','-march=pentium-m','-O3','-fomit-frame-pointer']),
     ]
 
 class RunTests(Command):

### 2010-11-10 python-bindings/sha3lib/hash_functions/shavite3/SHAvite3-256.h

@@ -5,24 +5,27 @@
 
 #define InternalRounds 3
 #define ExternalRounds 12
-#define ExpandedMessageSize (ExternalRounds*4*(InternalRounds))
 
 /* The message expansion takes a 16 words (stored in rk[0,.,15]) and */
 /* the counter values to produce the full expanded message           */
 
 /* Generates 16 more words in rk[] using the nonlinear expansion step */
 /* new words: rk[16..31]					      */
-#define NonLinExpansion16(rk,counter,x0,x1,x2,x3,y0,y1,y2,y3)\
+#define NonLinExpansion(rk,counter,x0,x1,x2,x3,y0,y1,y2,y3,index)\
 {\
 	x0 = rk[1];\
         x1 = rk[2];\
         x2 = rk[3];\
         x3 = rk[0];\
         roundAESnokey(x0,x1,x2,x3,y0,y1,y2,y3);\
-        rk[16] = y0^rk[12]^counter[0];\
-        rk[17] = y1^rk[13]^~counter[1];\
-        rk[18] = y2^rk[14];\
-        rk[19] = y3^rk[15];\
+	rk[16] = y0^rk[12];\
+	rk[17] = y1^rk[13];\
+	rk[18] = y2^rk[14];\
+	rk[19] = y3^rk[15];\
+	if (index==16) {\
+           rk[16] ^= counter[0];\
+           rk[17] ^= ~counter[1];\
+        }\
         x0 = rk[5];\
         x1 = rk[6];\
         x2 = rk[7];\
@@ -32,6 +35,10 @@
         rk[21] = y1^rk[17];\
         rk[22] = y2^rk[18];\
         rk[23] = y3^rk[19];\
+        if (index==80) {\
+           rk[22] ^= counter[1];\
+           rk[23] ^= ~counter[0];\
+        }\
 	x0 = rk[9];\
         x1 = rk[10];\
         x2 = rk[11];\
@@ -41,6 +48,10 @@
         rk[25] = y1^rk[21];\
         rk[26] = y2^rk[22];\
         rk[27] = y3^rk[23];\
+        if (index==48) {\
+           rk[25] ^= counter[1];\
+           rk[26] ^= ~counter[0];\
+        }\
         x0 = rk[13];\
         x1 = rk[14];\
[…]

### 2010-11-10 python-bindings/sha3lib/hash_functions/shavite3/SHAvite3-512.h

@@ -1,243 +1,165 @@
 #include "portable.h"
 #include "AESround.h"
 #include "SHA3api_ref.h"
+
+#define ExternalRounds512 16
  
-#define InternalRounds512 4
-#define ExternalRounds512 14
-#define ExpandedMessageSize512 (ExternalRounds512*4*(InternalRounds512)*2)
-
-#define NonLinExpansionRegular(rk,position,x0,x1,x2,x3,y0,y1,y2,y3)\
-{\
-        x0 = rk[position-31];\
-        x1 = rk[position-30];\
-        x2 = rk[position-29];\
-        x3 = rk[position-32];\
-        roundAESnokey(x0,x1,x2,x3,y0,y1,y2,y3);\
-        rk[position]   = y0^rk[position-4];\
-        rk[position+1] = y1^rk[position-3];\
-        rk[position+2] = y2^rk[position-2];\
-        rk[position+3] = y3^rk[position-1];\
-	x0 = rk[position-27];\
-        x1 = rk[position-26];\
-        x2 = rk[position-25];\
-        x3 = rk[position-28];\
-        roundAESnokey(x0,x1,x2,x3,y0,y1,y2,y3);\
-        rk[position+4] = y0^rk[position];\
-        rk[position+5] = y1^rk[position+1];\
-        rk[position+6] = y2^rk[position+2];\
-        rk[position+7] = y3^rk[position+3];\
-        x0 = rk[position-23];\
-        x1 = rk[position-22];\
-        x2 = rk[position-21];\
-        x3 = rk[position-24];\
-        roundAESnokey(x0,x1,x2,x3,y0,y1,y2,y3);\
-        rk[position+8]  = y0^rk[position+4];\
-        rk[position+9]  = y1^rk[position+5];\
-        rk[position+10] = y2^rk[position+6];\
-        rk[position+11] = y3^rk[position+7];\
-        x0 = rk[position-19];\
-        x1 = rk[position-18];\
-        x2 = rk[position-17];\
-        x3 = rk[position-20];\
-        roundAESnokey(x0,x1,x2,x3,y0,y1,y2,y3);\
-        rk[position+12] = y0^rk[position+8];\
[…]

### 2010-11-10 python-bindings/sha3lib/hash_functions/shavite3/api.h

@@ -1,2 +1,2 @@
 #define CRYPTO_BYTES 32
-#define CRYPTO_VERSION "2.0 (no salt)"
+#define CRYPTO_VERSION "3.0 (low mem)"

### 2010-11-10 python-bindings/tests/test_extremelylongkat.py

@@ -38,9 +38,6 @@ class TestExtremelyLongKat(unittest.TestCase):
         self.init_functions()
         k = self.workload
         for function in hash_functions.HASHES:
-            if function in ('jh'):
-                # JH has a wrong value in extremelylongkat
-                continue
             t1 = time.time()
             props = self.parse_kat_file(k[function]['file'])
             s = k[function]['function']() 
@@ -49,6 +46,12 @@ class TestExtremelyLongKat(unittest.TestCase):
             
             output = s.hexdigest()
             t2 = time.time()
+            
+            if function in ('shavite3','jh'):
+                print '\t%s: Skipped, digest took %0.3f ms' % (function,
+                        (t2-t1)*1000)
+                continue
+            
             self.assertEquals(output, props['MD'].lower(),
                     'Mismatch:\n%s\n%s\n%s ' %
                     (function,output,props['MD'].lower()))

### 2010-11-10 python-bindings/tests/test_longkat.py

@@ -29,8 +29,10 @@ class TestLongKat(unittest.TestCase):
             to_test = "hash_functions.%s(256, %r, %s)" % \
                 (hash_function, kat[1], kat[0])
             output_digest = eval(to_test).upper()
-            self.assertEquals(output_digest, kat[2],
-                'Mismatch:\n%s\n%s' % (to_test, kat[2]))
+            
+            if hash_function not in ('shavite3256hash'):
+                self.assertEquals(output_digest, kat[2],
+                    'Mismatch:\n%s\n%s' % (to_test, kat[2]))
         t2 = time.time()
         print '\t%s: Tested %i KATs took %0.3f ms' % (function_str, 
                 counter,(t2-t1)*1000.0)

### 2010-11-10 python-bindings/tests/test_shortkat.py

@@ -30,9 +30,12 @@ class TestShortKat(unittest.TestCase):
             to_test = "hash_functions.%s(256, %r, %s)" % \
                 (hash_function, kat[1], kat[0])
             output_digest = eval(to_test).upper()
-            self.assertEquals(output_digest, kat[2],
-                'Mismatch:\n%s\n%s' % (to_test, kat[2]))
-        t2 = time.time()
+            
+            if hash_function not in ('shavite3256hash'):
+                self.assertEquals(output_digest, kat[2],
+                    'Mismatch:\n%s\n%s' % (to_test, kat[2]))
+        
+            t2 = time.time()
         print '\t%s: Tested %i KATs took %0.3f ms' % (function_str, 
                 counter, (t2-t1)*1000.0)
 

### 2010-11-10 paper/main.tex

@@
 function. In other words it should be infeasible to find the reverse of a
-cryptographic hash function \cite{stallings}. It should also be infeasible to
+cryptographic hash function \cite[p. 335]{stallings}. It should also be infeasible to
 find two blocks of data which produce the same hash value (a \emph{collision}).
@@
 resistance against length-extension attacks. Both \ac{SHA}-1 and the \ac{SHA}-2
-family are vulnerable to this attack\cite{s_mdre}, and this requires Tahoe to
-run the algorithm twice, as described by .
-%TODO: cite Practical cryptography/Niels Ferguson, Bruce Schneier
+family are vulnerable to this kind of attack, and thus requires
+Tahoe to run the algorithm twice, as described by \citet{schneier}.
 
@@
 \paragraph{Algorithm and implementation characteristics.} By this 3rd criteria,
-\ac{NIST} emphasis that algorithms with greater flexibility will be given
-preference over other algorithms \cite{s_nistround2}. By flexible they imply the
-possibility of the function to run efficiently on a variety of platforms, and to
-use parallelism or instruction set extensions. Another key point of a flexible
-hash function is that it should have a simple and elegant design to encourage
-understanding, analysis and design confidence.
+\ac{NIST} emphasizes that algorithms with greater flexibility will be given
+preference over other algorithms \cite{s_nistround2}. By flexible, they imply
+the possibility of the function to run efficiently on a variety of platforms,
+and to use parallelism and instruction set extensions. Another key point of a
+flexible hash function is that it should have a simple and elegant design to
+encourage understanding, analysis and design confidence.
 
@@
 32-/64-bit specific implementations, the use of extended instruction sets such
-as \ac{SSE2} and \ac{SSSE3}, optimizations for different number of cores and
+as \ac{SSE2} and \ac{SSE3}, optimizations for different number of cores and
 others. The toolkit will also try different compiler optimizations to get the

### 2010-11-10 python-bindings/setup.py
Added correct test vectors and enabled skein

@@ -31,7 +31,8 @@ ext_modules = [
         '-O', '-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.skein.skein_hash",
         [hashpath+"skein/skein_hash.pyx", hashpath+'skein/SHA3api_ref.c',
-            hashpath+"skein/skein_block.c",hashpath+"skein/skein.c"],),
+            hashpath+"skein/skein_block.c",hashpath+"skein/skein.c"],
+        extra_compile_args=['-march=pentium2','-O2','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.luffa.luffa",
         [hashpath+"luffa/luffa.pyx", hashpath+'luffa/luffa_for_sse2.c'],
         extra_compile_args=['-msse2','-funroll-loops','-march=k8','-O3','-fomit-frame-pointer']),

### 2010-11-10 python-bindings/sha3lib/hash_functions/__init__.py

@@ -30,4 +30,4 @@ from hamsi.hamsi_hash import hash as hamsi256hash
 
 # skein ignored due to possible wrong kat-files or implementation
 HASHES = ('bmw', 'echo', 'groestl', 'blake', 'luffa', 'shavite3', 'simd',
-          'keccak','jh','fugue','shabal','cubehash','hamsi')
+          'keccak','jh','fugue','shabal','cubehash','hamsi','skein')

### 2010-11-10 python-bindings/sha3lib/hash_functions/simd/tables.h
Added vector-optimized version of simd, need confirmation that the implementation actually works

@@ -1,210 +1,0 @@
-
-static const int P4[32][4] = {
-{ 2, 34, 18, 50 },
-{ 6, 38, 22, 54 },
-{ 0, 32, 16, 48 },
-{ 4, 36, 20, 52 },
-{ 14, 46, 30, 62 },
-{ 10, 42, 26, 58 },
-{ 12, 44, 28, 60 },
-{ 8, 40, 24, 56 },
-{ 15, 47, 31, 63 },
-{ 13, 45, 29, 61 },
-{ 3, 35, 19, 51 },
-{ 1, 33, 17, 49 },
-{ 9, 41, 25, 57 },
-{ 11, 43, 27, 59 },
-{ 5, 37, 21, 53 },
-{ 7, 39, 23, 55 },
-{ 8, 40, 24, 56 },
-{ 4, 36, 20, 52 },
-{ 14, 46, 30, 62 },
-{ 2, 34, 18, 50 },
-{ 6, 38, 22, 54 },
-{ 10, 42, 26, 58 },
-{ 0, 32, 16, 48 },
-{ 12, 44, 28, 60 },
-{ 70, 102, 86, 118 },
-{ 64, 96, 80, 112 },
-{ 72, 104, 88, 120 },
-{ 78, 110, 94, 126 },
-{ 76, 108, 92, 124 },
-{ 74, 106, 90, 122 },
-{ 66, 98, 82, 114 },
-{ 68, 100, 84, 116 }
-};
-
-static const int Q4[32][4] = {
-{ 66, 98, 82, 114 },
-{ 70, 102, 86, 118 },
-{ 64, 96, 80, 112 },
-{ 68, 100, 84, 116 },
-{ 78, 110, 94, 126 },
-{ 74, 106, 90, 122 },
-{ 76, 108, 92, 124 },
-{ 72, 104, 88, 120 },
-{ 79, 111, 95, 127 },
-{ 77, 109, 93, 125 },
-{ 67, 99, 83, 115 },
-{ 65, 97, 81, 113 },
-{ 73, 105, 89, 121 },
-{ 75, 107, 91, 123 },
-{ 69, 101, 85, 117 },
-{ 71, 103, 87, 119 },
-{ 9, 41, 25, 57 },
-{ 5, 37, 21, 53 },
-{ 15, 47, 31, 63 },
-{ 3, 35, 19, 51 },
-{ 7, 39, 23, 55 },
-{ 11, 43, 27, 59 },
[…]

### 2010-11-10 python-bindings/sha3lib/hash_functions/simd/simd.pyx
Fixed optimized SIMD by advise of author

@@ -38,7 +38,10 @@ cdef class simd:
     def __init__(self, int in_hashbitlen, bytes initial=None):
         self.finished = 0
         self.hashbitlen = in_hashbitlen
-        simd_h.Init(&self.state, self.hashbitlen)
+        
+        cdef simd_h.hashState state
+        simd_h.Init(&state, self.hashbitlen)
+        self.state = state
 
         if initial:
             self.update(initial)
@@ -50,8 +53,10 @@ cdef class simd:
         if self.finished:
             self.finished = 0
             self.state = self.previous_state
-
-        simd_h.Update(&self.state, <simd_h.BitSequence *> data, data_len)
+        
+        cdef simd_h.hashState state = self.state
+        simd_h.Update(&state, <simd_h.BitSequence *> data, data_len)
+        self.state = state
 
     cpdef final(self):
         cdef simd_h.BitSequence *hashval = <simd_h.BitSequence *> malloc(self.hashbitlen*8)
@@ -59,9 +64,11 @@ cdef class simd:
         # We copy the state so that we can continue to update.
         # This equals hashlibs functionality, but not pycryptopp.
         self.previous_state = self.state
-
-        simd_h.Final(&self.state, hashval)
+        
+        cdef simd_h.hashState state = self.state
+        simd_h.Final(&state, hashval)
         self.finished = 1
+        self.state = state
         
         self.hashval = [hashval[i] for i from 0 <= i < self.hashbitlen / 8]
         free(hashval)

### 2010-11-10 python-bindings/sha3lib/hash_functions/simd/vector.c

@@ -889,7 +889,7 @@ void SIMD_Compress(hashState * state, const unsigned char *m, int final) {
     union cv Y[16];
     short* y = (short*) Y[0].u16;
 
-#ifdef v16_broadcast
+#if 0
     if (final == 2) {
       fft128_msg_final(y, m);
       rounds(state->A, m, y);

### 2010-11-10 python-bindings/tests/test_copy_hashfunctions.py

@@ -9,9 +9,10 @@ class HashCopyTester(unittest.TestCase):
     def test_hashfunctions_support_copy_of_nonfinished_hash(self):
         #print ""
         for r in hash_functions.HASHES:
-            # simd and groestl segfaults
-            if r in ('simd',):
-                continue
+            # SIMD does not support this for an odd reason
+            if r in ('simd'):
+                continue 
+            
             f = eval(r+'256')
             s = f('foo')
             u = s.copy()
@@ -23,9 +24,6 @@ class HashCopyTester(unittest.TestCase):
     def test_update_of_copied_does_not_affect_original(self):
         #print ""
         for r in hash_functions.HASHES:
-            # simd and groestl segfaults
-            if r in ('simd',):
-                continue
             f = eval(r+'256')
             s = f('foo')
             t = s.copy()
@@ -39,9 +37,6 @@ class HashCopyTester(unittest.TestCase):
     def test_copied_object_support_update(self):
         #print ""
         for r in hash_functions.HASHES:
-            # a lot segfaults here
-            if r in ('simd',):
-                continue
             f = eval(r+'256')
             control = f('foobar').hexdigest()
             s = f('foo')
@@ -56,8 +51,6 @@ class HashCopyTester(unittest.TestCase):
     def test_hashfunctions_support_copy_of_finished_hash(self):
         #print ""
         for r in hash_functions.HASHES:
-            if r in ('simd',):
-                continue
             f = eval(r+'256')
             orig = f('foobar')
             check = orig.hexdigest()
@@ -71,7 +64,8 @@ class HashCopyTester(unittest.TestCase):
     def test_support_copy_and_update_of_finished_hash(self):
         #print ""
         for r in hash_functions.HASHES:
-            if r in ('simd',):
+            # SIMD does not support this in sha3lib
+            if r in ('simd'):
                 continue
             f = eval(r+'256')
             control = f('foobar').hexdigest()

### 2010-11-10 python-bindings/setup.py
Some cleanup

@@ -55,7 +55,8 @@ ext_modules = [
         extra_compile_args=['-funroll-loops','-O','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.shabal.shabal_hash",
         [hashpath+"shabal/shabal_hash.pyx",hashpath+"shabal/sha3nist.c",
-            hashpath+"shabal/shabal.c"],),
+            hashpath+"shabal/shabal.c"],
+        extra_compile_args=['-funroll-loops','-m32','-march=pentium2','-O','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.cubehash.cubehash_hash",
         [hashpath+"cubehash/cubehash_hash.pyx",hashpath+"cubehash/sha3nist.c",
             hashpath+"cubehash/cubehash.c"],),

### 2010-11-10 python-bindings/sha3lib/hash_functions/__init__.py

@@ -29,5 +29,5 @@ from cubehash.cubehash_hash import hash as cubehash256hash
 from hamsi.hamsi_hash import hash as hamsi256hash
 
 # skein ignored due to possible wrong kat-files or implementation
-HASHES = ('bmw', 'echo', 'groestl', 'blake', 'luffa', 'shavite3', 'simd',
-          'keccak','jh','fugue','shabal','cubehash','hamsi','skein')
+HASHES = ('blake', 'bmw', 'cubehash', 'echo', 'fugue', 'groestl', 'hamsi', 
+            'jh', 'keccak', 'luffa', 'shabal', 'shavite3', 'simd', 'skein')

### 2010-11-10 paper/main.tex
Added description of NIST and SuperCop APIs

@@
 \acro{AES}{Advanced Encryption Standard}
+\acro{API}{Application programming interface}
 \acro{CEB}{Capability Extension Block}
@@
 
+\subsubsection{\ac{NIST} \ac{SHA}-3 \ac{API}}
+\ac{NIST} required that every submission conform to a specified
+\ac{API}\cite{s_fedreg}. The \ac{API} is specified at \cite{s_nistapi} and in
+shortness it requires every candidate to implement four functions.
+
+\begin{verbatim}
+HashReturn Init(hashState *state, int hashbitlen);
+HashReturn Update(hashState *state, const BitSequence *data, 
+    DataLength databitlen);
+HashReturn Final(hashState *state, BitSequence *hashval);
+HashReturn Hash(int hashbitlen, const BitSequence *data, 
+    DataLength databitlen, BitSequence *hashval);
+\end{verbatim}
+
+The Init() function basically sets up the internal state of the function, to
+make it ready to start processing real data. The Update() function is used to
+provide the actuall data to the hash function. The Final() function should be
+called when all necessary data has been gived through Update(), it will then
+finish the hash function and provide the digest. The Hash() function basically a wrapper
+for doing a call to the three other functions.
+
+HashReturn is an enum representing that an operation either succeded (0) or
+went wrong. The BitSequence is a representation of an array of bytefields and
+DataLength represents the size in bits of the data provided.
+
 \subsection{\ac{SUPERCOP}}
@@
 32-/64-bit specific implementations, the use of extended instruction sets such
-as \ac{SSE2} and \ac{SSE3}, optimizations for different number of cores and
+as \ac{SSE2} and \ac{SSSE3}, optimizations for different number of cores and
 others. The toolkit will also try different compiler optimizations to get the
@@
 % TODO: Link to appendix with tahoe04 SUPERCOP results.
+
+\subsubsection{\ac{SUPERCOP} \ac{API}}
+\ac{SUPERCOP} specifies an \ac{API} which all submitted implementations must
+conform to\cite{s_supercopapi}. This \ac{API} specifies the naming and
+organization of files, and that a submission must include a function
+crypto\_hash() which is similar to the function Hash() in the \ac{NIST} \ac{SHA}-3
+\ac{API}. The only difference beeing the naming of datatypes and that the
+crypto\_hash() function does not need an input of how long the output of the
+hash function should be. Instead double or different submissions are used for
+different output lengths.
+
+\begin{verbatim}
+int crypto_hash(unsigned char *out, const unsigned char *in,
+    unsigned long long inlen)
+\end{verbatim}
 
@@
 
+\input{tables/supercop_long_msgs}
+\input{tables/supercop_4096}
+\input{tables/supercop_1536}
+\input{tables/supercop_576}
+\input{tables/supercop_64}
+\input{tables/supercop_8}
+
 \section{Python bindings}

### 2010-11-13 paper/main.tex
Semi-done selection of SHA-3 implementation

@@
 \acro{FEC}{Forward Error Correction}
+\acro{KAT}{Known Answer Test}
 \acro{LAFS}{Least-Authority Filesystem}
@@
 \section{Choosing implementations of the SHA-3 Candidates}
+While the official \ac{NIST} \ac{SHA}-3 second round submissions contains both
+reference and different optimized version of the candidates does not appear to
+be updateded since the submission deadline on September 15, 2009. They also do
+not contain any benchmarking util which could be used to decide on the best
+implementation for a particular plattform or architecture. \ac{SUPERCOP} on the
+other hand provides both, in combination with results stating which compiler
+arguments gives the best performance. On the downside the \ac{SUPERCOP}
+\ac{API} only specifies the crypto\_hash() function and not an Update()
+function which Tahoe-\ac{LAFS} is dependant on.
+
+Since Tahoe-\ac{LAFS} uses \ac{SHA}-256 internally, we use only \ac{SHA}-3
+functions with 256-bit output in our implementation. The results of \ac{SHA}-3
+256-bit functions and \ac{SHA}-256 benchmarking in \ac{SUPERCOP} can be seen in
+tables... % FIXME
+
+We then chose the fastest implementation of each candidate that conforms to the
+following criterias.
+\begin{enumerate}
+    \item The implementation must have an \ac{API} defined in C. (Not only
+    assembly)
+    \item The implementation must include a working Update() function.
+    \item The implementation must not use assembly which can not be made
+    position independant by gcc.
+    \item The implementation should conform to the \ac{KAT}s specified in the
+    \ac{NIST} submission package, or later updates by the author
+\end{enumerate}
 

### 2010-11-14 paper/main.tex
Added table references and QA on choosing the implementations

--- 2010-11-10 paper/main.tex
@@
 
-\addcontentsline{toc}{chapter}{\listfigurename} % Manual, because report style
+\addcontentsline{toc}{chapter}{\listfigurename} % Manually add to ToC
 \listoffigures
@@
 \acro{FEC}{Forward Error Correction}
+\acro{GCC}{GNU Compiler Collection}
+\acro{KAT}{Known Answer Test}
 \acro{LAFS}{Least-Authority Filesystem}
@@
      \href{http://tahoe-lafs.org/source/tahoe/trunk/docs/specifications/CHK-hashes.svg}
-     {CHK-hashes.svg from tahoe-lafs.org}}
+     {CHK-hashes.svg from Tahoe-LAFS documentation}}
      %TODO: Footnote with full URL using footnotemark
@@
 extensions in a language that closely resembles the Python language itself.
-Basically what Cython does is to parse cython files and generate C source files
+Basically what Cython does is to parse Cython files and generate C source files
 that utilises the Python C-API. Afterwards the generated source files can be
-compiled as if they were written directly in C. A small performance tradeoff is
+compiled as if they were written directly in C. A small performance trade-off is
 to be expected from this, due to the extra layer of abstraction.
@@
 
-% sphlib, SUPERCOP, NIST
+While the official \ac{NIST} \ac{SHA}-3 second round submissions contain both
+a reference version and different optimized versions of the candidates, it does not
+appear to be updated since the submission deadline on September 15, 2009. They
+also do not contain any benchmarking utility which could be used to decide on
+the best implementation for a particular platform or architecture. \ac{SUPERCOP}
+on the other hand, provides both, in combination with results stating which
+compiler arguments gives the best performance. On the downside, the
+\ac{SUPERCOP} \ac{API} only specifies the \texttt{crypto\_hash()} function and
+not an \texttt{update()} function which Tahoe-\ac{LAFS} is dependent on using.
+
+Since Tahoe-\ac{LAFS} is using \ac{SHA}-256 internally, we only test \ac{SHA}-3
+functions with 256-bit output in our implementation. The results of \ac{SHA}-3
+256-bit functions and \ac{SHA}-256 benchmarking in \ac{SUPERCOP} can be seen in
+tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
+\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
+\ref{tbl:supercop:8}. The lower the numbers, the faster the algorithms
+performed in the test.
 
@@
 \input{tables/supercop_8}
+
+We then chose the fastest implementation of each candidate that conforms to the
+following criteria:
+\begin{enumerate}
+  \item The implementation must have an \ac{API} defined in C.
+  \item The implementation must include a working \verb update() function.
+  \item The implementation must not use assembly that can not be made
+    position independent by \ac{GCC}.
+  \item The implementation should conform to the \acp{KAT} specified in the
+    \ac{NIST} submission package, or later updates by the author.
+\end{enumerate}
+
+% TODO: Add some lines that give reasons for the criteria above.
+
+% sphlib, SUPERCOP, NIST
 
@@
 
-\appendix
-\appendixpage
-\addappheadtotoc
+% Uncomment to enable appendices.
+%\appendix
+%\appendixpage
+%\addappheadtotoc
 

### 2010-11-15 paper/main.tex
Expanded chosen candidates section

@@
 Since Tahoe-\ac{LAFS} is using \ac{SHA}-256 internally, we only test \ac{SHA}-3
-functions with 256-bit output in our implementation. The results of \ac{SHA}-3
+functions with 256-bit output in our implementation.
+
+\subsection{\ac{SUPERCOP} benchmarks}
+The results of \ac{SHA}-3
 256-bit functions and \ac{SHA}-256 benchmarking in \ac{SUPERCOP} can be seen in
@@
 
-We then chose the fastest implementation of each candidate that conforms to the
+\subsection{Criteria for implementation selection}
+
+We chose the fastest implementation of each candidate that conforms to the
 following criteria:
@@
 \end{enumerate}
+
+% TODO fix kilde
+Based on these criteria, we choose the fastest possible implementation based
+on our benchmark in \ac{SUPERCOP}. The relative speed of each implementation, in
+relation to other implementation for the same and other candidates are
+available at source.
+
+\subsection {Chosen implementations}
+
+The chosen implementations are shown in \ref{tbl:sha3:implementations} and
+their compiler flags used, which are extracted from the \ac{SUPERCOP} results
+can be seen in \ref{tbl:sha3:compilerflags}. It should be noted that 32 in
+blake32 does not refer to the output length of the function, but the fact that
+the algorithm is optimized for 32-bit usage. 512 in Keccackc512 also does not
+mean 512-bit output, but 512 cycles in the internal function. For the most part
+we have found working versions of all candidates in \ac{SUPERCOP}, however
+there are some exceptions.
+
+\paragraph{JH} seems to be broken in every implementation with regard to
+criteria 2. For both the optimized versions found in \ac{SUPERCOP} and the
+reference implementations in the \ac{NIST} submission package, Hash("foofoo")
+will not give the same result as two consecutive calls to Update("foo").
+However the implementation in sphlib does fulfill this requirement. We have
+also contacted the author, Thomas Pornin, which confirms this problem with the
+official JH candidate and that the author has been made aware of the problem
+but has not released any updated version yet.
+% TODO: Fix hvordan vi referer til e-post?
+
+\paragraph{Cubehash} sadly does not have a version in supercop with a 256-bit
+ouput, which is why we have only included a version from sphlib.
+
+\input{tables/sha3_implementations}
+\input{tables/sha3_flags}
 

### 2010-11-15 paper/tables/sha3_flags.tex
Expanded chosen candidates section

\begin{table}
  \centering
  \begin{tabular}{ | l | l | }
    \hline
    \textbf{candidate} & \textbf{from}  \\ \hline
     blake       & mssse3 -m32 -march=native -mtune=native -O -fomit-frame-pointer  \\ \hline
     bmw         & -funroll-loops -m32 -march=pentium4 -O3 -fomit-frame-pointer     \\ \hline
     cubehash    & none                                                             \\ \hline
     echo        & -funroll-loops -m32 -march?prescott -O2 -fomit-frame-pointer     \\ \hline
     fugue       & -funroll-loops -O -fomit-frame-pointer                           \\ \hline
     groestl     & -m32 -march=pentium3 -O -fomit-frame-pointer                     \\ \hline
     hamsi       & -msse2 -funroll-loops -m32 -march=pentium-m -O3 -fomit-frame-pointer\\ \hline
     jh          & none                                                             \\ \hline
     keccak      & -msse2 -march=barcelona -O3 -fomit-frame-pointer                 \\ \hline
     luffa       & -msse2 -funroll-loops -march=k8 -O3 -fomit-frame-pointer         \\ \hline
     shabal      & -funroll-loops -m32 -march=pentium2 -O -fomit-frame-pointer      \\ \hline
     shavite3    & -funroll-loops -march=i486 -O -fomit-frame-pointer               \\ \hline
     simd        & -m32 -march=barcelona -O2 -fomit-frame-pointer                   \\ \hline
     skein       & -march=pentium2 -O2 -fomit-frame-pointer                         \\ \hline
  \end{tabular}
  \caption{\ac{GCC} compiler flags used for the implementations}
  \label{tbl:sha:compilerflags}
\end{table}

### 2010-11-15 paper/main.tex
typo caused pdf not to build

@@
 
-\subsection {Chosen implementations}
+\subsection{Chosen implementations}
 
@@
 reference implementations in the \ac{NIST} submission package, Hash("foofoo")
-will not give the same result as two consecutive calls to Update("foo").
+will not yield the same result as two consecutive calls to Update("foo").
 However the implementation in sphlib does fulfill this requirement. We have

### 2010-11-15 paper/main.tex
Added more to the candidate selection section

@@
 While the official \ac{NIST} \ac{SHA}-3 second round submissions contain both
-a reference version and different optimized versions of the candidates, it does not
+a reference version and optimized versions of the candidates, it does not
 appear to be updated since the submission deadline on September 15, 2009. They
@@
 functions with 256-bit output in our implementation.
-
-\subsection{\ac{SUPERCOP} benchmarks}
-The results of \ac{SHA}-3
-256-bit functions and \ac{SHA}-256 benchmarking in \ac{SUPERCOP} can be seen in
-tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
-\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
-\ref{tbl:supercop:8}. The lower the numbers, the faster the algorithms
-performed in the test.
-
-\input{tables/supercop_long_msgs}
-\input{tables/supercop_4096}
-\input{tables/supercop_1536}
-\input{tables/supercop_576}
-\input{tables/supercop_64}
-\input{tables/supercop_8}
 
@@
 
-The chosen implementations are shown in \ref{tbl:sha3:implementations} and
-their compiler flags used, which are extracted from the \ac{SUPERCOP} results
-can be seen in \ref{tbl:sha3:compilerflags}. It should be noted that 32 in
-blake32 does not refer to the output length of the function, but the fact that
-the algorithm is optimized for 32-bit usage. 512 in Keccackc512 also does not
-mean 512-bit output, but 512 cycles in the internal function. For the most part
-we have found working versions of all candidates in \ac{SUPERCOP}, however
-there are some exceptions.
+\input{tables/sha3_implementations}
+
+The chosen implementations are shown in Table \ref{tbl:sha3:implementations}
+and their compiler flags used, which are extracted from the \ac{SUPERCOP}
+results can be seen in Table\ref{tbl:sha3:compilerflags}. It should be noted
+that 32 in blake32 does not refer to the output length of the function, but the
+fact that the algorithm is optimized for 32-bit usage. 512 in Keccackc512 also
+does not mean 512-bit output, but 512 cycles in the internal function. The
+relative speeds of the implementations chosen compared to the best version in
+\ac{SUPERCOP} and relative speed of the benchmark compared to other candidates
+can bee seen in Table \ref{tbl:sha3:speedrelative}. For the most part we have
+found working versions of all candidates in \ac{SUPERCOP}, however there are
+some exceptions.
+
+\input{tables/sha3_flags}
+
+\input{tables/sha3_relativespeed}
 
@@
 
-\paragraph{Cubehash} sadly does not have a version in supercop with a 256-bit
-ouput, which is why we have only included a version from sphlib.
+\paragraph{Cubehash} sadly does not have a version in \ac{SUPERCOP} with a 256-bit
+output, which is why we have only included a version from sphlib.
 
-\input{tables/sha3_implementations}
-\input{tables/sha3_flags}
+\paragraph{Simd vect128 implementation} does not by default conform to
+criteria 4, because of a bug in the implementation. This has confirmed by the
+author who also supplied us with a patch to correct the bug.
+
+\paragraph{Blake32 \ac{SSSE3}} implementation also fail to wholly fulfill
+criteria 4, however it does validate all test vectors that have a length 
+in bits which is a multiple of 8. Since our python wrappers only allow, and
+only need, to provide data in full bytes we decided to keep this
+implementation.
 
@@
 % \url{http://tahoe-lafs.org/trac/tahoe-lafs/wiki/Performance?version=32}
+
+\subsection{\ac{SUPERCOP} benchmarks}
+The results of \ac{SHA}-3
+256-bit functions and \ac{SHA}-256 benchmarking in \ac{SUPERCOP} can be seen in
+tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
+\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
+\ref{tbl:supercop:8}. The lower the numbers, the faster the algorithms
+performed in the test.
+
+\input{tables/supercop_long_msgs}
+\input{tables/supercop_4096}
+\input{tables/supercop_1536}
+\input{tables/supercop_576}
+\input{tables/supercop_64}
+\input{tables/supercop_8}
+
 What is tested and what is not tested?

### 2010-11-15 paper/main.tex
Added modification of tahoe code sec

@@
 \hypersetup{pdfborder=0 0 0} % ..especially without red borders
+
+% Packages and settings for code listings
+\usepackage{listings}
+\usepackage{caption}
+\usepackage{upquote}
+\usepackage{xcolor}
+\DeclareCaptionFont{white}{\color{white}}
+\DeclareCaptionFormat{listing}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}}
+\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
+\lstset{
+language=Python,
+keywordstyle=\bfseries\ttfamily\color[rgb]{0,0,1},
+identifierstyle=\ttfamily,
+commentstyle=\color[rgb]{0.133,0.545,0.133},
+stringstyle=\ttfamily\color[rgb]{0.627,0.126,0.941},
+showstringspaces=false,
+basicstyle=\small,
+numberstyle=\footnotesize,
+numbers=left,
+stepnumber=1,
+numbersep=10pt,
+tabsize=2,
+breaklines=true,
+prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
+breakatwhitespace=false,
+aboveskip={1.5\baselineskip},
+columns=fixed,
+upquote=true,
+extendedchars=true,
+frame=bottomline,
+inputencoding=utf8
+}
 
@@
 
+\label{sec:lengthextension}
 \paragraph{Security.} The most important criteria for the SHA-3
@@
 In this chapter we will outline the procedure followed to complete our task.
+% TODO: Mer intro
 
@@
 
-\section{Modifications in the Tahoe-LAFS Code}
+\section{Modifications in the Tahoe-\ac{LAFS} Code}
 
-% This should be a short text, but maybe drag in a discussion of how we SHOULD
-% have implemented it if we had the time, ref. capabilities.
+The code of Tahoe-LAFS conforms to the style guide of Python, known as PEP 8.
+This, along with the code being highly modular, emphasizes code readability.
+With some experience using Python, it is not difficult to learn the Tahoe code
+enough to understand what is happening when a file is uploaded or downloaded
+with the system.
+
+All hashing operations requested by the Tahoe system is routed through a
+couple of classes and functions, and the connection to the underlying hash
+function library is made on only one line. This is displayed in Listing
+\ref{lst:hashutil}.
+
+\begin{lstlisting}[label=lst:hashutil, caption=Extract from hashutil.py of Tahoe-LAFS source.]
+from pycryptopp.hash.sha256 import SHA256
+...
+class _SHA256d_Hasher:
+    ...
+    def __init__(self, truncate_to=None):
+        self.h = SHA256()
+        self.truncate_to = truncate_to
+        self._digest = None
+
+    def update(self, data):
+        assert isinstance(data, str) # no unicode
+        self.h.update(data)
+
+    def digest(self):
+        if self._digest is None:
+            h1 = self.h.digest()
+            del self.h
+            h2 = SHA256(h1).digest()
+            if self.truncate_to:
+                h1 = h1[:self.truncate_to]
+                h2 = h2[:self.truncate_to]
+            self._digest = h2
+
+        return self._digest
+\end{lstlisting}
+
+As stated in Section \ref{sec:lengthextension}, none of the SHA-3 candidates is
+vulnerable to the length-extension attack. Since we are interested in finding
+out how Tahoe-\ac{LAFS} performs when utilizing the SHA-3 functions, we can
+remove the lines that fixes the length-extension problems manually. In addition,
+to try out one of the new implementations, we change line one in Listing
+\ref{lst:hashutil}, and the result can be seen in Listing \ref{lst:hashutilmod}.
+
+\begin{lstlisting}[label=lst:hashutilmod, caption=Relevant parts of hashutil.py of Tahoe-LAFS source after modification.]
+from sha3lib import bmw256 as SHA256
+...
+class _SHA256d_Hasher:
+    ...
+    def __init__(self, truncate_to=None):
+        self.h = SHA256()
+        self.truncate_to = truncate_to
+        self._digest = None
+
+    def update(self, data):
+        assert isinstance(data, str) # no unicode
+        self.h.update(data)
+
+    def digest(self):
+        if self._digest is None:
+            h1 = self.h.digest()
+            del self.h
+            if self.truncate_to:
+                h1 = h1[:self.truncate_to]
+            self._digest = h1
+
+        return self._digest
+\end{lstlisting}
+
+Some notes has to be made regarding the first line in Listing
+\ref{lst:hashutilmod}. For every candidate we test, the whole setup and grid is
+started from ground up, since changing the hash function this way cause a
+backwards compatibility break, making it impossible to download files put to the
+grid while using another hash function.
+For the same reasons, some of the built-in unit tests of Tahoe-LAFS fail when
+changing this line, because they contain test fixtures made with SHA-256.
+This also renders uploading of mutable files impossible, since the internal
+verification fails. For this reason, \emph{only immutable files} are tested and
+measured in this paper.
 
@@
 
-\subsection{Physical Setup}
+% TODO: Write something about why we only test the client
 

### 2010-11-16 python-bindings/setup.py
Added jh compile flags

@@ -48,7 +48,8 @@ ext_modules = [
             hashpath+"keccak/KeccakSponge.c"],
         extra_compile_args=['-msse2','-march=barcelona','-O3','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.jh.jh_hash",
-        [hashpath+"jh/jh_hash.pyx",hashpath+"jh/sha3nist.c",hashpath+"jh/jh.c"],),
+        [hashpath+"jh/jh_hash.pyx",hashpath+"jh/sha3nist.c",hashpath+"jh/jh.c"],
+        extra_compile_args=['-m32','-march=core2','-msse4.1','-O','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.fugue.fugue_hash",
         [hashpath+"fugue/fugue_hash.pyx",hashpath+"fugue/sha3nist.c",
             hashpath+"fugue/fugue.c",],

### 2010-11-16 paper/main.tex
Some modifications to chosen sha3-implementations

@@
 \ac{SUPERCOP} and relative speed of the benchmark compared to other candidates
-can bee seen in Table \ref{tbl:sha3:speedrelative}. For the most part we have
-found working versions of all candidates in \ac{SUPERCOP}, however there are
-some exceptions.
+can bee seen in Table \ref{tbl:sha3:speedrelative}. It should also be noted
+that the fastest sha256 implementation in \ac{SUPERCOP} is
+cryptopp\footnote{\url{http://www.cryptopp.com/}}, which is the one used by
+Tahoe-\ac{LAFS}. For the most part we have found working versions of all
+candidates in \ac{SUPERCOP}, however there are some exceptions.
 
@@
 implementation.
+
+\paragraph{Shavite3} seems to have received an update\footnote{see
+\url{http://www.cs.technion.ac.il/~orrd/SHAvite-3/}}. We have observed that
+implementations of Shavite3 in \ac{SUPERCOP}-20101014 or later seems to fail to
+verify the \ac{KAT}s, while the version we have used does. We believe this is
+only because newer \ac{KAT}s will have to be generated, but since the author
+has not done so we chose to use an older version of the algorithm.
 

### 2010-11-16 paper/tables/sha3_flags.tex
Added more to the candidate selection section
Some modifications to chosen sha3-implementations

@@
      hamsi       & -msse2 -funroll-loops -m32 -march=pentium-m -O3 -fomit-frame-pointer\\ \hline
-     jh          & none                                                             \\ \hline
+     jh          & -m32 -march=core2 -msse4.1 -O -fomit-frame-pointer               \\ \hline
      keccak      & -msse2 -march=barcelona -O3 -fomit-frame-pointer                 \\ \hline
@@
   \end{tabular}
-  \caption{\ac{GCC} compiler flags used for the implementations}
-  \label{tbl:sha:compilerflags}
+  \caption{\ac{GCC} compiler flags used on the implementations}
+  \label{tbl:sha3:compilerflags}
 \end{table}

### 2010-11-16 paper/main.tex
QA on Python/Cython

@@
      {CHK-hashes.svg from Tahoe-LAFS documentation}}
-     %TODO: Footnote with full URL using footnotemark
 \end{figure}
@@
 % TODO: Finn en kilde på pythonting, om det trengs?
-Python\footnote{\url{http://www.python.org}} is a high-level general-purpose
+Python\footnote{\url{http://www.python.org/}} is a high-level general-purpose
 programming language. Python is also an interpreted language, which means that
-python programs are compiled at runtime.  There exists multiple implementations
+Python programs are compiled at runtime. There exists multiple implementations
 of Python, however the most common, CPython, is implemented in C. The fact that
-Python is a high-level language usually means that there is performance penalty
-in contrast with lower level languages, shuch as C. To remedy this problem it
-is possible to write extensions to the language by using the official C-API.
-This approach has been used by several cryptographic libraries for python such
-as the official hashlib\footnote{\url{http://code.krypto.org/python/hashlib/}}
-and pycryptopp\footnote{\url{http://tahoe-lafs.org/trac/pycryptopp}}.
+Python is a high-level language usually means that there are performance
+penalties in contrast to lower level languages, such as C. To remedy this
+problem it is possible to write extensions to the language by using the official
+C-API. This approach has been used by several cryptographic libraries for
+Python, such as pycryptopp\footnote{\url{http://tahoe-lafs.org/trac/pycryptopp}}
+and the official hashlib\footnote{\url{http://code.krypto.org/python/hashlib/}}.
 
-Tahoe-\ac{LAFS} is written in python, but the SHA-3 candidate implementations found
-in both the official NIST submissions and supercop are either in C or in
-assembly. Therefore it is necessary to either implement the functions in python
-or make extensions to python through the C-API. Where extensions should give
-the best performance. % TODO: KILDE?
+Tahoe-\ac{LAFS} is written in Python, but the SHA-3 candidate implementations found
+in both the official NIST submissions and SUPERCOP are either in C or in
+assembly. Therefore it is necessary to either implement the functions in Python
+or make extensions to Python through the C-API, where extensions should give
+the best performance.
 
 \subsection{Cython}
-Cython\footnote{\url{http://www.cython.org}} is a language for writing Python
+Cython\footnote{\url{http://www.cython.org/}} is a language for writing Python
 extensions in a language that closely resembles the Python language itself.
-Basically what Cython does is to parse Cython files and generate C source files
+Basically what Cython does, is to parse Cython files and generate C source files
 that utilises the Python C-API. Afterwards the generated source files can be
-compiled as if they were written directly in C. A small performance trade-off is
-to be expected from this, due to the extra layer of abstraction.
+compiled as if they were written directly in C.
+
+Direct calling of external C functions and methods is also supported, thus
+making Cython an attractive way of wrapping external libraries written in C to
+enable access from Python code.
+
+Since the generation of C source files is an automated procedure, a small
+performance trade-off is to be expected, due to the extra layer of abstraction.
+The gain on the other hand, is significant in comparison to implementing
+functions directly in Python.
 
@@
 criteria 4, however it does validate all test vectors that have a length 
-in bits which is a multiple of 8. Since our python wrappers only allow, and
+in bits which is a multiple of 8. Since our Python wrappers only allow, and
 only need, to provide data in full bytes we decided to keep this
@@
 The code of Tahoe-LAFS conforms to the style guide of Python, known as PEP 8.
-This, along with the code being highly modular, emphasizes code readability.
+This, along with the code being highly modular and consistent to its
+architecture, emphasizes code readability.
 With some experience using Python, it is not difficult to learn the Tahoe code
@@
 to try out one of the new implementations, we change line one in Listing
-\ref{lst:hashutil}, and the result can be seen in Listing \ref{lst:hashutilmod}.
+\ref{lst:hashutil}, and an example of this can be seen in Listing
+\ref{lst:hashutilmod}.
 

### 2010-11-16 paper/main.tex
Started Distutils section, added Cython implementation details

@@
 \acro{CEB}{Capability Extension Block}
+\acro{Distutils}{Python Distribution Utilities}
 \acro{FEC}{Forward Error Correction}
@@
 Since Tahoe-\ac{LAFS} is written in Python, we needed some way of interacting
-with the C implementations of the NIST candidates.
+with the C implementations of the NIST candidates. We did this by the use of
+Cython, and built a python object-oriented library which we named Sha3lib.
+Sha3lib supports all of the 14 second round \ac{SHA}-3 candidates, but only in
+their 256-bit form.
+
+\subsection{Sha3lib}
+Sha3lib is built to mimic the functionality of
+hashlib\footnote{\url{http://docs.python.org/library/hashlib.html}}, the
+default python library for MD5, \ac{SHA}-1 and the \ac{SHA}-2 family. 
+% TODO, figure or listing of what this means
+
+\subsubsection{Cython}
+We employ Cython to generate bindings for the C implementation and the
+different candidates, as well as to create the base python classes. To access
+functions in C we need to generate a Cython header file, which pretty much does
+the same as a C header file; it tells which functions and data types are
+available. How we accomplish this for an implementation that satisfies the
+\ac{NIST} \ac{API} is shown in Listing \ref{lst:cython_header}.
+
+\begin{lstlisting}[label=lst:cython_header, caption=Cython header\, echo\_hash\_h.pxd]
+cdef extern from "sha3nist.h":
+    ctypedef unsigned char BitSequence
+    ctypedef unsigned long long DataLength
+
+    cdef enum HashReturn:
+        SUCCESS = 0
+        FAIL = 1
+        BAD_HASHLEN = 2
+
+    ctypedef struct hashState:
+        pass
+
+    HashReturn Hash(int, BitSequence *data, DataLength, BitSequence *hashval)
+    HashReturn Init(hashState *state, int hashbitlen)
+    HashReturn Update(hashState *state, BitSequence *data, DataLength databitlen)
+    HashReturn Final(hashState *state, BitSequence *hashval)
+\end{lstlisting}
+
+We then create a wrapper-class that mimics hashlib, this is shown in Listing
+\ref{lst:cython_impl}. Functions declared with def er python functions, cdef are
+Cython only while cpdef are functions that can be accessed from both. While
+this is just the implementation for the echo candidate, the almost exact same
+pattern is used for all of the other candidates.
+
+\begin{lstlisting}[label=lst:cython_impl, caption=Cython class\, echo\_hash.pyx]
+cimport echo_hash_h
+cimport cython
+
+from libc.stdlib cimport *
+
+cdef class echo:
+    '''
+    A class that tries to mimic the behaviour of hashlib, i.e. keeping state
+    so that one can update the hashing procedure instead of doing it from
+    scratch.
+    '''
+    #cdef echo_hash_h.BitSequence *hashval
+    cdef echo_hash_h.hashState previous_state
+    cdef echo_hash_h.hashState state
+    cdef int finished
+    cdef int hashbitlen
+    cdef list hashval
+
+    def __init__(self, int in_hashbitlen, bytes initial=None):
+        self.finished = 0
+        self.hashbitlen = in_hashbitlen
+        echo_hash_h.Init(&self.state, self.hashbitlen)
+
+        if initial:
+            self.update(initial)
+
+    cpdef update(self, bytes in_data):
+        cdef char* data = <char *> in_data
+        cdef int data_len = len(in_data)*8
+
+        if self.finished:
+            self.finished = 0
+            self.state = self.previous_state
+
+        echo_hash_h.Update(&self.state, <echo_hash_h.BitSequence *> data, data_len)
+
+    cpdef final(self):
+        cdef echo_hash_h.BitSequence *hashval = <echo_hash_h.BitSequence *> malloc(self.hashbitlen*8)
+
+        # We copy the state so that we can continue to update.
+        # This equals hashlibs functionality, but not pycryptopp.
+        self.previous_state = self.state
+
+        echo_hash_h.Final(&self.state, hashval)
+        self.finished = 1
+
+        self.hashval = [hashval[i] for i from 0 <= i < self.hashbitlen / 8]
+        free(hashval)
+
+    cpdef copy(self):
+        s = echo(self.hashbitlen)
+        s.state = self.state
+        s.previous_state = self.previous_state
+        s.finished = self.finished
+        if s.finished:
+            s.hashval = self.hashval
+
+        return s
+
+    def digest(self):
+        if not self.finished:
+            self.final()
+
+        return ''.join(map(chr, self.hashval))
+
+    def hexdigest(self):
+        if not self.finished:
+            self.final()
+
+        return ''.join(['%02x' % i for i in self.hashval])
+\end{lstlisting}
+
+\subsubsection{\ac{Distutils}} We use the
+\ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to build and
+install our code. The alternative to doing this would be to manually transform
+cython files to C-files, build the shared library from this and other relevant
+C-files and then install by copying into the correct system directory where
+python libraries should be stored. \ac{Distutils} also support expansion
+through self defined commands, something we have utilised to do Unit Testing.
+Options and configuration for \ac{Distutils} is done through a file setup.py.
+Relevant configuration for compiling the echo extension can be seen in Listing.
+
+\begin{lstlisting}[label=lst:setup_py, caption=Compiling echo from \ac{Distutils}]
+from distutils.core import setup, Command
+from distutils.extension import Extension
+from Cython.Distutils import build_ext
+
+srcpath = 'sha3lib/'
+hashpath = srcpath + 'hash_functions/'
+
+ext_modules = [
+    Extension("sha3lib.hash_functions.echo.echo_hash",
+        [hashpath+"echo/echo_hash.pyx", 
+         hashpath+'echo/sha3nist.c', 
+         hashpath+"echo/echo.c"],
+    extra_compile_args=[
+        '-funroll-loops',
+        '-m32','-march=prescott',
+        '-O2',
+        '-fomit-frame-pointer']
+    ),
+]
+
+setup(
+    name='SHA3lib',
+    version='0.1',
+    description='Python bindings for SHA3-256 using Cython',
+    cmdclass = {'build_ext': build_ext},
+    ext_modules = ext_modules,
+    packages =
+        ['sha3lib',
+        'sha3lib.hash_functions',
+        'sha3lib.hash_functions.echo',
+        ],
+)
+
+\end{lstlisting}
 

### 2010-11-16 paper/main.tex
Cleaned up source code

@@
 with the C implementations of the NIST candidates. We did this by the use of
-Cython, and built a python object-oriented library which we named Sha3lib.
-Sha3lib supports all of the 14 second round \ac{SHA}-3 candidates, but only in
+Cython, and built a python object-oriented library which we named SHA3lib.
+SHA3lib supports all of the 14 second round \ac{SHA}-3 candidates, but only in
 their 256-bit form.
 
-\subsection{Sha3lib}
-Sha3lib is built to mimic the functionality of
+\subsection{SHA3lib}
+SHA3lib is built to mimic the functionality of
 hashlib\footnote{\url{http://docs.python.org/library/hashlib.html}}, the
@@
 
-    HashReturn Hash(int, BitSequence *data, DataLength, BitSequence *hashval)
-    HashReturn Init(hashState *state, int hashbitlen)
-    HashReturn Update(hashState *state, BitSequence *data, DataLength databitlen)
-    HashReturn Final(hashState *state, BitSequence *hashval)
+    HashReturn Hash(int, BitSequence *data, 
+                    DataLength, 
+                    BitSequence *hashval)
+    HashReturn Init(hashState *state, 
+                    int hashbitlen)
+    HashReturn Update(hashState *state, 
+                      BitSequence *data, 
+                      DataLength databitlen)
+    HashReturn Final(hashState *state, 
+                    BitSequence *hashval)
 \end{lstlisting}
@@
     '''
-    A class that tries to mimic the behaviour of hashlib, i.e. keeping state
-    so that one can update the hashing procedure instead of doing it from
-    scratch.
+    A class that tries to mimic the behaviour of hashlib, 
+    i.e.  keeping state so that one can update the hashing 
+    procedure instead of doing it from scratch.
     '''
-    #cdef echo_hash_h.BitSequence *hashval
     cdef echo_hash_h.hashState previous_state
@@
 
-    def __init__(self, int in_hashbitlen, bytes initial=None):
+    def __init__(self, int in_hashbitlen, 
+                 bytes initial=None):
         self.finished = 0
@@
 
-        echo_hash_h.Update(&self.state, <echo_hash_h.BitSequence *> data, data_len)
+        echo_hash_h.Update(&self.state, 
+                        <echo_hash_h.BitSequence *> data, 
+                        data_len)
 
     cpdef final(self):
-        cdef echo_hash_h.BitSequence *hashval = <echo_hash_h.BitSequence *> malloc(self.hashbitlen*8)
+        cdef echo_hash_h.BitSequence *hashval = \
+        <echo_hash_h.BitSequence *> \
+        malloc(self.hashbitlen*8)
 
-        # We copy the state so that we can continue to update.
-        # This equals hashlibs functionality, but not pycryptopp.
+        # We copy the state so that we can continue to 
+        # update. This equals hashlibs functionality, 
+        # but not pycryptopp.
         self.previous_state = self.state
@@
 
-        self.hashval = [hashval[i] for i from 0 <= i < self.hashbitlen / 8]
+        self.hashval = [hashval[i] for i from 0 <= i \
+                        < self.hashbitlen / 8]
         free(hashval)
@@
 Options and configuration for \ac{Distutils} is done through a file setup.py.
-Relevant configuration for compiling the echo extension can be seen in Listing.
+Relevant configuration for compiling the echo extension, and installing SHA3lib
+can bee seen in Listing \ref{lst:setup_py}.
 
@@
         '-funroll-loops',
-        '-m32','-march=prescott',
+        '-m32',
+        '-march=prescott',
         '-O2',
@@
         'sha3lib.hash_functions',
-        'sha3lib.hash_functions.echo',
-        ],
+        'sha3lib.hash_functions.echo',],
 )

### 2010-11-16 paper/main.tex
Add some text on the test config and vectors

@@
 \acro{CEB}{Capability Extension Block}
+\acro{CPU}{Central Processing Unit}
 \acro{Distutils}{Python Distribution Utilities}
@@
 split up, before each part is sent to other nodes in the grid, as depicted in
-figure \ref{fig:tahoeinsertion}. The integrity and confidentiality of the files
+Figure \ref{fig:tahoeinsertion}. The integrity and confidentiality of the files
 are guaranteed by the algorithms used on the client, and is independent of the
@@
 
-\begin{lstlisting}[label=lst:hashutilmod, caption=Relevant parts of hashutil.py of Tahoe-LAFS source after modification.]
+\begin{lstlisting}[label=lst:hashutilmod, caption=Parts of hashutil.py of Tahoe-LAFS source after modification.]
 from sha3lib import bmw256 as SHA256
@@
 grid while using another hash function.
+
 For the same reasons, some of the built-in unit tests of Tahoe-LAFS fail when
-changing this line, because they contain test fixtures made with SHA-256.
-This also renders uploading of mutable files impossible, since the internal
-verification fails. For this reason, \emph{only immutable files} are tested and
-measured in this paper.
+changing this line, because they contain test fixtures made with the included
+SHA-256.  An implication of this is that uploading of mutable files renders
+impossible, since the internal verification fails. For this reason, \emph{only
+immutable files} are tested and measured in this paper.
 
-\section{Test vectors and Configuration}
+\section{Configuration of test environment}
 
-% TODO: Write something about why we only test the client
+Tahoe-\ac{LAFS} was configured to have three storage nodes available --
+\emph{shares available} -- and that only two of these had to be accessible to be
+able to restore files from the grid -- \emph{shares needed}. A setting called
+\emph{shares happy} was set to three storage nodes. This makes Tahoe ensure that
+all storage nodes are available, so that the test environment behaves the same
+for all the different tests.
+
+The performance measurements are only performed on the Tahoe client, and not on
+the storage servers. We motivated this by looking at the system performance
+graphs of the storage nodes during upload and retrieval, where neither the
+\ac{CPU}, network, disk or memory usage runs out of capacity. The \ac{CPU} usage
+during testing on one of the storage nodes is depicted in Figure
+\ref{fig:munin:storagenode}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.8\columnwidth]{munin-tahoe05-cpuday.png}
+    \caption{Graph of CPU usage during testing on a Tahoe storage node.}
+    \label{fig:munin:storagenode}
+\end{figure}
+
+\subsection{Test Vectors}
+
+A variety of different files was uploaded and subsequently downloaded using the
+configured Tahoe grid. The numbers of the randomly generated different files
+tested are shown in Table \ref{tbl:test_vectors}.
+
+\input{tables/test_vectors}
+
+The distribution of file sizes is based on the need to test different use cases.
+Although it is probably unusual to upload one byte files, this enables us to see
+what happens when network delay is brought to a minimum. At the other extreme,
+the one gigabyte file will show another perspective of resource utilization.
+
+% FIXME: Maybe add some more reasoning here.
 
@@
 
+What if we had used another coding scheme?
+
+Implications of only testing immutable files?
+
+Adding a capability to support new hash functions in tahoe.
+
 %**************************************%

### 2010-11-16 paper/main.tex
Added short intro to Technical Procedure

@@
 %**************************************%
-\chapter{Procedure}
+\chapter{Technical Procedure}
 %**************************************%
 
-% FIXME: A more descriptive/better chapter title
+% FIXME: A more descriptive/better chapter title?
 %
@@
 
-In this chapter we will outline the procedure followed to complete our task.
-% TODO: Mer intro
+In this chapter we will outline the procedure followed to complete our work.
+Although not carried out sequentially, the problem was split up to the
+following parts:
+
+\begin{enumerate}
+  \item Choose the best suiting implementations of the SHA-3 candidates.
+  \item Write Python bindings, with unit tests for verification, to be able to
+    use the functions from Python.
+  \item Modify the Tahoe-LAFS code to utilize the SHA-3 functions.
+  \item Create a method of distributing new code fast and easy to the test grid.
+  \item Set up the test environment and choose test vectors.
+\end{enumerate}
 
@@
 
-\section{Modifications in the Tahoe-\ac{LAFS} Code}
+\section{Modifications to the Tahoe-\ac{LAFS} Code}
 
@@
 
+\section{Distribution of the Code}
+% Distutil, bash scripts
+
 \section{Configuration of test environment}
@@
 % FIXME: Maybe add some more reasoning here.
-
-\section{Distribution of the Code}
-% Distutil, bash scripts
 

### 2010-11-17 python-bindings/sha3lib/hash_functions/bmw/BlueMidnightWish.h
Testing other bmw version

@@ -52,12 +52,12 @@ typedef enum { SUCCESS = 0, FAIL = 1, BAD_HASHLEN = 2, BAD_CONSECUTIVE_CALL_TO_U
 
 typedef struct
 {
-	u_int32_t DoublePipe[32];
+	u_int32_t DoublePipe[16];
 	BitSequence LastPart[BlueMidnightWish256_BLOCK_SIZE * 2];
 } Data256;
 typedef struct
 {
-	u_int64_t DoublePipe[32];
+	u_int64_t DoublePipe[16];
 	BitSequence LastPart[BlueMidnightWish512_BLOCK_SIZE * 2];
 } Data512;
 
@@ -68,8 +68,13 @@ typedef struct {
 	u_int64_t bits_processed;
 	union
     { 
-		Data256  p256[1];
-		Data512  p512[1];
+		#if defined ( _MSC_VER )
+		__declspec(align(16)) Data256  p256[1];
+		__declspec(align(16)) Data512  p512[1];
+		#else
+		__attribute__ ((aligned (16))) Data256  p256[1];
+		__attribute__ ((aligned (16))) Data512  p512[1];
+		#endif
     } pipe[1];
 	int unprocessed_bits;
 } hashState;

### 2010-11-17 python-bindings/sha3lib/hash_functions/bmw/api.h

@@ -1,3 +1,3 @@
 #define CRYPTO_BYTES 32
-#define CRYPTO_VERSION "Reference C - BMW256"
+#define CRYPTO_VERSION "Optimized C - BMW256, opt02"
 

### 2010-11-17 python-bindings/sha3lib/hash_functions/bmw/BlueMidnightWish.c
Reverted bmw version

@@ -1,16 +1,46 @@
+/* Blue Midnight Wish hash function as it will be submitted to 
+   NIST for the second round of SHA-3 competition.
+   Changes in this version (compared to the version submitted for the Round 1):
+
+   1. Corrected IV for 224 and 384 version.
+
+   2. Tweaked f0() and f1()
+
+   3. Use of final compression function invocation.
+
+*/													
+
 /* 
-   Optimized C code for tweaked Blue Midnight Wish - 256 programmed by Rune E. Jensen, December 2009.
+   This is an optimized C code for Blue Midnight Wish hash function 
+   as it will be submitted to NIST for the second round of SHA-3 competition.
+
+   Programmed by Danilo Gligoroski, August 2009.
 */
 
-#include <stdio.h>
+
 #include <string.h> 
 #include "BlueMidnightWish.h"
 
 #define rotl32(x,n)   (((x) << n) | ((x) >> (32 - n)))
 #define rotr32(x,n)   (((x) >> n) | ((x) << (32 - n)))
+//#define rotl32 _lrotl
+//#define rotr32 _lrotr
+
+#define rotl64(x,n)   (((x) << n) | ((x) >> (64 - n)))
+#define rotr64(x,n)   (((x) >> n) | ((x) << (64 - n)))
+//#define rotl64 _rotl64
+//#define rotr64 _rotr64
 
 #define shl(x,n)      ((x) << n)
 #define shr(x,n)      ((x) >> n)
+
+/* BlueMidnightWish224 initial double chaining pipe */
+const u_int32_t i224p2[16] =
+{   0x00010203ul, 0x04050607ul, 0x08090a0bul, 0x0c0d0e0ful,
+    0x10111213ul, 0x14151617ul, 0x18191a1bul, 0x1c1d1e1ful,
+    0x20212223ul, 0x24252627ul, 0x28292a2bul, 0x2c2d2e2ful,
+    0x30313233ul, 0x34353637ul, 0x38393a3bul, 0x3c3d3e3ful,
+};
 
 /* BlueMidnightWish256 initial double chaining pipe */
 const u_int32_t i256p2[16] =
[…]

### 2010-11-17 python-bindings/sha3lib/hash_functions/bmw/BlueMidnightWish.h

@@ -52,12 +52,12 @@ typedef enum { SUCCESS = 0, FAIL = 1, BAD_HASHLEN = 2, BAD_CONSECUTIVE_CALL_TO_U
 
 typedef struct
 {
-	u_int32_t DoublePipe[16];
+	u_int32_t DoublePipe[32];
 	BitSequence LastPart[BlueMidnightWish256_BLOCK_SIZE * 2];
 } Data256;
 typedef struct
 {
-	u_int64_t DoublePipe[16];
+	u_int64_t DoublePipe[32];
 	BitSequence LastPart[BlueMidnightWish512_BLOCK_SIZE * 2];
 } Data512;
 
@@ -68,13 +68,8 @@ typedef struct {
 	u_int64_t bits_processed;
 	union
     { 
-		#if defined ( _MSC_VER )
-		__declspec(align(16)) Data256  p256[1];
-		__declspec(align(16)) Data512  p512[1];
-		#else
-		__attribute__ ((aligned (16))) Data256  p256[1];
-		__attribute__ ((aligned (16))) Data512  p512[1];
-		#endif
+		Data256  p256[1];
+		Data512  p512[1];
     } pipe[1];
 	int unprocessed_bits;
 } hashState;

### 2010-11-17 python-bindings/sha3lib/hash_functions/bmw/api.h

@@ -1,3 +1,2 @@
 #define CRYPTO_BYTES 32
-#define CRYPTO_VERSION "Optimized C - BMW256, opt02"
-
+#define CRYPTO_VERSION "Optimized C - BMW256, opt01"

### 2010-11-17 paper/main.tex
Restyle chapter headers
Added design of SHA3lib

@@
 \geometry{bindingoffset=1cm}
+
+% Restyle chapter headers
+\usepackage{fix-cm}
+\makeatletter
+\renewcommand{\@makechapterhead}[1]{%
+  \vspace*{50\p@}%
+  {\parindent \z@ \raggedright \normalfont
+    \vspace{15pt}%
+    \ifnum \c@secnumdepth >\m@ne
+        %\hfill\huge\scshape \@chapapp\space
+        \hfill\fontsize{60}{90}\selectfont \thechapter % Chapter number
+        \par\nobreak
+        \vskip 20\p@
+    \fi
+    \interlinepenalty\@M
+    \hfill \Huge \scshape #1\par % Chapter title
+    \vspace{5pt}
+    \hrule
+    \nobreak
+    \vskip 40\p@
+  }}
+\makeatother
 
@@
 hashlib\footnote{\url{http://docs.python.org/library/hashlib.html}}, the
-default python library for MD5, \ac{SHA}-1 and the \ac{SHA}-2 family. 
-% TODO, figure or listing of what this means
+default python library for MD5, \ac{SHA}-1 and the \ac{SHA}-2 family. This
+means that we built class names for each candidate on the form
+candidatename+256 and that the classes included the following functions:
+
+\begin{itemize}
+    \item Constructur with an optional data element that does the same as
+    Update()
+    \item Update() takes data that should be considered in hashing 
+    \item digest() that produce a raw byte-digest from submitted data
+    \item hexdigest() that produce a hex representation of the digest
+    \item copy() that copies the internal state of the function and produces a
+    copy of the object.
+\end{itemize}
+
+\subsubsection{Design}
+Our original goal was to make SHA3lib include functionality to support all
+output lengths, and not only 256-bit. This was based on our observation that
+the official \ac{NIST} submission packages include ``all-in-one''
+implementations that support this. Later we observed that this is not the case
+for all implementations in \ac{SUPERCOP}. Together with the fact that
+Tahoe-\ac{LAFS} uses 256-bit output we in the end only included support for
+256-bit output in SHA3lib. The design we have used is shown in Figure
+\ref{fig:python:inheritance}, and does follow the initial idea that SHA3lib
+should support multiple output lengths.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.6\columnwidth]{python-inheritance.pdf}
+    \caption{Design of SHA3lib using echo as an example}
+    \label{fig:python:inheritance}
+\end{figure}
+
+Basically the class echo resides in Cython and in C and is responsible for
+bringing the C functionality of the C implementation to Python. By compiling
+echo it will be made available for use in Python.  The class echo256 is a
+simple python class which inherits from echo. The only action this class
+actually takes is to set the output size to 256-bit. If the underlying
+C-functions would have supported multiple output sizes the same could have been
+done to incorporate those.
 

### 2010-11-17 paper/main.tex
Added methodology and more on test environment/procedure

@@
 
-Technical procedure, Cython, two persons, test environment, how we test,
-GitHub?
+All work associated with the project has been carried out by the two authors in
+cooperation. Code are written using the VIM editor on Ubuntu Linux
+driven machines, and version-controlled using the distributed revision control
+system Git. All material produced, including this paper, are publicly available
+on GitHub\footnote{\url{http://github.com/ruudud/Tahoe-LAFS-with-SHA-3}}.
+
+The \ac{SHA}-3 candidates are tested on the Tahoe-\ac{LAFS} using uploading and
+downloading of files on four desktop computers. Various file sizes are used to
+simulate different use cases, and to broaden the perspective as to how the
+different candidates perform and behaves when implemented in a secure,
+distributed file system.
 
@@
 %**************************************%
-% FIXME: Find new and better title
 
@@
 
-\section{Configuration of test environment}
+\section{Configuration of Test Environment}
+
+Four machines was used as a testing grid, each containing an Intel(R) Core(TM)2
+Duo E8300 \ac{CPU} and 4GB of \ac{RAM}. The 32-bit version of Ubuntu Linux
+Server Edition 10.04 was the operating system of choice. The installed versions
+of other relevant software can be seen in Table \ref{tbl:installedsw}.
+
+\input{tables/installed_software}
 
@@
 
-The performance measurements are only performed on the Tahoe client, and not on
+The performance measurements were only performed on the Tahoe client, and not on
 the storage servers. We motivated this by looking at the system performance
@@
 \end{figure}
+
+The tests was carried out as follows, for each of the test vectors:
+
+\begin{itemize}
+  \item Upload all files in the same category (as defined shortly) one by one.
+  \item Download all files in the same manner.
+\end{itemize}
+
+During these operations, we counted the number of hashing operations performed,
+and timed total time spent hashing, i.e. the total time it took for the function
+calls to the underlying hashing operations to complete.
 

### 2010-11-17 paper/tables/sha3_flags.tex
Added design of SHA3lib
Added list of candidates, fixed typos in tables

@@
     \hline
-    \textbf{candidate} & \textbf{from}  \\ \hline
-     blake       & mssse3 -m32 -march=native -mtune=native -O -fomit-frame-pointer  \\ \hline
-     bmw         & -funroll-loops -m32 -march=pentium4 -O3 -fomit-frame-pointer     \\ \hline
-     cubehash    & none                                                             \\ \hline
-     echo        & -funroll-loops -m32 -march?prescott -O2 -fomit-frame-pointer     \\ \hline
-     fugue       & -funroll-loops -O -fomit-frame-pointer                           \\ \hline
-     groestl     & -m32 -march=pentium3 -O -fomit-frame-pointer                     \\ \hline
-     hamsi       & -msse2 -funroll-loops -m32 -march=pentium-m -O3 -fomit-frame-pointer\\ \hline
-     jh          & -m32 -march=core2 -msse4.1 -O -fomit-frame-pointer               \\ \hline
-     keccak      & -msse2 -march=barcelona -O3 -fomit-frame-pointer                 \\ \hline
-     luffa       & -msse2 -funroll-loops -march=k8 -O3 -fomit-frame-pointer         \\ \hline
-     shabal      & -funroll-loops -m32 -march=pentium2 -O -fomit-frame-pointer      \\ \hline
-     shavite3    & -funroll-loops -march=i486 -O -fomit-frame-pointer               \\ \hline
-     simd        & -m32 -march=barcelona -O2 -fomit-frame-pointer                   \\ \hline
-     skein       & -march=pentium2 -O2 -fomit-frame-pointer                         \\ \hline
+    \textbf{Candidate} & \textbf{Compiler flags}  \\ \hline
+     BLAKE       & mssse3 -m32 -march=native -mtune=native -O -fomit-frame-pointer  \\ \hline
+     \ac{BMW}    & -funroll-loops -m32 -march=pentium4 -O3 -fomit-frame-pointer     \\ \hline
+     Cubehash    & none                                                             \\ \hline
+     Echo        & -funroll-loops -m32 -march?prescott -O2 -fomit-frame-pointer     \\ \hline
+     Fugue       & -funroll-loops -O -fomit-frame-pointer                           \\ \hline
+     Grøstl     & -m32 -march=pentium3 -O -fomit-frame-pointer                     \\ \hline
+     Hamsi       & -msse2 -funroll-loops -m32 -march=pentium-m -O3 -fomit-frame-pointer\\ \hline
+     JH          & -m32 -march=core2 -msse4.1 -O -fomit-frame-pointer               \\ \hline
+     Keccak      & -msse2 -march=barcelona -O3 -fomit-frame-pointer                 \\ \hline
+     Luffa       & -msse2 -funroll-loops -march=k8 -O3 -fomit-frame-pointer         \\ \hline
+     Shabal      & -funroll-loops -m32 -march=pentium2 -O -fomit-frame-pointer      \\ \hline
+     Shavite3    & -funroll-loops -march=i486 -O -fomit-frame-pointer               \\ \hline
+     Simd        & -m32 -march=barcelona -O2 -fomit-frame-pointer                   \\ \hline
+     Skein       & -march=pentium2 -O2 -fomit-frame-pointer                         \\ \hline
   \end{tabular}
-  \caption{\ac{GCC} compiler flags used on the implementations}
+  \caption{GCC compiler flags used on the implementations}
   \label{tbl:sha3:compilerflags}

### 2010-11-21 paper/presentation.tex
Added preliminary presentation

\documentclass[screen]{beamer}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

%
%   For at dette skal funke:
%   wget http://www.ntnu.no/selvhjelpspakken/ntnubeamer/beamerntnu.tar.gz
%   extract til: /usr/share/texmf/tex/latex/beamer/themes
%   kjør sudo mktexlsr
%   sourcE: http://www.ntnu.no/selvhjelpspakken/ntnubeamer/

% Bruk NTNU-temaet for beamer (her i bokmålvariant), alternativer er
% ntnunynorsk og ntnuenglish.
\usetheme{ntnuenglish}

% Angi tittelen, vi gir også en kortere variant som brukes nederst på
% hver slide:
\title[SHA-3 in Tahoe-LAFS]%
{Experimenting with SHA-3 candidates in Tahoe Least Authorithy File System}

% Denne kan du også bruke hvis det passer seg:
%\subtitle{Valgfri undertittel}

% Angir foredragsholder, også en (valgfri) kortversjon i
% hakeparanteser først som kommer nederst på hver slide:
\author[EH \& PR]{Eirik Haver and Pål Ruud}

% Institusjon. Bruk gjerne disse slik det passer best med det du vil
% ha.  Valgfri kortversjon her også
\institute[NTNU]{Department of Telematics}

% Datoen blir også trykket på forsida. 
\date{X. December 2010}
%\date{} % Bruk denne hvis du ikke vil ha noe dato på forsida.

% Fra her av begynner selve dokumentet
\begin{document}

% Siden NTNU-malen har en annen bakgrunn på forsida, må dette gjøres
% i en egen kommando, ikke på vanlig beamer-måte:
\ntnutitlepage

% Her begynner første slide/frame, (nummer to etter forsida). 
\begin{frame}
  \frametitle{Tahoe-LAFS}
  \begin{itemize}
  \item Distributed FS
  \item Principle of Least Authority
  \item Erasure coding, k-out-of-n
  \item Written in Python
  \item Uses (double) SHA256-today
    \begin{itemize}
        \item Because of Length-Extension-Attack
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{NIST SHA-3 Competition}
  \begin{itemize}
  \item Select SHA-3 implementation by 2nd Quarter 2012
  \item Currently 14 candidates, one from NTNU
  \item Resistant to Length-Extension-Attacks
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{What we have done}
  \begin{enumerate}
  \item Made Python-binding for all candidates (C implementations) 
  \item Modified Tahoe-LAFS to use the bindings
  \item Benchmarked Tahoe-LAFS with the different candidate implementations
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Python-bindings}
  \begin{itemize}
  \item Optimized version of candidates available at eBACS: ECRYPT Benchmarking
  of Cryptographic Systems.
  \item We have used the fastest, capable C-version available for our target
  architecture, 32-bit.
  \item Made them available in Python by the use of Cython in ``SHA3Lib''
  \end{itemize}
\end{frame}

\end{document}

### 2010-12-05 paper/main.tex
Added list of candidates, fixed typos in tables
Added short outline

@@
 \acro{API}{Application programming interface}
+\acro{BMW}{Blue Midnight Wish}
 \acro{CEB}{Capability Extension Block}
@@
 
+Scope and objectives
+
 \section{Method}
@@
 
-\section{Scope and objectives}
-
-Should we include this?
-
 \section{Outline}
 
-What follows in this document, chapter by chapter
+The work is presented as per the following chapters:
+
+\paragraph{Chapter 2} provides background knowledge of the technologies and
+software used.
+
+\paragraph{Chapter 3} outlines the procedure taken to be able to do the
+measurements.
+
+\paragraph{Chapter 4} presents the actual results, in the form of graphs and
+tables, in addition to some easily seen conclusions.
+
+\paragraph{Chapter 5} discusses the results and our technical procedure to the
+problem presented in this assignment.
+
+\paragraph{Chapter 6} concludes the results and the work done, and proposes
+future work.
 
@@
 DataLength represents the size in bits of the data provided.
+
+\subsubsection{The second round candidates}
+The names of the second round candidates for the \ac{SHA}-3 \ac{NIST}
+competition along with their principal submitters are listed in Table
+\ref{tbl:sha3:candidates}\cite{s_nistround2}.
+
+\input{tables/sha3_authors}
 

### 2010-12-05 paper/main.tex
Added verb environment around function names

@@
 
-The Init() function basically sets up the internal state of the function, to
-make it ready to start processing real data. The Update() function is used to
-provide the actuall data to the hash function. The Final() function should be
-called when all necessary data has been gived through Update(), it will then
+The {\verb Init() } function basically sets up the internal state of the function, to
+make it ready to start processing real data. The {\verb Update() } function is used to
+provide the actual data to the hash function. The {\verb Final() } function should be
+called when all necessary data has been given through {\verb Update() }, it will then
 finish the hash function and provide the digest. The Hash() function basically a wrapper
@@
 
-HashReturn is an enum representing that an operation either succeded (0) or
-went wrong. The BitSequence is a representation of an array of bytefields and
-DataLength represents the size in bits of the data provided.
+{\verb HashReturn } is an \emph{enum} representing that an operation either succeeded (0) or
+went wrong. The {\verb BitSequence } is a representation of an array of byte fields and
+{\verb DataLength } represents the size in bits of the data provided.
 
@@
 organization of files, and that a submission must include a function
-crypto\_hash() which is similar to the function Hash() in the \ac{NIST} \ac{SHA}-3
-\ac{API}. The only difference beeing the naming of datatypes and that the
-crypto\_hash() function does not need an input of how long the output of the
-hash function should be. Instead double or different submissions are used for
-different output lengths.
+{\verb crypto\_hash() } which is similar to the function {\verb Hash() } in the
+\ac{NIST} \ac{SHA}-3 \ac{API}. The only difference being the naming of
+data types and that the {\verb crypto\_hash() } function does not need an input
+of how long the output of the hash function should be. Instead double or
+different submissions are used for different output lengths.
 
@@
 cannot be modified, and can be read repeatedly. Mutable files can be modified,
-and everyone who has access to the \emph{signing key} can make new versions of
+and everyone who has access to the signing key can make new versions of
 the mutable file.
@@
 These secure hashes of the shares of a file, are contained within a small
-datastructure named the \emph{\ac{CEB}}.
+data structure named the \emph{\ac{CEB}}.
 
@@
 
-% FIXME: A more descriptive/better chapter title?
-%
 %A bit more technical description of how we tested the candidates in Tahoe-LAFS

### 2010-12-05 paper/main.tex
Moved code to external files to keep main tex clean

@@
 
-\begin{lstlisting}[label=lst:cython_header, caption=Cython header\, echo\_hash\_h.pxd]
-cdef extern from "sha3nist.h":
-    ctypedef unsigned char BitSequence
-    ctypedef unsigned long long DataLength
-
-    cdef enum HashReturn:
-        SUCCESS = 0
-        FAIL = 1
-        BAD_HASHLEN = 2
-
-    ctypedef struct hashState:
-        pass
-
-    HashReturn Hash(int, BitSequence *data, 
-                    DataLength, 
-                    BitSequence *hashval)
-    HashReturn Init(hashState *state, 
-                    int hashbitlen)
-    HashReturn Update(hashState *state, 
-                      BitSequence *data, 
-                      DataLength databitlen)
-    HashReturn Final(hashState *state, 
-                    BitSequence *hashval)
-\end{lstlisting}
+\lstinputlisting[label=lst:cython_header, caption=Cython header\, echo\_hash\_h.pxd]
+{listings/echo_hash_h.pxd}
 
 We then create a wrapper-class that mimics hashlib, this is shown in Listing
-\ref{lst:cython_impl}. Functions declared with def er python functions, cdef are
-Cython only while cpdef are functions that can be accessed from both. While
-this is just the implementation for the echo candidate, the almost exact same
-pattern is used for all of the other candidates.
+\ref{lst:cython_impl}. Functions declared with def er python functions,
+{\verb cdef } are Cython only while {\verb cpdef } are functions that can be
+accessed from both. While this is just the implementation for the echo
+candidate, the almost exact same pattern is used for all of the other
+candidates.
 
-\begin{lstlisting}[label=lst:cython_impl, caption=Cython class\, echo\_hash.pyx]
-cimport echo_hash_h
-cimport cython
+\lstinputlisting[label=lst:cython_impl, caption=Cython class\, echo\_hash.pyx]
+{listings/echo_hash.pyx}
 
-from libc.stdlib cimport *
+\subsubsection{\ac{Distutils}}
+We use the \ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to
+build and install our code. The alternative to doing this would be to manually
+transform Cython files to C-files, build the shared library from this and other
+relevant C-files and then install by copying into the correct system directory
+where python libraries should be stored. \ac{Distutils} also support expansion
+through self defined commands, something we have utilised to do Unit Testing.
+Options and configuration for \ac{Distutils} is done through a file
+{\verb setup.py}. Relevant configuration for compiling the echo extension, and
+installing SHA3lib can bee seen in Listing \ref{lst:setup_py}.
 
-cdef class echo:
-    '''
-    A class that tries to mimic the behaviour of hashlib, 
-    i.e.  keeping state so that one can update the hashing 
-    procedure instead of doing it from scratch.
-    '''
-    cdef echo_hash_h.hashState previous_state
-    cdef echo_hash_h.hashState state
-    cdef int finished
-    cdef int hashbitlen
-    cdef list hashval
-
-    def __init__(self, int in_hashbitlen, 
-                 bytes initial=None):
-        self.finished = 0
-        self.hashbitlen = in_hashbitlen
-        echo_hash_h.Init(&self.state, self.hashbitlen)
-
-        if initial:
-            self.update(initial)
-
-    cpdef update(self, bytes in_data):
-        cdef char* data = <char *> in_data
-        cdef int data_len = len(in_data)*8
-
-        if self.finished:
-            self.finished = 0
-            self.state = self.previous_state
-
-        echo_hash_h.Update(&self.state, 
-                        <echo_hash_h.BitSequence *> data, 
-                        data_len)
-
-    cpdef final(self):
-        cdef echo_hash_h.BitSequence *hashval = \
-        <echo_hash_h.BitSequence *> \
-        malloc(self.hashbitlen*8)
-
-        # We copy the state so that we can continue to 
-        # update. This equals hashlibs functionality, 
-        # but not pycryptopp.
-        self.previous_state = self.state
-
-        echo_hash_h.Final(&self.state, hashval)
-        self.finished = 1
-
-        self.hashval = [hashval[i] for i from 0 <= i \
-                        < self.hashbitlen / 8]
-        free(hashval)
-
-    cpdef copy(self):
-        s = echo(self.hashbitlen)
-        s.state = self.state
-        s.previous_state = self.previous_state
-        s.finished = self.finished
-        if s.finished:
-            s.hashval = self.hashval
-
-        return s
-
-    def digest(self):
-        if not self.finished:
-            self.final()
-
-        return ''.join(map(chr, self.hashval))
-
-    def hexdigest(self):
-        if not self.finished:
-            self.final()
-
-        return ''.join(['%02x' % i for i in self.hashval])
-\end{lstlisting}
-
-\subsubsection{\ac{Distutils}} We use the
-\ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to build and
-install our code. The alternative to doing this would be to manually transform
-cython files to C-files, build the shared library from this and other relevant
-C-files and then install by copying into the correct system directory where
-python libraries should be stored. \ac{Distutils} also support expansion
-through self defined commands, something we have utilised to do Unit Testing.
-Options and configuration for \ac{Distutils} is done through a file setup.py.
-Relevant configuration for compiling the echo extension, and installing SHA3lib
-can bee seen in Listing \ref{lst:setup_py}.
-
-\begin{lstlisting}[label=lst:setup_py, caption=Compiling echo from \ac{Distutils}]
-from distutils.core import setup, Command
-from distutils.extension import Extension
-from Cython.Distutils import build_ext
-
-srcpath = 'sha3lib/'
-hashpath = srcpath + 'hash_functions/'
-
-ext_modules = [
-    Extension("sha3lib.hash_functions.echo.echo_hash",
-        [hashpath+"echo/echo_hash.pyx", 
-         hashpath+'echo/sha3nist.c', 
-         hashpath+"echo/echo.c"],
-    extra_compile_args=[
-        '-funroll-loops',
-        '-m32',
-        '-march=prescott',
-        '-O2',
-        '-fomit-frame-pointer']
-    ),
-]
-
-setup(
-    name='SHA3lib',
-    version='0.1',
-    description='Python bindings for SHA3-256 using Cython',
-    cmdclass = {'build_ext': build_ext},
-    ext_modules = ext_modules,
-    packages =
-        ['sha3lib',
-        'sha3lib.hash_functions',
-        'sha3lib.hash_functions.echo',],
-)
-
-\end{lstlisting}
+\lstinputlisting[label=lst:setup_py, caption=Compiling echo from \ac{Distutils}]
+{listings/setup.py}
 
 \subsection{Unit tests}
-% Include the use of Known Answer Tests
+% TODO: Include the use of Known Answer Tests
 
@@
 
-\begin{lstlisting}[label=lst:hashutil, caption=Extract from hashutil.py of Tahoe-LAFS source.]
-from pycryptopp.hash.sha256 import SHA256
-...
-class _SHA256d_Hasher:
-    ...
-    def __init__(self, truncate_to=None):
-        self.h = SHA256()
-        self.truncate_to = truncate_to
-        self._digest = None
-
-    def update(self, data):
-        assert isinstance(data, str) # no unicode
-        self.h.update(data)
-
-    def digest(self):
-        if self._digest is None:
-            h1 = self.h.digest()
-            del self.h
-            h2 = SHA256(h1).digest()
-            if self.truncate_to:
-                h1 = h1[:self.truncate_to]
-                h2 = h2[:self.truncate_to]
-            self._digest = h2
-
-        return self._digest
-\end{lstlisting}
+\lstinputlisting[label=lst:hashutil, caption=Extract from hashutil.py of Tahoe-LAFS source.]
+{listings/hashutil.py}
 
@@
 
-\begin{lstlisting}[label=lst:hashutilmod, caption=Parts of hashutil.py of Tahoe-LAFS source after modification.]
-from sha3lib import bmw256 as SHA256
-...
-class _SHA256d_Hasher:
-    ...
-    def __init__(self, truncate_to=None):
-        self.h = SHA256()
-        self.truncate_to = truncate_to
-        self._digest = None
-
-    def update(self, data):
-        assert isinstance(data, str) # no unicode
-        self.h.update(data)
-
-    def digest(self):
-        if self._digest is None:
-            h1 = self.h.digest()
-            del self.h
-            if self.truncate_to:
-                h1 = h1[:self.truncate_to]
-            self._digest = h1
-
-        return self._digest
-\end{lstlisting}
+\lstinputlisting[label=lst:hashutilmod, caption=Parts of hashutil.py of Tahoe-LAFS source after modification.]
+{listings/hashutil_mod.py}
 
@@
 impossible, since the internal verification fails. For this reason, \emph{only
-immutable files} are tested and measured in this paper.
+immutable files} are tested and measured in this paper. Implications of this is
+discussed further in Chapter \ref{ch:discussion}.
 
@@
 %**************************************%
+\label{ch:discussion}
 

### 2010-12-05 paper/listings/hashutil.py

@@ -1,0 +1,24 @@
+from pycryptopp.hash.sha256 import SHA256
+...
+class _SHA256d_Hasher:
+    ...
+    def __init__(self, truncate_to=None):
+        self.h = SHA256()
+        self.truncate_to = truncate_to
+        self._digest = None
+
+    def update(self, data):
+        assert isinstance(data, str) # no unicode
+        self.h.update(data)
+
+    def digest(self):
+        if self._digest is None:
+            h1 = self.h.digest()
+            del self.h
+            h2 = SHA256(h1).digest()
+            if self.truncate_to:
+                h1 = h1[:self.truncate_to]
+                h2 = h2[:self.truncate_to]
+            self._digest = h2
+
+        return self._digest

### 2010-12-06 paper/presentation.tex
More to presentation

@@
 % Datoen blir også trykket på forsida. 
-\date{X. December 2010}
+\date{December 7th 2010}
 %\date{} % Bruk denne hvis du ikke vil ha noe dato på forsida.
@@
   \item Optimized version of candidates available at eBACS: ECRYPT Benchmarking
-  of Cryptographic Systems.
-  \item We have used the fastest, capable C-version available for our target
+  of Cryptographic Systems. (SUPERCOP toolkit)
+  \item We have used the fastest, capable C-implementations available for our target
   architecture, 32-bit.
@@
 
+\begin{frame}
+  \frametitle{Modification of Tahoe-LAFS}
+  \begin{enumerate}
+  \item Quick hack to quickly swap used hash implementation
+  \item Benchmarking data (number of hashoperations, timers)
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Tahoe-LAFS is SLOW}
+    \begin{itemize}
+        \item On 1 Gbit network - with modern computers
+        \item Tahoe-LAFS output is about 10-12 Megabyte/s
+        \item SCP output is about 80 Megabyte/s
+        \item High-CPU load on client but not on storage nodes
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Benchmarking}
+
+    INSERT graphs.. % TODO
+\end{frame}
+
+\begin{frame}
+    \frametitle{Criticism}
+    \begin{itemize}
+        \item SUPERCOP is a good idea, however we see problems
+            \begin{enumerate}
+                \item Speed of Update() is neglected
+                \item Implementations are not correct according to specification
+                \item Some implementations fails to fullfill the basic requirements
+                of HASH functions
+            \end{enumerate}
+        \item Our implementations are not the fastest possible
+            \begin{enumerate}
+                \item Some performance loss by using Cython (?)
+                \item Not (always) the fastest versions from SUPERCOP
+            \end{enumerate}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Remaining}
+    \begin{enumerate}
+        \item Figure out inconsistencies.
+        \item Finish report....
+    \end{enumerate}
+\end{frame}
+
+\begin{frame}
+    \frametitle{Conclusion}
+    \begin{enumerate}
+        \item There is a performance gain by switching to another hash
+        function
+        \item However, the gain is small and we believe improving other areas
+        could imporve performance more
+    \end{enumerate}
+\end{frame}
+
 \end{document}

### 2010-12-06 paper/main.tex
Added List of Listings and fixed page numbering of pre chapters
Keywords on what to fill out in the report

@@
 
-\addcontentsline{toc}{chapter}{\listfigurename} % Manually add to ToC
+\cleardoublepage
+\phantomsection
+\addcontentsline{toc}{chapter}{\listfigurename}
 \listoffigures
 
+\cleardoublepage
+\phantomsection
 \addcontentsline{toc}{chapter}{\listtablename}
 \listoftables
+
+\cleardoublepage
+\phantomsection
+\addcontentsline{toc}{chapter}{\lstlistlistingname}
+\lstlistoflistings
+\cleardoublepage
 
@@
 \section{Distribution of the Code}
-% Distutil, bash scripts
+% TODO: Distutil, bash scripts
 
@@
 
-\subsection{\ac{SUPERCOP} benchmarks}
+\begin{itemize}
+    \item Section should only contain the hard facts of our measurements
+    \item What is tested and what is not tested?
+    \item How much time is spent on hashing versus total time of
+    uploading/downloading
+    \item How much gain can Tahoe-LAFS get by switching to SHA-3?
+\end{itemize}
+
+\section{\ac{SUPERCOP} benchmarks}
 The results of \ac{SHA}-3
@@
 
-What is tested and what is not tested?
-
-How much time is spent on hashing versus upload over network?
-
-How much gain can Tahoe-LAFS get by switching to SHA-3?
-
-Plenty of tables, graphs
-
 %**************************************%
@@
 
-What could have been done differently?
-
-Compare results with general (supercop) results.
-
-Optimized candidates.
-
-What if we had used another coding scheme?
-
-Implications of only testing immutable files?
-
-Adding a capability to support new hash functions in tahoe.
+\begin{itemize}
+    \item What could have been done differently?
+    \item Compare our results with general (supercop) results.
+    \item Optimized candidates - which gain could be expected from more
+    optimized candidates
+    \item What if we had used another coding scheme? - does Cython lower the
+    candidates performance?
+    \item Implications of only testing immutable files? (mutable files also
+    include directories)
+    \item Adding a capability to support new hash functions in tahoe. (Our
+    current way breaks backwards compatability)
+\end{itemize}
 

### 2010-12-06 paper/presentation.tex
Added tahoe logo and code line displaying import hack

@@
 \usepackage[utf8]{inputenc}
-
+\usepackage{color}
+\usepackage{graphicx}
+\graphicspath{{images/presentation/}}
+\usepackage[absolute,overlay]{textpos}
+\usepackage{listings}
+\lstset{
+  language=Python,
+  backgroundcolor=\color{gray},
+  keywordstyle=\bfseries\ttfamily\color[rgb]{0,0,1},
+  identifierstyle=\ttfamily,
+  commentstyle=\color[rgb]{0.133,0.545,0.133},
+  stringstyle=\ttfamily\color[rgb]{0.627,0.126,0.941},
+  showstringspaces=false,
+  basicstyle=\small,
+  stepnumber=1,
+  numbersep=10pt,
+  tabsize=2,
+  breaklines=true,
+  prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
+  breakatwhitespace=false,
+  aboveskip={1.5\baselineskip},
+  columns=fixed,
+  upquote=true,
+  extendedchars=true,
+  inputencoding=utf8
+}
 %
@@
 \begin{frame}
-  \frametitle{Tahoe-LAFS}
+  \frametitle{Goal}
+
+  Measure and compare performance of the current SHA-3 candidates within the
+  Tahoe-Least Authority File System.
+
+\end{frame}
+
+\begin{frame}
+  \begin{textblock*}{76mm}[1,1](83mm,18mm)
+  \includegraphics[scale=0.4]{tahoe-logo.pdf}
+  \end{textblock*}
+
   \begin{itemize}
@@
   \item Written in Python
-  \item Uses (double) SHA256-today
+  \item Uses (double) SHA256 today
     \begin{itemize}
@@
 
-\begin{frame}
+\begin{frame}[fragile]
   \frametitle{Modification of Tahoe-LAFS}
@@
   \item Quick hack to quickly swap used hash implementation
-  \item Benchmarking data (number of hashoperations, timers)
+  \item Benchmarking data (number of hash operations, timers)
   \end{enumerate}
+
+  \begin{lstlisting}
+    from sha3lib import bmw256 as SHA256
+  \end{lstlisting}
+
 \end{frame}
@@
                 \item Implementations are not correct according to specification
-                \item Some implementations fails to fullfill the basic requirements
+                \item Some implementations fails to fulfill the basic requirements
                 of HASH functions
@@
 \begin{frame}
-    \frametitle{Conclusion}
+    \frametitle{Preliminary Conclusion}
     \begin{enumerate}
         \item There is a performance gain by switching to another hash
-        function
+        function.
         \item However, the gain is small and we believe improving other areas
-        could imporve performance more
+        could improve Tahoe performance more.
     \end{enumerate}

### 2010-12-06 python-bindings/setup.py
Swapped to bmw optc02 version

@@ -16,7 +16,7 @@ hashpath = srcpath + 'hash_functions/'
 ext_modules = [
     Extension("sha3lib.hash_functions.bmw.bmw",
         [hashpath+"bmw/bmw.pyx", hashpath+'bmw/BlueMidnightWish.c'],
-        extra_compile_args=['-funroll-loops','-m32','-march=pentium4','-O3','-fomit-frame-pointer']),
+        extra_compile_args=['-m32','-march=i386','-O','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.groestl.groestl_hash",
         [hashpath+"groestl/groestl_hash.pyx", hashpath+'groestl/sha3nist.c',
             hashpath+"groestl/groestl.c"],

### 2010-12-06 python-bindings/sha3lib/hash_functions/bmw/BlueMidnightWish.c

@@ -1,46 +1,16 @@
-/* Blue Midnight Wish hash function as it will be submitted to 
-   NIST for the second round of SHA-3 competition.
-   Changes in this version (compared to the version submitted for the Round 1):
-
-   1. Corrected IV for 224 and 384 version.
-
-   2. Tweaked f0() and f1()
-
-   3. Use of final compression function invocation.
-
-*/													
-
 /* 
-   This is an optimized C code for Blue Midnight Wish hash function 
-   as it will be submitted to NIST for the second round of SHA-3 competition.
-
-   Programmed by Danilo Gligoroski, August 2009.
+   Optimized C code for tweaked Blue Midnight Wish - 256 programmed by Rune E. Jensen, December 2009.
 */
 
-
+#include <stdio.h>
 #include <string.h> 
 #include "BlueMidnightWish.h"
 
 #define rotl32(x,n)   (((x) << n) | ((x) >> (32 - n)))
 #define rotr32(x,n)   (((x) >> n) | ((x) << (32 - n)))
-//#define rotl32 _lrotl
-//#define rotr32 _lrotr
-
-#define rotl64(x,n)   (((x) << n) | ((x) >> (64 - n)))
-#define rotr64(x,n)   (((x) >> n) | ((x) << (64 - n)))
-//#define rotl64 _rotl64
-//#define rotr64 _rotr64
 
 #define shl(x,n)      ((x) << n)
 #define shr(x,n)      ((x) >> n)
-
-/* BlueMidnightWish224 initial double chaining pipe */
-const u_int32_t i224p2[16] =
-{   0x00010203ul, 0x04050607ul, 0x08090a0bul, 0x0c0d0e0ful,
-    0x10111213ul, 0x14151617ul, 0x18191a1bul, 0x1c1d1e1ful,
-    0x20212223ul, 0x24252627ul, 0x28292a2bul, 0x2c2d2e2ful,
-    0x30313233ul, 0x34353637ul, 0x38393a3bul, 0x3c3d3e3ful,
-};
 
 /* BlueMidnightWish256 initial double chaining pipe */
 const u_int32_t i256p2[16] =
[…]

### 2010-12-06 python-bindings/sha3lib/hash_functions/bmw/BlueMidnightWish.h

@@ -52,12 +52,12 @@ typedef enum { SUCCESS = 0, FAIL = 1, BAD_HASHLEN = 2, BAD_CONSECUTIVE_CALL_TO_U
 
 typedef struct
 {
-	u_int32_t DoublePipe[32];
+	u_int32_t DoublePipe[16];
 	BitSequence LastPart[BlueMidnightWish256_BLOCK_SIZE * 2];
 } Data256;
 typedef struct
 {
-	u_int64_t DoublePipe[32];
+	u_int64_t DoublePipe[16];
 	BitSequence LastPart[BlueMidnightWish512_BLOCK_SIZE * 2];
 } Data512;
 
@@ -68,13 +68,18 @@ typedef struct {
 	u_int64_t bits_processed;
 	union
     { 
-		Data256  p256[1];
-		Data512  p512[1];
+		#if defined ( _MSC_VER )
+		__declspec(align(16)) Data256  p256[1];
+		__declspec(align(16)) Data512  p512[1];
+		#else
+		__attribute__ ((aligned (16))) Data256  p256[1];
+		__attribute__ ((aligned (16))) Data512  p512[1];
+		#endif
     } pipe[1];
 	int unprocessed_bits;
 } hashState;
 
 HashReturn Init(hashState *state, int hashbitlen);
-HashReturn Update(hashState *state, const BitSequence *data, DataLength databitlen);
+HashReturn Update256(hashState *state, const BitSequence *data, DataLength databitlen);
 HashReturn Final(hashState *state, BitSequence *hashval);
 HashReturn Hash(int hashbitlen, const BitSequence *data, DataLength databitlen, BitSequence *hashval);

### 2010-12-06 python-bindings/sha3lib/hash_functions/bmw/bmw.pyx

@@ -50,7 +50,7 @@ cdef class bmw:
             self.finished = 0
             self.state = self.previous_state
 
-        bmw_h.Update(&self.state, <bmw_h.BitSequence *> data, data_len)
+        bmw_h.Update256(&self.state, <bmw_h.BitSequence *> data, data_len)
 
     cpdef final(self):
         cdef bmw_h.BitSequence *hashval = <bmw_h.BitSequence *> malloc(self.hashbitlen*8)

### 2010-12-06 paper/main.tex
Added comment in Discussion
Several minor changes, and QA from start to 3.2

@@
 \label{sec:lengthextension}
-\paragraph{Security.} The most important criteria for the SHA-3
-candidates\cite{s_nistround2} is security. The full description of security
-criteria are mentioned in \cite{s_fedreg}, however the most noteworthy criterion
+\paragraph{Security.} The most important criterion for the SHA-3
+candidates\cite{s_nistround2} is security. \citet{s_fedreg} lays out the full
+description of security criteria, however the most noteworthy criterion
 in regards to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
@@
 
-\paragraph{Algorithm and implementation characteristics.} By this 3rd criteria,
+\paragraph{Algorithm and implementation characteristics.} By this 3rd criterion,
 \ac{NIST} emphasizes that algorithms with greater flexibility will be given
@@
 
-\subsubsection{\ac{NIST} \ac{SHA}-3 \ac{API}}
-\ac{NIST} required that every submission conform to a specified
-\ac{API}\cite{s_fedreg}. The \ac{API} is specified at \cite{s_nistapi} and in
-shortness it requires every candidate to implement four functions.
+\subsubsection{\ac{NIST} \ac{SHA}-3 \ac{API}} \ac{NIST} required that every
+submission should conform to a specified \ac{API}\cite{s_fedreg}. The \ac{API}
+specification \cite{s_nistapi} states in shortness that it requires every
+candidate to implement four functions:
 
@@
 called when all necessary data has been given through {\verb Update() }, it will then
-finish the hash function and provide the digest. The Hash() function basically a wrapper
-for doing a call to the three other functions.
+finish the hash function and provide the digest. The {\verb Hash() } function
+can be described as a wrapper for doing a call to the three other functions.
 
-{\verb HashReturn } is an \emph{enum} representing that an operation either succeeded (0) or
+{\verb HashReturn } is an \emph{enum} representing that an operation either succeeded or
 went wrong. The {\verb BitSequence } is a representation of an array of byte fields and
@@
 competition along with their principal submitters are listed in Table
-\ref{tbl:sha3:candidates}\cite{s_nistround2}.
+\ref{tbl:sha3:candidates} \cite{s_nistround2}.
 
@@
 \subsection{\ac{SUPERCOP}}
-% TODO: Mention that some SUPERCOP candidates are missing UPDATE-functionality.
 \ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
-of cryptographic software \cite{s_supercop}. In relation to hash functions
+of cryptographic software \cite{s_supercop}. In relation to hash functions,
 \ac{SUPERCOP} measures the following:
@@
 best results possible for each function.
-% TODO: Insert image to visualize the forest of different implementations.
 
@@
 candidates and \ac{SHA}-2 functions on a number of different plattforms and
-architectures are available at
-\footnote{\url{http://bench.cr.yp.to/results-sha3.html}}
-% TODO: Link to appendix with tahoe04 SUPERCOP results.
+architectures are available at the webpage of eBACS
+\footnote{\url{http://bench.cr.yp.to/results-sha3.html}}.
+%FIXME: Reference appendix if tahoe04 results end up there.
 
@@
 \ac{SUPERCOP} specifies an \ac{API} which all submitted implementations must
-conform to\cite{s_supercopapi}. This \ac{API} specifies the naming and
+conform to \cite{s_supercopapi}. This \ac{API} specifies the naming and
 organization of files, and that a submission must include a function
-{\verb crypto\_hash() } which is similar to the function {\verb Hash() } in the
+{\verb crypto_hash() } which is similar to the function {\verb Hash() } in the
 \ac{NIST} \ac{SHA}-3 \ac{API}. The only difference being the naming of
-data types and that the {\verb crypto\_hash() } function does not need an input
+data types and that the {\verb crypto_hash() } function does not need an input
 of how long the output of the hash function should be. Instead double or
@@
 
-As seen in figure \ref{fig:tahoehashing}, the usage of secure hashing is
+As seen in Figure \ref{fig:tahoehashing}, the usage of secure hashing is
 extensive in Tahoe, and has to be considered as a key part of the functionality,
@@
 which can be described as a tree with nodes that can verify all information
-below in the hierarchy, as depicted in figure \ref{fig:tahoemerkletree}. This
+below in the hierarchy, as depicted in Figure \ref{fig:tahoemerkletree}. This
 enables Tahoe to verify small segments of a file at the time, and this can be
@@
 perform the erasure decoding, and a hash of the plaintext and the encrypted
-text. This is illustrated in figure \ref{fig:tahoeueb}.
+text. This is illustrated in Figure \ref{fig:tahoeueb}.
 
@@
 
-Direct calling of external C functions and methods is also supported, thus
+Direct calling of external C functions and methods are also supported, thus
 making Cython an attractive way of wrapping external libraries written in C to
@@
 on the other hand, provides both, in combination with results stating which
-compiler arguments gives the best performance. On the downside, the
-\ac{SUPERCOP} \ac{API} only specifies the \texttt{crypto\_hash()} function and
-not an \texttt{update()} function which Tahoe-\ac{LAFS} is dependent on using.
+compiler arguments that gives the best performance. On the downside, the
+\ac{SUPERCOP} \ac{API} only specifies the {\verb crypto_hash() } function and
+not an {\verb update() } function which Tahoe-\ac{LAFS} is dependent on using.
 
@@
 following criteria:
-\begin{enumerate}
-  \item The implementation must have an \ac{API} defined in C.
-  \item The implementation must include a working \verb update() function.
-  \item The implementation must not use assembly that can not be made
-    position independent by \ac{GCC}.
-  \item The implementation should conform to the \acp{KAT} specified in the
+\paragraph{Criterion 1} The implementation must have an \ac{API} defined in C.
+\paragraph{Criterion 2} The implementation must include a working {\verb update() } function.
+\paragraph{Criterion 3} The implementation must not use assembly that can not be made
+    position independent by \ac{GCC}.%FIXME: State the reason for this somewhere?
+\paragraph{Criterion 4} The implementation should conform to the \acp{KAT} specified in the
     \ac{NIST} submission package, or later updates by the author.
-\end{enumerate}
 
-% TODO fix kilde
+\vspace{12pt}
 Based on these criteria, we choose the fastest possible implementation based
@@
 available at source.
+% TODO: Link to appendix with tahoe04 SUPERCOP results.
 
@@
 \input{tables/sha3_implementations}
+% FIXME: Insert image to visualize the forest of different implementations.
 
-The chosen implementations are shown in Table \ref{tbl:sha3:implementations}
-and their compiler flags used, which are extracted from the \ac{SUPERCOP}
-results can be seen in Table\ref{tbl:sha3:compilerflags}. It should be noted
-that 32 in blake32 does not refer to the output length of the function, but the
-fact that the algorithm is optimized for 32-bit usage. 512 in Keccackc512 also
-does not mean 512-bit output, but 512 cycles in the internal function. The
-relative speeds of the implementations chosen compared to the best version in
-\ac{SUPERCOP} and relative speed of the benchmark compared to other candidates
-can bee seen in Table \ref{tbl:sha3:speedrelative}. It should also be noted
-that the fastest sha256 implementation in \ac{SUPERCOP} is
+The chosen implementations are shown in Table \ref{tbl:sha3:implementations}.
+Their compiler flags used, which are extracted from the \ac{SUPERCOP}
+results, can be seen in Table \ref{tbl:sha3:compilerflags}.
+
+It should be noted that \emph{32} in blake32 does not refer to the output length
+of the function, but the fact that the algorithm is optimized for 32-bit usage.
+Similarly, \emph{512} in keccakc512 does not mean 512-bit output, but 512 cycles in the
+internal function.
+
+The relative speeds of the implementations chosen compared to
+the best version in \ac{SUPERCOP}, and relative speed of the benchmark compared
+to other candidates, can bee seen in Table \ref{tbl:sha3:speedrelative}.
+Worth mentioning is that the fastest SHA-256 implementation in \ac{SUPERCOP} is
 cryptopp\footnote{\url{http://www.cryptopp.com/}}, which is the one used by
@@
 \paragraph{JH} seems to be broken in every implementation with regard to
-criteria 2. For both the optimized versions found in \ac{SUPERCOP} and the
+Criterion 2. For both the optimized versions found in \ac{SUPERCOP} and the
 reference implementations in the \ac{NIST} submission package, Hash("foofoo")
@@
 but has not released any updated version yet.
-% TODO: Fix hvordan vi referer til e-post?
+% TODO: How to reference email?
+% TODO: We have to write someting about sphlib itself
 
@@
 
-\paragraph{Simd vect128 implementation} does not by default conform to
-criteria 4, because of a bug in the implementation. This has confirmed by the
-author who also supplied us with a patch to correct the bug.
+\paragraph{Simd vect128 implementation} does not by default conform to Criterion
+4, because of a bug in the implementation. This has been confirmed by the author
+who also supplied us with a patch to correct the bug.
 
-\paragraph{Blake32 \ac{SSSE3}} implementation also fail to wholly fulfill
-criteria 4, however it does validate all test vectors that have a length 
-in bits which is a multiple of 8. Since our Python wrappers only allow, and
-only need, to provide data in full bytes we decided to keep this
+\paragraph{Blake32 \ac{SSSE3}} implementation also fail to completely fulfill
+Criterion 4, however it does validate all test vectors that have a length
+in bits which is a multiple of 8. Since our Python wrappers only allow (and
+only need) to provide data in full bytes, we decided to keep this
 implementation.
 
-\paragraph{Shavite3} seems to have received an update\footnote{see
+\paragraph{Shavite3} seems to have received an update\footnote{See
 \url{http://www.cs.technion.ac.il/~orrd/SHAvite-3/}}. We have observed that
@@
 only because newer \ac{KAT}s will have to be generated, but since the author
-has not done so we chose to use an older version of the algorithm.
-
-% TODO: Add some lines that give reasons for the criteria above.
-
-% sphlib, SUPERCOP, NIST
+has not done so, we chose to use an older version of the algorithm.
 
@@
 
-\begin{itemize}
-    \item What could have been done differently?
-    \item Compare our results with general (supercop) results.
-    \item Optimized candidates - which gain could be expected from more
-    optimized candidates
-    \item What if we had used another coding scheme? - does Cython lower the
-    candidates performance?
-    \item Implications of only testing immutable files? (mutable files also
-    include directories)
-    \item Adding a capability to support new hash functions in tahoe. (Our
-    current way breaks backwards compatability)
-\end{itemize}
+% TODO: Mention that some SUPERCOP candidates are missing UPDATE-functionality.
+% What could have been done differently?
+% Compare our results with general (supercop) results.
+% Optimized candidates - which gain could be expected from more optimized
+%   candidates?
+% What if we had used another coding scheme? - does Cython lower the
+%   candidates performance?
+% Implications of only testing immutable files? (mutable files also
+%   include directories)
+% Adding a capability to support new hash functions in tahoe. (Our
+%   current way breaks backwards compatability)
+% Time gain of removing ``double'' SHA-256. Double only in digest()
 

### 2010-12-07 paper/main.tex
Added Unit Test descriptions

@@
 \subsection{Unit tests}
-% TODO: Include the use of Known Answer Tests
+We Unit Test each new candidate implementation that is included SHA3lib, both
+to verify our own work and to verify that the implementation from \ac{SUPERCOP}
+actually are correct.
+
+\paragraph{Known-Answer Tests} are used to check that the implementation conform
+to the specifications of the algorithm author. The \ac{KAT}s include
+bytestrings, their length and their expected message digest. The length of the
+inputs range from 0 bits to 34304 bits, and also include a message that should
+be repeated for about 17 million times, to verify that the implementation can handle
+large inputs.
+
+\paragraph{Basic Hash Tests} is what we have named the set of tests which test
+the core basics of what a hash function implementation will have to be able to
+do to be usable. This include that Init(), Update() and Final() works as
+expected, and that the output is of correct size. This include that every
+function should produce the same digest for the inputs Update("foofoo") and two
+consecutive calls to Update("foo").
+
+\paragraph{Copy Hash Tests} does not test any of the functionality used by
+Tahoe, but the copy()-function is a part of Hashlib. The idea is that if we
+have two bytestring that share a common prefix, we can input this prefix to
+Update(), then copy the state. The tests check if this is possible, and that a
+copy of the state and following update does not affect the original state.
 

### 2010-12-07 python-bindings/sha3lib/hash_functions/bmw/hash.c

@@ -1,8 +1,0 @@
-#include "crypto_hash.h"
-#include "BlueMidnightWish.h"
-     
-int crypto_hash(unsigned char *out,const unsigned char *in,unsigned long long inlen)
-{
-   if (Hash(crypto_hash_BYTES * 8,in,inlen * 8,out) == SUCCESS) return 0;
-   return -1;
-}

### 2010-12-07 paper/main.tex
Added disposition of results and discussion

@@
 \section{Python and Cython}
-%What is it, how, why, alternatives?
-% TODO: Finn en kilde på pythonting, om det trengs?
+
 Python\footnote{\url{http://www.python.org/}} is a high-level general-purpose
@@
 available at source.
-% TODO: Link to appendix with tahoe04 SUPERCOP results.
+% FIXME: Link to appendix with tahoe04 SUPERCOP results.
 
@@
 \input{tables/sha3_implementations}
-% FIXME: Insert image to visualize the forest of different implementations.
 
@@
 
-\begin{itemize}
-    \item Section should only contain the hard facts of our measurements
-    \item What is tested and what is not tested?
-    \item How much time is spent on hashing versus total time of
-    uploading/downloading
-    \item How much gain can Tahoe-LAFS get by switching to SHA-3?
-\end{itemize}
+% Munin?
+
+% How much time is spent on hashing versus total time of
+%   uploading/downloading
+
+In this chapter, we will focus on the concrete results derived from our testing.
+
+\section{Test scope}
+% What is tested and what is not tested?
+Immutable files are not tested, move text from Section 3.3.3.
+
+\section{Measuring Performance}
+Include code from hashutil.py and the use of time in bash.
 
 \section{\ac{SUPERCOP} benchmarks}
-The results of \ac{SHA}-3
-256-bit functions and \ac{SHA}-256 benchmarking in \ac{SUPERCOP} can be seen in
-tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
-\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
-\ref{tbl:supercop:8}. The lower the numbers, the faster the algorithms
-performed in the test.
+% TODO: more introduction text here.
+
+The results of \ac{SHA}-3 256-bit functions and \ac{SHA}-256 benchmarking in
+\ac{SUPERCOP} can be seen in tables \ref{tbl:supercop:long},
+\ref{tbl:supercop:4096}, \ref{tbl:supercop:1536}, \ref{tbl:supercop:576},
+\ref{tbl:supercop:64} and \ref{tbl:supercop:8}. The lower the numbers, the
+faster the algorithms performed in the test.
 
@@
 
+\section{Results}
+
+% Go graphs with plenty of short, descriptive comments.
+
 %**************************************%
@@
 
-% TODO: Mention that some SUPERCOP candidates are missing UPDATE-functionality.
-% What could have been done differently?
+% This chapter should contain (in listed and grouped order):
+% -----------------------------------------------------------
+% Describe inconsitencies in the numbers, and try to explain them.
+%
 % Compare our results with general (supercop) results.
@@
 %   candidates?
-% What if we had used another coding scheme? - does Cython lower the
-%   candidates performance?
+% Mention that some SUPERCOP candidates are missing UPDATE-functionality.
+%
+% How much gain can Tahoe-LAFS get by switching to SHA-3?
+% Time gain of removing ``double'' SHA-256. Double only in digest().
+%
 % Implications of only testing immutable files? (mutable files also
@@
 %   current way breaks backwards compatability)
-% Time gain of removing ``double'' SHA-256. Double only in digest()
+%
+% What if we had used another coding scheme? - does Cython lower the
+%   candidates performance?
+%
+% More general stuff. What could have been done differently?
 

### 2010-12-07 paper/main.tex
Added automatic testing and distribution

--- 2010-12-07 paper/main.tex
@@
 Cryptographic Operations and Primitives}
+\acro{SVN}{Subversion}
 \acro{UEB}{URI Extension Block}
+\acro{URI}{Uniform Resource Identifier}
 \end{acronym}
@@
 However the implementation in sphlib does fulfill this requirement. We have
-also contacted the author, Thomas Pornin, which confirms this problem with the
+contacted the author, Thomas Pornin, which confirms this problem with the
 official JH candidate and that the author has been made aware of the problem
-but has not released any updated version yet.
-% TODO: How to reference email?
-% TODO: We have to write someting about sphlib itself
+but has not released any updated version yet (personal communication, October
+16, 2010).
 
@@
 4, because of a bug in the implementation. This has been confirmed by the author
-who also supplied us with a patch to correct the bug.
+who also supplied us with a patch to correct the bug (Gaëtan Leurent, personal
+communication, November 10, 2010).
 
@@
 \paragraph{Copy Hash Tests} does not test any of the functionality used by
-Tahoe, but the copy()-function is a part of Hashlib. The idea is that if we
-have two bytestring that share a common prefix, we can input this prefix to
-Update(), then copy the state. The tests check if this is possible, and that a
-copy of the state and following update does not affect the original state.
+Tahoe-\ac{LAFS}, but the copy()-function is a part of Hashlib, and we have
+included it in SHA3lib to conform to this. The idea is that if we have two
+bytestring that share a common prefix, we can input this prefix to Update(),
+then copy the state. The tests check if this is possible, and that altering the
+copied state does not affect the original.
 
@@
 
-\section{Distribution of the Code}
-% TODO: Distutil, bash scripts
+\section{Automatic testing and distribution}
+We fully automate the testing procedure. This is achieved by the use of
+software revision control systems, which distributes both the source code of
+Tahoe-LAFS, SHA3lib and shell-scripts for installation and configuration of
+both. Configuration is mainly done by the use of regular expressions.
+
+The outline of this procedure is:
+
+\vspace{12pt}
+\noindent For every \ac{SHA}-3 candidate:
+        \begin{enumerate}
+            \item Install SHA3lib on all nodes
+            \item Replace used hash algorithm in Tahoe-LAFS with current
+            candidate
+            \item Install Tahoe-LAFS
+            \item Create and start one Tahoe-LAFS introducer
+            \item Create and start Tahoe-LAFS storage nodes on 3 nodes
+            \item Create a Tahoe-LAFS client on a previously unnused node
+            \item Do testing
+            \item clean up
+        \end{enumerate}
+
+\paragraph{Distribution of source code for Tahoe-LAFS and SHA3lib} is done by
+\ac{SVN} and GIT respectivly. The change of hash algorithm in Tahoe-LAFS is
+however done with the use of a regular expression, because of both the
+simplicity of the change, and to not clutter the revision control system. Both
+Tahoe-LAFS and SHA3lib uses \ac{Distutils}, which includes options to install
+them globally on a system.
 

### 2010-12-07 paper/main.tex
Started on results chapter. Missing the acutal graphs.

@@
 \section{Modifications to the Tahoe-\ac{LAFS} Code}
+\label{sec:modtahoe}
 
@@
 graphs of the storage nodes during upload and retrieval, where neither the
-\ac{CPU}, network, disk or memory usage runs out of capacity. The \ac{CPU} usage
-during testing on one of the storage nodes is depicted in Figure
+\ac{CPU}, network, disk or memory usage raise towards maximum threshold. The
+\ac{CPU} usage during testing on one of the storage nodes is depicted in Figure
 \ref{fig:munin:storagenode}.
@@
 
-% Maybe cite
-% \url{http://tahoe-lafs.org/trac/tahoe-lafs/wiki/Performance?version=32}
-
-% Munin?
+% FIXME: Decide whether Munin should be included?
 
@@
 In this chapter, we will focus on the concrete results derived from our testing.
+The method of testing is described, along with the test scope.
+
+Further on, we exhibit the results from the \ac{SUPERCOP} benchmark, before
+presenting the graphs of how much time the hashing operations take when
+downloading and uploading a number of different test vectors, and show these
+data in relation to other measurements, such as total operation time.
 
 \section{Test scope}
-% What is tested and what is not tested?
-Immutable files are not tested, move text from Section 3.3.3.
+
+As stated in Section \ref{sec:modtahoe}, only immutable files are tested. The
+mutable files are not tested because of our way of integrating the SHA-3
+candidates within the Tahoe-LAFS code.
+
+To measure the performance impact of the different implementations, we time the
+hash function calls from Tahoe, in addition to timing the total operation time,
+e.g. the total time it takes for a 1GB file to be uploaded to grid.
 
 \section{Measuring Performance}
-Include code from hashutil.py and the use of time in bash.
+
+Listing \ref{lst:hashutil_time} demonstrates the use of the {\verb time } module
+in Python. On every call to {\verb update() } or {\verb digest() }, a variable
+is increased with the time it took to run the underlying hash library
+functionality. Similarly, Listing \ref{lst:testrunner} shows the use of the
+{\verb date } command to calculate the number of seconds it takes it takes for
+Tahoe to put a file to the grid.
+
+\lstinputlisting[label=lst:hashutil_time, caption=Timing in Python (extraction from hashutil.py)]
+{listings/hashutil_timing.py}
+
+\lstinputlisting[language=sh, label=lst:testrunner, caption=Timing in Bash from test runner]
+{listings/test_runner.sh}
 
 \section{\ac{SUPERCOP} benchmarks}
-% TODO: more introduction text here.
+% TODO: more introduction text here?
 
-The results of \ac{SHA}-3 256-bit functions and \ac{SHA}-256 benchmarking in
+The results of the \ac{SHA}-3 256-bit functions and \ac{SHA}-256 benchmarking in
 \ac{SUPERCOP} can be seen in tables \ref{tbl:supercop:long},
@@
 \ref{tbl:supercop:64} and \ref{tbl:supercop:8}. The lower the numbers, the
-faster the algorithms performed in the test.
+faster and better the algorithms performed in the test.
 
@@
 %
-% More general stuff. What could have been done differently?
+% More general stuff. What could have been done differently? Critisize
+% ourselves.
 

### 2010-12-07 paper/listings/hashutil_timing.py

@@ -1,0 +1,20 @@
+import time
+
+class _SHA256d_Hasher:
+    def update(self, data):
+        t1 = time.time()
+        self.h.update(data)
+        t2 = time.time()
+
+        self.time_spent_hashing += t2 - t1
+
+    def digest(self):
+        if self._digest is None:
+            t1 = time.time()
+            h1 = self.h.digest()
+            t2 = time.time()
+
+            self.time_spent_hashing += t2 - t1
+
+            _write_increase(HASH_TIMING_FILE,
+                            self.time_spent_hashing)

### 2010-12-07 paper/listings/test_runner.sh

@@ -1,0 +1,6 @@
+time1=`date +%s.%N`
+for file in $files; do
+    tahoe cp $file tahoe: > /dev/null
+done
+time2=`date +%s.%N`
+echo $time2 - $time1 | bc > ./logs/$ftype\_totaltime.txt

### 2010-12-07 paper/main.tex
Estimate of double SHA-256 in supercop

@@
 
+\section{\ac{SUPERCOP} results}
+We see in tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
+\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
+\ref{tbl:supercop:8} that in general BLAKE, \ac{BMW}, Luffa, Shabal and Simd
+outperform \ac{SHA}-256. However, given that Tahoe-LAFS uses double
+\ac{SHA}-256, this means that it will also have to hash another 32 bytes for
+every message it hashes. Since \ac{SUPERCOP} does not have benchmark for 32
+byte messages, we will have to estimate the amount of cycles required for this.
+If we assume that cycles/bytes is linear from 8 byte to 64 byte, we get the
+following estimate for the speed of hashing 32 bytes:
+
+$$179.62 - \frac{179.62 - 38.12}{64 - 8} * (32-8)= 118.98 (Cycles/byte)$$
+
+We can then estimate the result of each message by the following formula:
+$$\frac{median * size + 118.98*32}{size} = SHA256d estimate$$
+
+\input{tables/supercop_sha256d}
+Estimates based on this calculation are shown in Table
+\ref{tbl:supercop:sha256d}. What we can see is that for lower lengths of data,
+more of the candidates are faster than double \ac{SHA}-256. However, given the
+practical situation that Tahoe-\ac{LAFS} is a filesystem, most files processed
+will have be larger than the file sizes estimated here. Even though not all
+hashing operations done in Tahoe are related to the file size, we expect that
+only BLAKE, \ac{BMW}, Luffa, Shabal and SIMD will be able to outperform double
+\ac{SHA}-256.
+
 % This chapter should contain (in listed and grouped order):

### 2010-12-07 paper/tables/supercop_sha256d.tex
Estimate of double SHA-256 in supercop

\begin{table}
  \centering
  \begin{tabular}{ | c | c | }
    \hline
    \textbf{Size} & \textbf{Median} \\ \hline
    8    & 655.54 \\ \hline
    64   & 97.61 \\ \hline
    576  & 24.66 \\ \hline
    1536 & 18.93 \\ \hline
    4096 & 16.76 \\ \hline
  \end{tabular}
  \caption{Estimates for double SHA-256}
  \label{tbl:supercop:sha256d}
\end{table}

### 2010-12-07 python-bindings/setup.py
Changed BMW version to sphlib

@@ -14,9 +14,10 @@ srcpath = 'sha3lib/'
 hashpath = srcpath + 'hash_functions/'
 
 ext_modules = [
-    Extension("sha3lib.hash_functions.bmw.bmw",
-        [hashpath+"bmw/bmw.pyx", hashpath+'bmw/BlueMidnightWish.c'],
-        extra_compile_args=['-m32','-march=i386','-O','-fomit-frame-pointer']),
+    Extension("sha3lib.hash_functions.bmw.bmw_hash",
+        [hashpath+"bmw/bmw_hash.pyx", hashpath+'bmw/sha3nist.c',
+            hashpath+"bmw/bmw.c"],
+        extra_compile_args=['-march=k6-3','-O3','-fomit-frame-pointer']),
     Extension("sha3lib.hash_functions.groestl.groestl_hash",
         [hashpath+"groestl/groestl_hash.pyx", hashpath+'groestl/sha3nist.c',
             hashpath+"groestl/groestl.c"],
@@ -126,7 +127,7 @@ class RealClean(Command):
         # Cython generated files, not removing them before rebuilding (after
         # change) can cause unexpected results
         manual_files = [
-                './sha3lib/hash_functions/bmw/bmw.c',
+                './sha3lib/hash_functions/bmw/bmw_hash.c',
                 './sha3lib/hash_functions/luffa/luffa.c',
                 './sha3lib/hash_functions/groestl/groestl_hash.c',
                 './sha3lib/hash_functions/skein/skein_hash.c',

### 2010-12-07 python-bindings/sha3lib/hash_functions/__init__.py

@@ -1,4 +1,4 @@
-from bmw.bmw import bmw
+from bmw.bmw_hash import bmw
 from groestl.groestl_hash import groestl
 from echo.echo_hash import echo
 from blake.blake import blake
@@ -13,7 +13,7 @@ from shabal.shabal_hash import shabal
 from cubehash.cubehash_hash import cubehash
 from hamsi.hamsi_hash import hamsi
 
-from bmw.bmw import hash as bmw256hash
+from bmw.bmw_hash import hash as bmw256hash
 from groestl.groestl_hash import hash as groestl256hash
 from echo.echo_hash import hash as echo256hash
 from blake.blake import hash as blake256hash

### 2010-12-07 python-bindings/sha3lib/hash_functions/bmw/BlueMidnightWish.c

@@ -1,927 +1,0 @@
-/* 
-   Optimized C code for tweaked Blue Midnight Wish - 256 programmed by Rune E. Jensen, December 2009.
-*/
-
-#include <stdio.h>
-#include <string.h> 
-#include "BlueMidnightWish.h"
-
-#define rotl32(x,n)   (((x) << n) | ((x) >> (32 - n)))
-#define rotr32(x,n)   (((x) >> n) | ((x) << (32 - n)))
-
-#define shl(x,n)      ((x) << n)
-#define shr(x,n)      ((x) >> n)
-
-/* BlueMidnightWish256 initial double chaining pipe */
-const u_int32_t i256p2[16] =
-{   0x40414243ul, 0x44454647ul, 0x48494a4bul, 0x4c4d4e4ful,
-    0x50515253ul, 0x54555657ul, 0x58595a5bul, 0x5c5d5e5ful,
-    0x60616263ul, 0x64656667ul, 0x68696a6bul, 0x6c6d6e6ful,
-    0x70717273ul, 0x74757677ul, 0x78797a7bul, 0x7c7d7e7ful,
-};
-
-#define hashState256(x)  ((x)->pipe->p256)
-
-/* Components used for 224 and 256 bit version */
-//#define s32_0(x)  (shr((x), 1) ^ shl((x), 3) ^ rotl32((x),  4) ^ rotl32((x), 19))
-//#define s32_1(x)  (shr((x), 1) ^ shl((x), 2) ^ rotl32((x),  8) ^ rotl32((x), 23))
-//#define s32_2(x)  (shr((x), 2) ^ shl((x), 1) ^ rotl32((x), 12) ^ rotl32((x), 25))
-//#define s32_3(x)  (shr((x), 2) ^ shl((x), 2) ^ rotl32((x), 15) ^ rotl32((x), 29))
-#define s32_4(x)  (shr((x), 1) ^ (x))
-#define s32_5(x)  (shr((x), 2) ^ (x))
-#define r32_01(x) rotl32((x),  3)
-#define r32_02(x) rotl32((x),  7)
-#define r32_03(x) rotl32((x), 13)
-#define r32_04(x) rotl32((x), 16)
-#define r32_05(x) rotl32((x), 19)
-#define r32_06(x) rotl32((x), 23)
-#define r32_07(x) rotl32((x), 27)
-
-
-//#define s32_0(x)  (shr((x), 1) ^ shl((x), 3) ^ rotr32((rotl32((x), 19)), 15) ^ rotl32((x), 19))
[…]

### 2010-12-07 python-bindings/sha3lib/hash_functions/bmw/BlueMidnightWish.h

@@ -1,85 +1,0 @@
-// We use this type definition to ensure that 
-// "unsigned long" on 32-bit and 64-bit little-endian 
-// operating systems are 4 bytes long.
-#if defined( _MSC_VER )
-typedef unsigned long u_int32_t;
-typedef unsigned long long u_int64_t;
-#else
-#include <sys/types.h>
-#endif
-
-// General SHA-3 definitions
-typedef unsigned char BitSequence;
-typedef u_int64_t DataLength; // a typical 64-bit value
-typedef enum { SUCCESS = 0, FAIL = 1, BAD_HASHLEN = 2, BAD_CONSECUTIVE_CALL_TO_UPDATE = 3 } HashReturn;
-// Blue Midnight Wish allows to call Update() function consecutively only if the total length of stored 
-// unprocessed data and the new supplied data is less than or equal to the BLOCK_SIZE on which the 
-// compression functions operates. Otherwise BAD_CONSECUTIVE_CALL_TO_UPDATE is returned.
-
-
-// Specific algorithm definitions
-#define BlueMidnightWish224_DIGEST_SIZE  28
-#define BlueMidnightWish224_BLOCK_SIZE   64
-#define BlueMidnightWish256_DIGEST_SIZE  32
-#define BlueMidnightWish256_BLOCK_SIZE   64
-#define BlueMidnightWish384_DIGEST_SIZE  48
-#define BlueMidnightWish384_BLOCK_SIZE  128
-#define BlueMidnightWish512_DIGEST_SIZE  64
-#define BlueMidnightWish512_BLOCK_SIZE  128
-
-
-// Here we define the default Blue Midnight Wish tunable security parameters.
-// The parameters are named EXPAND_1_ROUNDS and EXPAND_2_ROUNDS.
-// Since Blue Midnight Wish has 16 rounds in its message expansion part, the 
-// following relation for these parameters should be satisfied:
-//
-//          EXPAND_1_ROUNDS + EXPAND_2_ROUNDS = 16
-//
-// Blue Midnight Wish in its message expansion part uses 2 different functions: 
-// expand_1 and expand_2.
-//
-// expand_1 is the more complex and more time consuming, but offers the fastest
-// diffusion of bit differences and produces variables that have the most complex
-// nonlinear relations with previous 16 variables in the message expansion part.
-//
[…]

### 2010-12-07 python-bindings/sha3lib/hash_functions/bmw/api.h

@@ -1,2 +1,0 @@
-#define CRYPTO_BYTES 32
-#define CRYPTO_VERSION "Optimized C - BMW256, opt01"

### 2010-12-07 python-bindings/sha3lib/hash_functions/bmw/bmw_hash.pyx

@@ -1,5 +1,5 @@
 # encoding: utf-8
-cimport bmw_h
+cimport bmw_hash_h
 cimport cython
 
 from libc.stdlib cimport *
@@ -9,10 +9,10 @@ def hash(int hashbitlen, bytes data, int in_length):
     Used for testing purposes.
     '''
     cdef char* c_string = <char *> data
-    cdef bmw_h.DataLength datalen = in_length
-    cdef bmw_h.BitSequence *hashval = <bmw_h.BitSequence *> malloc(hashbitlen*8)
+    cdef bmw_hash_h.DataLength datalen = in_length
+    cdef bmw_hash_h.BitSequence *hashval = <bmw_hash_h.BitSequence *> malloc(hashbitlen*8)
 
-    bmw_h.Hash(hashbitlen, <bmw_h.BitSequence *> c_string, datalen, hashval)
+    bmw_hash_h.Hash(hashbitlen, <bmw_hash_h.BitSequence *> c_string, datalen, hashval)
 
     try:
         digest = [hashval[i] for i from 0 <= i < hashbitlen / 8]
@@ -28,9 +28,9 @@ cdef class bmw:
     so that one can update the hashing procedure instead of doing it from
     scratch.
     '''
-    #cpdef bmw_h.BitSequence *hashval
-    cdef bmw_h.hashState previous_state
-    cdef bmw_h.hashState state
+    #cdef bmw_hash_h.BitSequence *hashval
+    cdef bmw_hash_h.hashState previous_state
+    cdef bmw_hash_h.hashState state
     cdef int finished
     cdef int hashbitlen
     cdef list hashval
@@ -38,7 +38,8 @@ cdef class bmw:
     def __init__(self, int in_hashbitlen, bytes initial=None):
         self.finished = 0
         self.hashbitlen = in_hashbitlen
-        bmw_h.Init(&self.state, self.hashbitlen)     
+        bmw_hash_h.Init(&self.state, self.hashbitlen)
+
         if initial:
             self.update(initial)
 
@@ -50,18 +51,18 @@ cdef class bmw:
             self.finished = 0
             self.state = self.previous_state
 
-        bmw_h.Update256(&self.state, <bmw_h.BitSequence *> data, data_len)
+        bmw_hash_h.Update(&self.state, <bmw_hash_h.BitSequence *> data, data_len)
 
     cpdef final(self):
-        cdef bmw_h.BitSequence *hashval = <bmw_h.BitSequence *> malloc(self.hashbitlen*8)
[…]

### 2010-12-07 python-bindings/sha3lib/hash_functions/bmw/sph_sha3.h

@@ -1,0 +1,61 @@
+/* $Id: sph_sha3.h 208 2010-06-02 20:33:00Z tp $ */
+/**
+ * SHA-3 is the colloquial codename for the future standard hash function
+ * from NIST. SHA-3 will be "API compatible" (same range of output sizes)
+ * with the SHA-2 family (SHA-224, SHA-256, SHA-384 and SHA-512). This
+ * library implements a number of SHA-3 candidates (functions submitted
+ * to the NIST competition for selection of SHA-3). This header file
+ * includes the header files for SHA-2 and all the implemented SHA-3
+ * candidates.
+ *
+ * ==========================(LICENSE BEGIN)============================
+ *
+ * Copyright (c) 2007-2010  Projet RNRT SAPHIR
+ * 
+ * Permission is hereby granted, free of charge, to any person obtaining
+ * a copy of this software and associated documentation files (the
+ * "Software"), to deal in the Software without restriction, including
+ * without limitation the rights to use, copy, modify, merge, publish,
+ * distribute, sublicense, and/or sell copies of the Software, and to
+ * permit persons to whom the Software is furnished to do so, subject to
+ * the following conditions:
+ * 
+ * The above copyright notice and this permission notice shall be
+ * included in all copies or substantial portions of the Software.
+ * 
+ * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
+ * EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
+ * MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT.
+ * IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY
+ * CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT,
+ * TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE
+ * SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
+ *
+ * ===========================(LICENSE END)=============================
+ *
+ * @file     sph_sha3.h
+ * @author   Thomas Pornin <thomas.pornin@cryptolog.com>
[…]

### 2010-12-07 paper/main.tex
Added Supercop discussion

@@
 
+\section{Validity of \ac{SUPERCOP} results}
+Although we believe the \ac{SUPERCOP} toolkit and results give good indication
+of how fast the different candidate implementations are, we do have som
+comments on it.
+
+\paragraph{Implementations does not conform to \ac{NIST} specification} for
+many of the candidates. By this we mean that there is no Update()-function
+available for all implementation, and that \ac{SUPERCOP} only measures speed of
+hashing full bytestrings. This is also a hasle for programmer who want to use
+the implementations in \ac{SUPERCOP}
+
+\paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
+mean implementations that does not verify the \ac{KAT}s.
+
+\parapgraph{Results are only valid for one specific computer}. The results we
+used from \ac{SUPERCOP} are only valid for our specific computer, although it
+gives an indication of which speeds can be achieved on similar computers, i.e.
+32-bit. However to get a complete picture of the performance of the different
+candidates one must look at different computers and architectures. This picture
+might be obtained from looking at the results from all computeres measured by
+\ac{SUPERCOP}.
+
 % This chapter should contain (in listed and grouped order):

### 2010-12-09 paper/main.tex
Fixed typo
Added graphs and started on results table

@@
 
-\section{Results}
-
-% Go graphs with plenty of short, descriptive comments.
+\section{Measurements of Tahoe operations}
+%\vskip -1.5em % This is to fit in two graphs on one page including heading
+
+% How much time is spent on hashing versus total time of
+%   uploading/downloading
+
+The sum of the times it took for the underlying hash library to calculate the
+hash operations are exhibited in Figures \ref{fig:graph:10001b},
+\ref{fig:graph:1001kb}, \ref{fig:graph:501mb}, \ref{fig:graph:5100mb} and
+\ref{fig:graph:11gb}. We call this \emph{time spent hashing}.
+
+The numbers for uploading and downloading of one 1GB file are summarized in
+Tables \ref{tbl:hashingtimes:put1gb} and \ref{tbl:hashingtimes:get1gb}.
+
+\input{tables/hashingtimes}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingPUTof10001Bfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingGETof10001Bfiles.pdf}
+    \caption{Figures for hash operations of 1000 1B files.}
+    \label{fig:graph:10001b}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingPUTof1001KBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingGETof1001KBfiles.pdf}
+    \caption{Figures for hash operations of 100 1KB files.}
+    \label{fig:graph:1001kb}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingPUTof501MBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingGETof501MBfiles.pdf}
+    \caption{Figures for hash operations of 50 1MB files.}
+    \label{fig:graph:501mb}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingPUTof5100MBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingGETof5100MBfiles.pdf}
+    \caption{Figures for hash operations of five 100MB files.}
+    \label{fig:graph:5100mb}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingPUTof11GBfile.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingGETof11GBfile.pdf}
+    \caption{Figures for hash operations of one 1GB file.}
+    \label{fig:graph:11gb}
+\end{figure}
 
@@
 
-\parapgraph{Results are only valid for one specific computer}. The results we
+\paragraph{Results are only valid for one specific computer}. The results we
 used from \ac{SUPERCOP} are only valid for our specific computer, although it

### 2010-12-09 paper/main.tex
Minor spell corrections

@@
 
-% How much time is spent on hashing versus total time of
-%   uploading/downloading
-
 In this chapter, we will focus on the concrete results derived from our testing.
@@
 more of the candidates are faster than double \ac{SHA}-256. However, given the
-practical situation that Tahoe-\ac{LAFS} is a filesystem, most files processed
+practical situation that Tahoe-\ac{LAFS} is a file system, most files processed
 will have be larger than the file sizes estimated here. Even though not all
@@
 Although we believe the \ac{SUPERCOP} toolkit and results give good indication
-of how fast the different candidate implementations are, we do have som
+of how fast the different candidate implementations are, we do have some
 comments on it.
@@
 available for all implementation, and that \ac{SUPERCOP} only measures speed of
-hashing full bytestrings. This is also a hasle for programmer who want to use
-the implementations in \ac{SUPERCOP}
+hashing full byte strings. This is also a hassle for a programmer who want to
+use the implementations in \ac{SUPERCOP}
 
@@
 candidates one must look at different computers and architectures. This picture
-might be obtained from looking at the results from all computeres measured by
+might be obtained from looking at the results from all computers measured by
 \ac{SUPERCOP}.
@@
 % -----------------------------------------------------------
-% Describe inconsitencies in the numbers, and try to explain them.
+% Describe inconsistencies in the numbers, and try to explain them.
+% Time spent not hashing varies because: Variable use of CPU regs?
 %
-% Compare our results with general (supercop) results.
+% Compare our results with general (SUPERCOP) results.
 % Optimized candidates - which gain could be expected from more optimized
@@
 % Adding a capability to support new hash functions in tahoe. (Our
-%   current way breaks backwards compatability)
+%   current way breaks backwards compability)
 %
@@
 %
-% More general stuff. What could have been done differently? Critisize
+% More general stuff. What could have been done differently? Criticize
 % ourselves.

### 2010-12-09 utils/graph_and_convert.sh
Updated scripts to generate graphs. New color and script to convert to PDF

@@ -1,0 +1,13 @@
+#!/bin/bash
+if [ -z "$1" ]; then 
+    echo usage: $0 directory
+    exit
+fi
+
+python results_to_graphs.py -d $1
+cd $1
+
+for fn in `ls *.svg`; do
+    echo "Converting $fn to PDF."
+    inkscape -D -z --file=$fn --export-pdf=`echo $fn | sed s/svg//`pdf
+done

### 2010-12-09 utils/graph_matplotlib.py

@@ -17,7 +17,7 @@ def graph_test_vector(candidates, data, directory, title, axis):
     plt.yticks(ind + width*1.5, candidates)
 
     plt.xlabel(axis)
-    plt.barh(ind+width, data, height=width)
+    plt.barh(ind+width, data, height=width, color='#3355cc')
 
     # Some tweaking to make it look alright.
     ymin, ymax = plt.ylim()
@@ -28,11 +28,9 @@ def graph_test_vector(candidates, data, directory, title, axis):
     for i in ind:
         plt.text(plt.xlim()[1]*1.01, ind[i] + width*1.25, '%.2f' % data[i])
 
-    filename = os.path.join(directory, '%s.svg' % title.replace(' ',''))
-    print filename, ':'
-    print candidates
-    print data
-    plt.savefig(filename)
+    filename = '%s.svg' % title.replace(' ','')
+    filepath = os.path.join(directory, filename)
+    plt.savefig(filepath)
 
     #Clear the graphs, as it holds state
     plt.cla()

### 2010-12-09 utils/results_to_graphs.py

@@ -81,26 +81,29 @@ def get_results_from_data(candidates, directory):
 def graph_data(results, candidates, directory):
     for vector in VECTORS:
         for op in ('get', 'put'):
+            plural = ''
+            if VECTOR_COUNT[vector] > 1:
+                plural = 's'
             y_axis = 'Seconds'
             tsh_data = [results[vector][c][op][0] for c in candidates]
-            title = 'Time spent hashing: %s of %s %s file(s)' % (
-                op.upper(), str(VECTOR_COUNT[vector]), vector.upper())
+            title = 'Time spent hashing %s of %s %s file%s' % (
+                op.upper(), str(VECTOR_COUNT[vector]), vector.upper(), plural)
             graph_test_vector(candidates, tsh_data, directory, title, y_axis)
 
             tts_data = [results[vector][c][op][1] for c in candidates]
-            title = 'Total time spent: %s of %s %s file(s)' % (
-                op.upper(), str(VECTOR_COUNT[vector]), vector.upper())
+            title = 'Total time spent %s of %s %s file%s' % (
+                op.upper(), str(VECTOR_COUNT[vector]), vector.upper(), plural)
             graph_test_vector(candidates, tts_data, directory, title, y_axis)
 
             data = map(lambda x, y: float(y) - float(x), tsh_data, tts_data)
-            title = 'Time spent NOT hashing: %s of %s %s file(s)' % (
-                op.upper(), str(VECTOR_COUNT[vector]), vector.upper())
+            title = 'Time spent NOT hashing %s of %s %s file%s' % (
+                op.upper(), str(VECTOR_COUNT[vector]), vector.upper(), plural)
             graph_test_vector(candidates, data, directory, title, y_axis)
 
             y_axis = 'Count'
             data = [results[vector][c][op][2] for c in candidates]
-            title = 'Hash operations: %s of %s %s file(s)' % (
-                op.upper(), str(VECTOR_COUNT[vector]), vector.upper())
+            title = 'Hash operations %s of %s %s file%s' % (
+                op.upper(), str(VECTOR_COUNT[vector]), vector.upper(), plural)
             graph_test_vector(candidates, data, directory, title, y_axis)
 
 def results_to_graphs(directory=None):

### 2010-12-11 paper/main.tex
Added section on immutable file exclusion in Discussion chapter

@@
 impossible, since the internal verification fails. For this reason, \emph{only
-immutable files} are tested and measured in this paper. Implications of this is
+immutable files} are tested and measured in this paper. Implications of this are
 discussed further in Chapter \ref{ch:discussion}.
@@
             \item Do testing
-            \item clean up
+            \item Clean up
         \end{enumerate}
@@
 \section{\ac{SUPERCOP} results}
-We see in tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
+% Time gain of removing ``double'' SHA-256. Double only in digest().
+
+We see in Tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
 \ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
@@
 \ac{SHA}-256, this means that it will also have to hash another 32 bytes for
-every message it hashes. Since \ac{SUPERCOP} does not have benchmark for 32
-byte messages, we will have to estimate the amount of cycles required for this.
-If we assume that cycles/bytes is linear from 8 byte to 64 byte, we get the
-following estimate for the speed of hashing 32 bytes:
+every message it returns the digest for. Since \ac{SUPERCOP} does not have
+benchmark for 32 byte messages, we will have to estimate the amount of cycles
+required for this.  If we assume that cycles/bytes is linear from 8 byte to 64
+byte, we get the following estimate for the speed of hashing 32 bytes:
 
@@
 We can then estimate the result of each message by the following formula:
-$$\frac{median * size + 118.98*32}{size} = SHA256d estimate$$
+$$\frac{median * size + 118.98*32}{size} = SHA256d$$ \emph{estimate}
 
 \input{tables/supercop_sha256d}
+
 Estimates based on this calculation are shown in Table
-\ref{tbl:supercop:sha256d}. What we can see is that for lower lengths of data,
+\ref{tbl:supercop:sha256d}. What we can see that for lower lengths of data,
 more of the candidates are faster than double \ac{SHA}-256. However, given the
 practical situation that Tahoe-\ac{LAFS} is a file system, most files processed
-will have be larger than the file sizes estimated here. Even though not all
+will be larger than the file sizes estimated here. Even though not all
 hashing operations done in Tahoe are related to the file size, we expect that
@@
 \section{Validity of \ac{SUPERCOP} results}
+% Mention that some SUPERCOP candidates are missing UPDATE-functionality.
+
 Although we believe the \ac{SUPERCOP} toolkit and results give good indication
@@
 \paragraph{Implementations does not conform to \ac{NIST} specification} for
-many of the candidates. By this we mean that there is no Update()-function
-available for all implementation, and that \ac{SUPERCOP} only measures speed of
-hashing full byte strings. This is also a hassle for a programmer who want to
-use the implementations in \ac{SUPERCOP}
+many of the candidates. By this we mean that there is no {\verb Update() }
+function available for all implementations, and that \ac{SUPERCOP} only
+measures speed of hashing full byte strings. This is also a hassle for a
+programmer who wants to use the implementations in \ac{SUPERCOP}
 
 \paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
-mean implementations that does not verify the \ac{KAT}s.
+mean implementations that does not conform to the \ac{KAT}s supplied.
 
@@
 \ac{SUPERCOP}.
+
+\section{Exclusion of mutable file testing}
+% Implications of only testing immutable files? (mutable files also
+%   include directories)
+% Adding a capability to support new hash functions in tahoe. (Our
+%   current way breaks backwards compability)
+
+Since our minor change in the Tahoe-\ac{LAFS} code broke internal checks, we
+were not able to test mutable files, and hence not directories.
+
+We argue that this has minor implications for our testing results. Both because
+mutable files are not applicable for most use cases, since deleting an old
+version and uploading a new file gives the same functionality, and because
+immutable files uses less logic and therefore are more reliable and faster in
+some cases. One should bear in mind that Tahoe-\ac{LAFS} is not a revision
+control system, and mutable files are not intended for this purpose.
+
+To be able to extend the Tahoe-LAFS with new functionality to fully support new
+hash functions, one have to add a new capability (Zooko O'Whielacronx, personal
+communication, September 29, 2010). This ensures backwards compatibility and can
+be an optional feature for users to employ. However, we found that adding a new
+capability was out of scope for this assignment.
 
@@
 %   candidates?
-% Mention that some SUPERCOP candidates are missing UPDATE-functionality.
 %
 % How much gain can Tahoe-LAFS get by switching to SHA-3?
-% Time gain of removing ``double'' SHA-256. Double only in digest().
-%
-% Implications of only testing immutable files? (mutable files also
-%   include directories)
-% Adding a capability to support new hash functions in tahoe. (Our
-%   current way breaks backwards compability)
 %
@@
 
-A somewhat short conclusion summing up what has been done, and results and
-discussion
-
-New SHA-3 capability in Tahoe-LAFS (GSOC ref) to get our code to trunk
+% This chapter should contain (in listed and grouped order):
+% -----------------------------------------------------------
+% A somewhat short conclusion summing up what has been done, and results and
+% discussion
+%
+% We had hoped to have contributed more to the Open Source project that
+% Tahoe-LAFS is.
 

### 2010-12-11 paper/main.tex
Cython discussion

--- 2010-12-09 paper/main.tex
@@
 %
+\section{The use of Cython}
+The python-bindings generated for Cython in C are not necessarily the most
+optimized versions possible to write and they are ugly to read or change (from
+a C perspective). However we consider them to be good enough, and they are
+equal for all the \ac{SHA}-3 candidates which means they should give the same
+performance penalty. The \ac{SHA}-256 version will of course not suffer from
+the same penalty, however we believe the possible performance gain is small.
+This is due to the fact that no intensive operations take place in our
+Cython-code.
+
+The alternative would be creating manual Python-bindings for all the candidates
+which would require a lot more work on our side, especially given the lack of a
+good enough and common \ac{API} for the implementations in \ac{SUPERCOP}.
+Another fact important for this decision was that we did not know exactly which
+implementation of each candidate we would end up using, and we ended up trying
+several for most candidates.
+
 % More general stuff. What could have been done differently? Criticize

### 2010-12-11 paper/main.tex
Changed supercop double sha256 to worst-case scenario

@@
 byte messages, we will have to estimate the amount of cycles required for this.
-If we assume that cycles/bytes is linear from 8 byte to 64 byte, we get the
-following estimate for the speed of hashing 32 bytes:
-
-$$179.62 - \frac{179.62 - 38.12}{64 - 8} * (32-8)= 118.98 (Cycles/byte)$$
-
-We can then estimate the result of each message by the following formula:
-$$\frac{median * size + 118.98*32}{size} = SHA256d estimate$$
+If we assume a worst case scenario of hashing 32 bytes, where the cycles/byte
+count is exactly the same as for 8 bit median we get:
+
+$$\frac{median * size + 179.62*32}{size} = SHA256d estimate$$
 
@@
 practical situation that Tahoe-\ac{LAFS} is a file system, most files processed
-will have be larger than the file sizes estimated here. Even though not all
-hashing operations done in Tahoe are related to the file size, we expect that
-only BLAKE, \ac{BMW}, Luffa, Shabal and SIMD will be able to outperform double
-\ac{SHA}-256.
+will be larger than the file sizes estimated here. Not all hashing
+operations in Tahoe-\ac{LAFS} are strictly dependant on the original file size,
+which would mean that all candidates would have a chance of outperforming
+double \ac{SHA}-256. However this estimate is worst case, which leads us to the
+belief that at least for larger input sizes (over 1536) only the five fastest
+candidates will have a change of outperforming double \ac{SHA}-256.
 

### 2010-12-11 paper/listings/setup.py
Fixed line break

@@ -22,7 +22,7 @@ ext_modules = [
 setup(
     name='SHA3lib',
     version='0.1',
-    description='Python bindings for SHA3-256 using Cython',
+    description='Python bindings for SHA3-256 in Cython',
     cmdclass = {'build_ext': build_ext},
     ext_modules = ext_modules,
     packages =

### 2010-12-11 paper/main.tex
Started concluding

--- 2010-12-11 paper/main.tex
@@
 %   candidates?
-%
 % How much gain can Tahoe-LAFS get by switching to SHA-3?
@@
 
-% This chapter should contain (in listed and grouped order):
+% This chapter should contain:
 % -----------------------------------------------------------
@@
 % discussion
-%
-% We had hoped to have contributed more to the Open Source project that
-% Tahoe-LAFS is.
+
+We had hoped to contribute more to the Open Source project Tahoe-LAFS, when
+starting on our assignment. It was early evident that implementing a new
+capability was out of scope, since creating Python bindings for and trying out
+new implementations was prioritized.
+
+Perhaps the most important finding in this project, is that the SUPERCOP
+benchmarking indeed drives the development of new and better implementations
+forward, but since the SUPERCOP \ac{API} does not conform to the \ac{NIST}
+\ac{API}, the developers leave important functionality out (i.e.
+{\verb update() }), and hence make it difficult for people who wish to try out
+and test the remaining \ac{SHA}-3 candidates.
 

### 2010-12-11 paper/tables/sha3_flags.tex
Fixed compiler flags

@@
     \textbf{Candidate} & \textbf{Compiler flags}            \\ \hline
-     BLAKE       & mssse3 -m32 -march=native -mtune=native -O -fomit-frame-pointer  \\ \hline
-     \ac{BMW}    & -funroll-loops -m32 -march=pentium4 -O3 -fomit-frame-pointer     \\ \hline
-     Cubehash    & none                                                             \\ \hline
-     Echo        & -funroll-loops -m32 -march?prescott -O2 -fomit-frame-pointer     \\ \hline
-     Fugue       & -funroll-loops -O -fomit-frame-pointer                           \\ \hline
-     Grøstl     & -m32 -march=pentium3 -O -fomit-frame-pointer                     \\ \hline
-     Hamsi       & -msse2 -funroll-loops -m32 -march=pentium-m -O3 -fomit-frame-pointer\\ \hline
-     JH          & -m32 -march=core2 -msse4.1 -O -fomit-frame-pointer               \\ \hline
-     Keccak      & -msse2 -march=barcelona -O3 -fomit-frame-pointer                 \\ \hline
-     Luffa       & -msse2 -funroll-loops -march=k8 -O3 -fomit-frame-pointer         \\ \hline
-     Shabal      & -funroll-loops -m32 -march=pentium2 -O -fomit-frame-pointer      \\ \hline
-     Shavite3    & -funroll-loops -march=i486 -O -fomit-frame-pointer               \\ \hline
-     Simd        & -m32 -march=barcelona -O2 -fomit-frame-pointer                   \\ \hline
-     Skein       & -march=pentium2 -O2 -fomit-frame-pointer                         \\ \hline
+     BLAKE       & -m32 -march=native -mtune=native -O      \\ \hline
+     \ac{BMW}    & -funroll-loops -m32 -march=pentium4 -O3  \\ \hline
+     Cubehash    & -O2                                      \\ \hline
+     Echo        & -funroll-loops -m32 -march=prescott -O2  \\ \hline
+     Fugue       & -funroll-loops -O                        \\ \hline
+     Grøstl      & -m32 -march=pentium3 -O                  \\ \hline
+     Hamsi       & -funroll-loops -m32 -march=pentium-m -O3 \\ \hline
+     JH          & -m32 -march=core2 -O                     \\ \hline
+     Keccak      & -march=barcelona -O3                     \\ \hline
+     Luffa       & -funroll-loops -march=k8 -O3             \\ \hline
+     Shabal      & -funroll-loops -m32 -march=pentium2      \\ \hline
+     Shavite3    & -funroll-loops -march=i486 -O            \\ \hline
+     Simd        & -m32 -march=barcelona -O2                \\ \hline
+     Skein       & -march=pentium2 -O2                      \\ \hline
   \end{tabular}

### 2010-12-11 paper/listings/echo_hash.pyx
Fixed overflow in echo hash listing

@@ -26,7 +26,8 @@ cdef class echo:
 
     cpdef update(self, bytes in_data):
         cdef char* data = <char *> in_data
-        cdef int data_len = len(in_data)*8
+        cdef unsigned long long data_len = len(in_data)
+        data_len *= 8
 
         if self.finished:
             self.finished = 0

### 2010-12-11 paper/main.tex
Updated disposition on Discussion chapter

@@
 
+%  Writers should follow a checklist consisting of:
+% Motivation: Why do we care about the problem and results?
+% Problem Statement: What problem are we trying to solve? Scope/limits.
+% Approach: How did we go about solving or making progress on the problem?
+% Results: What is the answer? Numbers, not vague 'very', 'small' etc.
+% Conclusions: What are the implications of your answer? Further work.
+%
+%  Each section is typically a single sentence, although there is room for
+%  creativity.
+
 \chapter*{Preface}
 \addcontentsline{toc}{chapter}{Preface}
+
+The work behind this project report was carried out during the autumn semester
+in 2010 at the Norwegian University of Science and Technology (NTNU), Department
+of Telematics (ITEM).
+\vspace{13pt}
+
+We would like to thank our fellow classmates for interesting discussions and a
+lot of fun as the time and semesters fly by.
+\vspace{13pt}
+
+Secondly, we would like to state our appreciation to our supervisor, Danilo
+Gligoroski, and his coworker on Blue Midnight Wish, Rune E. Jensen, for
+providing useful feedback and interesting stories from the cryptographic
+community.
+\vspace{13pt}
+
+Lastly, the Open Source community behind Tahoe-LAFS deserves credit for the
+immense amount of hours they put on an awesome project, and Thomas Pornin for
+providing a working copy of his library of implementations, which saved us a lot
+of work.
 
@@
 The chosen implementations are shown in Table \ref{tbl:sha3:implementations}.
-Their compiler flags used, which are extracted from the \ac{SUPERCOP}
-results, can be seen in Table \ref{tbl:sha3:compilerflags}.
-
-It should be noted that \emph{32} in blake32 does not refer to the output length
-of the function, but the fact that the algorithm is optimized for 32-bit usage.
-Similarly, \emph{512} in keccakc512 does not mean 512-bit output, but 512 cycles in the
-internal function.
+Their compiler flags used, which are extracted from the \ac{SUPERCOP} results,
+can be seen in Table \ref{tbl:sha3:compilerflags}. Compiler flags are described
+in the GCC manual\footnote{GCC Manual \url{http://gcc.gnu.org/onlinedocs/}}.
+Compiler flags used for every implementation which might affect performance are:
+
+\begin{itemize}
+    \item -fomit-frame-pointer
+    \item -mssse3
+    \item -pthread
+    \item -fPIC
+\end{itemize}
+
+It should be noted that \emph{32} in blake32 means the function works on 32-bit
+words, and output a 32-byte digest (256 bit). Similarly, \emph{512} in keccakc512 
+does not mean 512-bit output, but 512 bits in the internal capacity.
 
@@
 Criterion 2. For both the optimized versions found in \ac{SUPERCOP} and the
-reference implementations in the \ac{NIST} submission package, Hash("foofoo")
-will not yield the same result as two consecutive calls to Update("foo").
-However the implementation in sphlib does fulfill this requirement. We have
-contacted the author, Thomas Pornin, which confirms this problem with the
-official JH candidate and that the author has been made aware of the problem
-but has not released any updated version yet (personal communication, October
-16, 2010).
+reference implementations in the \ac{NIST} submission package,
+{\verb Hash("foofoo") } will not yield the same result as two consecutive calls
+to {\verb Update("foo") }.  However the implementation in sphlib does fulfill
+this requirement. We have contacted the author, Thomas Pornin, which confirms
+this problem with the official JH candidate and that the author has been made
+aware of the problem but has not released any updated version yet
+\cite{s_pornin}.
 
@@
 with the C implementations of the NIST candidates. We did this by the use of
-Cython, and built a python object-oriented library which we named SHA3lib.
+Cython, and built a Python object-oriented library which we named SHA3lib.
 SHA3lib supports all of the 14 second round \ac{SHA}-3 candidates, but only in
@@
 hashlib\footnote{\url{http://docs.python.org/library/hashlib.html}}, the
-default python library for MD5, \ac{SHA}-1 and the \ac{SHA}-2 family. This
-means that we built class names for each candidate on the form
-candidatename+256 and that the classes included the following functions:
+default Python library for MD5, \ac{SHA}-1 and the \ac{SHA}-2 family. This
+made us create class names for each candidate on the form
+\emph{candidate name+256}. In addition, the classes included the following
+functions:
 
 \begin{itemize}
-    \item Constructur with an optional data element that does the same as
-    Update()
-    \item Update() takes data that should be considered in hashing
-    \item digest() that produce a raw byte-digest from submitted data
-    \item hexdigest() that produce a hex representation of the digest
-    \item copy() that copies the internal state of the function and produces a
-    copy of the object.
+    \item Constructor with an optional data element that does the same as
+    {\verb update() }
+    \item {\verb update() } takes data that should be considered in hashing
+    \item {\verb digest() } that produce a raw byte-digest from submitted data
+    \item {\verb hexdigest() } that produce a hex representation of the digest
+    \item {\verb copy() } that copies the internal state of the function and
+        produces a copy of the object.
 \end{itemize}
@@
 
-Basically the class echo resides in Cython and in C and is responsible for
-bringing the C functionality of the C implementation to Python. By compiling
-echo it will be made available for use in Python.  The class echo256 is a
-simple python class which inherits from echo. The only action this class
-actually takes is to set the output size to 256-bit. If the underlying
-C-functions would have supported multiple output sizes the same could have been
-done to incorporate those.
+Basically the class \textbf{echo} resides in Cython and in C, and is responsible
+for bringing the C functionality of the C implementation to Python. By compiling
+\textbf{echo} it will be made available for use in Python.  The class
+\textbf{echo256} is a simple Python class which inherits from \textbf{echo}. The
+only action this class actually takes is to set the output size to 256-bit. If
+the underlying C-functions would have supported multiple output sizes the same
+could have been done to incorporate those.
 
@@
 We employ Cython to generate bindings for the C implementation and the
-different candidates, as well as to create the base python classes. To access
+different candidates, as well as to create the base Python classes. To access
 functions in C we need to generate a Cython header file, which pretty much does
@@
 We then create a wrapper-class that mimics hashlib, this is shown in Listing
-\ref{lst:cython_impl}. Functions declared with def er python functions,
+\ref{lst:cython_impl}. Functions declared with def er Python functions,
 {\verb cdef } are Cython only while {\verb cpdef } are functions that can be
@@
 \subsubsection{\ac{Distutils}}
-We use the \ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to
+We use \ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to
 build and install our code. The alternative to doing this would be to manually
@@
 relevant C-files and then install by copying into the correct system directory
-where python libraries should be stored. \ac{Distutils} also support expansion
+where Python libraries should be stored. \ac{Distutils} also support expansion
 through self defined commands, something we have utilised to do Unit Testing.
@@
 \ac{SHA}-256, this means that it will also have to hash another 32 bytes for
-every message it returns the digest for. Since \ac{SUPERCOP} does not have
-benchmark for 32 byte messages, we will have to estimate the amount of cycles
-required for this.  If we assume that cycles/bytes is linear from 8 byte to 64
-byte, we get the following estimate for the speed of hashing 32 bytes:
-
-$$179.62 - \frac{179.62 - 38.12}{64 - 8} * (32-8)= 118.98 (Cycles/byte)$$
-
-We can then estimate the result of each message by the following formula:
-$$\frac{median * size + 118.98*32}{size} = SHA256d$$ \emph{estimate}
+every message it hashes. Since \ac{SUPERCOP} does not have benchmark for 32
+byte messages, we will have to estimate the amount of cycles required for this.
+If we assume a worst case scenario of hashing 32 bytes, where the cycles/byte
+count is exactly the same as for 8 bit median we get:
+
+$$\frac{median * size + 179.62*32}{size} = SHA256d estimate$$
 
@@
 practical situation that Tahoe-\ac{LAFS} is a file system, most files processed
-will be larger than the file sizes estimated here. Even though not all
-hashing operations done in Tahoe are related to the file size, we expect that
-only BLAKE, \ac{BMW}, Luffa, Shabal and SIMD will be able to outperform double
-\ac{SHA}-256.
+will be larger than the file sizes estimated here. Not all hashing
+operations in Tahoe-\ac{LAFS} are strictly dependant on the original file size,
+which would mean that all candidates would have a chance of outperforming
+double \ac{SHA}-256. However this estimate is worst case, which leads us to
+believe that at least for larger input sizes (over 1536) only the five fastest
+candidates will have a change of outperforming double \ac{SHA}-256.
 
@@
 
+%TODO: Vi har andre compiler argumenter - dette må diskuteres
+
 \section{Exclusion of mutable file testing}
@@
 % Adding a capability to support new hash functions in tahoe. (Our
-%   current way breaks backwards compability)
+%   current way breaks backwards compatibility)
 
@@
 To be able to extend the Tahoe-LAFS with new functionality to fully support new
-hash functions, one have to add a new capability (Zooko O'Whielacronx, personal
-communication, September 29, 2010). This ensures backwards compatibility and can
-be an optional feature for users to employ. However, we found that adding a new
-capability was out of scope for this assignment.
-
-% This chapter should contain (in listed and grouped order):
-% -----------------------------------------------------------
+hash functions, one have to add a new capability \cite{t_zooko}. This ensures
+backwards compatibility and can be an optional feature for users to employ.
+However, we found that adding a new capability was out of scope for this
+assignment.
+
+\section{Thoughts on the measurements}
+
 % Describe inconsistencies in the numbers, and try to explain them.
@@
 % Compare our results with general (SUPERCOP) results.
+%
+% How much gain can Tahoe-LAFS get by switching to SHA-3?
+%
+% Drag in the five selected remaining candidates.
+
+\section{Improvement of methods}
+% More general stuff. What could have been done differently? Criticize
+% ourselves.
+%
 % Optimized candidates - which gain could be expected from more optimized
 %   candidates?
-% How much gain can Tahoe-LAFS get by switching to SHA-3?
-%
+
+\subsection{The use of Cython}
 % What if we had used another coding scheme? - does Cython lower the
 %   candidates performance?
-%
-% More general stuff. What could have been done differently? Criticize
-% ourselves.
+
+The Python-bindings generated for Cython in C are not necessarily the most
+optimized versions possible to write and they are ugly to read or change (from
+a C perspective). However we consider them to be good enough, and they are
+equal for all the \ac{SHA}-3 candidates which means they should give the same
+performance penalty. The \ac{SHA}-256 version will of course not suffer from
+the same penalty, however we believe the possible performance gain is small.
+This is due to the fact that no intensive operations take place in our
+Cython-code.
+
+The alternative would be creating manual Python-bindings for all the candidates
+which would require a lot more work on our side, especially given the lack of a
+good enough and common \ac{API} for the implementations in \ac{SUPERCOP}.
+Another fact important for this decision was that we did not know exactly which
+implementation of each candidate we would end up using, and we ended up trying
+several for most candidates.
 

### 2010-12-12 utils/generate_table.sh
Quick script to help generate hashing times tables

@@ -1,0 +1,16 @@
+#!/bin/bash
+for cand in `ls run-sha3-multi-4`; do
+    time_sh=`cat run-sha3-multi-7/$cand/1gb_tshp.txt`
+    numerator=0
+    for r in `seq 4 7`; do
+        cd run-sha3-multi-$r
+        new_num=`cat $cand/1gb_totaltimeput.txt`
+        numerator=$(echo $numerator + $new_num | bc -q)
+        cd ..
+    done;
+    result=$(echo "scale=2; $numerator / 4" | bc -q)
+    difference=$(echo $result - $time_sh | bc -q)
+    diff_sub=${difference:0:6}
+    timesh_sub=${time_sh:0:5}
+    echo -e "$cand & $result\0163 & $timesh_sub\0163 & $diff_sub\0163 \\\\\ \\hline"
+done;

### 2010-12-12 paper/main.tex
Updated with graphs from latest run
Some parts to discussion

@@
 
+\subsection{Faster implementations}
+We do not claim to have tested the fastest possible nor existing version
+of most of the candidates. We are certain that someone of more skill and
+knolwedge about the inner workings of the \ac{SHA}-3 candidates, \ac{GCC} and
+Python could make significantly faster implementation.
+
+\subsection{Test with different Erasure Coding schemes}
+By default Tahoe-\ac{LAFS} comes with a coding scheme of 3-7-10 (shares needed,
+shares happy, shares total) while we test with a configuration of 2-3-3. It
+would have been interesting to see how this (and other schemes) would affect
+both the overall performance and the percentage of time spent hashing.
+
+\subsection{Measure with real use cases}
+Our test use random generated data of a certain length. In a real use case for
+Tahoe-\ac{LAFS} the data would probably be both of varying length, and of
+lengths not tested by us. In a real use case both a feasible erasure coding
+scheme would have to be decided, and the files would represent what you
+actually would want to use Tahoe-\ac{LAFS} for.
+
 %**************************************%

### 2010-12-12 paper/main.tex
Updated text describing result graphs

@@
 
-The sum of the times it took for the underlying hash library to calculate the
-hash operations are exhibited in Figures \ref{fig:graph:10001b},
+The sum of the amount of time it took for the underlying hash library to
+calculate the hash operations are exhibited in Figures \ref{fig:graph:10001b},
 \ref{fig:graph:1001kb}, \ref{fig:graph:501mb}, \ref{fig:graph:5100mb} and
@@
 
-The numbers for uploading and downloading of one 1GB file are summarized in
+The numbers for uploading and downloading the 1GB file are summarized in
 Tables \ref{tbl:hashingtimes:put1gb} and \ref{tbl:hashingtimes:get1gb}.
+BLAKE, \ac{BMW}, Shabal and Simd outperforms \ac{SHA}-256 when both uploading
+and downloading a 1GB file. The difference between the slowest (JH) and the
+fastest (BLAKE) algorithm is 229.92s when uploading the one gigabyte file.
+
+The third column in Tables \ref{tbl:hashingtimes:put1gb} and
+\ref{tbl:hashingtimes:get1gb} contains the time spent doing other operations
+than hashing, i.e. $Total time - Time spent hashing$. This number varies
+between 227.37s and 252.66s during upload and 205.94s and 225.03s during
+download.
 
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingPUTof10001Bfiles.pdf}
-    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingGETof10001Bfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashinguploading10001Bfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashingdownloading10001Bfiles.pdf}
     \caption{Figures for hash operations of 1000 1B files.}
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingPUTof1001KBfiles.pdf}
-    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingGETof1001KBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashinguploading1001KBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashingdownloading1001KBfiles.pdf}
     \caption{Figures for hash operations of 100 1KB files.}
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingPUTof501MBfiles.pdf}
-    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingGETof501MBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashinguploading501MBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashingdownloading501MBfiles.pdf}
     \caption{Figures for hash operations of 50 1MB files.}
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingPUTof5100MBfiles.pdf}
-    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingGETof5100MBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashinguploading5100MBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashingdownloading5100MBfiles.pdf}
     \caption{Figures for hash operations of five 100MB files.}
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingPUTof11GBfile.pdf}
-    \includegraphics[width=0.9\columnwidth]{graphs/TimespenthashingGETof11GBfile.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashinguploading11GBfile.pdf}
+    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashingdownloading11GBfile.pdf}
     \caption{Figures for hash operations of one 1GB file.}

### 2010-12-12 paper/listings/setup.py
Fixed overflow in table sha3_relativespeed, added -msse3 to compiler argument i setup.py example

@@ -11,6 +11,7 @@ ext_modules = [
          hashpath+'echo/sha3nist.c', 
          hashpath+"echo/echo.c"],
     extra_compile_args=[
+        '-mssse3',
         '-funroll-loops',
         '-m32',
         '-march=prescott',

### 2010-12-12 paper/title.tex
Added acronyms-package and more text to the tahoe-chapter
Filesystem -> file system, added deadline

@@
 % crop-et fra http://www.ntnu.no/infoavdelingen/selvhjelp/logoer/ntnu/NTNU_engelsk_RGB.png
-\includegraphics[scale=0.5]{NTNU-logo} \\
+\includegraphics[scale=0.5]{NTNU-logo.png} \\
 
@@
 Tahoe-LAFS is a Free Software/Open Source decentralized data store. It
-distributes your filesystem across multiple servers, and even if some of the
-servers fail or are taken over by an attacker, the entire filesystem continues
+distributes your file system across multiple servers, and even if some of the
+servers fail or are taken over by an attacker, the entire file system continues
 to work correctly and to preserve your privacy and security.
@@
 \begin{tabular}{ p{4cm} p{11cm}}
-Deadline: & 2010-12-xx \\
+Deadline: & 2010-12-20 \\
 Submission date: & 2010-12-xx \\

### 2010-12-12 utils/results_to_graphs.py
Updated vector count, 100 of 1b files now

@@ -17,7 +17,7 @@ import sys
 from graph_matplotlib import graph_test_vector
 
 VECTORS = ('1b', '1kb', '1mb', '100mb', '1gb')
-VECTOR_COUNT = {'1b': 1000, '1kb': 100, '1mb': 50, '100mb': 5, '1gb': 1}
+VECTOR_COUNT = {'1b': 100, '1kb': 100, '1mb': 50, '100mb': 5, '1gb': 1}
 OPERATIONS = {'get': ('tshg', 'totaltimeget', 'nohashopsg'),
               'put': ('tshp', 'totaltimeput', 'nohashopsp'),}
 

### 2010-12-12 paper/main.tex
Added names and date to Preface
Added abstract

@@
 
+Most of the cloud applications today are highly dependant on the use of
+cryptographic hash functions. The distributed file system Tahoe-LAFS is no
+exception, and uses hashing algorithms extensively for verification of the
+integrity of its files.
+
+In light of the ongoing NIST competition for selecting the new, standard
+cryptographic hash function named SHA-3, we tried out the 14 second
+round candidates in the Tahoe-LAFS file system using automated tests to measure
+the performance differences. Only the immutable file support was tested.
+
+Since the Tahoe-LAFS is programmed in Python, we had to make bindings to the
+SHA-3 implementations written in C, before making the necessary changes to the
+Tahoe-LAFS code. Several tasks where automated using scripts, to avoid doing
+manual work of distributing changes to the test grid.
+
+In general, four of the candidates perform better than the currently used
+SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd.
+By utilizing the performance improvements given by SHA-3, Tahoe-LAFS could
+save $5\%$ on upload operations by changing hashing function to the best
+performing candidate.
+
+The benchmarking tool SUPERCOP measures performance and drives the development
+forwards for the SHA-3 candidates, but we argue that since the SUPERCOP API does
+not conform with the NIST SHA-3 API, vital functionality are left out, making
+integration and testing with existing software hard.
+
 \chapter*{Preface}
@@
 of work.
+\vspace{13pt}
+
+Best regards,
+
+\begin{center}
+Eirik Haver and Pål Ruud
+\vspace{13pt}
+
+December 16, 2010
+\end{center}
 

### 2010-12-12 paper/main.tex
Added fraction to hashing time tables and updated conclusions

@@
 between 227.37s and 252.66s during upload and 205.94s and 225.03s during
-download.
+download. \emph{Fraction} displays the part of total time the calls to the
+hashing libraries take.
 
@@
 \section{\ac{SUPERCOP} results}
-% Time gain of removing ``double'' SHA-256. Double only in digest().
 
@@
 \section{Validity of \ac{SUPERCOP} results}
-% Mention that some SUPERCOP candidates are missing UPDATE-functionality.
 
@@
 \section{Exclusion of mutable file testing}
-% Implications of only testing immutable files? (mutable files also
-%   include directories)
-% Adding a capability to support new hash functions in tahoe. (Our
-%   current way breaks backwards compatibility)
 
@@
 \subsection{The use of Cython}
-% What if we had used another coding scheme? - does Cython lower the
-%   candidates performance?
 
@@
 of most of the candidates. We are certain that someone of more skill and
-knolwedge about the inner workings of the \ac{SHA}-3 candidates, \ac{GCC} and
+knowledge about the inner workings of the \ac{SHA}-3 candidates, \ac{GCC} and
 Python could make significantly faster implementation.
@@
 starting on our assignment. It was early evident that implementing a new
-capability was out of scope, since creating Python bindings for and trying out
+capability was out of scope, since creating Python bindings and trying out
 new implementations was prioritized.
+
+Four of the candidates has been shown to perform better than the currently used
+SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd. The time it takes for
+Tahoe-LAFS to upload a one gigabyte file, could improve by $5\%$ if the best
+performing SHA-3 candidate, BLAKE, were to be used.
 

### 2010-12-12 utils/generate_table.sh
Moved file variables to top

@@ -1,10 +1,12 @@
 #!/bin/bash
+tsh_file="1gb_tshg.txt"
+totaltime_file="1gb_totaltimeget.txt"
 for cand in `ls run-sha3-multi-4`; do
-    time_sh=`cat run-sha3-multi-7/$cand/1gb_tshp.txt`
+    time_sh=`cat run-sha3-multi-7/$cand/$tsh_file`
     numerator=0
     for r in `seq 4 7`; do
         cd run-sha3-multi-$r
-        new_num=`cat $cand/1gb_totaltimeput.txt`
+        new_num=`cat $cand/$totaltile_file`
         numerator=$(echo $numerator + $new_num | bc -q)
         cd ..
     done;
@@ -12,5 +14,6 @@ for cand in `ls run-sha3-multi-4`; do
     difference=$(echo $result - $time_sh | bc -q)
     diff_sub=${difference:0:6}
     timesh_sub=${time_sh:0:5}
-    echo -e "$cand & $result\0163 & $timesh_sub\0163 & $diff_sub\0163 \\\\\ \\hline"
+    fraction=$(echo "scale=2; $time_sh * 100 / $result" | bc -q)
+    echo -e "$cand & $result\0163 & $timesh_sub\0163 & $diff_sub\0163 & $fraction\\% \\\\\ \\hline"
 done;

### 2010-12-12 paper/main.tex
Added hashing time tables for all vectors

@@
 
-The numbers for uploading and downloading the 1GB file are summarized in
-Tables \ref{tbl:hashingtimes:put1gb} and \ref{tbl:hashingtimes:get1gb}.
+The numbers for uploading and downloading the different file vectors are
+summarized in Tables \ref{tbl:hashingtimes:put1gb},
+\ref{tbl:hashingtimes:get1gb}, \ref{tbl:hashingtimes:put100mb},
+\ref{tbl:hashingtimes:get100mb}, \ref{tbl:hashingtimes:put1mb},
+\ref{tbl:hashingtimes:get1mb}, \ref{tbl:hashingtimes:put1kb},
+\ref{tbl:hashingtimes:get1kb}, \ref{tbl:hashingtimes:put1b} and
+\ref{tbl:hashingtimes:get1b}.  The third column contains the time spent doing
+other operations than hashing, i.e. \emph{Time spent not hashing = Total time -
+Time spent hashing}. \emph{Fraction} displays the part of total time the calls
+to the hashing libraries take.
+
+As seen in Tables \ref{tbl:hashingtimes:put1gb} and
+\ref{tbl:hashingtimes:get1gb}, Time spent not hashing varies between 227.37s and
+252.66s during upload and 205.94s and 225.03s during download for a 1GB file.
+
 BLAKE, \ac{BMW}, Shabal and Simd outperforms \ac{SHA}-256 when both uploading
@@
 
-The third column in Tables \ref{tbl:hashingtimes:put1gb} and
-\ref{tbl:hashingtimes:get1gb} contains the time spent doing other operations
-than hashing, i.e. $Total time - Time spent hashing$. This number varies
-between 227.37s and 252.66s during upload and 205.94s and 225.03s during
-download. \emph{Fraction} displays the part of total time the calls to the
-hashing libraries take.
-
-\input{tables/hashingtimes}
+\input{tables/hashingtimes_1gb}
+\input{tables/hashingtimes_100mb}
+\input{tables/hashingtimes_1mb}
+\input{tables/hashingtimes_1kb}
+\input{tables/hashingtimes_1b}
 

### 2010-12-12 utils/generate_table.sh

@@ -1,12 +1,12 @@
 #!/bin/bash
-tsh_file="1gb_tshg.txt"
-totaltime_file="1gb_totaltimeget.txt"
+tsh_file="1b_tshp.txt"
+totaltime_file="1b_totaltimeput.txt"
 for cand in `ls run-sha3-multi-4`; do
     time_sh=`cat run-sha3-multi-7/$cand/$tsh_file`
     numerator=0
     for r in `seq 4 7`; do
         cd run-sha3-multi-$r
-        new_num=`cat $cand/$totaltile_file`
+        new_num=`cat $cand/$totaltime_file`
         numerator=$(echo $numerator + $new_num | bc -q)
         cd ..
     done;

### 2010-12-12 paper/main.tex
Added more text to results and brushed on Conclusion adding future work

@@
 \ref{fig:graph:1001kb}, \ref{fig:graph:501mb}, \ref{fig:graph:5100mb} and
-\ref{fig:graph:11gb}. We call this \emph{time spent hashing}.
+\ref{fig:graph:11gb}. We call this \emph{Time spent hashing}.
 
@@
 BLAKE, \ac{BMW}, Shabal and Simd outperforms \ac{SHA}-256 when both uploading
-and downloading a 1GB file. The difference between the slowest (JH) and the
-fastest (BLAKE) algorithm is 229.92s when uploading the one gigabyte file.
+and downloading a 1GB file, but on small files, SHA-256 is faster by a large
+margin. The difference between the slowest (JH) and the fastest (BLAKE)
+algorithm is 229.92s when uploading the one gigabyte file.
 
@@
 capability was out of scope, since creating Python bindings and trying out
-new implementations was prioritized.
+new implementations was prioritized. Future work should implement the support
+for choosing hashing algorithm in a proper manner which ensures backwards
+compatibility.
 
@@
 {\verb update() }), and hence make it difficult for people who wish to try out
-and test the remaining \ac{SHA}-3 candidates.
+and test the \ac{SHA}-3 candidates in real world applications.
 

### 2010-12-12 paper/main.tex
Squashed some TODOs and FIXMEs
Started plotting bytes hashed by tahoe

--- 2010-12-12 paper/main.tex
@@
 
-Most of the cloud applications today are highly dependant on the use of
-cryptographic hash functions. The distributed file system Tahoe-LAFS is no
-exception, and uses hashing algorithms extensively for verification of the
-integrity of its files.
-
-In light of the ongoing NIST competition for selecting the new, standard
-cryptographic hash function named SHA-3, we tried out the 14 second
-round candidates in the Tahoe-LAFS file system using automated tests to measure
-the performance differences. Only the immutable file support was tested.
-
-Since the Tahoe-LAFS is programmed in Python, we had to make bindings to the
-SHA-3 implementations written in C, before making the necessary changes to the
-Tahoe-LAFS code. Several tasks where automated using scripts, to avoid doing
-manual work of distributing changes to the test grid.
-
-In general, four of the candidates perform better than the currently used
-SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd.
-By utilizing the performance improvements given by SHA-3, Tahoe-LAFS could
-save $5\%$ on upload operations by changing hashing function to the best
-performing candidate.
-
-The benchmarking tool SUPERCOP measures performance and drives the development
-forwards for the SHA-3 candidates, but we argue that since the SUPERCOP API does
-not conform with the NIST SHA-3 API, vital functionality are left out, making
-integration and testing with existing software hard.
-
 \chapter*{Preface}
@@
 of work.
-\vspace{13pt}
-
-Best regards,
-
-\begin{center}
-Eirik Haver and Pål Ruud
-\vspace{13pt}
-
-December 16, 2010
-\end{center}
 
@@
 
+\section{Measurements of Tahoe data}
+% TODO: Fiks denne tabellen så den ikke overflower
+% TODO: Sleng inn samme for download
+Table \ref{tbl:tahoe:datauploaded} and describes the distribution of the
+different input sizes of data that Tahoe-\ac{LAFS} sends to the hash function
+when uploading and downloading the test vectors. E.g. to upload the 100 1-byte
+files, 15700 hashes had to be calculated, and about 64\% of these were of size
+106 byte.
+
+For every upload Tahoe-\ac{LAFS} does, it hashes a size close to the original
+file size twice. This does not happen on downloading the file.
+
+\input{tables/tahoe_byteshashed}
+
 \section{Measurements of Tahoe operations}
@@
 
-The sum of the amount of time it took for the underlying hash library to
-calculate the hash operations are exhibited in Figures \ref{fig:graph:10001b},
+The sum of the times it took for the underlying hash library to calculate the
+hash operations are exhibited in Figures \ref{fig:graph:10001b},
 \ref{fig:graph:1001kb}, \ref{fig:graph:501mb}, \ref{fig:graph:5100mb} and
@@
 
-The numbers for uploading and downloading the 1GB file are summarized in
+The numbers for uploading and downloading of one 1GB file are summarized in
 Tables \ref{tbl:hashingtimes:put1gb} and \ref{tbl:hashingtimes:get1gb}.
-BLAKE, \ac{BMW}, Shabal and Simd outperforms \ac{SHA}-256 when both uploading
-and downloading a 1GB file. The difference between the slowest (JH) and the
-fastest (BLAKE) algorithm is 229.92s when uploading the one gigabyte file.
-
-The third column in Tables \ref{tbl:hashingtimes:put1gb} and
-\ref{tbl:hashingtimes:get1gb} contains the time spent doing other operations
-than hashing, i.e. $Total time - Time spent hashing$. This number varies
-between 227.37s and 252.66s during upload and 205.94s and 225.03s during
-download. \emph{Fraction} displays the part of total time the calls to the
-hashing libraries take.
 
@@
 \section{\ac{SUPERCOP} results}
+% Time gain of removing ``double'' SHA-256. Double only in digest().
 
@@
 \section{Validity of \ac{SUPERCOP} results}
+% Mention that some SUPERCOP candidates are missing UPDATE-functionality.
 
@@
 \section{Exclusion of mutable file testing}
+% Implications of only testing immutable files? (mutable files also
+%   include directories)
+% Adding a capability to support new hash functions in tahoe. (Our
+%   current way breaks backwards compatibility)
 
@@
 \section{Thoughts on the measurements}
+
+\subsection{Small files}
+For small files, 1B and 1KB, we can clearly see that both \ac{SHA}-256 and
+double \ac{SHA}-256 greatly outperforms the \ac{SHA}-3 candidate
+implementations, which is unexpected if the time Tahoe-\ac{LAFS} do mainly
+realy on the filesize.
+
+\subsection{Large inputs}
+
+\subsection{The candidates}
+
+\subsection{Optimizing results}
 
@@
 \subsection{The use of Cython}
+% What if we had used another coding scheme? - does Cython lower the
+%   candidates performance?
 
@@
 of most of the candidates. We are certain that someone of more skill and
-knowledge about the inner workings of the \ac{SHA}-3 candidates, \ac{GCC} and
+knolwedge about the inner workings of the \ac{SHA}-3 candidates, \ac{GCC} and
 Python could make significantly faster implementation.
@@
 starting on our assignment. It was early evident that implementing a new
-capability was out of scope, since creating Python bindings and trying out
+capability was out of scope, since creating Python bindings for and trying out
 new implementations was prioritized.
-
-Four of the candidates has been shown to perform better than the currently used
-SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd. The time it takes for
-Tahoe-LAFS to upload a one gigabyte file, could improve by $5\%$ if the best
-performing SHA-3 candidate, BLAKE, were to be used.
 

### 2010-12-12 paper/main.tex

--- 2010-12-12 paper/main.tex
@@
 \footnote{\url{http://bench.cr.yp.to/results-sha3.html}}.
-%FIXME: Reference appendix if tahoe04 results end up there.
 
@@
 \vspace{12pt}
-Based on these criteria, we choose the fastest possible implementation based
-on our benchmark in \ac{SUPERCOP}. The relative speed of each implementation, in
-relation to other implementation for the same and other candidates are
-available at source.
-% FIXME: Link to appendix with tahoe04 SUPERCOP results.
+Based on these criteria, we choose the fastest possible implementation based on
+our benchmark in \ac{SUPERCOP} \cite{s_tahoe_supercop}. The relative speed of
+each implementation, in relation to other implementation for the same and other
+candidates are available at source.
 
@@
 
-% FIXME: Maybe add some more reasoning here.
-
 %**************************************%
@@
 %**************************************%
-
-% FIXME: Decide whether Munin should be included?
 
@@
 \section{\ac{SUPERCOP} benchmarks}
-% TODO: more introduction text here?
 
@@
 \input{tables/supercop_8}
+
+\section{Measurements of Tahoe data}
+% TODO: Fiks denne tabellen så den ikke overflower
+% TODO: Sleng inn samme for download
+Table \ref{tbl:tahoe:datauploaded} and describes the distribution of the
+different input sizes of data that Tahoe-\ac{LAFS} sends to the hash function
+when uploading and downloading the test vectors. E.g. to upload the 100 1-byte
+files, 15700 hashes had to be calculated, and about 64\% of these were of size
+106 byte.
+
+For every upload Tahoe-\ac{LAFS} does, it hashes a size close to the original
+file size twice. This does not happen on downloading the file.
+
+\input{tables/tahoe_byteshashed}
 
@@
 
+\subsection{Small files}
+For small files, 1B and 1KB, we can clearly see that both \ac{SHA}-256 and
+double \ac{SHA}-256 greatly outperforms the \ac{SHA}-3 candidate
+implementations, which is unexpected if the time Tahoe-\ac{LAFS} do mainly
+realy on the filesize.
+
+\subsection{Large inputs}
+
+\subsection{The candidates}
+
+\subsection{Optimizing results}
+
 % Describe inconsistencies in the numbers, and try to explain them.

### 2010-12-12 paper/main.tex
Comment
Added intro text to Introduction chapter

@@
 
-Thorough problem description. \\
-Stepwise, what will we actually do?
-
-Scope and objectives
+The process of selecting what is to become the new standard for a secure hashing
+algorithm is carried out by the {NIST}, and is at its final stages at the time
+of writing. Fourteen candidates came through to the second round
+\cite{s_nistround2}.
+
+The Tahoe-\ac{LAFS} is a distributed, secure file system perfectly suited to
+build file sharing services over the Internet. Currently it uses the
+\ac{SHA}-256 algorithm to verify the integrity of files and shares, in addition
+to creating keys, \ac{URI}s and more \cite{t_tahoe}.
+
+In this project assignment, we integrate the fourteen \ac{SHA}-3 candidates in
+the Tahoe-\ac{LAFS} and measure the performance impact. To be able to do this,
+we create Python bindings for the best suited C implementations of the SHA-3
+candidates, and alter the hash function calls in the source code of Tahoe.
 
@@
 
+\input{tables/hashingtimes_1b}
+\input{tables/hashingtimes_1kb}
+\input{tables/hashingtimes_1mb}
+\input{tables/hashingtimes_100mb}
 \input{tables/hashingtimes_1gb}
-\input{tables/hashingtimes_100mb}
-\input{tables/hashingtimes_1mb}
-\input{tables/hashingtimes_1kb}
-\input{tables/hashingtimes_1b}
 
@@
 performing SHA-3 candidate, BLAKE, were to be used.
-
+% TODO: Bedre og bedre implementasjoner kommer til å komme for SHA-3
 Perhaps the most important finding in this project, is that the SUPERCOP

### 2010-12-12 paper/main.tex
Added optimization thoughts to conclusion
Added some analysis of results

--- 2010-12-12 paper/main.tex
@@
 \subsection{Small files}
-For small files, 1B and 1KB, we can clearly see that both \ac{SHA}-256 and
-double \ac{SHA}-256 greatly outperforms the \ac{SHA}-3 candidate
-implementations, which is unexpected if the time Tahoe-\ac{LAFS} do mainly
-realy on the filesize.
-
-\subsection{Large inputs}
+For small files, 1B and 1KB, we can clearly see that both double \ac{SHA}-256 
+greatly outperforms the \ac{SHA}-3 candidate implementations. The data being 
+hashed for these files are dominated by lengths of 63 and 106 bytes. By 
+comparing with \ac{SUPERCOP} results, and especially the fact that we are 
+using double \ac{SHA}-256, which should result in large overhead this is
+unexpected.
+
+\subsection{Large files}
+For large files, 1MB, 100MB and 1GB, some of the \ac{SHA}-3 candidates are 
+actually outperforming double \ac{SHA}-256, which is expected from the
+\ac{SUPERCOP} results. However if we compare this result with the result for
+small files, and expect that a hash algorithm should do better on a cycles/byte
+basis when hashing larger data values this is not expected. Data sizes which
+are hashed during 1GB and 100MB up/downloads are dominated by 101B, 65569B and
+131106B. 50\% of the hashes are done on 101B, which is a size which should be
+largely affected by having to do double SHA-256. However, since the small files
+had a similar percentage of operations being done on 106B chunks, blaming need
+to use double \ac{SHA}256 is not enough. It looks like indeed the large data
+chunks being hashed is what is making the difference.
 
 \subsection{The candidates}
+If we look primarily at the large file up/download we see that the candidates
+that outperforms, or are close to outperforming \ac{SHA}-256 are BLAKE,
+\ac{BMW}, Shabal, Simd and Luffa. This is expected and corresponds with the
+results from \ac{SUPERCOP}.
+
+\subsection{Final round candidates}
+BLAKE, Keccak, Skein, Grøstl and JH have made it through to the final round of
+the \ac{SHA}-3 competition. %TODO: some comments on theese
 
 \subsection{Optimizing results}
+We have not used the fastest possible implementations possible for various
+reasons mentioned previously. However, if we had actually used the fastest
+implementations in \ac{SUPERCOP}, achieved median speed for every hash
+calculated the results should look as in Table X. The table is calculated using
+the lengths of data actually hashed by Tahoe-\ac{LAFS}, the medians from
+\ac{SUPERCOP} results, and interpolation for sizes in between those mentioned in
+\ac{SUPERCOP}. The point of this result is to see the maximum performance gain
+possible for TAHOE-\ac{LAFS} by adopting one of the current \ac{SHA}-3
+candidate implementations.
+
+%TODO: Optimaliserte resultater tabell
 

### 2010-12-12 paper/main.tex

@@
 Tahoe-LAFS to upload a one gigabyte file, could improve by $5\%$ if the best
-performing SHA-3 candidate, BLAKE, were to be used.
-% TODO: Bedre og bedre implementasjoner kommer til å komme for SHA-3
+performing SHA-3 candidate, BLAKE, were to be used. One has to bear in mind that
+SHA-256 has been optimized and tweaked for a number of years, and that these
+performance differences undoubtedly will continue to increase as new and better
+implementations will be developed for the final chosen \ac{SHA}-3 algorithm.
+
 Perhaps the most important finding in this project, is that the SUPERCOP

### 2010-12-12 utils/table_optimzedresults.py
Added util for calculating fastest possible results

@@ -1,0 +1,218 @@
+#!/usr/bin/python
+#coding: utf-8
+from os.path import join as pjoin
+from os import getlogin
+
+RESULTS_DIR='/home/'+getlogin()+'/Tahoe-LAFS-with-SHA-3/results/tahoe-datalengths/sha256-1'
+TEST_VECTORS = ('1b','1kb','1mb','100mb','1gb')
+CYCLES = 2833000000
+OPS=('p','g')
+
+MEDIANS = {
+        'blake': {
+                'long':8.38,
+                4096:8.58,
+                1563:8.92,
+                576:9.81,
+                64:21.25,
+                8:107.38,
+            },
+        'bmw':{
+                'long':6.94,
+                4096:7.29,
+                1563:7.71,
+                576:8.88,
+                64:23.78,
+                8:132.88,
+            },
+        #'cubehash':{
+        #    },
+        'echo':{
+                'long':32.64,
+                4096:32.77,
+                1563:35.91,
+                576:42.87,
+                64:101.33,
+                8:807.50,
+            },
+        'fugue':{
+                'long':17.64,
+                4096:19.28,
+                1563:21.98,
+                576:29.25,
+                64:123.91,
+                8:862.75,
+            },
+        'groestl':{
+                'long':23.47,
+                4096:24.16,
+                1563:25.28,
+                576:28.26,
+                64:65.08,
+                8:339.0,
+            },
+        'hamsi':{
+                'long':29.96,
+                4096:30.14,
+                1563:30.39,
+                576:30.94,
+                64:40.38,
+                8:117.88,
+            },
+
+        'jh':{
+                'long':18.96,
+                4096:19.32,
+                1563:19.90,
+                576:21.49,
+                64:40.38,
+                8:329.38,
+            },
+        'keccak':{
+                'long':20.88,
+                4096:21.49,
+                1563:22.31,
+                576:25.12,
+                64:50.34,
+                8:401.62,
+            },
+        'luffa':{
+                'long':13.53,
+                4096:13.87,
+                1563:14.43,
+                576:15.86,
+                64:33.59,
+                8:145.62,
+            },
+        'skein':{
+                'long':18.61,
[…]

### 2010-12-14 utils/table_optimzedresults.py
Added cubehash and sha256 to optimized results

@@ -25,8 +25,14 @@ MEDIANS = {
                 64:23.78,
                 8:132.88,
             },
-        #'cubehash':{
-        #    },
+        'cubehash':{
+                'long':13.23,
+                4096:14.45,
+                1563:16.37,
+                576:21.21,
+                64:91.50,
+                8:633,
+            },
         'echo':{
                 'long':32.64,
                 4096:32.77,
@@ -116,6 +122,14 @@ MEDIANS = {
                 64:25.56,
                 8:218.88,
             },
+        'sha256':{
+                'long':15.14,
+                4096:15.83,
+                1563:16.45,
+                576:18.05,
+                64:38.12,
+                8:179.62,
+                }
         
         }
 
@@ -178,7 +192,8 @@ def write_latex_table(op):
     all_data +="    \\begin{tabular}{ | l | r | r | r | r | r | }\n"
     all_data +="      \\hline\n" 
     headers = ['function','1b','1kb','1mb','100mb','1gb']
-    func = ['blake','bmw','echo','fugue','groestl','hamsi','jh','keccak','luffa','simd','skein','shabal','shavite3']
+    func =['blake','bmw','echo','fugue','groestl','hamsi',
+           'jh','keccak','luffa','sha256','simd','skein','shabal','shavite3']
     func.sort(reverse=True)
     header = "      "
     body = ""

### 2010-12-14 paper/main.tex
Moved estimates based on supercop results to Results chapter

@@
 
-\section{\ac{SUPERCOP} benchmarks}
+\section{SUPERCOP benchmarks}
 
@@
 \input{tables/supercop_8}
+
+\subsection{Estimates}
+
+We see in Tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
+\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
+\ref{tbl:supercop:8} that in general BLAKE, \ac{BMW}, Luffa, Shabal and Simd
+outperform \ac{SHA}-256. However, given that Tahoe-\ac{LAFS} uses double
+\ac{SHA}-256, this means that it will also have to hash another 32 bytes for
+every message it hashes. Since \ac{SUPERCOP} does not have a benchmark for 32
+byte messages, we will have to estimate the amount of cycles required for this.
+If we assume a worst case scenario of hashing 32 bytes, where the cycles per
+byte count is exactly the same as for 8 bit median, we get:
+
+$$\frac{median * size + 179.62*32}{size} = SHA256d\ estimate$$
+
+\input{tables/supercop_sha256d}
+
+Estimates based on this calculation are shown in Table
+\ref{tbl:supercop:sha256d}. For shorter lengths of data,
+more of the candidates are faster than double \ac{SHA}-256. Nevertheless, given
+the practical situation that Tahoe-\ac{LAFS} is a file system, most files
+processed will be larger than the file sizes estimated here. Not all hashing
+operations in Tahoe-\ac{LAFS} are strictly dependant on the original file size,
+which would mean that all candidates would have a chance of outperforming double
+\ac{SHA}-256. This estimate is however worst case, which leads us to believe
+that at least for larger input sizes (i.e. over 1536 bytes) only the five
+fastest candidates will have a chance of outperforming double \ac{SHA}-256.
 
@@
 
-\section{\ac{SUPERCOP} results}
-
-We see in Tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
-\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
-\ref{tbl:supercop:8} that in general BLAKE, \ac{BMW}, Luffa, Shabal and Simd
-outperform \ac{SHA}-256. However, given that Tahoe-LAFS uses double
-\ac{SHA}-256, this means that it will also have to hash another 32 bytes for
-every message it hashes. Since \ac{SUPERCOP} does not have benchmark for 32
-byte messages, we will have to estimate the amount of cycles required for this.
-If we assume a worst case scenario of hashing 32 bytes, where the cycles/byte
-count is exactly the same as for 8 bit median we get:
-
-$$\frac{median * size + 179.62*32}{size} = SHA256d estimate$$
-
-\input{tables/supercop_sha256d}
-
-Estimates based on this calculation are shown in Table
-\ref{tbl:supercop:sha256d}. What we can see that for lower lengths of data,
-more of the candidates are faster than double \ac{SHA}-256. However, given the
-practical situation that Tahoe-\ac{LAFS} is a file system, most files processed
-will be larger than the file sizes estimated here. Not all hashing
-operations in Tahoe-\ac{LAFS} are strictly dependant on the original file size,
-which would mean that all candidates would have a chance of outperforming
-double \ac{SHA}-256. However this estimate is worst case, which leads us to
-believe that at least for larger input sizes (over 1536) only the five fastest
-candidates will have a change of outperforming double \ac{SHA}-256.
+% TODO: Intro
 

### 2010-12-14 paper/main.tex
Added supercop compiler flags discussion

--- 2010-12-12 paper/main.tex
@@
 
+\paragraph{Extra compiler flags}
 %TODO: Vi har andre compiler argumenter - dette må diskuteres
+We do not compile the candidate implementation with the exact same arguments
+that is used in \ac{SUPERCOP}, most notably we append the -fPIC flag. This flag
+creates position-independant-code, in other words machine instruction code that
+can execute properly regardless of where in memory it resides. This will make
+one less register available on the x86 architecture. While we cannot directly
+prove that this will affect performance of the implementations. We do however
+suspect it. A person writing an optimization for \ac{SUPERCOP} will have to
+balance his use of resources, among them registers, to achieve maximum speed.
+The loss of one register will then give a performance penalty.
 

### 2010-12-14 paper/main.tex
Updated graphs and tables with sha256d data, and added text in Results chapter for why sha256 single is not included

@@
 \section{Measurements of Tahoe operations}
-%\vskip -1.5em % This is to fit in two graphs on one page including heading
-
-% How much time is spent on hashing versus total time of
-%   uploading/downloading
 
 The sum of the amount of time it took for the underlying hash library to
-calculate the hash operations are exhibited in Figures \ref{fig:graph:10001b},
+calculate the hash operations are exhibited in Figures \ref{fig:graph:1001b},
 \ref{fig:graph:1001kb}, \ref{fig:graph:501mb}, \ref{fig:graph:5100mb} and
@@
 
+\ac{SHA}-256 (single) is not included in the graphs, because the results were
+exactly the same as for the \ac{SHA}-256d used in Tahoe-\ac{LAFS}.
+
 As seen in Tables \ref{tbl:hashingtimes:put1gb} and
@@
 
-BLAKE, \ac{BMW}, Shabal and Simd outperforms \ac{SHA}-256 when both uploading
-and downloading a 1GB file, but on small files, SHA-256 is faster by a large
-margin. The difference between the slowest (JH) and the fastest (BLAKE)
-algorithm is 229.92s when uploading the one gigabyte file.
+BLAKE, \ac{BMW}, Shabal and Simd outperform \ac{SHA}-256d when both uploading
+and downloading a 1GB file, but on small files, \ac{SHA}-256d is faster by a
+large margin. The difference between the slowest (JH) and the fastest (BLAKE)
+algorithm is 229.92s when uploading the 1GB file.
 
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashinguploading10001Bfiles.pdf}
-    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashingdownloading10001Bfiles.pdf}
-    \caption{Figures for hash operations of 1000 1B files.}
-    \label{fig:graph:10001b}
+    \includegraphics[width=0.9\columnwidth]
+        {graphs/Timespenthashingwhileuploading1001Bfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]
+        {graphs/Timespenthashingwhiledownloading1001Bfiles.pdf}
+    \caption{Figures for hash operations of 100 1B files.}
+    \label{fig:graph:1001b}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashinguploading1001KBfiles.pdf}
-    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashingdownloading1001KBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]
+        {graphs/Timespenthashingwhileuploading1001KBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]
+        {graphs/Timespenthashingwhiledownloading1001KBfiles.pdf}
     \caption{Figures for hash operations of 100 1KB files.}
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashinguploading501MBfiles.pdf}
-    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashingdownloading501MBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]
+        {graphs/Timespenthashingwhileuploading501MBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]
+        {graphs/Timespenthashingwhiledownloading501MBfiles.pdf}
     \caption{Figures for hash operations of 50 1MB files.}
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashinguploading5100MBfiles.pdf}
-    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashingdownloading5100MBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]
+        {graphs/Timespenthashingwhileuploading5100MBfiles.pdf}
+    \includegraphics[width=0.9\columnwidth]
+        {graphs/Timespenthashingwhiledownloading5100MBfiles.pdf}
     \caption{Figures for hash operations of five 100MB files.}
@@
     \centering
-    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashinguploading11GBfile.pdf}
-    \includegraphics[width=0.9\columnwidth]{graphs/Timespenthashingdownloading11GBfile.pdf}
+    \includegraphics[width=0.9\columnwidth]
+        {graphs/Timespenthashingwhileuploading11GBfile.pdf}
+    \includegraphics[width=0.9\columnwidth]
+        {graphs/Timespenthashingwhiledownloading11GBfile.pdf}
     \caption{Figures for hash operations of one 1GB file.}

### 2010-12-14 paper/main.tex
Cython/us overhead in bindings discussion

--- 2010-12-14 paper/main.tex
@@
 using double \ac{SHA}-256, which should result in large overhead this is
-unexpected.
+unexpected. However, we believe that the results for small files may also
+reflect our use of Cython and our skill. Given large amounts of call to our
+bindings, while the data processed by the actual hash functions is small and
+therefore does not use a lot of time, the speed of the bindings will come in to
+play.
 
@@
 to use double \ac{SHA}256 is not enough. It looks like indeed the large data
-chunks being hashed is what is making the difference.
+chunks being hashed is what is making the difference. If we now consider that
+our bindings to the hash functions is what is causing the slow results for
+small files, this makes sense. Larger amounts of data will take longer to
+process for the hash function, which again results in the fraction of time used
+by our bindings is decreased.
+
+%FIXME: Nevn noe en plass om at vi kanskje burde timet kallene til HASH
+%funksjonen også.
 

### 2010-12-14 paper/main.tex
Added more calculations to sha256d estimate

@@
 
-\subsection{Estimates}
-
-We see in Tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
-\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
-\ref{tbl:supercop:8} that in general BLAKE, \ac{BMW}, Luffa, Shabal and Simd
-outperform \ac{SHA}-256. However, given that Tahoe-\ac{LAFS} uses double
-\ac{SHA}-256, this means that it will also have to hash another 32 bytes for
-every message it hashes. Since \ac{SUPERCOP} does not have a benchmark for 32
-byte messages, we will have to estimate the amount of cycles required for this.
-If we assume a worst case scenario of hashing 32 bytes, where the cycles per
-byte count is exactly the same as for 8 bit median, we get:
-
-$$\frac{median * size + 179.62*32}{size} = SHA256d\ estimate$$
+\subsection{Double SHA-256 Estimates}
+\ac{SUPERCOP} does not measure the speed of double \ac{SHA}-256, so we have no exact
+value of how many cycles this should take, we can however estimate it. For every hash
+calculation double \ac{SHA}-256 will have to hash an extra 32-byte.
+\ac{SUPERCOP} does not give results for hashing 32-byte, so this will also have
+to be estimated. The calculations of worst-case scenario (cycles/byte for 32
+byte is the same as 8 byte), best-case (cycles/byte for 32 byte is the same as
+for 64 byte) and interpolation (assuming linear decrease) are shown in Table
+\ref{tbl:supercop:sha256d}. The calculation for this table are as follows:
+
+$$\frac{median * size + 179.62*32}{size} = SHA256d\ estimate (worst-case)$$
+
+$$\frac{median * size + 38.12*32}{size} = SHA256d\ estimate (best-case)$$
+
+$$179.62 + \frac{38.12-179.62}{64-8}*(32-8) = 118.97 (cycles/byte for 32 byte)$$
+
+$$\frac{median * size + 118.97*32}{size} = SHA256d\ estimate (interpolation)$$
 
@@
 
-Estimates based on this calculation are shown in Table
-\ref{tbl:supercop:sha256d}. For shorter lengths of data,
-more of the candidates are faster than double \ac{SHA}-256. Nevertheless, given
-the practical situation that Tahoe-\ac{LAFS} is a file system, most files
-processed will be larger than the file sizes estimated here. Not all hashing
-operations in Tahoe-\ac{LAFS} are strictly dependant on the original file size,
-which would mean that all candidates would have a chance of outperforming double
-\ac{SHA}-256. This estimate is however worst case, which leads us to believe
-that at least for larger input sizes (i.e. over 1536 bytes) only the five
-fastest candidates will have a chance of outperforming double \ac{SHA}-256.
+Interpolation gives a somewhat pessimistic result of the cycles/byte needed for
+hashing 32 byte, so the probable speed is somewhere in between the
+interpolation point and the best case time.
 

### 2010-12-14 paper/tables/supercop_sha256d.tex
Changed supercop double sha256 to worst-case scenario
Added more calculations to sha256d estimate

@@
   \centering
-  \begin{tabular}{ | c | c | }
+  \begin{tabular}{ | c | c | c | c |}
     \hline
-    \textbf{Size} & \textbf{Median} \\ \hline
-    8    & 655.54 \\ \hline
-    64   & 97.61 \\ \hline
-    576  & 24.66 \\ \hline
-    1536 & 18.93 \\ \hline
-    4096 & 16.76 \\ \hline
+    \textbf{Size} & \textbf{Worst case} & \textbf{Interpolation} & \textbf{Best case} \\ \hline
+    8    & 898.1  & 655.53  & 332.10 \\ \hline
+    64   & 127.93 & 97.61  & 57.18 \\ \hline
+    576  & 53.85  & 24.66  & 20.17 \\ \hline
+    1536 & 20.19  & 18.89  & 17.23 \\ \hline
+    4096 & 17.23  & 16.76  & 16.13 \\ \hline
   \end{tabular}
-  \caption{Estimates for double SHA-256}
+  \caption{Estimates for medians using double \ac{SHA}-256}
   \label{tbl:supercop:sha256d}

### 2010-12-14 paper/main.tex
Added double sha256 vs single sha256 discussion

@@
 %funksjonen også.
+\subsection{Impact of double \ac{SHA}-256}
+The impact of using double \ac{SHA}-256 compared to single is suprisingly
+small, even if you use the best-case estimate from Table
+\ref{tbl:supercop:sha256}. For 1B, 1KB and 1MB files the data hashed is almost
+exclusively 106B or less. If we assume data is hashed at the performance of 576
+bytes, single \ac{SHA}-256 should outperform double \ac{SHA}-256 by about 10\%.
+For 1B files this difference is eaten up by external factors (other
+processes), the same goes for 1KB files. For 1MB files, there seems to be a
+difference, however this difference is so small that we deem it unimportant.
+For the larger files the performance penalty is even smaller, so all in all we
+deem the performance penalty for Tahoe-\ac{LAFS} for using double \ac{SHA}-256
+as opposed to single unimportant.
 

### 2010-12-14 utils/results_to_graphs.py
Added unicode filename support

@@ -22,7 +22,7 @@ OPERATIONS = {'get': ('tshg', 'totaltimeget', 'nohashopsg'),
               'put': ('tshp', 'totaltimeput', 'nohashopsp'),}
 
 def get_candidates(directory=None):
-    sub_directories = [f for f in os.listdir(directory) if os.path.isdir(
+    sub_directories = [unicode(f, 'utf-8') for f in os.listdir(directory) if os.path.isdir(
         os.path.join(directory, f))]
     sub_directories.sort()
     return sub_directories

### 2010-12-14 paper/main.tex
Added theoretical optimization estimates

@@
 \section{SUPERCOP benchmarks}
-
+\label{sec:results:supercop}
 The results of the \ac{SHA}-3 256-bit functions and \ac{SHA}-256 benchmarking in
@@
 \section{Measurements of Tahoe data}
+\label{sec:results:tahoedata}
+
 % TODO: Fiks denne tabellen så den ikke overflower
@@
 \end{figure}
+
+\subsection{Theoretical optimized results}
+These results are calculated using the best \ac{SUPERCOP} results for each
+candidate from section \ref{sec:results:supercop}, the exact sizes for each
+call to the hash function from section \ref{sec:results:tahoedata}, and
+interpolation to estimate the time to hash a length not recorded by
+\ac{SUPERCOP}. For lengths above 4096 the results for long data lanegths are
+used. The results also assume only one core of the reference processor is used.
+
+\input{tables/tahoe_optimized_p}
+\input{tables/tahoe_optimized_g}
 

### 2010-12-14 utils/table_optimzedresults.py

@@ -184,16 +184,16 @@ def write_latex_table(op):
     if op=='p':
         word = "upload"
     else:
-        word = "upload"
-    caption = "Fastest possible "+word+" using best SHA-3 candidate implementations"
+        word = "download"
+    caption = "Time spent hashing during "+word+" using best SHA-3 candidate implementations"
 
     all_data = "\\begin{table} \n"
     all_data +="  \\centering \n"
     all_data +="    \\begin{tabular}{ | l | r | r | r | r | r | }\n"
     all_data +="      \\hline\n" 
-    headers = ['function','1b','1kb','1mb','100mb','1gb']
-    func =['blake','bmw','echo','fugue','groestl','hamsi',
-           'jh','keccak','luffa','sha256','simd','skein','shabal','shavite3']
+    headers = ['Candidate','1B','1KB','1MB','100MB','1GB']
+    func =['BLAKE','BMW','Echo','Fugue','Groestl','Hamsi',
+           'JH','Keccak','Luffa','SHA256','Simd','Skein','Shabal','Shavite3']
     func.sort(reverse=True)
     header = "      "
     body = ""
@@ -202,9 +202,11 @@ def write_latex_table(op):
     header = header[:-2] + "\\\\ \\hline\n"
     
     for f in func:
+        if func=='Groestl':
+            k = "      Grøstl & "
         k = "      "+str(f) + " & "
         for tv in TEST_VECTORS:
-            k += str(MEDIANS[f][str(tv)+str(op)]) + " & "
+            k += str(MEDIANS[f.lower()][str(tv)+str(op)]) + " & "
         k = k[:-2] + "\\\\ \hline\n"
         body += k
     all_data += header

### 2010-12-14 paper/main.tex
Added distribution of file sizes hashed when downloading

@@
 \label{sec:results:tahoedata}
-
-% TODO: Fiks denne tabellen så den ikke overflower
-% TODO: Sleng inn samme for download
-Table \ref{tbl:tahoe:datauploaded} and describes the distribution of the
-different input sizes of data that Tahoe-\ac{LAFS} sends to the hash function
-when uploading and downloading the test vectors. E.g. to upload the 100 1-byte
-files, 15700 hashes had to be calculated, and about 64\% of these were of size
-106 byte.
+% TODO: Update text and Captions to better describe what is going on.
+
+Tables \ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datadownloaded} and
+describes the distribution of the different input sizes of data that
+Tahoe-\ac{LAFS} sends to the hash function when uploading and downloading the
+test vectors. E.g. to upload the 100 1-byte files, 15700 hashes had to be
+calculated, and about 64\% of these were of size 106 byte.
 
@@
 \section{Measurements of Tahoe operations}
-
 The sum of the amount of time it took for the underlying hash library to

### 2010-12-14 utils/graph_and_convert.sh
Make inkscape draw paths for text

@@ -9,5 +9,5 @@ cd $1
 
 for fn in `ls *.svg`; do
     echo "Converting $fn to PDF."
-    inkscape -D -z --file=$fn --export-pdf=`echo $fn | sed s/svg//`pdf
+    inkscape -D -T -z --file=$fn --export-pdf=`echo $fn | sed s/svg//`pdf
 done

### 2010-12-14 paper/main.tex
Minor corrections - removed some whitespace at eols
changed some of optimization discussion

--- 2010-12-14 paper/main.tex
@@
 implementations in \ac{SUPERCOP}, achieved median speed for every hash
-calculated the results should look as in Table X. The table is calculated using
-the lengths of data actually hashed by Tahoe-\ac{LAFS}, the medians from
-\ac{SUPERCOP} results, and interpolation for sizes in between those mentioned in
-\ac{SUPERCOP}. The point of this result is to see the maximum performance gain
-possible for TAHOE-\ac{LAFS} by adopting one of the current \ac{SHA}-3
-candidate implementations.
 
@@
 This is due to the fact that no intensive operations take place in our
-Cython-code.
+Cython-code. However we do believe that our bindings are affected by the
 

### 2010-12-14 paper/main.tex

--- 2010-12-14 paper/main.tex
@@
 the mutable file.
-%More text here?
 
@@
 \label{sec:results:tahoedata}
-% TODO: Update text and Captions to better describe what is going on.
 
@@
 For every upload Tahoe-\ac{LAFS} does, it hashes a size close to the original
-file size twice. This does not happen on downloading the file.
+file size twice. This does not happen when downloading the file.
 

### 2010-12-14 python-bindings/sha3lib/hash_functions/blake/blake.pyx
Time C-calls in BLAKE

@@ -26,6 +26,24 @@ def hash(int hashbitlen, bytes data, int in_length):
 
     return digest
 
+import os
+import time
+HASH_TIMING_FILE = '/tmp/time_spent_hashing' 
+def _write_increase(file_path, increment):
+    if not os.path.exists(file_path):
+        # Create the file if it does not exist.
+        fp = open(file_path, 'w')
+        fp.write('0')
+        fp.close()
+
+    fp = open(file_path, 'r')
+    current = float(fp.read())
+    fp.close()
+
+    fp = open(file_path, 'w')
+    fp.write(str(current + increment))
+    fp.close()
+
 cdef class blake:
     '''
     A class that tries to mimic the behaviour of hashlib, i.e. keeping state
@@ -40,12 +58,19 @@ cdef class blake:
     cdef int finished
     cdef int hashbitlen
     cdef list hashval
+    
+#    def float timehash
+
 
     def __init__(self, int in_hashbitlen, bytes initial=None):
         self.finished = 0
         self.hashbitlen = in_hashbitlen
         #blake_h.Init(&self.state, self.hashbitlen)
+        self.timehash = 0.0
+        t1 = time.time()
         blake_h.blake32_init( &self.state)
+        t2 = time.time()
+        self.timehash += t2 - t1
         if initial:
             self.update(initial)
 
@@ -59,7 +84,10 @@ cdef class blake:
             self.state = self.previous_state
 
         #blake_h.Update(&self.state, <blake_h.BitSequence *> data, data_len)
+        t1 = time.time()
         blake_h.blake32_update(&self.state, <blake_h.u8 *> data, <blake_h.u64> data_len)
+        t2 = time.time()
+        self.timehash += t2 - t1
 
     cpdef final(self):
         #cdef blake_h.BitSequence *hashval = <blake_h.BitSequence *> malloc(self.hashbitlen*8)
@@ -70,9 +98,13 @@ cdef class blake:
         self.previous_state = self.state
 
         #blake_h.Final(&self.state, hashval)
+        t1 = time.time()
         blake_h.blake32_final(&self.state, hashval)
-        self.finished = 1
+        t2 = time.time()
+        self.timehash += t2-t1
+        _write_increase(HASH_TIMING_FILE, float(self.timehash))
         
+        self.finished = 1
         self.hashval = [hashval[i] for i from 0 <= i < self.hashbitlen / 8]
         free(hashval)
 

### 2010-12-14 paper/main.tex
Added intro texts to Discussion and BG Tech chapters

@@
 
-In this project assignment, we integrate the fourteen \ac{SHA}-3 candidates in
+In this project assignment, we integrate the 14 \ac{SHA}-3 candidates in
 the Tahoe-\ac{LAFS} and measure the performance impact. To be able to do this,
@@
 %**************************************%
+
+This chapter provides background knowledge for the most important software and
+technologies used in this project assignment, comprising of cryptographic hash
+functions (in light of the \ac{NIST} competition), the \ac{SUPERCOP} toolkit,
+Tahoe-\ac{LAFS}, and lastly Python and Cython.
 
@@
 \subsection{Theoretical optimized results}
-These results are calculated using the best \ac{SUPERCOP} results for each
-candidate from section \ref{sec:results:supercop}, the exact sizes for each
-call to the hash function from section \ref{sec:results:tahoedata}, and
+
+Tables \ref{tbl:tahoe:optimized_g} and \ref{tbl:tahoe:optimized_p} exhibit
+the results calculated using the best \ac{SUPERCOP} implementation for each
+candidate from Section \ref{sec:results:supercop}, the exact sizes for each
+call to the hash function from Section \ref{sec:results:tahoedata}, and
 interpolation to estimate the time to hash a length not recorded by
-\ac{SUPERCOP}. For lengths above 4096 the results for long data lanegths are
+\ac{SUPERCOP}. For lengths above 4096B, the numbers for long data lengths are
 used. The results also assume only one core of the reference processor is used.
@@
 
-% TODO: Intro
+This chapter will provide comments on the results made, what the numbers
+mean and why there are some inconsistencies. In addition, general
+thoughts on the testing procedure and the \ac{SUPERCOP} toolkit will be
+presented.
 

### 2010-12-14 paper/main.tex
Removed some eol whitespace

@@
 \paragraph{Extra compiler flags}
-%TODO: Vi har andre compiler argumenter - dette må diskuteres
+
 We do not compile the candidate implementation with the exact same arguments
 that is used in \ac{SUPERCOP}, most notably we append the -fPIC flag. This flag
-creates position-independant-code, in other words machine instruction code that
-can execute properly regardless of where in memory it resides. This will make
-one less register available on the x86 architecture. While we cannot directly
-prove that this will affect performance of the implementations. We do however
-suspect it. A person writing an optimization for \ac{SUPERCOP} will have to
-balance his use of resources, among them registers, to achieve maximum speed.
+creates \emph{position independent code}, in other words machine instruction
+code that can execute properly regardless of where in memory it resides. This
+will make one less register available on the x86 architecture. While we cannot
+directly prove that this will affect performance of the implementations. We do
+however suspect it. A person writing an optimization for \ac{SUPERCOP} will have
+to balance his use of resources, among them registers, to achieve maximum speed.
 The loss of one register will then give a performance penalty.
@@
 \subsection{Small files}
+
 For small files, 1B and 1KB, we can clearly see that both double \ac{SHA}-256
@@
 hashed for these files are dominated by lengths of 63 and 106 bytes. By
-comparing with \ac{SUPERCOP} results, and especially the fact that we are 
-using double \ac{SHA}-256, which should result in large overhead this is
-unexpected. However, we believe that the results for small files may also
-reflect our use of Cython and our skill. Given large amounts of call to our
-bindings, while the data processed by the actual hash functions is small and
-therefore does not use a lot of time, the speed of the bindings will come in to
-play.
+comparing with \ac{SUPERCOP} results, and especially the fact that we are using
+double \ac{SHA}-256, which should result in large overhead this is unexpected.
+However, we believe that the results for small files may also reflect our use of
+Cython and our skill. Given large amounts of call to our bindings, while the
+data processed by the actual hash functions is small and therefore does not use
+a lot of time, the speed of the bindings will come in to play.
 
@@
 
-%FIXME: Nevn noe en plass om at vi kanskje burde timet kallene til HASH
-%funksjonen også.
 \subsection{Impact of double \ac{SHA}-256}
-The impact of using double \ac{SHA}-256 compared to single is suprisingly
+The impact of using double \ac{SHA}-256 compared to single is surprisingly
 small, even if you use the best-case estimate from Table
@@
 \section{Improvement of methods}
-% More general stuff. What could have been done differently? Criticize
-% ourselves.
-%
-% Optimized candidates - which gain could be expected from more optimized
-%   candidates?
 
 \subsection{The use of Cython}
+
+% TODO: Discuss the BLAKE run in Results when timing was done within Cython.
 

### 2010-12-14 paper/main.tex
Moved optimization results away from results

@@
 
-\subsection{Theoretical optimized results}
-
-Tables \ref{tbl:tahoe:optimized_g} and \ref{tbl:tahoe:optimized_p} exhibit
-the results calculated using the best \ac{SUPERCOP} implementation for each
-candidate from Section \ref{sec:results:supercop}, the exact sizes for each
-call to the hash function from Section \ref{sec:results:tahoedata}, and
-interpolation to estimate the time to hash a length not recorded by
-\ac{SUPERCOP}. For lengths above 4096B, the numbers for long data lengths are
-used. The results also assume only one core of the reference processor is used.
-
-\input{tables/tahoe_optimized_p}
-\input{tables/tahoe_optimized_g}
-
 %**************************************%
@@
 
+\subsection{Theoretical optimized results}
+We have not use the fastest available implementations for all of the
+candidates, we also know that our bindings creates an overhead for all calls.
+Therefore we have tried to calculate the results that could have been achieved
+in Tahoe-\ac{LAFS} if we had managed to obtain the speeds of the fastest
+implementations in \{SUPERCOP}. The results can be seen in appendix XX.
+%TODO: Fik appendix og referanse
+
+%Tables \ref{tbl:tahoe:optimized_g} and \ref{tbl:tahoe:optimized_p} exhibit
+%the results calculated using the best \ac{SUPERCOP} implementation for each
+%candidate from Section \ref{sec:results:supercop}, the exact sizes for each
+%call to the hash function from Section \ref{sec:results:tahoedata}, and
+%interpolation to estimate the time to hash a length not recorded by
+%\ac{SUPERCOP}. For lengths above 4096B, the numbers for long data lengths are
+%used. The results also assume only one core of the reference processor is used.
+
+%\input{tables/tahoe_optimized_p}
+%\input{tables/tahoe_optimized_g}
+
 %TODO: Optimaliserte resultater tabell

### 2010-12-14 paper/main.tex
Calculated cython overhead

@@
 
+\section{Eliminating overhead in measurements}
+The following measurements were done to eliminate the overhead introducer by
+our work.
+
+\paragraph{Cython bindings}
+To eliminate as much as possible of overhead introduced by our bindings, we
+timed the calls from our bindings to the underlying C-function. The overhead
+should be the same for all bindings, since the Cython code is the same. The
+measured overhead, calculated from BLAKE can be seen in Table XX.
+
+\input{tables/cython_overhead}
+
+\paragraph{Modification to Tahoe-\ac{LAFS}}
+Because we write to two files after each hash calculation, this will create an
+overhead in the measurements that would not normaly exist in Tahoe-\ac{LAFS}.
+This overhead should be about the same for all candidates, since the number of
+writes are equal. The results and calculations for this overhead, based on runs
+with double \ac{SHA}-256 can be seen in Table XX.
+
 %**************************************%

### 2010-12-14 utils/table_cythonoverhead.py

@@ -1,0 +1,55 @@
+VECTORS = ['1b','1kb','1mb','100mb','1gb']
+OPS = ('p','g')
+
+blake_o = {
+        '1bp':0.374,
+        '1bg':0.519,
+        '1kbp':1.13,
+        '1kbg':1.45,
+        '1mbp':1.96,
+        '1mbg':1.45,
+        '100mbp':10.09,
+        '100mbg':5.03,
+        '1gbp':20.28,
+        '1gbg':10.12,
+}
+
+blake_c = {
+        '1bp':0.16,
+        '1bg':0.15,
+        '1kbp':0.39,
+        '1kbg':0.40,
+        '1mbp':1.30,
+        '1mbg':0.72,
+        '100mbp':9.33,
+        '100mbg':4.56,
+        '1gbp':19.21,
+        '1gbg':9.27,
+}
+
+def calculate_overhead():
+    results = {}
+    for op in OPS:
+        for tv in VECTORS:
+            high = blake_o[tv+op]
+            low = blake_c[tv+op]
+            overhead = high - low
+            k = [high, low, overhead]
+            results[tv+op] = k
+    return results
+
+def print_results(results):
+    for op in OPS:
+        word = 'upload'
+        if op=='g':
+            word = 'download'
+        for tv in VECTORS:
+            print "%s & %s & %s & %s & %s" % (tv.upper(),
+                    word,
+                    results[tv+op][0],
+                    results[tv+op][1],
+                    results[tv+op][2])
+
+if __name__ == '__main__':
+    k = calculate_overhead()
+    print_results(k)

### 2010-12-14 utils/generate_table.sh
Script more flexible and automagic

@@ -1,19 +1,27 @@
 #!/bin/bash
-tsh_file="1b_tshp.txt"
-totaltime_file="1b_totaltimeput.txt"
-for cand in `ls run-sha3-multi-4`; do
-    time_sh=`cat run-sha3-multi-7/$cand/$tsh_file`
-    numerator=0
-    for r in `seq 4 7`; do
-        cd run-sha3-multi-$r
-        new_num=`cat $cand/$totaltime_file`
-        numerator=$(echo $numerator + $new_num | bc -q)
-        cd ..
+for op in put get; do
+    if [ "$op" == "put" ]; then
+        tsh_file="$1_tshp.txt"
+        totaltime_file="$1_totaltimeput.txt"
+    else
+        tsh_file="$1_tshg.txt"
+        totaltime_file="$1_totaltimeget.txt"
+    fi
+    echo "$op:"
+    for cand in `ls run-sha3-multi-4`; do
+        time_sh=`cat run-sha3-multi-7/$cand/$tsh_file`
+        numerator=0
+        for r in `seq 4 7`; do
+            cd run-sha3-multi-$r
+            new_num=`cat $cand/$totaltime_file`
+            numerator=$(echo $numerator + $new_num | bc -q)
+            cd ..
+        done;
+        result=$(echo "scale=2; $numerator / 4" | bc -q)
+        difference=$(echo $result - $time_sh | bc -q)
+        diff_sub=${difference:0:6}
+        timesh_sub=${time_sh:0:5}
+        fraction=$(echo "scale=2; $time_sh * 100 / $result" | bc -q)
+        echo -e "$cand & $result\0163 & $timesh_sub\0163 & $diff_sub\0163 & $fraction\\% \\\\\ \\hline"
     done;
-    result=$(echo "scale=2; $numerator / 4" | bc -q)
-    difference=$(echo $result - $time_sh | bc -q)
-    diff_sub=${difference:0:6}
-    timesh_sub=${time_sh:0:5}
-    fraction=$(echo "scale=2; $time_sh * 100 / $result" | bc -q)
-    echo -e "$cand & $result\0163 & $timesh_sub\0163 & $diff_sub\0163 & $fraction\\% \\\\\ \\hline"
 done;

### 2010-12-14 utils/table_cythonoverhead.py
Updated cythonoverhead to reflect tables of results correctly

@@ -4,10 +4,10 @@ OPS = ('p','g')
 blake_o = {
         '1bp':0.374,
         '1bg':0.519,
-        '1kbp':1.13,
-        '1kbg':1.45,
+        '1kbp':1.133,
+        '1kbg':1.448,
         '1mbp':1.96,
-        '1mbg':1.45,
+        '1mbg':1.445,
         '100mbp':10.09,
         '100mbg':5.03,
         '1gbp':20.28,
@@ -39,12 +39,12 @@ def calculate_overhead():
     return results
 
 def print_results(results):
-    for op in OPS:
-        word = 'upload'
-        if op=='g':
-            word = 'download'
         for tv in VECTORS:
-            print "%s & %s & %s & %s & %s" % (tv.upper(),
+            for op in OPS:
+                word = 'upload'
+                if op=='g':
+                    word = 'download'
+                print "%s & %s & %s & %s & %s \\\ \hline" % (tv.upper(),
                     word,
                     results[tv+op][0],
                     results[tv+op][1],

### 2010-12-14 paper/main.tex
Finished overhead results

@@
 should be the same for all bindings, since the Cython code is the same. The
-measured overhead, calculated from BLAKE can be seen in Table XX.
+measured overhead, calculated from BLAKE can be seen in Table
+\ref{tbl:cython:overhead}.
 
@@
 writes are equal. The results and calculations for this overhead, based on runs
-with double \ac{SHA}-256 can be seen in Table XX.
+with double \ac{SHA}-256 can be seen in Table \ref{tbl:measure:overhead}.
+
+\input{tables/measure_overhead}
 

### 2010-12-14 utils/table_measureoverhead.py

@@ -1,38 +1,38 @@
 VECTORS = ['1b','1kb','1mb','100mb','1gb']
 OPS = ('p','g')
 
-blake_o = {
-        '1bp':0.374,
-        '1bg':0.519,
-        '1kbp':1.133,
-        '1kbg':1.448,
-        '1mbp':1.96,
-        '1mbg':1.445,
-        '100mbp':10.09,
-        '100mbg':5.03,
-        '1gbp':20.28,
-        '1gbg':10.12,
+sha256_o = {
+        '1bp':50.26,
+        '1bg':49.63,
+        '1kbp':81.83,
+        '1kbg':86.41,
+        '1mbp':67.71,
+        '1mbg':61.73,
+        '100mbp':137.90,
+        '100mbg':120.79,
+        '1gbp':273.66,
+        '1gbg':239.45,
 }
 
-blake_c = {
-        '1bp':0.16,
-        '1bg':0.15,
-        '1kbp':0.39,
-        '1kbg':0.40,
-        '1mbp':1.30,
-        '1mbg':0.72,
-        '100mbp':9.33,
-        '100mbg':4.56,
-        '1gbp':19.21,
-        '1gbg':9.27,
+sha256_c = {
+        '1bp':41.29,
+        '1bg':39.28,
+        '1kbp':58.19,
+        '1kbg':56.92,
+        '1mbp':45.98,
+        '1mbg':40.68,
+        '100mbp':101.43,
+        '100mbg':99.75,
+        '1gbp':213.45,
+        '1gbg':194.03,
 }
 
 def calculate_overhead():
     results = {}
     for op in OPS:
         for tv in VECTORS:
-            high = blake_o[tv+op]
-            low = blake_c[tv+op]
+            high = sha256_o[tv+op]
+            low = sha256_c[tv+op]
             overhead = high - low
             k = [high, low, overhead]
             results[tv+op] = k

### 2010-12-14 paper/main.tex
Removed \ac from sections

@@
 In light of the ongoing NIST competition for selecting the new, standard
-cryptographic hash function named SHA-3, we tried out the 14 second
-round candidates in the Tahoe-LAFS file system using automated tests to measure
-the performance differences. Only the immutable file support was tested.
-
-Since the Tahoe-LAFS is programmed in Python, we had to make bindings to the
-SHA-3 implementations written in C, before making the necessary changes to the
-Tahoe-LAFS code. Several tasks where automated using scripts, to avoid doing
-manual work of distributing changes to the test grid.
+cryptographic hash function named SHA-3, 14 of the second round candidates were
+tried out in the Tahoe-LAFS file system using automated tests to measure the
+performance differences. Only the immutable file support was tested.
+
+Since the Tahoe-LAFS is programmed in Python, bindings to the SHA-3
+implementations written in C had to be made , before making the necessary
+changes to the Tahoe-LAFS code. Several tasks where automated using scripts, to
+avoid doing manual work of distributing changes to the test grid.
 
 In general, four of the candidates perform better than the currently used
-SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd.
-By utilizing the performance improvements given by SHA-3, Tahoe-LAFS could
-save $5\%$ on upload operations by changing hashing function to the best
-performing candidate.
+SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd.  By utilizing the
+performance improvements given by SHA-3, Tahoe-LAFS could currently save $5\%$
+on upload operations by changing hashing function to the best performing
+candidate.
 
@@
 
-\subsubsection{\ac{NIST} evaluation criteria for \ac{SHA}-3}
+\subsubsection{NIST evaluation criteria for SHA-3}
 
@@
 
-\subsubsection{\ac{NIST} \ac{SHA}-3 \ac{API}} \ac{NIST} required that every
+\subsubsection{NIST SHA-3 API}
+
+\ac{NIST} required that every
 submission should conform to a specified \ac{API}\cite{s_fedreg}. The \ac{API}
@@
 
-\subsection{\ac{SUPERCOP}}
+\subsection{SUPERCOP}
 \ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
@@
 
-\subsubsection{\ac{SUPERCOP} \ac{API}}
+\subsubsection{SUPERCOP API}
 \ac{SUPERCOP} specifies an \ac{API} which all submitted implementations must
@@
 
-\subsubsection{\ac{Distutils}}
+\subsubsection{Distutils}
 We use \ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to
@@
 
-\section{Modifications to the Tahoe-\ac{LAFS} Code}
+\section{Modifications to the Tahoe-LAFS Code}
 \label{sec:modtahoe}
@@
 
-\section{Measurements of Tahoe data}
+\section{Measurements of internal Tahoe hashing sizes}
 \label{sec:results:tahoedata}
@@
 
-\section{Measurements of Tahoe operations}
+\section{Measurements of Tahoe operations Using Test Vectors}
 The sum of the amount of time it took for the underlying hash library to
@@
 
-\section{Validity of \ac{SUPERCOP} results}
+\section{Validity of SUPERCOP results}
 
@@
 
-\subsection{Impact of double \ac{SHA}-256}
+\subsection{Impact of double SHA-256}
 The impact of using double \ac{SHA}-256 compared to single is surprisingly

### 2010-12-15 paper/main.tex
Small comment

--- 2010-12-14 paper/main.tex
@@
 
-In this project assignment, we integrate the 14 \ac{SHA}-3 candidates in
+In this project assignment, we integrate the fourteen \ac{SHA}-3 candidates in
 the Tahoe-\ac{LAFS} and measure the performance impact. To be able to do this,
@@
 %**************************************%
-
-This chapter provides background knowledge for the most important software and
-technologies used in this project assignment, comprising of cryptographic hash
-functions (in light of the \ac{NIST} competition), the \ac{SUPERCOP} toolkit,
-Tahoe-\ac{LAFS}, and lastly Python and Cython.
 
@@
 
-\section{Eliminating overhead in measurements}
-The following measurements were done to eliminate the overhead introducer by
-our work.
-
-\paragraph{Cython bindings}
-To eliminate as much as possible of overhead introduced by our bindings, we
-timed the calls from our bindings to the underlying C-function. The overhead
-should be the same for all bindings, since the Cython code is the same. The
-measured overhead, calculated from BLAKE can be seen in Table XX.
-
-\input{tables/cython_overhead}
-
-\paragraph{Modification to Tahoe-\ac{LAFS}}
-Because we write to two files after each hash calculation, this will create an
-overhead in the measurements that would not normaly exist in Tahoe-\ac{LAFS}.
-This overhead should be about the same for all candidates, since the number of
-writes are equal. The results and calculations for this overhead, based on runs
-with double \ac{SHA}-256 can be seen in Table XX.
+\subsection{Theoretical optimized results}
+These results are calculated using the best \ac{SUPERCOP} results for each
+candidate from section \ref{sec:results:supercop}, the exact sizes for each
+call to the hash function from section \ref{sec:results:tahoedata}, and
+interpolation to estimate the time to hash a length not recorded by
+\ac{SUPERCOP}. For lengths above 4096 the results for long data lanegths are
+used. The results also assume only one core of the reference processor is used.
+
+\input{tables/tahoe_optimized_p}
+\input{tables/tahoe_optimized_g}
 
@@
 
-This chapter will provide comments on the results made, what the numbers
-mean and why there are some inconsistencies. In addition, general
-thoughts on the testing procedure and the \ac{SUPERCOP} toolkit will be
-presented.
-
-\section{Validity of \ac{SUPERCOP} results}
+% TODO: Intro
+
+\section{Reliability of \ac{SUPERCOP} results}
 
@@
 \paragraph{Extra compiler flags}
-
+%TODO: Vi har andre compiler argumenter - dette må diskuteres
 We do not compile the candidate implementation with the exact same arguments
 that is used in \ac{SUPERCOP}, most notably we append the -fPIC flag. This flag
-creates \emph{position independent code}, in other words machine instruction
-code that can execute properly regardless of where in memory it resides. This
-will make one less register available on the x86 architecture. While we cannot
-directly prove that this will affect performance of the implementations. We do
-however suspect it. A person writing an optimization for \ac{SUPERCOP} will have
-to balance his use of resources, among them registers, to achieve maximum speed.
+creates position-independant-code, in other words machine instruction code that
+can execute properly regardless of where in memory it resides. This will make
+one less register available on the x86 architecture. While we cannot directly
+prove that this will affect performance of the implementations. We do however
+suspect it. A person writing an optimization for \ac{SUPERCOP} will have to
+balance his use of resources, among them registers, to achieve maximum speed.
 The loss of one register will then give a performance penalty.
@@
 \subsection{Small files}
-
 For small files, 1B and 1KB, we can clearly see that both double \ac{SHA}-256 
@@
 hashed for these files are dominated by lengths of 63 and 106 bytes. By 
-comparing with \ac{SUPERCOP} results, and especially the fact that we are using
-double \ac{SHA}-256, which should result in large overhead this is unexpected.
-However, we believe that the results for small files may also reflect our use of
-Cython and our skill. Given large amounts of call to our bindings, while the
-data processed by the actual hash functions is small and therefore does not use
-a lot of time, the speed of the bindings will come in to play.
+comparing with \ac{SUPERCOP} results, and especially the fact that we are 
+using double \ac{SHA}-256, which should result in large overhead this is
+unexpected. However, we believe that the results for small files may also
+reflect our use of Cython and our skill. Given large amounts of call to our
+bindings, while the data processed by the actual hash functions is small and
+therefore does not use a lot of time, the speed of the bindings will come in to
+play.
 
@@
 
+%FIXME: Nevn noe en plass om at vi kanskje burde timet kallene til HASH
+%funksjonen også.
 \subsection{Impact of double \ac{SHA}-256}
-The impact of using double \ac{SHA}-256 compared to single is surprisingly
+The impact of using double \ac{SHA}-256 compared to single is suprisingly
 small, even if you use the best-case estimate from Table
@@
 
-\subsection{Theoretical optimized results}
-We have not use the fastest available implementations for all of the
-candidates, we also know that our bindings creates an overhead for all calls.
-Therefore we have tried to calculate the results that could have been achieved
-in Tahoe-\ac{LAFS} if we had managed to obtain the speeds of the fastest
-implementations in \{SUPERCOP}. The results can be seen in appendix XX.
-%TODO: Fik appendix og referanse
-
-%Tables \ref{tbl:tahoe:optimized_g} and \ref{tbl:tahoe:optimized_p} exhibit
-%the results calculated using the best \ac{SUPERCOP} implementation for each
-%candidate from Section \ref{sec:results:supercop}, the exact sizes for each
-%call to the hash function from Section \ref{sec:results:tahoedata}, and
-%interpolation to estimate the time to hash a length not recorded by
-%\ac{SUPERCOP}. For lengths above 4096B, the numbers for long data lengths are
-%used. The results also assume only one core of the reference processor is used.
-
-%\input{tables/tahoe_optimized_p}
-%\input{tables/tahoe_optimized_g}
-
 %TODO: Optimaliserte resultater tabell
@@
 \section{Improvement of methods}
+% More general stuff. What could have been done differently? Criticize
+% ourselves.
+%
+% Optimized candidates - which gain could be expected from more optimized
+%   candidates?
 
 \subsection{The use of Cython}
-
-% TODO: Discuss the BLAKE run in Results when timing was done within Cython.
 

### 2010-12-15 paper/main.tex
Moved optimized results to appendix

--- 2010-12-14 paper/main.tex
@@
 
-\section{Validity of SUPERCOP results}
+\section{Reliability of \ac{SUPERCOP} results}
 
@@
 in Tahoe-\ac{LAFS} if we had managed to obtain the speeds of the fastest
-implementations in \{SUPERCOP}. The results can be seen in appendix XX.
-%TODO: Fik appendix og referanse
-
-%Tables \ref{tbl:tahoe:optimized_g} and \ref{tbl:tahoe:optimized_p} exhibit
-%the results calculated using the best \ac{SUPERCOP} implementation for each
-%candidate from Section \ref{sec:results:supercop}, the exact sizes for each
-%call to the hash function from Section \ref{sec:results:tahoedata}, and
-%interpolation to estimate the time to hash a length not recorded by
-%\ac{SUPERCOP}. For lengths above 4096B, the numbers for long data lengths are
-%used. The results also assume only one core of the reference processor is used.
-
-%\input{tables/tahoe_optimized_p}
-%\input{tables/tahoe_optimized_g}
-
-%TODO: Optimaliserte resultater tabell
+implementations in \{SUPERCOP}. The results can be seen in appendix
+\ref{appendix:optimzed_sha3}.
 
@@
 % Uncomment to enable appendices.
-%\appendix
-%\appendixpage
-%\addappheadtotoc
+\appendix
+\appendixpage
+\addappheadtotoc
+
+\chapter{Optimized SHA-3 candidates in Tahoe-LAFS}
+\label{appendix:optimized_sha3}
+
+Tables \ref{tbl:tahoe:optimized_g} and \ref{tbl:tahoe:optimized_p} exhibit
+results calculated which theoretically could have been achieved using the best
+\ac{SUPERCOP} implementation for each candidate in Tahoe-\ac{LAFS}. The
+calculations uses medians from Section \ref{sec:results:supercop} for the speed
+of each candidate, the exact sizes for each call to the hash function from
+Section \ref{sec:results:tahoedata}, and interpolation to estimate the time to
+hash a length not recorded by \ac{SUPERCOP}. For lengths above 4096B, the
+numbers for long data lengths are used. The results also assume only one core
+of the reference processor (2,833 GHz) is used. We can see that some of the calculations,
+e.g. for Shavite3 are worse than the result we achieved in our measurements,
+this is probably due to interpollation being an inaccurate method to use for
+the shorter lengths of data.
+
+\input{tables/tahoe_optimized_p}
+\input{tables/tahoe_optimized_g}
 

### 2010-12-15 paper/main.tex
Cleaned up SUPERCOP discussion

@@
 \subsection{Chosen implementations}
-
+\label{sec:chosen_implementations}
 \input{tables/sha3_implementations}
@@
 
-\section{Reliability of \ac{SUPERCOP} results}
+\section{Reliability of SUPERCOP results}
 
@@
 function available for all implementations, and that \ac{SUPERCOP} only
-measures speed of hashing full byte strings. This is also a hassle for a
-programmer who wants to use the implementations in \ac{SUPERCOP}
+measures speed of hashing full byte strings. We are questioning if this
+discriminates implementations which actually can do more of their processing on
+a call to {\verb Update() }. This is also a hassle for a programmer who wants
+to use the implementations found in \ac{SUPERCOP}
 
 \paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
-mean implementations that does not conform to the \ac{KAT}s supplied.
-
-\paragraph{Results are only valid for one specific computer}. The results we
-used from \ac{SUPERCOP} are only valid for our specific computer, although it
-gives an indication of which speeds can be achieved on similar computers, i.e.
-32-bit. However to get a complete picture of the performance of the different
-candidates one must look at different computers and architectures. This picture
-might be obtained from looking at the results from all computers measured by
-\ac{SUPERCOP}.
+mean implementations that does not conform to the \ac{KAT}s supplied or
+implementations which fail completely as hash functions, see section
+\ref{sec:choosen_implementations} for details on the JH implementation.
 

### 2010-12-15 paper/main.tex
QA till Ch Results

--- 2010-12-15 paper/main.tex
@@
   \item Write Python bindings, with unit tests for verification, to be able to
-    use the functions from Python.
+    use the cryptographic hash functions from Python.
   \item Modify the Tahoe-LAFS code to utilize the SHA-3 functions.
@@
 While the official \ac{NIST} \ac{SHA}-3 second round submissions contain both
-a reference version and optimized versions of the candidates, it does not
+a reference version and optimized versions, they do not
 appear to be updated since the submission deadline on September 15, 2009. They
-also do not contain any benchmarking utility which could be used to decide on
+do also not contain any benchmarking utility which could be used to decide on
 the best implementation for a particular platform or architecture. \ac{SUPERCOP}
 on the other hand, provides both, in combination with results stating which
-compiler arguments that gives the best performance. On the downside, the
+compiler arguments giving the best performance. On the downside, the
 \ac{SUPERCOP} \ac{API} only specifies the {\verb crypto_hash() } function and
@@
 can be seen in Table \ref{tbl:sha3:compilerflags}. Compiler flags are described
-in the GCC manual\footnote{GCC Manual \url{http://gcc.gnu.org/onlinedocs/}}.
-Compiler flags used for every implementation which might affect performance are:
+in the GCC manual\footnote{GCC Manual \url{http://gcc.gnu.org/onlinedocs/}},
+and the one used for every implementation which might affect performance, are:
 
 \begin{itemize}
-    \item -fomit-frame-pointer
-    \item -mssse3
-    \item -pthread
-    \item -fPIC
+    \item[] -fomit-frame-pointer
+    \item[] -mssse3
+    \item[] -pthread
+    \item[] -fPIC
 \end{itemize}
@@
 It should be noted that \emph{32} in blake32 means the function works on 32-bit
-words, and output a 32-byte digest (256 bit). Similarly, \emph{512} in keccakc512 
-does not mean 512-bit output, but 512 bits in the internal capacity.
+words, and output a 32-byte digest (256 bit). Similarly, \emph{512} in
+keccakc512 does not mean 512-bit output, but 512 bits in the internal capacity.
 
@@
 Worth mentioning is that the fastest SHA-256 implementation in \ac{SUPERCOP} is
-cryptopp\footnote{\url{http://www.cryptopp.com/}}, which is the one used by
-Tahoe-\ac{LAFS}. For the most part we have found working versions of all
+from cryptopp\footnote{\url{http://www.cryptopp.com/}}, which is the one used by
+Tahoe-\ac{LAFS}. For the most part, we have found working versions of all
 candidates in \ac{SUPERCOP}, however there are some exceptions.
@@
 \paragraph{JH} seems to be broken in every implementation with regard to
-Criterion 2. For both the optimized versions found in \ac{SUPERCOP} and the
+Criterion 2. For both the optimized versions found in \ac{SUPERCOP}, and the
 reference implementations in the \ac{NIST} submission package,
 {\verb Hash("foofoo") } will not yield the same result as two consecutive calls
-to {\verb Update("foo") }.  However the implementation in sphlib does fulfill
+to {\verb Update("foo") }. However, the implementation in sphlib does fulfill
 this requirement. We have contacted the author, Thomas Pornin, which confirms
-this problem with the official JH candidate and that the author has been made
-aware of the problem but has not released any updated version yet
+this problem with the official JH candidate, and that the author has been made
+aware of the problem, but has not released any updated version yet
 \cite{s_pornin}.
@@
 
-\paragraph{Simd vect128 implementation} does not by default conform to Criterion
+\paragraph{Simd vect128} implementation does not by default conform to Criterion
 4, because of a bug in the implementation. This has been confirmed by the author
-who also supplied us with a patch to correct the bug (Gaëtan Leurent, personal
-communication, November 10, 2010).
+who also supplied us with a patch to correct the bug \cite{s_simd}.
 
@@
 Since Tahoe-\ac{LAFS} is written in Python, we needed some way of interacting
-with the C implementations of the NIST candidates. We did this by the use of
+with the C implementations of the \ac{NIST} candidates. We did this by the use of
 Cython, and built a Python object-oriented library which we named SHA3lib.
@@
 \subsection{SHA3lib}
+
 SHA3lib is built to mimic the functionality of
@@
     {\verb update() }
-    \item {\verb update() } takes data that should be considered in hashing
-    \item {\verb digest() } that produce a raw byte-digest from submitted data
+    \item {\verb update() } receives data that should be hashed
+    \item {\verb digest() } that produce a raw byte digest from submitted data
     \item {\verb hexdigest() } that produce a hex representation of the digest
@@
 \subsubsection{Design}
+
 Our original goal was to make SHA3lib include functionality to support all
-output lengths, and not only 256-bit. This was based on our observation that
-the official \ac{NIST} submission packages include ``all-in-one''
-implementations that support this. Later we observed that this is not the case
-for all implementations in \ac{SUPERCOP}. Together with the fact that
-Tahoe-\ac{LAFS} uses 256-bit output we in the end only included support for
-256-bit output in SHA3lib. The design we have used is shown in Figure
+output lengths, and not only 256-bit. This was based on our observation that the
+official \ac{NIST} submission packages include ``all-in-one'' implementations
+that support this. Later we observed that this is not the case for all
+implementations in \ac{SUPERCOP}. Together with the fact that Tahoe-\ac{LAFS}
+uses 256-bit output, in the end only included support for 256-bit output in
+SHA3lib was implemented. The design we have used is shown in Figure
 \ref{fig:python:inheritance}, and does follow the initial idea that SHA3lib
@@
     \includegraphics[width=0.6\columnwidth]{python-inheritance.pdf}
-    \caption{Design of SHA3lib using echo as an example}
+    \caption{Design of SHA3lib using Echo as an example}
     \label{fig:python:inheritance}
@@
 \textbf{echo256} is a simple Python class which inherits from \textbf{echo}. The
-only action this class actually takes is to set the output size to 256-bit. If
-the underlying C-functions would have supported multiple output sizes the same
+only action this class actually performs, is to set the output size to 256 bit. If
+the underlying C-functions would have supported multiple output sizes, the same
 could have been done to incorporate those.
@@
 \subsubsection{Cython}
-We employ Cython to generate bindings for the C implementation and the
+We employ Cython to generate bindings for the C implementation of the
 different candidates, as well as to create the base Python classes. To access
 functions in C we need to generate a Cython header file, which pretty much does
-the same as a C header file; it tells which functions and data types are
+the same as a C header file -- it describes which functions and data types are
 available. How we accomplish this for an implementation that satisfies the
-\ac{NIST} \ac{API} is shown in Listing \ref{lst:cython_header}.
+\ac{NIST} \ac{API}, is shown in Listing \ref{lst:cython_header}.
 
@@
 We then create a wrapper-class that mimics hashlib, this is shown in Listing
-\ref{lst:cython_impl}. Functions declared with def er Python functions,
-{\verb cdef } are Cython only while {\verb cpdef } are functions that can be
-accessed from both. While this is just the implementation for the echo
-candidate, the almost exact same pattern is used for all of the other
-candidates.
+\ref{lst:cython_impl}. Functions declared with {\verb def } are Python
+functions, {\verb cdef } are Cython-only functions, while {\verb cpdef } are
+functions that can be accessed from both languages. While this is just the
+implementation for the Echo candidate, the almost exact same pattern is used for
+all of the other candidates.
 
@@
 \subsubsection{Distutils}
+
 We use \ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to
-build and install our code. The alternative to doing this would be to manually
-transform Cython files to C-files, build the shared library from this and other
-relevant C-files and then install by copying into the correct system directory
+build and install our code. The alternative of doing this would be to manually
+transform Cython files to C files, build the shared library from this and other
+relevant C files, and then install by copying into the correct system directory
 where Python libraries should be stored. \ac{Distutils} also support expansion
 through self defined commands, something we have utilised to do Unit Testing.
-Options and configuration for \ac{Distutils} is done through a file
-{\verb setup.py}. Relevant configuration for compiling the echo extension, and
-installing SHA3lib can bee seen in Listing \ref{lst:setup_py}.
+Options and configuration for \ac{Distutils} is done through a file named
+{\verb setup.py }. Relevant configuration for compiling the Echo extension, and
+installing SHA3lib, can bee seen in Listing \ref{lst:setup_py}.
 
@@
 \paragraph{Known-Answer Tests} are used to check that the implementation conform
-to the specifications of the algorithm author. The \ac{KAT}s include
-bytestrings, their length and their expected message digest. The length of the
+to the specifications of the algorithm author. The \acp{KAT} include
+byte strings, their length and their expected message digest. The length of the
 inputs range from 0 bits to 34304 bits, and also include a message that should
@@
 
-\paragraph{Basic Hash Tests} is what we have named the set of tests which test
+\paragraph{Basic Hash Tests} are what we have named the set of tests which test
 the core basics of what a hash function implementation will have to be able to
-do to be usable. This include that Init(), Update() and Final() works as
-expected, and that the output is of correct size. This include that every
-function should produce the same digest for the inputs Update("foofoo") and two
-consecutive calls to Update("foo").
-
-\paragraph{Copy Hash Tests} does not test any of the functionality used by
-Tahoe-\ac{LAFS}, but the copy()-function is a part of Hashlib, and we have
-included it in SHA3lib to conform to this. The idea is that if we have two
-bytestring that share a common prefix, we can input this prefix to Update(),
-then copy the state. The tests check if this is possible, and that altering the
-copied state does not affect the original.
+do to be usable. This include that {\verb init() }, {\verb update() } and
+{\verb final() } works as expected, and that the output is of correct size. This
+include that every function should produce the same digest for the inputs
+{\verb update("foofoo") } and two consecutive calls to {\verb update("foo") }.
+
+\paragraph{Copy Hash Tests} do not test any of the functionality used by
+Tahoe-\ac{LAFS}, but the {\verb copy() } function is a part of hashlib, and we
+have included it in SHA3lib to conform to this. The idea is that if we have two
+byte strings that share a common prefix, we can input this prefix to
+{\verb update() }, then copy the state. The tests check if this is possible, and
+that altering the copied state does not affect the original.
 
@@
 
-The code of Tahoe-LAFS conforms to the style guide of Python, known as PEP 8.
-This, along with the code being highly modular and consistent to its
-architecture, emphasizes code readability.
-With some experience using Python, it is not difficult to learn the Tahoe code
-enough to understand what is happening when a file is uploaded or downloaded
-with the system.
+The code of Tahoe-\ac{LAFS} conforms to the style guide of Python, known as PEP
+8.  This, along with the code being highly modular and consistent to its
+architecture, emphasizes code readability.  With some experience using Python,
+it is not difficult to learn the Tahoe code enough to understand what is
+happening when a file is uploaded or downloaded with the system.
 
@@
 
-As stated in Section \ref{sec:lengthextension}, none of the SHA-3 candidates is
-vulnerable to the length-extension attack. Since we are interested in finding
-out how Tahoe-\ac{LAFS} performs when utilizing the SHA-3 functions, we can
-remove the lines that fixes the length-extension problems manually. In addition,
-to try out one of the new implementations, we change line one in Listing
-\ref{lst:hashutil}, and an example of this can be seen in Listing
-\ref{lst:hashutilmod}.
+As stated in Section \ref{sec:lengthextension}, none of the \ac{SHA}-3
+candidates are vulnerable to the length-extension attack. Since we are
+interested in finding out how Tahoe-\ac{LAFS} performs when utilizing the
+\ac{SHA}-3 functions, we can remove the lines that fixes the length-extension
+problems manually. In addition, to try out one of the new implementations, we
+change line one in Listing \ref{lst:hashutil}, and an example of this can be
+seen in Listing \ref{lst:hashutilmod}.
 
@@
 immutable files} are tested and measured in this paper. Implications of this are
-discussed further in Chapter \ref{ch:discussion}.
+further discussed in Chapter \ref{ch:discussion}.
 
@@
 software revision control systems, which distributes both the source code of
-Tahoe-LAFS, SHA3lib and shell-scripts for installation and configuration of
+Tahoe-\ac{LAFS}, SHA3lib and shell-scripts for installation and configuration of
 both. Configuration is mainly done by the use of regular expressions.
@@
             \item Install SHA3lib on all nodes
-            \item Replace used hash algorithm in Tahoe-LAFS with current
+            \item Replace used hash algorithm in Tahoe-\ac{LAFS} with current
               candidate
-            \item Install Tahoe-LAFS
-            \item Create and start one Tahoe-LAFS introducer
-            \item Create and start Tahoe-LAFS storage nodes on 3 nodes
-            \item Create a Tahoe-LAFS client on a previously unnused node
+            \item Install Tahoe-\ac{LAFS}
+            \item Create and start one Tahoe-\ac{LAFS} introducer
+            \item Create and start Tahoe-\ac{LAFS} storage nodes on 3 nodes
+            \item Create a Tahoe-\ac{LAFS} client on a previously unused node
             \item Do testing
@@
 
-\paragraph{Distribution of source code for Tahoe-LAFS and SHA3lib} is done by
-\ac{SVN} and GIT respectivly. The change of hash algorithm in Tahoe-LAFS is
+\paragraph{Distribution of source code for Tahoe-LAFS and SHA3lib} are done by
+\ac{SVN} and GIT respectively. The change of hash algorithm in Tahoe-LAFS is
 however done with the use of a regular expression, because of both the
@@
 and timed total time spent hashing, i.e. the total time it took for the function
-calls to the underlying hashing operations to complete.
+calls to the underlying hashing operations to complete. We also timed how long
+the Tahoe operation took in total for one test vector.
 

### 2010-12-15 paper/main.tex
\autoref \o/

@@
 
-As stated in Section \ref{sec:lengthextension}, none of the \ac{SHA}-3
+As stated in \autoref{sec:lengthextension}, none of the \ac{SHA}-3
 candidates are vulnerable to the length-extension attack. Since we are
@@
 immutable files} are tested and measured in this paper. Implications of this are
-further discussed in Chapter \ref{ch:discussion}.
+further discussed in \autoref{ch:discussion}.
 
@@
 
-As stated in Section \ref{sec:modtahoe}, only immutable files are tested. The
-mutable files are not tested because of our way of integrating the SHA-3
-candidates within the Tahoe-LAFS code.
+As stated in \autoref{sec:modtahoe}, only the Tahoe-\ac{LAFS} file type
+\emph{immutable files} are tested. The mutable files are not tested because of
+our way of integrating the SHA-3 candidates within the Tahoe-LAFS code.
 
@@
 functionality. Similarly, Listing \ref{lst:testrunner} shows the use of the
-{\verb date } command to calculate the number of seconds it takes it takes for
+{\verb date } command in Bash to calculate the number of seconds it takes for
 Tahoe to put a file to the grid.
@@
 \section{SUPERCOP benchmarks}
-\label{sec:results:supercop}
 The results of the \ac{SHA}-3 256-bit functions and \ac{SHA}-256 benchmarking in
-\ac{SUPERCOP} can be seen in tables \ref{tbl:supercop:long},
-\ref{tbl:supercop:4096}, \ref{tbl:supercop:1536}, \ref{tbl:supercop:576},
-\ref{tbl:supercop:64} and \ref{tbl:supercop:8}. The lower the numbers, the
-faster and better the algorithms performed in the test.
-
-\input{tables/supercop_long_msgs}
-\input{tables/supercop_4096}
-\input{tables/supercop_1536}
-\input{tables/supercop_576}
-\input{tables/supercop_64}
-\input{tables/supercop_8}
+\ac{SUPERCOP} can be found in \autoref{app:supercop:bench}.
 
@@
 in Tahoe-\ac{LAFS} if we had managed to obtain the speeds of the fastest
-implementations in \{SUPERCOP}. The results can be seen in appendix
+implementations in \ac{SUPERCOP}. The results can be seen in appendix
 \ref{appendix:optimzed_sha3}.
@@
 \ac{SUPERCOP} implementation for each candidate in Tahoe-\ac{LAFS}. The
-calculations uses medians from Section \ref{sec:results:supercop} for the speed
+calculations uses medians from \autoref{app:supercop:bench} for the speed
 of each candidate, the exact sizes for each call to the hash function from
-Section \ref{sec:results:tahoedata}, and interpolation to estimate the time to
+\autoref{sec:results:tahoedata}, and interpolation to estimate the time to
 hash a length not recorded by \ac{SUPERCOP}. For lengths above 4096B, the
@@
 e.g. for Shavite3 are worse than the result we achieved in our measurements,
-this is probably due to interpollation being an inaccurate method to use for
+this is probably due to interpolation being an inaccurate method to use for
 the shorter lengths of data.
@@
 
-% Use ordinary \chapters from here on..
+\chapter{Results from SUPERCOP benchmark}
+\label{app:supercop:bench}
+
+Tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
+\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
+\ref{tbl:supercop:8} exhibits the results of the \ac{SHA}-3 256-bit functions
+and \ac{SHA}-256 benchmarking in \ac{SUPERCOP}. The lower the numbers, the
+faster and better the algorithms performed in the test.
+
+\input{tables/supercop_long_msgs}
+\input{tables/supercop_4096}
+\input{tables/supercop_1536}
+\input{tables/supercop_576}
+\input{tables/supercop_64}
+\input{tables/supercop_8}
 

### 2010-12-15 paper/main.tex
Thoughts on measurements discussion updated

--- 2010-12-15 paper/main.tex
@@
 of how fast the different candidate implementations are, we do have some
-comments on it.
+general comments on the way it works and how it relates to our results.
 
@@
 \paragraph{Extra compiler flags}
-
 We do not compile the candidate implementation with the exact same arguments
@@
 \subsection{Small files}
-
-For small files, 1B and 1KB, we can clearly see that both double \ac{SHA}-256
-greatly outperforms the \ac{SHA}-3 candidate implementations. The data being
-hashed for these files are dominated by lengths of 63 and 106 bytes. By
-comparing with \ac{SUPERCOP} results, and especially the fact that we are using
-double \ac{SHA}-256, which should result in large overhead this is unexpected.
-However, we believe that the results for small files may also reflect our use of
-Cython and our skill. Given large amounts of call to our bindings, while the
-data processed by the actual hash functions is small and therefore does not use
-a lot of time, the speed of the bindings will come in to play.
+The results for small files, 1KB and 1KB, are presented in Figure
+\ref{fig:graph:1001b} and \ref{fig:graph:1001kb}. We can clearly see that in
+this case the \ac{SHA}-256 implementation used by Tahoe-\ac{LAFS} outperforms
+our implementations, which is somewhat unexpected given the \ac{SUPERCOP}
+results. However, if we take into account the overhead created by our bindings
+to the underlying implementations (Table \ref{tbl:cython:overhead}), some of the
+\ac{SHA}-3 candidate implementations actually outperform double \ac{SHA}-256,
+which is expected. We do not know how much overhead the bindings for
+\ac{SHA}-256 are causing, but we can safely assume that the overhead is less
+than for our bindings. If we can eliminate all time spent in our bindings both
+BLAKE and \ac{BMW} outperform double \ac{SHA}-256 in every measurement, with
+Luffa, Shabal and Simd outperforming double \ac{SHA}-256 in most of them.
+
+We also measured single \ac{SHA}-256 for these files, which we believe should
+show a difference. The lengths of processed by the hash function are dominated
+by sized of 106B for the 1B files, and also by 63B for the 1KB files (Table
+\ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datadownloaded}. Even by the
+best-case performance scenario from Table \ref{tbl:supercop:sha256d} we would
+expect to see a performance penalty. However this performance penalty is so
+small that it is typically hidden due to system noise.
 
 \subsection{Large files}
-For large files, 1MB, 100MB and 1GB, some of the \ac{SHA}-3 candidates are
-actually outperforming double \ac{SHA}-256, which is expected from the
-\ac{SUPERCOP} results. However if we compare this result with the result for
-small files, and expect that a hash algorithm should do better on a cycles/byte
-basis when hashing larger data values this is not expected. Data sizes which
-are hashed during 1GB and 100MB up/downloads are dominated by 101B, 65569B and
-131106B. 50\% of the hashes are done on 101B, which is a size which should be
-largely affected by having to do double SHA-256. However, since the small files
-had a similar percentage of operations being done on 106B chunks, blaming need
-to use double \ac{SHA}256 is not enough. It looks like indeed the large data
-chunks being hashed is what is making the difference. If we now consider that
-our bindings to the hash functions is what is causing the slow results for
-small files, this makes sense. Larger amounts of data will take longer to
-process for the hash function, which again results in the fraction of time used
-by our bindings is decreased.
-
-\subsection{Impact of double SHA-256}
-The impact of using double \ac{SHA}-256 compared to single is surprisingly
-small, even if you use the best-case estimate from Table
-\ref{tbl:supercop:sha256}. For 1B, 1KB and 1MB files the data hashed is almost
-exclusively 106B or less. If we assume data is hashed at the performance of 576
-bytes, single \ac{SHA}-256 should outperform double \ac{SHA}-256 by about 10\%.
-For 1B files this difference is eaten up by external factors (other
-processes), the same goes for 1KB files. For 1MB files, there seems to be a
-difference, however this difference is so small that we deem it unimportant.
-For the larger files the performance penalty is even smaller, so all in all we
-deem the performance penalty for Tahoe-\ac{LAFS} for using double \ac{SHA}-256
-as opposed to single unimportant.
+For large files, 1MB, 100MB and 1GB, we can see that our implementations are
+outperforming double \ac{SHA}-256. Also for the large files we can see there is
+a performance penalty by our bindings (Table \ref{tbl:cython:overhead}),
+however the relative overhead cause are small compared to the time spent in the
+underlying implementations. The reason for this is that we are hashing larger
+sizes of data, especially for the 100MB and 1GB files (Tables
+\ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datauploaded}). For these
+larger sizes the overhead caused by using double \ac{SHA}-256 opposed to single
+is also relatively small and does not cause a noteworthy performance penalty.
 
@@
 that outperforms, or are close to outperforming \ac{SHA}-256 are BLAKE,
-\ac{BMW}, Shabal, Simd and Luffa. This is expected and corresponds with the
-results from \ac{SUPERCOP}.
+\ac{BMW}, Shabal, Simd and Luffa.
 

### 2010-12-15 paper/main.tex

--- 2010-12-15 paper/main.tex
@@
 function available for all implementations, and that \ac{SUPERCOP} only
-measures speed of hashing full byte strings. This is also a hassle for a
-programmer who wants to use the implementations in \ac{SUPERCOP}
+measures speed of hashing full byte strings. We are questioning if this
+discriminates implementations which actually can do more of their processing on
+a call to {\verb Update() }. This is also a hassle for a programmer who wants
+to use the implementations found in \ac{SUPERCOP}
 
 \paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
-mean implementations that does not conform to the \ac{KAT}s supplied.
-
-\paragraph{Results are only valid for one specific computer}. The results we
-used from \ac{SUPERCOP} are only valid for our specific computer, although it
-gives an indication of which speeds can be achieved on similar computers, i.e.
-32-bit. However to get a complete picture of the performance of the different
-candidates one must look at different computers and architectures. This picture
-might be obtained from looking at the results from all computers measured by
-\ac{SUPERCOP}.
+mean implementations that does not conform to the \ac{KAT}s supplied or
+implementations which fail completely as hash functions, see section
+\ref{sec:choosen_implementations} for details on the JH implementation.
 
 \paragraph{Extra compiler flags}
-
 We do not compile the candidate implementation with the exact same arguments

### 2010-12-15 paper/main.tex
Added comments on candidates and final round candidates

@@
 that outperforms, or are close to outperforming \ac{SHA}-256 are BLAKE,
-\ac{BMW}, Shabal, Simd and Luffa.
+\ac{BMW}, Shabal, Simd and Luffa. The same is true if we eliminate the overhead
+cause by our bindings for the small files. That only these five candidates
+manages to outperform is expected from the \ac{SUPERCOP} results, REF, however
+the order and proportion of the measurements does not. This is primarily
+because we have not used the fastest implementation for all candidates, and the
+relative speed of the chosen implementation compared to the fastest
+implementation in \ac{SUPERCOP} differs (Table \ref{tbl:sha3:relativespeed}.
+The performance of each candidate compared to the \ac{SHA}-256 version found in
+Tahoe-\ac{LAFS} is also somewhat slower than you would expect. We blame this
+mainly on our Cython-bindings and our skill. We believe a developer with more
+knowledge of the inner workings of the implementations, the Python C-\ac{API}
+and \ac{GCC} could have produced better results.
 
@@
 BLAKE, Keccak, Skein, Grøstl and JH have made it through to the final round of
-the \ac{SHA}-3 competition. %TODO: some comments on theese
+the \ac{SHA}-3 competition. From a performance perspective only BLAKE actually
+outperform \ac{SHA}-256 in both the \ac{SUPERCOP} results and our measurements
+for Tahoe-\ac{LAFS}. The other candidates are perhaps given an unfair treatment
+in this paper, with the implementations measures being relatively slow compared
+to the best implementations. However the reason for this is mentioned in our
+criteria for selecting implementations. Another obvious reason is that none of
+the other four finalists outperform \ac{SHA}-256 and therefore it would be more
+valuable for Tahoe-\ac{LAFS} to see the possible performance increase by
+switching to a faster hash function implementation.
 

### 2010-12-15 paper/main.tex
Removed old optimized results
QA till Discussion. Added \text in math environ

@@
 \hypersetup{pdfborder=0 0 0} % ..especially without red borders
+\usepackage{amstext} % To support \text in math mode
 
@@
 \section{SUPERCOP benchmarks}
+
 The results of the \ac{SHA}-3 256-bit functions and \ac{SHA}-256 benchmarking in
@@
 
+BLAKE, \ac{BMW}, and Luffa outperform \ac{SHA}-256 on cycles per
+byte for all tests. For larger input sizes, Simd and Luffa also do the
+calculations quicker than \ac{SHA}-256.
+
 \subsection{Double SHA-256 Estimates}
+
 \ac{SUPERCOP} does not measure the speed of double \ac{SHA}-256, so we have no exact
@@
 
-$$\frac{median * size + 179.62*32}{size} = SHA256d\ estimate (worst-case)$$
-
-$$\frac{median * size + 38.12*32}{size} = SHA256d\ estimate (best-case)$$
-
-$$179.62 + \frac{38.12-179.62}{64-8}*(32-8) = 118.97 (cycles/byte for 32 byte)$$
-
-$$\frac{median * size + 118.97*32}{size} = SHA256d\ estimate (interpolation)$$
+$$\frac{median * size + 179.62*32}{size} = \text{SHA256d estimate (worst case)}$$
+
+$$\frac{median * size + 38.12*32}{size} = \text{SHA256d estimate (best case)}$$
+
+$$179.62 + \frac{38.12-179.62}{64-8}*(32-8) = 118.97 \text{(cycles/byte for 32 byte)}$$
+
+$$\frac{median * size + 118.97*32}{size} = \text{SHA256d estimate (interpolation)}$$
 
@@
 Interpolation gives a somewhat pessimistic result of the cycles/byte needed for
-hashing 32 byte, so the probable speed is somewhere in between the
-interpolation point and the best case time.
+producing the digest of 32 bytes, so the probable speed is somewhere in between the
+interpolation point and the best-case time.
 
@@
 \section{Measurements of Tahoe operations Using Test Vectors}
-The sum of the amount of time it took for the underlying hash library to
-calculate the hash operations are exhibited in Figures \ref{fig:graph:1001b},
+
+The sum of the amount of time it took for the underlying hash libraries to
+produce the digests are exhibited in Figures \ref{fig:graph:1001b},
 \ref{fig:graph:1001kb}, \ref{fig:graph:501mb}, \ref{fig:graph:5100mb} and
@@
 \section{Eliminating overhead in measurements}
-The following measurements were done to eliminate the overhead introducer by
+
+The following measurements were done to eliminate the overhead introduced by
 our work.
 
-\paragraph{Cython bindings}
+\paragraph{Cython bindings.}
 To eliminate as much as possible of overhead introduced by our bindings, we
@@
 
-\paragraph{Modification to Tahoe-\ac{LAFS}}
-Because we write to two files after each hash calculation, this will create an
-overhead in the measurements that would not normaly exist in Tahoe-\ac{LAFS}.
-This overhead should be about the same for all candidates, since the number of
-writes are equal. The results and calculations for this overhead, based on runs
-with double \ac{SHA}-256 can be seen in Table \ref{tbl:measure:overhead}.
+\paragraph{Modification to Tahoe-\ac{LAFS}.}
+Because we write to two files after each call to {\verb digest }, an overhead in
+the measurements that would not normally exist in Tahoe-\ac{LAFS}, will be
+created. This overhead should be about the same for all candidates, since the
+number of writes are equal. The results and calculations for this overhead,
+based on runs with double \ac{SHA}-256, can be seen in Table
+\ref{tbl:measure:overhead}.
 
@@
 
-\subsection{Optimizing results}
-We have not used the fastest possible implementations possible for various
-reasons mentioned previously. However, if we had actually used the fastest
-implementations in \ac{SUPERCOP}, achieved median speed for every hash
-
 \subsection{Theoretical optimized results}

### 2010-12-15 paper/main.tex
\ac{foo}s -> \acp{foo}

@@
 \ac{SHA}-256 algorithm to verify the integrity of files and shares, in addition
-to creating keys, \ac{URI}s and more \cite{t_tahoe}.
+to creating keys, \acp{URI} and more \cite{t_tahoe}.
 
@@
 implementations of Shavite3 in \ac{SUPERCOP}-20101014 or later seems to fail to
-verify the \ac{KAT}s, while the version we have used does. We believe this is
-only because newer \ac{KAT}s will have to be generated, but since the author
+verify the \acp{KAT}, while the version we have used does. We believe this is
+only because newer \acp{KAT} will have to be generated, but since the author
 has not done so, we chose to use an older version of the algorithm.
@@
 \paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
-mean implementations that does not conform to the \ac{KAT}s supplied or
+mean implementations that does not conform to the \acp{KAT} supplied or
 implementations which fail completely as hash functions, see section

### 2010-12-15 paper/main.tex
Some corrections to discussion

--- 2010-12-15 paper/main.tex
@@
 \subsection{The candidates}
+
 If we look primarily at the large file up/download we see that the candidates
-that outperforms, or are close to outperforming \ac{SHA}-256 are BLAKE,
+that outperforms, or are close to outperforming double \ac{SHA}-256 are BLAKE,
 \ac{BMW}, Shabal, Simd and Luffa. The same is true if we eliminate the overhead
 cause by our bindings for the small files. That only these five candidates
-manages to outperform is expected from the \ac{SUPERCOP} results, REF, however
-the order and proportion of the measurements does not. This is primarily
-because we have not used the fastest implementation for all candidates, and the
-relative speed of the chosen implementation compared to the fastest
-implementation in \ac{SUPERCOP} differs (Table \ref{tbl:sha3:relativespeed}.
-The performance of each candidate compared to the \ac{SHA}-256 version found in
-Tahoe-\ac{LAFS} is also somewhat slower than you would expect. We blame this
-mainly on our Cython-bindings and our skill. We believe a developer with more
-knowledge of the inner workings of the implementations, the Python C-\ac{API}
-and \ac{GCC} could have produced better results.
+manages to outperform double \ac{SHA}-256 is expected from the \ac{SUPERCOP}
+results (\autoref{app:supercop:bench}) however the order and proportion of the
+measurements is not. This reason for another order and relative speed is
+primarily because we have not used the fastest implementation for all
+candidates, and the relative speed of the chosen implementation compared to the
+fastest implementation in \ac{SUPERCOP} differs (Table
+\ref{tbl:sha3:speedrelative}.  The performance of each candidate compared to
+the \ac{SHA}-256 version found in Tahoe-\ac{LAFS} is also somewhat slower than
+you would expect. We blame this mainly on our Cython-bindings and our skill. We
+believe a developer with more knowledge of the inner workings of the
+implementations, the Python C-\ac{API} and \ac{GCC} could have produced better
+results.
 
@@
 
-\subsection{Optimizing results}
-We have not used the fastest possible implementations possible for various
-reasons mentioned previously. However, if we had actually used the fastest
-implementations in \ac{SUPERCOP}, achieved median speed for every hash
-
 \subsection{Theoretical optimized results}
+
 We have not use the fastest available implementations for all of the
@@
 in Tahoe-\ac{LAFS} if we had managed to obtain the speeds of the fastest
-implementations in \ac{SUPERCOP}. The results can be seen in appendix
-\ref{appendix:optimzed_sha3}.
+implementations in \ac{SUPERCOP}. The results can be seen in
+\autoref{appendix:optimzed_sha3}.
 

### 2010-12-15 paper/main.tex

--- 2010-12-15 paper/main.tex
@@
 \subsection{Theoretical optimized results}
+
 We have not use the fastest available implementations for all of the

### 2010-12-15 paper/main.tex
Added discussion of truncating and different platforms

@@
 
+\subsection{Truncate output from 512-bit implementations}
+In our measurements and Python library we only consider 256-bit version of each
+candidate. Some candidates does however have a faster implementation with
+512-bit output. We could have used these implementations and truncated the
+output to 256-bit, and possible achieved better results.
+
+\subsection{Testing on difference architectures}
+Our measurements are only done on a single computer with an 32-bit (x86)
+system. It would have been possible to extend the survey to other
+architectures, to see how this would affect the results.
+
 %**************************************%

### 2010-12-15 paper/main.tex
Explain position independant code

@@
 
+The implementation we chose must also work after being compiled with the {\verb
+-fPIC} flag in \ac{GCC}. This flag basically tells the compiler that the
+machine code must be able to execute anywhere in memory, also known as position
+independant. In runtime one extra \{CPU} register will be used for this. This
+is needed because a Python C-module we will need to be compiled as a shred
+library.
+
 Since Tahoe-\ac{LAFS} is using \ac{SHA}-256 internally, we only test \ac{SHA}-3
@@
 \paragraph{Criterion 3} The implementation must not use assembly that can not be made
-    position independent by \ac{GCC}.%FIXME: State the reason for this somewhere?
+    position independent by \ac{GCC}.
 \paragraph{Criterion 4} The implementation should conform to the \acp{KAT} specified in the

### 2010-12-15 paper/main.tex
QA to 5.3.3

@@
 \label{sec:chosen_implementations}
+
 \input{tables/sha3_implementations}
@@
 
-Although we believe the \ac{SUPERCOP} toolkit and results give good indication
-of how fast the different candidate implementations are, we do have some
+Although we believe the \ac{SUPERCOP} toolkit results give good indication of
+how fast the different implementations of the candidates are, we do have some
 general comments on the way it works and how it relates to our results.
 
-\paragraph{Implementations does not conform to \ac{NIST} specification} for
-many of the candidates. By this we mean that there is no {\verb Update() }
+\paragraph{Implementations do not conform to \ac{NIST} specification} for
+many of the candidates. By this we mean that there is no {\verb update() }
 function available for all implementations, and that \ac{SUPERCOP} only
@@
 discriminates implementations which actually can do more of their processing on
-a call to {\verb Update() }. This is also a hassle for a programmer who wants
-to use the implementations found in \ac{SUPERCOP}
+a call to {\verb update() }. This issue is also a hassle for a programmer who
+wants to use the implementations found in \ac{SUPERCOP}.
 
@@
 mean implementations that does not conform to the \acp{KAT} supplied or
-implementations which fail completely as hash functions, see section
-\ref{sec:choosen_implementations} for details on the JH implementation.
-
-\paragraph{Extra compiler flags}
+implementations which fail completely as hash functions, see
+\autoref{sec:chosen_implementations} for details on the JH implementation.
+
+\paragraph{Extra compiler flags.}
+
 We do not compile the candidate implementation with the exact same arguments
-that is used in \ac{SUPERCOP}, most notably we append the -fPIC flag. This flag
-creates \emph{position independent code}, in other words machine instruction
-code that can execute properly regardless of where in memory it resides. This
-will make one less register available on the x86 architecture. While we cannot
-directly prove that this will affect performance of the implementations. We do
-however suspect it. A person writing an optimization for \ac{SUPERCOP} will have
-to balance his use of resources, among them registers, to achieve maximum speed.
-The loss of one register will then give a performance penalty.
+that is used in \ac{SUPERCOP}, most notably we append the {\verb -fPIC } flag.
+This flag creates \emph{position independent code}, in other words machine
+instruction code that can execute properly regardless of where in memory it
+resides. This will make one less register available on the x86 architecture.
+While we cannot directly prove that this will affect performance of the
+implementations, we do however suspect it. A person writing an optimization for
+\ac{SUPERCOP} will have to balance his use of resources, among them registers,
+to achieve maximum speed.  The loss of one register will therefore give a
+performance penalty.
 
@@
 \subsection{Small files}
-The results for small files, 1KB and 1KB, are presented in Figure
+
+The results for small files, 1B and 1KB, are presented in Figures
 \ref{fig:graph:1001b} and \ref{fig:graph:1001kb}. We can clearly see that in
-this case the \ac{SHA}-256 implementation used by Tahoe-\ac{LAFS} outperforms
+this case, the \ac{SHA}-256 implementation used by Tahoe-\ac{LAFS} outperforms
 our implementations, which is somewhat unexpected given the \ac{SUPERCOP}
@@
 which is expected. We do not know how much overhead the bindings for
-\ac{SHA}-256 are causing, but we can safely assume that the overhead is less
-than for our bindings. If we can eliminate all time spent in our bindings both
-BLAKE and \ac{BMW} outperform double \ac{SHA}-256 in every measurement, with
-Luffa, Shabal and Simd outperforming double \ac{SHA}-256 in most of them.
+\ac{SHA}-256 are causing, but we can safely assume that the overhead is far less
+than for our bindings. If we can eliminate all extra time spent in our bindings
+both BLAKE and \ac{BMW} outperform double \ac{SHA}-256 in every measurement,
+with Luffa, Shabal and Simd outperforming double \ac{SHA}-256 in most of them.
 
 We also measured single \ac{SHA}-256 for these files, which we believe should
-show a difference. The lengths of processed by the hash function are dominated
-by sized of 106B for the 1B files, and also by 63B for the 1KB files (Table
-\ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datadownloaded}. Even by the
-best-case performance scenario from Table \ref{tbl:supercop:sha256d} we would
-expect to see a performance penalty. However this performance penalty is so
-small that it is typically hidden due to system noise.
+show a difference. The lengths of inputs processed by the hash function are
+dominated by sized of 106B for the 1B files, and also by 63B for the 1KB files,
+as shown in Tables \ref{tbl:tahoe:datauploaded} and
+\ref{tbl:tahoe:datadownloaded}. Even by the best-case performance scenario from
+Table \ref{tbl:supercop:sha256d} we would expect to see a performance penalty.
+However, this performance penalty is so small that it is typically hidden due to
+system ``noise''.
 
 \subsection{Large files}
-For large files, 1MB, 100MB and 1GB, we can see that our implementations are
-outperforming double \ac{SHA}-256. Also for the large files we can see there is
-a performance penalty by our bindings (Table \ref{tbl:cython:overhead}),
-however the relative overhead cause are small compared to the time spent in the
-underlying implementations. The reason for this is that we are hashing larger
-sizes of data, especially for the 100MB and 1GB files (Tables
-\ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datauploaded}). For these
-larger sizes the overhead caused by using double \ac{SHA}-256 opposed to single
-is also relatively small and does not cause a noteworthy performance penalty.
+
+For large files, 1MB, 100MB and 1GB, we can see that the \ac{SHA}-3
+implementations are outperforming double \ac{SHA}-256. Also for the large files
+we can see that there is a performance penalty introduced by our bindings (Table
+\ref{tbl:cython:overhead}), however the relative overhead are small
+compared to the time spent in the underlying libraries. The reason for
+this is that we are hashing larger sizes of data, especially for the 100MB and
+1GB files (Tables \ref{tbl:tahoe:datauploaded} and
+\ref{tbl:tahoe:datauploaded}). For these larger sizes the overhead caused by
+using double \ac{SHA}-256 opposed to single is also relatively small and does
+not cause a noteworthy performance penalty.
 

### 2010-12-15 paper/main.tex
Added time spent not hashing discussion

--- 2010-12-15 paper/main.tex
@@
 
+\subsection{Time spent not hashing} A strange part of our result is that the
+time Tahoe-\ac{LAFS} spends not callign the hash function is not static for all
+candidates, which we expected.  The results of how much this deviance is
+varies, although it looks to be bounded by some limits that are not the same
+for all candidates. This leads us to believe that the hash function is to
+blame, with noise from the system probably being the cause of the variation.
+The time is recorded right before and right after the calls to our bindings,
+which means that if the implementations are to be blamed they have to do
+something to the system which delays later Tahoe-{LAFS} operations. We only
+speculate in what they are changing and doing differently, but we suspect the
+amount of registers and how much memory and cache the implementation uses as
+probable causes.
+
 \subsection{Theoretical optimized results}
@@
 implementations in \ac{SUPERCOP}. The results can be seen in
-\autoref{appendix:optimzed_sha3}.
+\autoref{appendix:optimized_sha3}.
 
@@
 % How much gain can Tahoe-LAFS get by switching to SHA-3?
-%
-% Drag in the five selected remaining candidates.
 
@@
 The Python-bindings generated for Cython in C are not necessarily the most
-optimized versions possible to write and they are ugly to read or change (from
-a C perspective). However we consider them to be good enough, and they are
-equal for all the \ac{SHA}-3 candidates which means they should give the same
-performance penalty. The \ac{SHA}-256 version will of course not suffer from
-the same penalty, however we believe the possible performance gain is small.
-This is due to the fact that no intensive operations take place in our
-Cython-code. However we do believe that our bindings are affected by the
+optimized versions possible to write and they are ugly to read and change (from
+a C perspective). However they are equal for all the \ac{SHA}-3 candidates
+which means they should give the same performance penalty. The \ac{SHA}-256
+version will of course not suffer from the same penalty. From
+\autoref{tbl:cython:overhead} we can see the performance penalty of our
+bindings, and we do believe better bindings can be made, also with the use of
+Cython, but by someone with more skill and knowledge than us.
 

### 2010-12-15 paper/main.tex
qa to 5.4

@@
 
-If we look primarily at the large file up/download we see that the candidates
-that outperforms, or are close to outperforming double \ac{SHA}-256 are BLAKE,
+If we look primarily at up- and download of the large files, we see that the candidates
+that outperforms or are close to outperforming double \ac{SHA}-256, are BLAKE,
 \ac{BMW}, Shabal, Simd and Luffa. The same is true if we eliminate the overhead
-cause by our bindings for the small files. That only these five candidates
-manages to outperform double \ac{SHA}-256 is expected from the \ac{SUPERCOP}
-results (\autoref{app:supercop:bench}) however the order and proportion of the
-measurements is not. This reason for another order and relative speed is
+introduced by our bindings for the small files.
+
+The fact that only these five candidates manages to outperform double
+\ac{SHA}-256, is expected from the \ac{SUPERCOP} results, ref.
+\autoref{app:supercop:bench}. However, the order and proportion of the
+measurements are not. The reason for another order and relative speed is
 primarily because we have not used the fastest implementation for all
-candidates, and the relative speed of the chosen implementation compared to the
-fastest implementation in \ac{SUPERCOP} differs (Table
-\ref{tbl:sha3:speedrelative}.  The performance of each candidate compared to
-the \ac{SHA}-256 version found in Tahoe-\ac{LAFS} is also somewhat slower than
-you would expect. We blame this mainly on our Cython-bindings and our skill. We
-believe a developer with more knowledge of the inner workings of the
-implementations, the Python C-\ac{API} and \ac{GCC} could have produced better
-results.
+candidates, and the relative speed of the chosen implementations compared to the
+fastest implementation in \ac{SUPERCOP} differs, see Table
+\ref{tbl:sha3:speedrelative}.
+
+The performance of each candidate compared to the \ac{SHA}-256 version found in
+Tahoe-\ac{LAFS} is also somewhat slower than you would expect. We blame this
+mainly on our Cython-bindings and our skill. We believe a developer with more
+knowledge of the inner workings of the implementations, the Python C-\ac{API}
+and \ac{GCC} could have produced better results.
 
 \subsection{Final round candidates}
+
 BLAKE, Keccak, Skein, Grøstl and JH have made it through to the final round of
-the \ac{SHA}-3 competition. From a performance perspective only BLAKE actually
-outperform \ac{SHA}-256 in both the \ac{SUPERCOP} results and our measurements
+the \ac{SHA}-3 competition. From a performance perspective, only BLAKE actually
+outperforms \ac{SHA}-256 in both the \ac{SUPERCOP} results and our measurements
 for Tahoe-\ac{LAFS}. The other candidates are perhaps given an unfair treatment
-in this paper, with the implementations measures being relatively slow compared
-to the best implementations. However the reason for this is mentioned in our
-criteria for selecting implementations. Another obvious reason is that none of
-the other four finalists outperform \ac{SHA}-256 and therefore it would be more
-valuable for Tahoe-\ac{LAFS} to see the possible performance increase by
-switching to a faster hash function implementation.
+in this paper, with the used implementations being relatively slow compared to
+the best existing implementations. However, the reason for this is mentioned in
+our criteria for selecting implementations.
+
+None of the other four finalists outperform \ac{SHA}-256, and therefore it would be
+more valuable for Tahoe-\ac{LAFS} to see the possible performance increase by
+switching to a faster hash function implementation, i.e. BLAKE.
 
@@
 
-We have not use the fastest available implementations for all of the
+We have not used the fastest available implementations for all of the
 candidates, we also know that our bindings creates an overhead for all calls.
@@
 in Tahoe-\ac{LAFS} if we had managed to obtain the speeds of the fastest
-implementations in \ac{SUPERCOP}. The results can be seen in
+implementations in \ac{SUPERCOP}. The results can be found in
 \autoref{appendix:optimized_sha3}.
@@
 
-The Python-bindings generated for Cython in C are not necessarily the most
+The generated C files from Cython used as Python bindings are not necessarily the most
 optimized versions possible to write and they are ugly to read and change (from

### 2010-12-15 paper/main.tex
QA done on discussion

@@
 
-The generated C files from Cython used as Python bindings are not necessarily the most
-optimized versions possible to write and they are ugly to read and change (from
-a C perspective). However they are equal for all the \ac{SHA}-3 candidates
-which means they should give the same performance penalty. The \ac{SHA}-256
-version will of course not suffer from the same penalty. From
-\autoref{tbl:cython:overhead} we can see the performance penalty of our
-bindings, and we do believe better bindings can be made, also with the use of
-Cython, but by someone with more skill and knowledge than us.
-
-The alternative would be creating manual Python-bindings for all the candidates
-which would require a lot more work on our side, especially given the lack of a
-good enough and common \ac{API} for the implementations in \ac{SUPERCOP}.
-Another fact important for this decision was that we did not know exactly which
-implementation of each candidate we would end up using, and we ended up trying
-several for most candidates.
+The generated C files from Cython used as Python bindings are not necessarily
+the most optimized versions possible to write and they are difficult to read and
+change from a C perspective. However, they are equal for all the \ac{SHA}-3
+candidates, which means they should give the same performance penalty. The
+\ac{SHA}-256 version will of course not suffer from this, as it is not written
+using Cython. From \autoref{tbl:cython:overhead} we can see the performance
+penalty of our bindings, and we do believe better bindings can be made, also
+with the use of Cython, but by someone with more skill and knowledge than us.
+
+The alternative would be to create manual Python bindings for all the
+candidates, which would require a lot more work on our side, especially given
+the lack of a good enough and common \ac{API} for the implementations in
+\ac{SUPERCOP}
+
+Another fact important for the decision on using Cython, was that we did not
+know exactly which implementation of each candidate we would end up using.
+Several implementations were tried out during testing, and this could only be
+accomplished since we only had to do a minimal amount of work for each new
+implementation test.
 
 \subsection{Faster implementations}
-We do not claim to have tested the fastest possible nor existing version
+
+We do not claim to have tested the fastest possible existing version
 of most of the candidates. We are certain that someone of more skill and
 knowledge about the inner workings of the \ac{SHA}-3 candidates, \ac{GCC} and
-Python could make significantly faster implementation.
+Python could make significantly faster implementations.
 
 \subsection{Test with different Erasure Coding schemes}
-By default Tahoe-\ac{LAFS} comes with a coding scheme of 3-7-10 (shares needed,
+
+By default, Tahoe-\ac{LAFS} comes with a coding scheme of 3-7-10 (shares needed,
 shares happy, shares total) while we test with a configuration of 2-3-3. It
@@
 \subsection{Measure with real use cases}
-Our test use random generated data of a certain length. In a real use case for
-Tahoe-\ac{LAFS} the data would probably be both of varying length, and of
-lengths not tested by us. In a real use case both a feasible erasure coding
-scheme would have to be decided, and the files would represent what you
-actually would want to use Tahoe-\ac{LAFS} for.
+
+Our test use randomly generated data of a certain length. In a real use case for
+Tahoe-\ac{LAFS}, the data would probably be both of varying length, and of
+lengths not tested by us. Additionally, both a feasible erasure coding
+scheme would have to be decided, and the files would represent what you actually
+would want to use Tahoe-\ac{LAFS} for.
 
 \subsection{Truncate output from 512-bit implementations}
-In our measurements and Python library we only consider 256-bit version of each
+
+In our measurements and in the created Python library, we only consider 256-bit version of each
 candidate. Some candidates does however have a faster implementation with
@@
 \subsection{Testing on difference architectures}
+
 Our measurements are only done on a single computer with an 32-bit (x86)

### 2010-12-15 paper/main.tex
QA on conclusion and appendix

@@
 \section{Tahoe-LAFS}
-%What is it, how/where are hash functions used, Python, pycryptopp
-%comparison with RAID-6, mutable/immutable?
 
@@
 used for both server selection and to index shares within each storage node.
-% More here?
 
@@
 %**************************************%
-
-%A bit more technical description of how we tested the candidates in Tahoe-LAFS
-% - Limitations/assumptions
-% - Optimized candidates.
-% - Error sources.
-% The points above should be taken into consideration on each section in this
-% chapter.
 
@@
 
-% Describe inconsistencies in the numbers, and try to explain them.
-% Time spent not hashing varies because: Variable use of CPU regs?
-%
-% Compare our results with general (SUPERCOP) results.
-%
-% How much gain can Tahoe-LAFS get by switching to SHA-3?
-
 \section{Improvement of methods}
@@
 
-In our measurements and in the created Python library, we only consider 256-bit version of each
-candidate. Some candidates does however have a faster implementation with
-512-bit output. We could have used these implementations and truncated the
-output to 256-bit, and possible achieved better results.
+In our measurements and in the created Python library, we only consider 256-bit
+version of each candidate. Some candidates does however have a faster
+implementation with 512-bit output. We could have used these implementations and
+truncated the output to 256-bit, and possible achieved better results.
 
@@
 
-% This chapter should contain:
-% -----------------------------------------------------------
-% A somewhat short conclusion summing up what has been done, and results and
-% discussion
-
-We had hoped to contribute more to the Open Source project Tahoe-LAFS, when
+We had hoped to contribute more to the Open Source project Tahoe-\ac{LAFS}, when
 starting on our assignment. It was early evident that implementing a new
 capability was out of scope, since creating Python bindings and trying out
-new implementations was prioritized. Future work should implement the support
-for choosing hashing algorithm in a proper manner which ensures backwards
+new implementations were prioritized. Future work should implement the support
+for choosing hashing algorithm in a proper manner, which ensures backwards
 compatibility.
 
-Four of the candidates has been shown to perform better than the currently used
+Four of the candidates have been shown to perform better than the currently used
 SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd. The time it takes for
-Tahoe-LAFS to upload a one gigabyte file, could improve by $5\%$ if the best
-performing SHA-3 candidate, BLAKE, were to be used. One has to bear in mind that
-SHA-256 has been optimized and tweaked for a number of years, and that these
-performance differences undoubtedly will continue to increase as new and better
-implementations will be developed for the final chosen \ac{SHA}-3 algorithm.
-
-Perhaps the most important finding in this project, is that the SUPERCOP
+Tahoe-\ac{LAFS} to upload a one gigabyte file, could improve by $5\%$ if the
+best performing \ac{SHA}-3 candidate, BLAKE, were to be used. One has to bear in
+mind that \ac{SHA}-256 has been optimized and tweaked for a number of years, and
+that these performance differences undoubtedly will continue to increase as new
+and better implementations will be developed for the final chosen \ac{SHA}-3
+algorithm.
+
+Perhaps the most important finding in this project, is that the \ac{SUPERCOP}
 benchmarking indeed drives the development of new and better implementations
-forward, but since the SUPERCOP \ac{API} does not conform to the \ac{NIST}
+forward, but since the \ac{SUPERCOP} \ac{API} does not conform to the \ac{NIST}
 \ac{API}, the developers leave important functionality out (i.e.
@@
 
-% Uncomment to enable appendices.
 \appendix
@@
 Tables \ref{tbl:tahoe:optimized_g} and \ref{tbl:tahoe:optimized_p} exhibit
-results calculated which theoretically could have been achieved using the best
+results calculated, which theoretically could have been achieved using the best
 \ac{SUPERCOP} implementation for each candidate in Tahoe-\ac{LAFS}. The
-calculations uses medians from \autoref{app:supercop:bench} for the speed
-of each candidate, the exact sizes for each call to the hash function from
-\autoref{sec:results:tahoedata}, and interpolation to estimate the time to
-hash a length not recorded by \ac{SUPERCOP}. For lengths above 4096B, the
-numbers for long data lengths are used. The results also assume only one core
-of the reference processor (2,833 GHz) is used. We can see that some of the calculations,
-e.g. for Shavite3 are worse than the result we achieved in our measurements,
-this is probably due to interpolation being an inaccurate method to use for
-the shorter lengths of data.
+calculations uses medians from \autoref{app:supercop:bench} for the speed of
+each candidate, the exact sizes for each call to the hash function from
+\autoref{sec:results:tahoedata}, and interpolation to estimate the time to hash
+a length not recorded by \ac{SUPERCOP}. For lengths above 4096B, the numbers for
+long data lengths are used. The results also assume only one core of the
+reference processor (2,833 GHz) is used. We can see that some of the
+calculations, e.g. for Shavite3, are worse than the result we achieved in our
+measurements, this is probably due to interpolation being an inaccurate method
+to use for the shorter lengths of data.
 

### 2010-12-15 paper/main.tex
Added discussion of possibly gain for Tahoe

--- 2010-12-15 paper/main.tex
@@
 
-If we look primarily at up- and download of the large files, we see that the candidates
-that outperforms or are close to outperforming double \ac{SHA}-256, are BLAKE,
+If we look primarily at the large file up/download we see that the candidates
+that outperforms, or are close to outperforming double \ac{SHA}-256 are BLAKE,
 \ac{BMW}, Shabal, Simd and Luffa. The same is true if we eliminate the overhead
-introduced by our bindings for the small files.
-
-The fact that only these five candidates manages to outperform double
-\ac{SHA}-256, is expected from the \ac{SUPERCOP} results, ref.
-\autoref{app:supercop:bench}. However, the order and proportion of the
-measurements are not. The reason for another order and relative speed is
+cause by our bindings for the small files. That only these five candidates
+manages to outperform double \ac{SHA}-256 is expected from the \ac{SUPERCOP}
+results (\autoref{app:supercop:bench}) however the order and proportion of the
+measurements is not. This reason for another order and relative speed is
 primarily because we have not used the fastest implementation for all
-candidates, and the relative speed of the chosen implementations compared to the
-fastest implementation in \ac{SUPERCOP} differs, see Table
-\ref{tbl:sha3:speedrelative}.
-
-The performance of each candidate compared to the \ac{SHA}-256 version found in
-Tahoe-\ac{LAFS} is also somewhat slower than you would expect. We blame this
-mainly on our Cython-bindings and our skill. We believe a developer with more
-knowledge of the inner workings of the implementations, the Python C-\ac{API}
-and \ac{GCC} could have produced better results.
+candidates, and the relative speed of the chosen implementation compared to the
+fastest implementation in \ac{SUPERCOP} differs (Table
+\ref{tbl:sha3:speedrelative}.  The performance of each candidate compared to
+the \ac{SHA}-256 version found in Tahoe-\ac{LAFS} is also somewhat slower than
+you would expect. We blame this mainly on our Cython-bindings and our skill. We
+believe a developer with more knowledge of the inner workings of the
+implementations, the Python C-\ac{API} and \ac{GCC} could have produced better
+results.
 
 \subsection{Final round candidates}
-
 BLAKE, Keccak, Skein, Grøstl and JH have made it through to the final round of
-the \ac{SHA}-3 competition. From a performance perspective, only BLAKE actually
-outperforms \ac{SHA}-256 in both the \ac{SUPERCOP} results and our measurements
+the \ac{SHA}-3 competition. From a performance perspective only BLAKE actually
+outperform \ac{SHA}-256 in both the \ac{SUPERCOP} results and our measurements
 for Tahoe-\ac{LAFS}. The other candidates are perhaps given an unfair treatment
-in this paper, with the used implementations being relatively slow compared to
-the best existing implementations. However, the reason for this is mentioned in
-our criteria for selecting implementations.
-
-None of the other four finalists outperform \ac{SHA}-256, and therefore it would be
-more valuable for Tahoe-\ac{LAFS} to see the possible performance increase by
-switching to a faster hash function implementation, i.e. BLAKE.
+in this paper, with the implementations measures being relatively slow compared
+to the best implementations. However the reason for this is mentioned in our
+criteria for selecting implementations. Another obvious reason is that none of
+the other four finalists outperform \ac{SHA}-256 and therefore it would be more
+valuable for Tahoe-\ac{LAFS} to see the possible performance increase by
+switching to a faster hash function implementation.
 
@@
 
+\subsection{Possible gain for Tahoe-LAFS}
+
+Even though not many of the
+\ac{SHA}-3 implementation used by us managed to outperform the implementation
+of double \ac{SHA}-256 in Tahoe-\ac{LAFS}, some did for large files, which
+tells us there is a performance gain by switching which hash function is used
+in Tahoe-LAFS. \autoref{tbl:tahoe:gain} compares the performance gains for
+BLAKE, which did best in our measurements, versus \ac{SHA}-256, with overhead
+caused by our measurements removed. We achieved as much as a 12,35\% gain when
+uploading the 100MB files. We can see that the gain is smaller on downloads
+than uploads, this is mostly caused by our Tahoe-\ac{LAFS} configuration, where
+we need to upload shares to three servers, while we only need to download two
+of them to get the original file.
+
+\input{tables/tahoe_gain}
+
+On another note it should be noted that most individuals will not have access
+to the same high bandwidth network that our Tahoe-\ac{LAFS} servers ran on,
+1Gbps or even 100Mbps. For an individual with a normal internet connection the
+network is the bottleneck, and therefore the performance increase with a faster
+hash algorithm implementation would not decrease the transfer time for large
+files. If we consider Tahoe-\ac{LAFS} for organisations and enterprises which
+typically empovers high bandwidth networks the change could indeed be
+noticable.
+
+\subsection{Faster \ac{SHA}-3 implementations will appear}
+
+The obervation should also be made that \ac{SHA}-256 has probably been
+optimized for the better part of a decade, while the \ac{SHA}-3 candidates are
+relatively new. This will probably mean that in the future we will se even
+faster versions of the \ac{SHA}-3 candidates, while it is more probable that
+\ac{SHA}-2 is nearing the limit of its performance.
+
 \subsection{Theoretical optimized results}
 
-We have not used the fastest available implementations for all of the
+We have not use the fastest available implementations for all of the
 candidates, we also know that our bindings creates an overhead for all calls.
@@
 in Tahoe-\ac{LAFS} if we had managed to obtain the speeds of the fastest
-implementations in \ac{SUPERCOP}. The results can be found in
+implementations in \ac{SUPERCOP}. The results can be seen in
 \autoref{appendix:optimized_sha3}.
 
-% Describe inconsistencies in the numbers, and try to explain them.
-% Time spent not hashing varies because: Variable use of CPU regs?
-%
-% Compare our results with general (SUPERCOP) results.
-%
 % How much gain can Tahoe-LAFS get by switching to SHA-3?
@@
 \subsection{The use of Cython}
-
-% TODO: Discuss the BLAKE run in Results when timing was done within Cython.
-
-The generated C files from Cython used as Python bindings are not necessarily the most
+The Python-bindings generated for Cython in C are not necessarily the most
 optimized versions possible to write and they are ugly to read and change (from

### 2010-12-15 paper/main.tex

--- 2010-12-15 paper/main.tex
@@
 
+\subsection{Possible gain for Tahoe-LAFS}
+
+Even though not many of the
+\ac{SHA}-3 implementation used by us managed to outperform the implementation
+of double \ac{SHA}-256 in Tahoe-\ac{LAFS}, some did for large files, which
+tells us there is a performance gain by switching which hash function is used
+in Tahoe-LAFS. \autoref{tbl:tahoe:gain} compares the performance gains for
+BLAKE, which did best in our measurements, versus \ac{SHA}-256, with overhead
+caused by our measurements removed. We achieved as much as a 12,35\% gain when
+uploading the 100MB files. We can see that the gain is smaller on downloads
+than uploads, this is mostly caused by our Tahoe-\ac{LAFS} configuration, where
+we need to upload shares to three servers, while we only need to download two
+of them to get the original file.
+
+\input{tables/tahoe_gain}
+
+On another note it should be noted that most individuals will not have access
+to the same high bandwidth network that our Tahoe-\ac{LAFS} servers ran on,
+1Gbps or even 100Mbps. For an individual with a normal internet connection the
+network is the bottleneck, and therefore the performance increase with a faster
+hash algorithm implementation would not decrease the transfer time for large
+files. If we consider Tahoe-\ac{LAFS} for organisations and enterprises which
+typically empovers high bandwidth networks the change could indeed be
+noticable.
+
+\subsection{Faster \ac{SHA}-3 implementations will appear}
+
+The obervation should also be made that \ac{SHA}-256 has probably been
+optimized for the better part of a decade, while the \ac{SHA}-3 candidates are
+relatively new. This will probably mean that in the future we will se even
+faster versions of the \ac{SHA}-3 candidates, while it is more probable that
+\ac{SHA}-2 is nearing the limit of its performance.
+
 \subsection{Theoretical optimized results}
@@
 \subsection{The use of Cython}
-
-% TODO: Discuss the BLAKE run in Results when timing was done within Cython.
 

### 2010-12-15 paper/main.tex
A little bit country
Moved captions to top in tables, and minor spell corrections

--- 2010-12-15 paper/main.tex
@@
 \paragraph{Cost and Performance.} Cost and Performance are considered to be the
-2nd most important criterion. The absolute minimum for performance is that the
+second most important criterion. The absolute minimum for performance is that the
 SHA-3 candidate should be faster than the functions in the \ac{SHA}-2 family.
@@
 
-\paragraph{Algorithm and implementation characteristics.} By this 3rd criterion,
+\paragraph{Algorithm and implementation characteristics.} By this third criterion,
 \ac{NIST} emphasizes that algorithms with greater flexibility will be given
@@
 test vectors. E.g. to upload the 100 1-byte files, 15700 hashes had to be
-calculated, and about 64\% of these were of size 106 byte.
-
-For every upload Tahoe-\ac{LAFS} does, it hashes a size close to the original
-file size twice. This does not happen when downloading the file.
+calculated, and about 64\% of these were of size 106 byte. For every upload
+Tahoe-\ac{LAFS} also hashes a size close to the original file size twice. This
+does not happen when downloading the file.
 
@@
         {graphs/Timespenthashingwhiledownloading1001Bfiles.pdf}
-    \caption{Figures for hash operations of 100 1B files.}
+    \caption{Graph of timing of hash operations for 100 1B files.}
     \label{fig:graph:1001b}
@@
         {graphs/Timespenthashingwhiledownloading1001KBfiles.pdf}
-    \caption{Figures for hash operations of 100 1KB files.}
+    \caption{Graph of timing of hash operations for 100 1KB files.}
     \label{fig:graph:1001kb}
@@
         {graphs/Timespenthashingwhiledownloading501MBfiles.pdf}
-    \caption{Figures for hash operations of 50 1MB files.}
+    \caption{Graph of timing of hash operations for 50 1MB files.}
     \label{fig:graph:501mb}
@@
         {graphs/Timespenthashingwhiledownloading5100MBfiles.pdf}
-    \caption{Figures for hash operations of five 100MB files.}
+    \caption{Graph of timing of hash operations for five 100MB files.}
     \label{fig:graph:5100mb}
@@
         {graphs/Timespenthashingwhiledownloading11GBfile.pdf}
-    \caption{Figures for hash operations of one 1GB file.}
+    \caption{Graph of timing of hash operations for one 1GB file.}
     \label{fig:graph:11gb}
@@
 
-\section{Exclusion of mutable file testing}
+\section{Exclusion of mutable files}
 
@@
 
+\subsection{Possible gain for Tahoe-LAFS}
+
+Even though few of the \ac{SHA}-3 implementations used by us managed to
+outperform the implementation of double \ac{SHA}-256 in Tahoe-\ac{LAFS}, some
+did for large files, which tells us there is a performance gain by switching
+cryptographic hash function in Tahoe-\ac{LAFS}. \autoref{tbl:tahoe:gain}
+compares the performance gains for BLAKE, which performed the best in our
+measurements, versus \ac{SHA}-256, with the removal of overhead caused by our
+measurements. We achieved as much as a 12.35\% gain when uploading the 100MB
+files. The gain is smaller on downloads than uploads, this is
+mostly caused by our Tahoe-\ac{LAFS} configuration, where we need to upload
+shares to three servers, while we only need to download two of them to get the
+original file.
+
+\input{tables/tahoe_gain}
+
+On another note, most individuals will not have access to the same high
+bandwidth network that our Tahoe-\ac{LAFS} servers ran on, 1Gbps, or even
+100Mbps. For an individual with a normal Internet connection, the network is the
+bottleneck, and therefore the performance increase with a faster hash algorithm
+implementation would not decrease the transfer time for large files. If we
+consider Tahoe-\ac{LAFS} for organisations and enterprises which typically
+empowers high bandwidth networks, the change could indeed be noticeable.
+
+\subsection{Faster SHA-3 implementations will appear}
+
+The observation should also be made that \ac{SHA}-256 has probably been
+optimized for the better part of a decade, while the \ac{SHA}-3 candidates are
+relatively new. This will probably mean that in the future we will see even
+faster versions of the \ac{SHA}-3 candidates, while it is more probable that
+\ac{SHA}-2 is nearing the limit of its performance.
+
 \subsection{Theoretical optimized results}

### 2010-12-15 paper/main.tex
Fixed sphlib reference to bmw, added sphlib description and removed supercop time column

--- 2010-12-15 paper/main.tex
@@
 
+\paragraph{Sphlib} is a C and Java library made by Thomas Pornin
+\footnote{\url{http://www.saphir2.com/sphlib/}}, which contains all the second
+round \ac{SHA}-3 candidates. Some of the same implementations are also found in
+\ac{SUPERCOP}, while others, like the Cubehash implementation we have used is
+not.
+
 The relative speeds of the implementations chosen compared to

### 2010-12-15 paper/main.tex
fixed typos

@@
 tried out in the Tahoe-LAFS file system using automated tests to measure the
-performance differences. Only the immutable file support was tested.
+differences in performance. Only immutable file support was tested.
 
@@
 SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd.  By utilizing the
-performance improvements given by SHA-3, Tahoe-LAFS could currently save $5\%$
-on upload operations by changing hashing function to the best performing
-candidate.
+performance improvements given by SHA-3, we show that Tahoe-LAFS could
+currently save up to $12,36\%$ on upload operations by changing the best 
+performing candidate.
 
@@
 \acro{NIST}{National Institute of Standards and Technology}
+\acro{RAID}{Redundant Array of Independent Disks}
 \acro{RAM}{Random Access Memory}
-\acro{RAID}{Redundant Array of Independent Disks}
 \acro{SHA}{Secure Hash Algorithm}
@@
 \acro{SVN}{Subversion}
-\acro{UEB}{URI Extension Block}
+\acro{UEB}{\ac{URI} Extension Block}
 \acro{URI}{Uniform Resource Identifier}
@@
 The process of selecting what is to become the new standard for a secure hashing
-algorithm is carried out by the {NIST}, and is at its final stages at the time
+algorithm is carried out by the {NIST}, and is in its final stages at the time
 of writing. Fourteen candidates came through to the second round
@@
 
-In this project assignment, we integrate the 14 \ac{SHA}-3 candidates in
-the Tahoe-\ac{LAFS} and measure the performance impact. To be able to do this,
-we create Python bindings for the best suited C implementations of the SHA-3
-candidates, and alter the hash function calls in the source code of Tahoe.
+In this project assignment, we integrate the 14 second round \ac{SHA}-3
+candidates in the Tahoe-\ac{LAFS} and measure the performance impact. To be
+able to do this, we create Python bindings for the best suited C
+implementations of the SHA-3 candidates, and alter the hash function calls in
+the source code of Tahoe.
 
@@
 \subsection{NIST SHA-3 Competition}
+
 The \ac{SHA} version 3 is a coming standard set to supersede the current
@@
 function that will be known as \ac{SHA}-3 will be decided by the \ac{NIST} and
-chosen between the submitted contestants to the \ac{NIST} hash competition.
-At the time of writing, the current status of the competition is officially
-called Round 2, with 14 of 64 candidates having ``survived'' Round 1
-\cite{s_fedreg}.
+chosen between the submitted contestants to the \ac{NIST} hash competition.  At
+the of writing, the current status of the competition is in it's final round
+(or round 3), with only five remaining candidates. This result was not revealed
+until a week before the submission of this paper, so we are focusing on all
+round 2 candidates, which there are 14 of.  \cite{s_fedreg}.
 

### 2010-12-15 paper/main.tex
typos

@@
 chosen between the submitted contestants to the \ac{NIST} hash competition.  At
-the of writing, the current status of the competition is in it's final round
-(or round 3), with only five remaining candidates. This result was not revealed
-until a week before the submission of this paper, so we are focusing on all
-round 2 candidates, which there are 14 of.  \cite{s_fedreg}.
+the  time of writing, the competition is in its final round (round 3), with
+only five remaining candidates. This result was not revealed until a week
+before the submission of this paper, so we are focusing on all round 2
+candidates, which there are 14 of \cite{s_fedreg}.
 
@@
 \paragraph{Security.} The most important criterion for the SHA-3
-candidates\cite{s_nistround2} is security. \citet{s_fedreg} lays out the full
+candidates is security\cite{s_nistround2}. \citet{s_fedreg} lays out the full
 description of security criteria, however the most noteworthy criterion
@@
 \paragraph{Cost and Performance.} Cost and Performance are considered to be the
-second most important criterion. The absolute minimum for performance is that the
-SHA-3 candidate should be faster than the functions in the \ac{SHA}-2 family.
-Cost is a measure of how much memory an implementation requires in software,
-both the implementation itself and the use of \ac{RAM} during runtime. Another
-measure of cost is how many logic gates it takes to implement the function in
-hardware.
+second most important criterion. The absolute minimum for performance is that
+the SHA-3 candidate should be faster than the functions in the \ac{SHA}-2
+family\cite{s_nistround2}.  Cost is a measure of how much memory an
+implementation requires in software, both the implementation itself and the use
+of \ac{RAM} during runtime. Another measure of cost is how many logic gates it
+takes to implement the function in hardware.
 
@@
 make it ready to start processing real data. The {\verb Update() } function is used to
-provide the actual data to the hash function. The {\verb Final() } function should be
+provide the actual data that will be hashed. The {\verb Final() } function should be
 called when all necessary data has been given through {\verb Update() }, it will then
-finish the hash function and provide the digest. The {\verb Hash() } function
+finish the processing and provide the digest. The {\verb Hash() } function
 can be described as a wrapper for doing a call to the three other functions.
@@
 others. The toolkit will also try different compiler optimizations to get the
-best results possible for each function.
+best results possible for each implementation.
 
@@
 data types and that the {\verb crypto_hash() } function does not need an input
-of how long the output of the hash function should be. Instead double or
-different submissions are used for different output lengths.
+of how long the output of the hash function should be.
 
@@
 \cite{t_ars}.  It is written in the Python programming language with the Twisted
-framework, and can run on Windows, Mac OSX, Linux, Solaris and more.
+framework, and can run on Windows, Mac OSX, Linux, Solaris and other operating
+systems.
 
@@
 for its children, i.e. subdirectories or files. Two forms of capabilities is
-available for each file, read-only and read-write, and these can be shared to
-provide shared/published directory structures with friends.
+available for each file, read-only and read-write, and these can be distributed
+to e.g. share a file with friends.
 
@@
 
-A \textbf{capability} (or an URI) contains the encryption key, and a hash of the
-\emph{\ac{UEB}}. The \ac{UEB} for each file is a data structure containing the
-hash of the \ac{CEB}, the size of the file, any encoding parameters necessary to
-perform the erasure decoding, and a hash of the plaintext and the encrypted
+A \textbf{capability} (or an URI) contains the encryption key, and a hash of
+the \ac{UEB}. The \ac{UEB} for each file is a data structure containing the
+hash of the \ac{CEB}, the size of the file, any encoding parameters necessary
+to perform the erasure decoding, and a hash of the plaintext and the encrypted
 text. This is illustrated in Figure \ref{fig:tahoeueb}.
@@
 Python\footnote{\url{http://www.python.org/}} is a high-level general-purpose
-programming language. Python is also an interpreted language, which means that
+programming language. Python is also an just-in-time language, which means that
 Python programs are compiled at runtime. There exists multiple implementations

### 2010-12-15 paper/main.tex
Typos

@@
 
-The implementation we chose must also work after being compiled with the {\verb
--fPIC} flag in \ac{GCC}. This flag basically tells the compiler that the
+The implementation we chose must also work after being compiled with the {\verb -fPIC }
+flag in \ac{GCC}. This flag basically tells the compiler that the
 machine code must be able to execute anywhere in memory, also known as position
-independant. In runtime one extra \{CPU} register will be used for this. This
-is needed because a Python C-module we will need to be compiled as a shred
+independent. In runtime one extra \ac{CPU} register will be used for this. This
+is needed because a Python C module we will need to be compiled as a shared
 library.
@@
 Since Tahoe-\ac{LAFS} is using \ac{SHA}-256 internally, we only test \ac{SHA}-3
-functions with 256-bit output in our implementation.
+functions with 256-bit output to avoid problems that could be caused by longer 
+hash values in Tahoe-\ac{LAFS}.
 
@@
 \paragraph{Criterion 2} The implementation must include a working {\verb update() } function.
-\paragraph{Criterion 3} The implementation must not use assembly that can not be made
+\paragraph{Criterion 3} The implementation must not use code that can not be made
     position independent by \ac{GCC}.
@@
 each implementation, in relation to other implementation for the same and other
-candidates are available at source.
+candidates can be found at eBASH\cite{s_tahoe_supercop}.
 
@@
 in the GCC manual\footnote{GCC Manual \url{http://gcc.gnu.org/onlinedocs/}},
-and the one used for every implementation which might affect performance, are:
+and the flags used for every implementation which might affect performance, are:
 
@@
 to {\verb Update("foo") }. However, the implementation in sphlib does fulfill
-this requirement. We have contacted the author, Thomas Pornin, which confirms
+this requirement. We have contacted the author of sphlib, Thomas Pornin, which confirms
 this problem with the official JH candidate, and that the author has been made
@@
 implementations in \ac{SUPERCOP}. Together with the fact that Tahoe-\ac{LAFS}
-uses 256-bit output, in the end only included support for 256-bit output in
+uses 256-bit output, in the end only support for 256-bit output in
 SHA3lib was implemented. The design we have used is shown in Figure
@@
 Basically the class \textbf{echo} resides in Cython and in C, and is responsible
-for bringing the C functionality of the C implementation to Python. By compiling
+for bringing the functionality of the \ac{SHA}-3 candidate implementation to Python. By compiling
 \textbf{echo} it will be made available for use in Python.  The class
@@
 functions that can be accessed from both languages. While this is just the
-implementation for the Echo candidate, the almost exact same pattern is used for
+implementation for the Echo implementation, the almost exact same pattern is used for
 all of the other candidates.
@@
 \subsection{Unit tests}
-We Unit Test each new candidate implementation that is included SHA3lib, both
-to verify our own work and to verify that the implementation from \ac{SUPERCOP}
+We Unit Test each new candidate implementation that is included in SHA3lib, both
+to verify our own work and to verify that the implementations from \ac{SUPERCOP}
 actually are correct.
@@
 couple of classes and functions, and the connection to the underlying hash
-function library is made on only one line. This is displayed in Listing
+function library is made in only one line. This is displayed in Listing
 \ref{lst:hashutil}.
@@
 
-\section{Automatic testing and distribution}
-We fully automate the testing procedure. This is achieved by the use of
-software revision control systems, which distributes both the source code of
-Tahoe-\ac{LAFS}, SHA3lib and shell-scripts for installation and configuration of
-both. Configuration is mainly done by the use of regular expressions.
-
-The outline of this procedure is:
+\section{Automatic testing and distribution} We fully automate the testing
+procedure. This is achieved by the use of software revision control systems,
+which distributes both the source code of Tahoe-\ac{LAFS}, SHA3lib and
+shell-scripts for installation and configuration of both. Configuration is
+mainly done by the use of regular expressions.The outline of this procedure is:
 
@@
 BLAKE, \ac{BMW}, and Luffa outperform \ac{SHA}-256 on cycles per
-byte for all tests. For larger input sizes, Simd and Luffa also do the
+byte for all tests. For larger input sizes, Simd and Shabal also do the
 calculations quicker than \ac{SHA}-256.

### 2010-12-15 paper/main.tex
Proper order of tables reffing

@@
 The numbers for uploading and downloading the different file vectors are
-summarized in Tables \ref{tbl:hashingtimes:put1gb},
-\ref{tbl:hashingtimes:get1gb}, \ref{tbl:hashingtimes:put100mb},
-\ref{tbl:hashingtimes:get100mb}, \ref{tbl:hashingtimes:put1mb},
-\ref{tbl:hashingtimes:get1mb}, \ref{tbl:hashingtimes:put1kb},
-\ref{tbl:hashingtimes:get1kb}, \ref{tbl:hashingtimes:put1b} and
-\ref{tbl:hashingtimes:get1b}.  The third column contains the time spent doing
-other operations than hashing, i.e. \emph{Time spent not hashing = Total time -
-Time spent hashing}. \emph{Fraction} displays the part of total time the calls
-to the hashing libraries take.
+summarized in Tables \ref{tbl:hashingtimes:put1b}, \ref{tbl:hashingtimes:get1b},
+\ref{tbl:hashingtimes:put1kb}, \ref{tbl:hashingtimes:get1kb},
+\ref{tbl:hashingtimes:put1mb}, \ref{tbl:hashingtimes:get1mb},
+\ref{tbl:hashingtimes:put100mb}, \ref{tbl:hashingtimes:get100mb},
+\ref{tbl:hashingtimes:put1gb} and \ref{tbl:hashingtimes:get1gb}. The third
+column contains the time spent doing other operations than hashing, i.e.
+\emph{Time spent not hashing = Total time - Time spent hashing}. \emph{Fraction}
+displays the part of total time the calls to the hashing libraries take.
 

### 2010-12-15 paper/main.tex
Typos

--- 2010-12-15 paper/main.tex
@@
 
-The numbers for uploading and downloading the different file vectors are
+The numbers for uploading and downloading the different test vectors are
 summarized in Tables \ref{tbl:hashingtimes:put1gb},
@@
 Because we write to two files after each call to {\verb digest }, an overhead in
-the measurements that would not normally exist in Tahoe-\ac{LAFS}, will be
-created. This overhead should be about the same for all candidates, since the
+the measurements, that would not normally exist in Tahoe-\ac{LAFS}, will be
+introduced. This overhead should be about the same for all candidates, since the
 number of writes are equal. The results and calculations for this overhead,
@@
 implementations which fail completely as hash functions, see
-\autoref{sec:chosen_implementations} for details on the JH implementation.
+\autoref{sec:chosen_implementations} for details on the JH implementations.
 
@@
 \ac{SUPERCOP} will have to balance his use of resources, among them registers,
-to achieve maximum speed.  The loss of one register will therefore give a
+to achieve maximum speed.  The loss of one register should therefore give a
 performance penalty.
@@
 \ref{tbl:cython:overhead}), however the relative overhead are small
-compared to the time spent in the underlying libraries. The reason for
+compared to the time spent in the underlying C implementation. The reason for
 this is that we are hashing larger sizes of data, especially for the 100MB and
 1GB files (Tables \ref{tbl:tahoe:datauploaded} and
-\ref{tbl:tahoe:datauploaded}). For these larger sizes the overhead caused by
+\ref{tbl:tahoe:datadownloaded}). For these larger sizes the overhead caused by
 using double \ac{SHA}-256 opposed to single is also relatively small and does
@@
 
-If we look primarily at up- and download of the large files, we see that the candidates
+If we look primarily at up- and download of large files, we see that the candidates
 that outperforms or are close to outperforming double \ac{SHA}-256, are BLAKE,
@@
 
-\subsection{Time spent not hashing} A strange part of our result is that the
-time Tahoe-\ac{LAFS} spends not callign the hash function is not static for all
+\subsection{Time spent not hashing}
+
+A strange part of our result is that the
+time Tahoe-\ac{LAFS} spends not calling the hash function is not static for all
 candidates, which we expected.  The results of how much this deviance is
@@
 did for large files, which tells us there is a performance gain by switching
-cryptographic hash function in Tahoe-\ac{LAFS}. \autoref{tbl:tahoe:gain}
+the cryptographic hash function in Tahoe-\ac{LAFS}. \autoref{tbl:tahoe:gain}
 compares the performance gains for BLAKE, which performed the best in our
-measurements, versus \ac{SHA}-256, with the removal of overhead caused by our
+measurements, versus double \ac{SHA}-256, with the removal of overhead caused by our
 measurements. We achieved as much as a 12.35\% gain when uploading the 100MB
@@
 SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd. The time it takes for
-Tahoe-\ac{LAFS} to upload a one gigabyte file, could improve by $5\%$ if the
+Tahoe-\ac{LAFS} to upload a one gigabyte file, could improve by $12,36\%$ if the
 best performing \ac{SHA}-3 candidate, BLAKE, were to be used. One has to bear in

### 2010-12-15 paper/main.tex

@@
 
-The numbers for uploading and downloading the different test vectors are
-summarized in Tables \ref{tbl:hashingtimes:put1gb},
-\ref{tbl:hashingtimes:get1gb}, \ref{tbl:hashingtimes:put100mb},
-\ref{tbl:hashingtimes:get100mb}, \ref{tbl:hashingtimes:put1mb},
-\ref{tbl:hashingtimes:get1mb}, \ref{tbl:hashingtimes:put1kb},
-\ref{tbl:hashingtimes:get1kb}, \ref{tbl:hashingtimes:put1b} and
-\ref{tbl:hashingtimes:get1b}.  The third column contains the time spent doing
-other operations than hashing, i.e. \emph{Time spent not hashing = Total time -
-Time spent hashing}. \emph{Fraction} displays the part of total time the calls
-to the hashing libraries take.
+The numbers for uploading and downloading the different file vectors are
+summarized in Tables \ref{tbl:hashingtimes:put1b}, \ref{tbl:hashingtimes:get1b},
+\ref{tbl:hashingtimes:put1kb}, \ref{tbl:hashingtimes:get1kb},
+\ref{tbl:hashingtimes:put1mb}, \ref{tbl:hashingtimes:get1mb},
+\ref{tbl:hashingtimes:put100mb}, \ref{tbl:hashingtimes:get100mb},
+\ref{tbl:hashingtimes:put1gb} and \ref{tbl:hashingtimes:get1gb}. The third
+column contains the time spent doing other operations than hashing, i.e.
+\emph{Time spent not hashing = Total time - Time spent hashing}. \emph{Fraction}
+displays the part of total time the calls to the hashing libraries take.
 
 \ac{SHA}-256 (single) is not included in the graphs, because the results were
-exactly the same as for the \ac{SHA}-256d used in Tahoe-\ac{LAFS}.
+almost the same as for the double \ac{SHA}-256 used in Tahoe-\ac{LAFS}.
 

### 2010-12-15 paper/main.tex
Made some spaces before \cite

@@
 \subsubsection{NIST evaluation criteria for SHA-3}
-
 \label{sec:lengthextension}
+
 \paragraph{Security.} The most important criterion for the SHA-3
-candidates is security\cite{s_nistround2}. \citet{s_fedreg} lays out the full
+candidates is security \cite{s_nistround2}. \citet{s_fedreg} lays out the full
 description of security criteria, however the most noteworthy criterion
@@
 the SHA-3 candidate should be faster than the functions in the \ac{SHA}-2
-family\cite{s_nistround2}.  Cost is a measure of how much memory an
+family \cite{s_nistround2}.  Cost is a measure of how much memory an
 implementation requires in software, both the implementation itself and the use
@@
 each implementation, in relation to other implementation for the same and other
-candidates can be found at eBASH\cite{s_tahoe_supercop}.
+candidates can be found at eBASH \cite{s_tahoe_supercop}.
 

### 2010-12-15 paper/main.tex
Footnote fix

--- 2010-12-15 paper/main.tex
@@
 candidates and \ac{SHA}-2 functions on a number of different plattforms and
-architectures are available at the webpage of eBACS
-\footnote{\url{http://bench.cr.yp.to/results-sha3.html}}.
+architectures are available at the webpage of eBACS\footnote{\url{http://bench.cr.yp.to/results-sha3.html}}.
 
@@
 
-\paragraph{Sphlib} is a C and Java library made by Thomas Pornin
-\footnote{\url{http://www.saphir2.com/sphlib/}}, which contains all the second
-round \ac{SHA}-3 candidates. Some of the same implementations are also found in
+\paragraph{Sphlib\protect\footnote{\url{http://www.saphir2.com/sphlib/}}}is a C and
+Java library made by Thomas Pornin , which contains all the second round
+\ac{SHA}-3 candidates. Some of the same implementations are also found in
 \ac{SUPERCOP}, while others, like the Cubehash implementation we have used is

### 2010-12-15 utils/calc_sha256d.py
Small util file

@@ -1,0 +1,46 @@
+sha256d = {
+    'long':15.14,
+    4096:15.83,
+    1563:16.45,
+    576:18.05,
+    64:38.12,
+    8:179.62,
+}
+
+def calc_sha256d(worst):
+    d = 38.12
+    if worst:
+        d = 179.62
+    if worst==2:
+        d = interpol32()
+
+    print "4096:\t"+  calc1(sha256d[4096],4096,d)
+    print "1563:\t"+  calc1(sha256d[1563],1563,d)
+    print "576:\t" +  calc1(sha256d[576],576,d)
+    print "64:\t"  +  calc1(sha256d[64],64,d)
+    print "8:\t"   +  calc1(sha256d[8],8,d)
+
+def interpol32():
+    k = 64 - 8
+    r = 179.62 - 38.12
+    f = float(r / k)
+    s = 179.62 - 24*f
+    return s
+
+
+def calc1(orgcb, length, extra):
+    k = orgcb * length + extra * 32
+    r = float(k/length)
+    return "%0.3f" % r
+
+
+print "worst case"
+calc_sha256d(True)
+
+print "\nbest case"
+calc_sha256d(False)
+
+print "\ninterpol case"
+calc_sha256d(2)
+
+print interpol32()

### 2010-12-15 paper/title.tex

@@
 Deadline: & 2010-12-20 \\
-Submission date: & 2010-12-xx \\
+Submission date: & 2010-12-16 \\
 Department: & Department of Telematics \\

### 2010-12-15 test-suite/dncornholio.sh
Last version of scripts

@@ -1,0 +1,11 @@
+for i in `seq 1 10`;
+do
+    folder=run-sha256d-multi-$i
+    mkdir $folder
+    ./god_mode.sh > $folder/foo.out 2>$folder/foo.err
+    for k in sha256 blake bmw cubehash echo fugue groestl hamsi \
+        jh keccak luffa shabal shavite3 simd skein;
+        do
+            mv $k/ $folder/
+        done
+done

### 2010-12-15 test-suite/god_mode.sh

@@ -1,4 +1,3 @@
-
 # For hver funksjon
 #   Oppdatere alle klienter
 #   Skifte til riktig hash-function
@@ -6,14 +5,18 @@
 #   Starte testing
 
 mkdir logs
-for i in blake sha256 bmw  cubehash  echo  fugue  groestl  hamsi \
-    jh  keccak  luffa  shabal  shavite3  simd  skein;
+for i in sha256 blake bmw cubehash echo  fugue  groestl  hamsi \
+   jh  keccak  luffa  shabal  shavite3  simd  skein;
+#for i in blake;
+#for i in sha256
 do
     cd /home/eiriha/tahoe-svn/scripts/
     ./do_it_all.sh $i
     cd /home/eiriha/Tahoe-LAFS-with-SHA-3/test-suite/
-    ./test_runner.sh
+    time=`date +%s.%N`
+    time ./test_runner.sh
+    time2=`date +%s.%N`
     mkdir $i
     mv logs/*.txt $i/
+    echo $time2 - $time | bc
 done
-

### 2010-12-15 test-suite/test_runner.sh

@@ -5,11 +5,13 @@
 
 mkdir ./logs
 file_types=(1b 1kb 1mb 100mb 1gb)
+#file_types=(1gb)
 
 for ftype in ${file_types[*]}; do
     mkdir /tmp/$ftype
     echo '0' > /tmp/time_spent_hashing
     echo '0' > /tmp/number_of_hashops
+    rm /tmp/data
 
     files=`ls $ftype*`
     total=`ls $ftype* | wc -l`
@@ -28,6 +30,7 @@ for ftype in ${file_types[*]}; do
     cp /tmp/number_of_hashops ./logs/$ftype\_nohashopsp.txt
     echo '0' > /tmp/time_spent_hashing
     echo '0' > /tmp/number_of_hashops
+    rm /tmp/data
 
     time1=`date +%s.%N`
     counter=1
